### 2022-12-09 README.Rmd
project generated by rrtools

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# minTEMU

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/UD3-Lab/minTEMU/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2022-12-12 README.Rmd
removed docekr files
Add title to README

@@
 # Please put your title here to include it in the file below.
-Title <- "Title of your paper goes here"
+Title <- "The Legacy of the European Postgraduate Master in Urbanism at TU Delft: A Text Mining Approach"
 ```
 
-# minTEMU
+# minTEMU - The Legacy of the European Postgraduate Master in Urbanism at TU Delft: A Text Mining Approach
 
@@
 You can download the compendium as a zip from from this URL:
-[master.zip](/archive/master.zip). After unzipping:
+[main.zip](/archive/main.zip). After unzipping:
 - open the `.Rproj` file in RStudio

### 2022-12-13 analysis/paper/paper.qmd
adapted DESCRIPTION and built the documentation for package with roxygen2

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r}
#| label: setup
library(minTEMU)

```

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

Figure @fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2022-12-13 R/convert.R

@@ -1,0 +1,33 @@
+#' Convert pdf to text
+#'
+#' Function returns a list of files from the directory chosen by the user and to be converted.
+#' Checks if the provided file path exists and whether it holds the adequate file formats (pdf).
+#'
+#' @param pdf_filename_vector  Vector of paths to input (pdf) files
+#' @return A vector of body of text one string per file
+#' @export
+convert_pdf_text <-function(pdf_filename_vector){
+  pdf_list<- as.list(pdf_filename_vector)
+  text_list <- convert_pdf(pdf_list)
+
+  # Activate the virtual environment for python
+  reticulate::use_virtualenv(here::here("R","pyvenv"))
+
+  # Load python script with functions to extract texts
+  reticulate::source_python(here::here("R","extract-text-pdf-python.py"))
+
+  text_list <- convert_pdf(pdf_list) %>% unlist()
+
+  test_list
+
+}
+
+
+# read csv file with pages that are useful for the analysis
+# emu_theses <- read_csv(here("data", "raw", "emu-theses.csv")) %>%
+#   mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
+#   mutate(text = "")
+# pdf_list<- as.list( here("data", "raw", emu_theses$"pdf_filename") )
+
+
+

### 2022-12-14 analysis/paper/paper.qmd
fixing description

@@
 #| label: setup
-library(minTEMU)
+library(dplyr)
+library(readr)
+devtools::load_all(".")
+
+# read csv file with pages that are useful for the analysis
+emu_theses <- read_csv(here::here("analysis","data", "raw_data", "emu-theses.csv")) %>%
+  mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
+  mutate(text = "")
+
+pdf_paths<- here::here("analysis","data", "raw_data", emu_theses$pdf_filename)
+
+emu_theses$text <- convert_pdf_text(pdf_paths)
 

### 2022-12-14 R/convert.R

@@ -3,12 +3,11 @@
 #' Function returns a list of files from the directory chosen by the user and to be converted.
 #' Checks if the provided file path exists and whether it holds the adequate file formats (pdf).
 #'
-#' @param pdf_filename_vector  Vector of paths to input (pdf) files
-#' @return A vector of body of text one string per file
+#' @param pdf_filenames  A vector of paths to input (pdf) files
+#' @return A vector of body of text; one string per pdf file
 #' @export
-convert_pdf_text <-function(pdf_filename_vector){
-  pdf_list<- as.list(pdf_filename_vector)
-  text_list <- convert_pdf(pdf_list)
+convert_pdf_text <-function(pdf_filenames){
+  pdf_list<- as.list(pdf_filenames)
 
   # Activate the virtual environment for python
   reticulate::use_virtualenv(here::here("R","pyvenv"))
@@ -18,16 +17,12 @@ convert_pdf_text <-function(pdf_filename_vector){
 
   text_list <- convert_pdf(pdf_list) %>% unlist()
 
-  test_list
+  text_list
 
 }
 
 
-# read csv file with pages that are useful for the analysis
-# emu_theses <- read_csv(here("data", "raw", "emu-theses.csv")) %>%
-#   mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
-#   mutate(text = "")
-# pdf_list<- as.list( here("data", "raw", emu_theses$"pdf_filename") )
+
 
 
 

### 2023-01-06 analysis/paper/paper.qmd
add cleaning files

@@
 
+# clean text
+  emu_theses$text_clean <- emu_theses$text %>%
+    str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+    str_squish() %>%
+    str_replace_all("- ", "") %>%       ### remove all hyphens
+    str_replace_all("[0-9]", "") %>%    ### remove all numbers
+    str_replace_all(emu_theses$author_firstname[i], "") %>%   ### remove author first name
+    str_replace_all(emu_theses$author_surname[i], "") %>%     ### remove author surname
+    str_replace_all(emu_theses$title[i], "") %>%              ### remove title
+    # str_replace_all("Claudiu Forgaci", "") %>%
+    # str_replace_all("Bucharest: Between North and South", "") %>%
+    str_replace_all("EMU", "")          ### remove common words
+
 ```

### 2023-01-06 analysis/paper/paper.qmd
adapted text cleaning

--- 2022-12-14 analysis/paper/paper.qmd
@@
 library(readr)
+library(stringr)
 devtools::load_all(".")
@@
 emu_theses$text <- convert_pdf_text(pdf_paths)
+
+# clean text
+  emu_theses$text_clean <- emu_theses$text %>%
+    str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+    str_squish() %>%
+    str_replace_all("- ", "") %>%       ### remove all hyphens
+    str_replace_all("[0-9]", "") %>%    ### remove all numbers
+    str_replace_all(emu_theses$author_firstname, "") %>%   ### remove author first name
+    str_replace_all(emu_theses$author_surname, "") %>%     ### remove author surname
+    str_replace_all(emu_theses$title, "") %>%              ### remove title
+    # str_replace_all("Claudiu Forgaci", "") %>%
+    # str_replace_all("Bucharest: Between North and South", "") %>%
+    str_replace_all("EMU", "")          ### remove common words
 

### 2023-01-09 analysis/paper/paper.qmd
adapted text cleaning
Update papeer metadata

--- 2022-12-14 analysis/paper/paper.qmd
@@
 ---
-title: "Title Goes Here"
+title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
 author:
@@
       orcid: 0000-0003-1689-0557
+      institute: tud
+  - John Q. Doe:
+      email: johndoe@fosg.org
+      orcid: 0000-0003-1689-0558
+      institute: tud
 institute:
-        - fosg
-        - fop
-  - John Q. Doe:
-      institute: fosg
-      orcid: 0000-0003-1689-0558
-  - Peder Ås:
-      institute: fosg
-      orcid: 0000-0003-1689-0559
-  - Juan Pérez:
-      orcid: 0000-0003-1689-0551
-      institute:
-        - name: Acme Corporation
-  - Max Mustermann:
-      orcid: 0000-0003-1689-0552
-institute:
-  - fosg:
-      name: Formatting Open Science Group
-      address: 23 Science Street, Eureka, Mississippi, USA
-  - fop: Federation of Planets
+  - tud:
+      name: Delft University of Technology
+      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
 title-block-published: "Last updated"  
@@
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  urbanism; education; post-master; text mining
 highlights: |
@@
 ---
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
 
@@
 #| label: setup
-library(dplyr)
-library(readr)
-devtools::load_all(".")
-
-# read csv file with pages that are useful for the analysis
-emu_theses <- read_csv(here::here("analysis","data", "raw_data", "emu-theses.csv")) %>%
-  mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
-  mutate(text = "")
-
-pdf_paths<- here::here("analysis","data", "raw_data", emu_theses$pdf_filename)
-
-emu_theses$text <- convert_pdf_text(pdf_paths)
-
+#| echo: false
+library(tidyverse)
+library(tmap)
 ```
@@
 
-Here is a citation [@Marwick2017]
+The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft.
+<!-- Complete paragraph about the program -->
 
-# Background
+In order to describe the legacy of the EMU program, including the distinctive features of the didactics employed in it, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
+To that end, we employed a text mining approach in which we analysed its output: 97 theses of 100+ pages each produced over the years for the entire duration of the program between 2007-2021.
 
@@
 
+## Data collection
+The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
+
+## Data analysis
+
+[@silge2017a]
+[@hvitfeldt2021a]
+
 # Results
+
+```{r}
+#| label: map-theses
+#| warning: false
+
+# Create dummy data
+location_names <- as.data.frame(c("Paris, France", "Bucharest, Romania", "Delft, The Netherlands"))
+
+# Geocode
+
+# Visualise thesis locations on world map
+world <- map_data("world")
+ggplot() +
+  geom_map(
+    data = world, map = world,
+    aes(long, lat, map_id = region),
+    color = "black", fill = "lightgray", size = 0.1
+  )
+```
 

### 2023-01-09 analysis/paper/paper.qmd

@@
 #| label: setup
-#| echo: false
-library(tidyverse)
-library(tmap)
+library(dplyr)
+library(readr)
+library(stringr)
+devtools::load_all(".")
+
+# read csv file with pages that are useful for the analysis
+emu_theses <- read_csv(here::here("analysis","data", "raw_data", "emu-theses.csv")) %>%
+  mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
+  mutate(text = "")
+
+pdf_paths<- here::here("analysis","data", "raw_data", emu_theses$pdf_filename)
+
+emu_theses$text <- convert_pdf_text(pdf_paths)
+
+# clean text 
+  emu_theses$text_clean <- emu_theses$text %>%
+    str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+    str_squish() %>%
+    str_replace_all("- ", "") %>%       ### remove all hyphens
+    str_replace_all("[0-9]", "") %>%    ### remove all numbers
+    str_replace_all(emu_theses$author_firstname, "") %>%   ### remove author first name
+    str_replace_all(emu_theses$author_surname, "") %>%     ### remove author surname
+    str_replace_all(emu_theses$title, "") %>%              ### remove title
+    # str_replace_all("Claudiu Forgaci", "") %>%
+    # str_replace_all("Bucharest: Between North and South", "") %>%
+    str_replace_all("EMU", "")          ### remove common words
 ```

### 2023-01-09 R/convert.R
Add code to create virtual environment if not present

@@ -9,6 +9,14 @@
 convert_pdf_text <-function(pdf_filenames){
   pdf_list<- as.list(pdf_filenames)
 
+  # Create virtual environment
+  if (!file.exists(here::here("R","pyvenv"))) {
+    reticulate::install_python(version = '3.9.7')
+    reticulate::virtualenv_create(envname = here::here("R","pyvenv"),
+                                  version = "3.9.7")
+    reticulate::py_install("PyMuPDF==1.21.0", envname = here::here("R","pyvenv"))
+  }
+
   # Activate the virtual environment for python
   reticulate::use_virtualenv(here::here("R","pyvenv"))
 
@@ -18,7 +26,6 @@ convert_pdf_text <-function(pdf_filenames){
   text_list <- convert_pdf(pdf_list) %>% unlist()
 
   text_list
-
 }
 
 

### 2023-01-09 R/clean.R
data cleaning functions

@@ -1,0 +1,70 @@
+#' Clean body of text
+#'
+#' Function removes white space, punctuation and  converts the strings to lowercase
+#'
+#' @param text A vector string to be converted
+#' @return A cleaned vector string
+#' @export
+clean_basic <-function(text){
+  out_text <- text  %>%
+    tolower() %>%
+    str_replace_all("\\s"," ") %>%
+    str_remove_all("[^0-9a-z ]") %>%
+    str_squish()
+  out_text
+
+}
+
+
+#' Find custom stop-phrases based on metadata file
+#'
+#' Function scans the dataset with metadata for fields that can be removed from the body fo text before the text analysis
+#' @param metadata A data frame with metadata information about the text
+#' @param stop_cols  Names of columns that should be used for extraction of the custom stop words
+#' @return A list of strings to be removed for each of the texts
+#' @export
+find_meta_stop <- function(metadata, stop_cols =  c('author_firstname', 'author_surname','title') ){
+
+  cols_meta <- names(metadata)
+
+  cols_meta <- cols_meta[cols_meta %in% stop_cols]
+
+  if (length(cols_meta) == 0 ) stop("The dataset has not stop word columns")
+
+
+
+}
+
+str_replace_all(emu_theses$author_firstname, "") %>%   ### remove author first name
+  str_replace_all(emu_theses$author_firstname, "") %>%     ### remove author surname
+  str_replace_all(emu_theses$title, "") %>%
+
+
+
+# 3. custom stopwords
+
+
+# custom stop words
+stop_words_custom <- emu_theses %>%
+  unnest_tokens(output = word, input = title) %>%   ### add the name of the author and words from the title
+  select("word") %>%
+  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis"))) %>%
+  rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>%
+  rbind(., data.frame(word = c("hab", "km")))
+
+
+# remove white space and punctuation
+
+
+# clean text
+emu_theses$text_clean <- emu_theses$text %>%
+  str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+  str_squish() %>%
+  str_replace_all("- ", "") %>%       ### remove all hyphens
+  str_replace_all("[0-9]", "") %>%    ### remove all numbers
+  str_replace_all(emu_theses$author_firstname, "") %>%   ### remove author first name
[…]

### 2023-01-09 analysis/supplementary-materials/create_rrtools.R

@@ -2,6 +2,6 @@ usethis::use_mit_license(copyright_holder = "Claudiu Forgaci")
 rrtools::use_readme_rmd()
 rrtools::use_analysis()
 renv::init()
-rrtools::use_dockerfile()
+# rrtools::use_dockerfile()
 rrtools::add_dependencies_to_description()
 devtools::document()

### 2023-01-11 analysis/supplementary-materials/merge-metadata-files.R
Update metadata script

@@ -12,8 +12,8 @@ theses_all <-
          -pdf_exists,
          -`Current Employmment`) %>%
   mutate(file_name = ifelse(is.na(last_name_2),
-                            paste0(grad_year, "-", last_name),
-                            paste0(grad_year, "-", last_name, "-", last_name_2)),
+                            paste0(grad_year, "-", last_name, ".pdf"),
+                            paste0(grad_year, "-", last_name, "-", last_name_2, ".pdf")),
          last_name_split = str_split(last_name, "-|\\s")) %>%
   unnest(last_name_split) %>%
   mutate(last_name_split_year = str_to_lower(paste0(last_name_split, "-", grad_year)))

### 2023-01-11 analysis/paper/paper.qmd
add filtering for files
Update indentation in setup code
data cleaning functions
Add existing script to the paper

--- 2023-01-09 analysis/paper/paper.qmd
@@
 #| label: setup
-#| echo: false
-library(tidyverse)
-library(tmap)
+
+# Load packages ----
+library(here)
+library(dplyr)
+library(readr)
+library(stringr)
+library(tidytext)
+library(pluralize)
+library(tm)
+devtools::load_all(".")
+
+# Read thesis metadata ----
+emu_theses <- read_csv(here::here("analysis","data", "raw_data", "theses-metadata.csv")) %>%
+  mutate(text = "")
+
+pdf_names <- dir(here::here("analysis", "data", "raw_data"), pattern = "*.pdf")
+
+emu_theses <- emu_theses %>%
+  filter(file_name %in% pdf_names)
+
+pdf_paths <- here::here("analysis","data", "raw_data", emu_theses$file_name)
+
+emu_theses$text <- convert_pdf_text(pdf_paths)
+
+# clean text 
+emu_theses$text_clean <- emu_theses$text %>%
+  str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+  str_replace_all("- ", "") %>%       ### remove all hyphens
+  str_replace_all("[0-9]", "") %>%    ### remove all numbers
+  str_replace_all(emu_theses$first_name, "") %>%   ### remove author first name
+  str_replace_all(emu_theses$last_name, "") %>%     ### remove author surname
+  str_replace_all(emu_theses$title, "") %>%              ### remove title
+  str_replace_all("EMU", "") %>%           ### remove common words
+  str_squish()
 ```
@@
   )
+```
+
+Stop words:
+- Custom stop words from titles and author names should only be removed from corresponding theses;
+- Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words;
+- All theses were written in English; for common English stop words the "snowball" lexicon was used.
+
+```{r}
+stop_words_custom <- emu_theses %>% 
+  unnest_tokens(output = word, input = title) %>%   ### add words from the title
+  select(word) %>% 
+  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis"))) %>% 
+  rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
+  rbind(., data.frame(word = c("hab", "km")))
+
+emu_theses_words <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") %>%  # remove stop words
+  anti_join(stop_words_custom, by = "word")
+```
+
+```{r}
+# find the most frequently used words
+emu_theses_top <- emu_theses_words %>%
+  group_by(title) %>%
+  count(word, sort = TRUE) %>%
+  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+  filter(n > 50) %>%              ### filter words with a minimum count
+  ungroup() %>% 
+  mutate(word = reorder(word, n))
+
+emu_theses_top[emu_theses_top$word == "ring",]
+
+emu_theses_top %>%
+  ggplot(aes(x = word, y = n, fill = title)) +
+  geom_col() +
+  xlab(NULL) +
+  coord_flip() +
+  labs(y = "Count",
+       x = "Unique words",
+       title = "Count of unique words found in thesis",
+       subtitle = "Stop words were removed from the list") +
+  facet_wrap( ~ title, scales = "free", ncol = 2)
+
+# # wip function ngrams
+# get_ngrams <- function(df, n) {
+#   if (!is.numeric(n)) {
+#     stop("n is not numeric")
+#   }
+#   emu_theses %>%
+#     mutate(author = paste(first_name, last_name)) %>% 
+#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
+#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+#     anti_join(stop_words, by = c("first" = "word")) %>%
+#     anti_join(stop_words, by = c("second" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
+#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+#     filter(first != second) %>% 
+#     filter(first != "source")
+# }
+# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
+
+# bigrams
+emu_theses_bigrams <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+  anti_join(stop_words, by = c("first" = "word")) %>%
+  anti_join(stop_words, by = c("second" = "word")) %>%
+  anti_join(stop_words_custom, by = c("first" = "word")) %>%
+  anti_join(stop_words_custom, by = c("second" = "word")) %>%
+  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+  filter(first != second) %>% 
+  filter(first != "source")
+
+emu_theses_bigrams_top <- emu_theses_bigrams %>%
+  group_by(title) %>%
+  count(bigram, sort = TRUE) %>%
+  slice_max(n, n = 10) %>%        ### top_n is superseded; better use slice_max instead
+  # filter(n > 50)                ### filter words with a minimum count
+  ungroup() %>% 
+  mutate(bigram = reorder(bigram, n))
+
+emu_theses_bigrams_top %>%
+  ggplot(aes(x = bigram, y = n, fill = title)) +
+  geom_col() +
+  xlab(NULL) +
+  coord_flip() +
+  labs(y = "Count",
+       x = "Unique words",
+       title = "Count of unique words found in thesis",
+       subtitle = "Stop words were removed from the list") +
+  facet_wrap( ~ title, scales = "free", ncol = 2)
+
+# find the most frequently used bigrams
+emu_thesis_bigrams_top <- emu_theses_bigrams %>%
+  group_by(title) %>% 
+  count(bigram, sort = TRUE) %>%
+  slice_max(n, n = 10) %>%                   ### change the number of top words to show
+  mutate(bigram = reorder(bigram, n)) %>% 
+  print()
 ```

### 2023-01-12 analysis/paper/paper.qmd
Update setup code in paper

@@
 # Load packages ----
+## Package for managing paths
 library(here)
-library(dplyr)
-library(readr)
-library(stringr)
+
+## Metapackage for data science
+library(tidyverse)
+
+## Packages for text processing and analysis
 library(tidytext)
-library(pluralize)
+library(SnowballC)
 library(tm)
+
+# Load analysis functions ----
 devtools::load_all(".")
@@
 # Read thesis metadata ----
-emu_theses <- read_csv(here::here("analysis","data", "raw_data", "theses-metadata.csv")) %>%
-  mutate(text = "")
+data_path <- here("analysis", "data", "raw_data")
+pdf_names <- dir(path, pattern = "*.pdf")
 
-pdf_names <- dir(here::here("analysis", "data", "raw_data"), pattern = "*.pdf")
+emu_theses <- 
+  read_csv(here(data_path, "theses-metadata.csv")) %>%
+  mutate(text = "") %>%
+  filter(!is.na(pdf_via))
 
-emu_theses <- emu_theses %>%
-  filter(file_name %in% pdf_names)
+pdf_paths <- here(data_path, emu_theses$file_name)
+```
 
-pdf_paths <- here::here("analysis","data", "raw_data", emu_theses$file_name)
+```{r}
+#| label: extract-pdf-text
+emu_theses$text <- convert_pdf_text(pdf_paths).
+```
 
-emu_theses$text <- convert_pdf_text(pdf_paths)
+```{r}
+#| label: clean-text
 
@@
 
+Word-level tokenisation was applied with the `unnest_tokens()` function of the 
+`tidytext` package. This was preferred over other tokenisers as it produced 
+results that work seamlessly with other `tidyverse` tools.
+
 Stop words:
-- Custom stop words from titles and author names should only be removed from corresponding theses;
-- Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words;
-- All theses were written in English; for common English stop words the "snowball" lexicon was used.
+- Three types of stop words: global, subject-specific and document stop words.
+- Global stop words:
+  - All theses were written in English, so for common English stop words the "snowball" lexicon was used
+  - Optionally, try creating a custom stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words
+- Subject-specific stop words:
+  - Manually constructed list, needs domain knowledge
+  - Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words
+- Document stop words:
+  - Stop words from titles and author names should only be removed from corresponding theses
 
 ```{r}
+#| label: stop-words
+
 stop_words_custom <- emu_theses %>% 
@@
   select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis"))) %>% 
-  rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
+  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
+                               "source", "author"))) %>% 
+  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
   rbind(., data.frame(word = c("hab", "km")))
@@
 
+Stemming is a pre-processing step that needs to be carefully thought through, as 
+it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
+
 ```{r}
-# find the most frequently used words
-emu_theses_top <- emu_theses_words %>%
+#| label: stemming
+
+emu_theses_words %>% 
+  mutate(stem = wordStem(words = word, 
+                         language = "porter")) %>%  # Implement the Porter stemming algorithm for English provided by the SnowballC package
+  count(stem, sort = TRUE)
+```
+
+```{r}
+# Which are the top 20 most frequently used words in each thesis?
+emu_theses_top_20 <- emu_theses_words %>%
   group_by(title) %>%
@@
   slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  filter(n > 50) %>%              ### filter words with a minimum count
+  # filter(n > 50) %>%              ### filter words with a minimum count
   ungroup() %>% 
@@
 
-emu_theses_top[emu_theses_top$word == "ring",]
+# Which are first 5 most frequently used words in each thesis?
+emu_theses_first_5 <- emu_theses_words %>%
+  count(title, word) %>% 
+  group_by(title) %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5)
 
-emu_theses_top %>%
-  ggplot(aes(x = word, y = n, fill = title)) +
+# Which theses contain "fabric" in the top 20 most frequently used words?
+emu_theses_top_20 %>% 
+  filter(word == "fabric")
+
+emu_theses_top_20 %>%
+  slice(1:40) %>% 
+  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
   geom_col() +

### 2023-01-23 analysis/paper/paper.qmd

@@
 #| label: setup
+#| echo: false
+#| warning: false
 
 # Load packages ----
-## Package for managing paths
+## Managing paths
 library(here)
 
-## Metapackage for data science
+## Data manipulation and visualisation
 library(tidyverse)
 
-## Packages for text processing and analysis
+## Text processing and analysis
 library(tidytext)
@@
 library(tm)
+
+## Working with spatial data
+library(tidygeocoder)
 
@@
 # Read thesis metadata ----
+## All theses
+theses_all <-
+  readxl::read_xlsx(path = here("analysis", "data", "raw_data", "theses-all.xlsx"))
+
+## Theses with PDF available
 data_path <- here("analysis", "data", "raw_data")
-pdf_names <- dir(path, pattern = "*.pdf")
+pdf_names <- dir(data_path, pattern = "*.pdf")
 
@@
 #| label: extract-pdf-text
-emu_theses$text <- convert_pdf_text(pdf_paths).
+#| echo: false
+emu_theses$text <- convert_pdf_text(pdf_paths)
 ```
@@
 #| label: clean-text
+#| echo: false
 
@@
   str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
-  str_replace_all("- ", "") %>%       ### remove all hyphens
-  str_replace_all("[0-9]", "") %>%    ### remove all numbers
+  # str_replace_all("- ", "") %>%       ### remove all hyphens
+  # str_replace_all("[0-9]", "") %>%    ### remove all numbers
   str_replace_all(emu_theses$first_name, "") %>%   ### remove author first name
@@
 
+```{r}
+#| label: preview
+#| include: false
+
+# Preview theses
+emu_theses %>% 
+  arrange(graduation_year) %>% 
+  pull(text_clean) %>% 
+  str_sub(1, 500)
+
+# Preview theses
+hist(str_length(emu_theses$text_clean))
+
+Encoding(emu_theses$text_clean)
+```
+
 # Introduction
@@
 The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft.
+The EMU was an advanced master ran jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia.
 <!-- Complete paragraph about the program -->
 
-In order to describe the legacy of the EMU program, including the distinctive features of the didactics employed in it, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: 97 theses of 100+ pages each produced over the years for the entire duration of the program between 2007-2021.
+In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(theses_all)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(theses_all$grad_year, na.rm = TRUE), "-", max(theses_all$grad_year, na.rm = TRUE))`.
+The first year of the program was not represented, as PDF files were only available from 2008 onward.
 
@@
 
-## Data analysis
-
-[@silge2017a]
-[@hvitfeldt2021a]
-
-# Results
+A total number of `r nrow(emu_theses)` were analysed.
 
 ```{r}
-#| label: map-theses
+#| label: fig-map-theses
 #| warning: false 
+#| echo: false
+#| fig-cap: "Location of EMU theses included in the analysis"
 
-# Create dummy data
-location_names <- as.data.frame(c("Paris, France", "Bucharest, Romania", "Delft, The Netherlands"))
+# Geocode thesis locations
+emu_theses <- emu_theses %>%
+  geocode(location, method = 'osm', lat = latitude , long = longitude)
 
-# Geocode
-
-# Visualise thesis locations on world map
+# Show thesis locations on world map
 world <- map_data("world")
@@
     color = "black", fill = "lightgray", size = 0.1
-  )
+  ) +
+  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
+  coord_fixed() +
+  theme_void()
 ```
 
+Figure @fig-map-theses shows the geographic spread of the theses included in the analysis.
+
+## Data analysis
+
+[@silge2017a]
+[@hvitfeldt2021a]
+
+### Tokenisation
 Word-level tokenisation was applied with the `unnest_tokens()` function of the 
@@
 
-Stop words:
+### Stop words
 - Three types of stop words: global, subject-specific and document stop words.
@@
 ```
+
+# Results
 
@@
 
-<!-- Here's some example analysis code: -->
-
-```{r}
-#| label: get-data
-#| eval: false
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
-```
-
-```{r}
-#| label: fig-demo-plot
-#| fig-cap: "A plot of random numbers"
-plot(rnorm(10))
-```
-
-Figure @fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`
-
-```{r}
-#| label: demo-inline-code
-#| echo: false
-x <- round(pi, 2)
-```
-
-Here is an example of inline code `r x` in the middle of a sentence.
-
 # Discussion

### 2023-01-25 analysis/paper/paper.qmd
finalisezed and applied cleaning functions'

--- 2023-01-11 analysis/paper/paper.qmd
@@
 #| label: setup
-
-# Load packages ----
-library(here)
 library(dplyr)
@@
 library(stringr)
-library(tidytext)
-library(pluralize)
-library(tm)
 devtools::load_all(".")
 
-# Read thesis metadata ----
-emu_theses <- read_csv(here::here("analysis","data", "raw_data", "theses-metadata.csv")) %>%
+# read csv file with pages that are useful for the analysis
+emu_theses <- read_csv(here::here("analysis","data", "raw_data", "emu-theses.csv")) %>%
+  mutate(pdf_filename = paste0(year, "-", author_surname, ".pdf")) %>%
   mutate(text = "")
 
-pdf_names <- dir(here::here("analysis", "data", "raw_data"), pattern = "*.pdf")
+pdf_names <- dir(here::here("analysis","data", "raw_data"), pattern = "*.pdf")
 
 emu_theses <- emu_theses %>%
-  filter(file_name %in% pdf_names)
+  filter(pdf_filename %in% pdf_names)
 
-pdf_paths <- here::here("analysis","data", "raw_data", emu_theses$file_name)
+pdf_paths<- here::here("analysis","data", "raw_data", emu_theses$pdf_filename)
 
@@
 
+#head_text(emu_theses$text)
+
 # clean text 
 emu_theses$text_clean <- emu_theses$text %>%
-  str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
-  str_replace_all("- ", "") %>%       ### remove all hyphens
-  str_replace_all("[0-9]", "") %>%    ### remove all numbers
-  str_replace_all(emu_theses$first_name, "") %>%   ### remove author first name
-  str_replace_all(emu_theses$last_name, "") %>%     ### remove author surname
-  str_replace_all(emu_theses$title, "") %>%              ### remove title
-  str_replace_all("EMU", "") %>%           ### remove common words
-  str_squish()
+  clean_basic() %>%
+  str_remove_all(find_meta_stopwords(emu_theses) )
+
 ```
@@
   )
-```
-
-Stop words:
-- Custom stop words from titles and author names should only be removed from corresponding theses;
-- Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words;
-- All theses were written in English; for common English stop words the "snowball" lexicon was used.
-
-```{r}
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis"))) %>% 
-  rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
-  rbind(., data.frame(word = c("hab", "km")))
-
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
-  anti_join(stop_words_custom, by = "word")
-```
-
-```{r}
-# find the most frequently used words
-emu_theses_top <- emu_theses_words %>%
-  group_by(title) %>%
-  count(word, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  filter(n > 50) %>%              ### filter words with a minimum count
-  ungroup() %>% 
-  mutate(word = reorder(word, n))
-
-emu_theses_top[emu_theses_top$word == "ring",]
-
-emu_theses_top %>%
-  ggplot(aes(x = word, y = n, fill = title)) +
-  geom_col() +
-  xlab(NULL) +
-  coord_flip() +
-  labs(y = "Count",
-       x = "Unique words",
-       title = "Count of unique words found in thesis",
-       subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
-
-# # wip function ngrams
-# get_ngrams <- function(df, n) {
-#   if (!is.numeric(n)) {
-#     stop("n is not numeric")
-#   }
-#   emu_theses %>%
-#     mutate(author = paste(first_name, last_name)) %>% 
-#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
-#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-#     anti_join(stop_words, by = c("first" = "word")) %>%
-#     anti_join(stop_words, by = c("second" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
-#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-#     filter(first != second) %>% 
-#     filter(first != "source")
-# }
-# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
-
-# bigrams
-emu_theses_bigrams <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-  anti_join(stop_words, by = c("first" = "word")) %>%
-  anti_join(stop_words, by = c("second" = "word")) %>%
-  anti_join(stop_words_custom, by = c("first" = "word")) %>%
-  anti_join(stop_words_custom, by = c("second" = "word")) %>%
-  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-  filter(first != second) %>% 
-  filter(first != "source")
-
-emu_theses_bigrams_top <- emu_theses_bigrams %>%
-  group_by(title) %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 10) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
-  ungroup() %>% 
-  mutate(bigram = reorder(bigram, n))
-
-emu_theses_bigrams_top %>%
-  ggplot(aes(x = bigram, y = n, fill = title)) +
-  geom_col() +
-  xlab(NULL) +
-  coord_flip() +
-  labs(y = "Count",
-       x = "Unique words",
-       title = "Count of unique words found in thesis",
-       subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
-
-# find the most frequently used bigrams
-emu_thesis_bigrams_top <- emu_theses_bigrams %>%
-  group_by(title) %>% 
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 10) %>%                   ### change the number of top words to show
-  mutate(bigram = reorder(bigram, n)) %>% 
-  print()
 ```

### 2023-01-25 R/clean.R

@@ -7,6 +7,7 @@
 #' @export
 #'
 clean_basic <- function(text) {
+
   out_text <- text |>
     tolower() |>
     str_replace_all("\\s{1,}", " ") |>
@@ -29,6 +30,7 @@ clean_basic <- function(text) {
 #' @export
 #'
 head_text <- function(text_vector, head_n = 10, sub_start = 1 , sub_end = 500 ){
+
   head_text <- text_vector |>
     head(head_n) |>
     stringr::str_sub(start = sub_start, end = sub_end)
@@ -40,13 +42,17 @@ head_text <- function(text_vector, head_n = 10, sub_start = 1 , sub_end = 500 ){
 #'
 #' Function scans the dataset with metadata for fields that can be removed from the body fo text before the text analysis
 #' @param metadata A data frame with metadata information about the text
-#' @param stop_cols  Names of columns that should be used for extraction of the custom stop words, the vectors need to be named as in the example
-#' @return A list of strings to be removed for each of the texts
+#' @param stop_cols  List of names of columns that should be used for extraction of the custom stop words, the vectors need to be named as in the example
+#' @param convert_to_regex A logical value determining if the final outcome should be a regex input. Default is TRUE
+#' @return A list with strings to be removed for each of the texts
 #' @export
 find_meta_stopwords <- function(metadata,
-                                stop_cols =  list('author1'= c('author_firstname', 'author_surname'),
-                                                  'author2'= c('author_firstname2', 'author_surname2'),
-                                                  'title'  = 'title') ) {
+                                stop_cols =  list(
+                                  'author1' = c('author_firstname', 'author_surname'),
+                                  'author2' = c('author_firstname2', 'author_surname2'),
+                                  'title'  = 'title'
+                                ),
+                                convert_to_regex = TRUE) {
   cols_meta <- names(metadata)
 
   cols_meta <- cols_meta[cols_meta %in% unlist(stop_cols)]
@@ -54,20 +60,32 @@ find_meta_stopwords <- function(metadata,
   if (length(cols_meta) == 0)
     stop("The dataset has no stopword columns")
 
-  used_stop_cols <- lapply(stop_cols, function(x) x[x %in% cols_meta])
-  used_stop_cols <- used_stop_cols[lengths(used_stop_cols) >0 ]
+  used_stop_cols <-
+    lapply(stop_cols, function(x)
[…]

### 2023-01-25 analysis/paper/paper.qmd

--- 2023-01-23 analysis/paper/paper.qmd
@@
 
+#head_text(emu_theses$text)
+
 # clean text 
 emu_theses$text_clean <- emu_theses$text %>%
-  str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
-  # str_replace_all("- ", "") %>%       ### remove all hyphens
-  # str_replace_all("[0-9]", "") %>%    ### remove all numbers
-  str_replace_all(emu_theses$first_name, "") %>%   ### remove author first name
-  str_replace_all(emu_theses$last_name, "") %>%     ### remove author surname
-  str_replace_all(emu_theses$title, "") %>%              ### remove title
-  str_replace_all("EMU", "") %>%           ### remove common words
-  str_squish()
+  clean_basic() %>%
+  str_remove_all(find_meta_stopwords(emu_theses) )
+
 ```

### 2023-02-13 analysis/paper/paper.qmd
Add research questions

--- 2023-01-25 analysis/paper/paper.qmd
@@
 
-pdf_names <- dir(here::here("analysis","data", "raw_data"), pattern = "*.pdf")
+pdf_names <- dir("analysis/data/raw_data", pattern = "*.pdf")
 
@@
 
-#head_text(emu_theses$text)
-
 # clean text 
 emu_theses$text_clean <- emu_theses$text %>%
-  clean_basic() %>%
-  str_remove_all(find_meta_stopwords(emu_theses) )
-
+  str_replace_all("\\s", " ") %>%  ### remove all carriage return and newline characters
+  str_squish() %>%
+  str_replace_all("- ", "") %>%       ### remove all hyphens
+  str_replace_all("[0-9]", "") %>%    ### remove all numbers
+  str_replace_all(emu_theses$author_firstname, "") %>%   ### remove author first name
+  str_replace_all(emu_theses$author_surname, "") %>%     ### remove author surname
+  str_replace_all(emu_theses$title, "") %>%              ### remove title
+  # str_replace_all("Claudiu Forgaci", "") %>%
+  # str_replace_all("Bucharest: Between North and South", "") %>%
+  str_replace_all("EMU", "")          ### remove common words
 ```
@@
 The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft.
+The EMU was an advanced master ran jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia.
 <!-- Complete paragraph about the program -->
@@
 
+Research questions:
+
+-   What were the main topics addressed in the EMU program?
+
+-   How did those topics evolve throughout the duration of the program between 2007-2021?
+
+-   To what extent were the topics and evolution thereof influenced by the assignments given to the students throughout their studies leading up to their thesis?
+
+-   To what extent were the topics and evolution thereof influenced by the exchange semester followed by the students?
+
 ## Data collection
+
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
@@
 
+[@silge2017a] [@hvitfeldt2021a]
+
+### Tokenisation
+
+Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
+This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools.
+
+### Stop words
+
+-   Three types of stop words: global, subject-specific and document stop words.
+-   Global stop words:
+    -   All theses were written in English, so for common English stop words the "snowball" lexicon was used
+    -   Optionally, try creating a custom stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words
+-   Subject-specific stop words:
+    -   Manually constructed list, needs domain knowledge
+    -   Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words
+-   Document stop words:
+    -   Stop words from titles and author names should only be removed from corresponding theses
+
 ```{r}
@@
 ```
+
+Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
 

### 2023-02-13 analysis/paper/paper.qmd

--- 2023-01-25 analysis/paper/paper.qmd
@@
 
+[@silge2017a] [@hvitfeldt2021a]
+
 ### Tokenisation
-Word-level tokenisation was applied with the `unnest_tokens()` function of the 
-`tidytext` package. This was preferred over other tokenisers as it produced 
-results that work seamlessly with other `tidyverse` tools.
+
+Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
+This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools.
 
 ### Stop words
+
 -   Three types of stop words: global, subject-specific and document stop words.
@@
 
-Stemming is a pre-processing step that needs to be carefully thought through, as 
-it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
+Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
 

### 2023-02-13 analysis/smp.qmd

---
title: "mintEMU Software Management Plan"
author: "Claudiu Forgaci"
format: pdf
editor: visual
bibliography: paper/references.bib
---

This Software Management Plan uses the guidelines for medium management level as described in the Practical guide to Software Management Plans [@martinez-ortiz2022].

**Please provide a brief description of your software, stating its purpose and intended audience.**

*mintEMU* is a research compendium created as an R package with `rrtools` in preparation of the scientific paper *The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach*. The compendum contains the data, code for data processing and analysis, as well as the paper presenting the results of the analysis. The intended audience is the scientific community interested in the research results, methods employed or data generated in the project.

The project was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft. In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this research aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employ a text mining approach in which we analyse its output: 90+ theses each with 100+ pages produced over the years for the duration of the program between 2007-2021.

The findings will be presented in a journal article, accompanied by a fully reproducible research compendium comprising an open dataset and corresponding software, as well as a dashboard to enable users to interact with the data and explore relations and patterns in the data beyond the findings presented in the paper. The project is carried out with in-kind support from the TU Delft Digital Competence Center.

**How will you manage versioning of your software?**

The software will be versioned with Git and hosted on GitHub to allow for collaborative development. Although the research compendium is not a software that will be versioned after publication, semantic versioning will be used in the process leading up to publication, as follows: development version 0.0.0.9000 in the research stage; patch version 0.0.1 with the publication of the preprint; minor version 0.1.0 and subsequent increments of minor version at the time of submission for review, and responses to reviews, respectively; and 1.0.0 for the published paper.

**How will your software be documented for users? Please provide a link to the documentation if available. How will you document your software's contribution guidelines and governance structure?**

The research compendium will be documented as an R package, that is, all functions used in data processing and analysis will have associated documentation available from the help interface of R. Once loaded as a package all documentation is available to the user. The analysis will be fully documented in a Quarto computational notebook using R Markdown with the knitr engine.

**What licence will you give your software? How will you check that it respects the licences of libraries and dependencies it uses?**

The software will use an MIT license.

**How will the installation requirements of your software be documented? Please provide a link to the installation documentation if available.**

Installation requirements are described in the README file of the repository on GitHub:

**How will users of your software be able to cite your software? Please provide a link to your software citation file (CFF) if available.**

**How will your software be documented for future developers?**

**How will your software be tested? Please provide a link to the (automated) testing results.**

**Do you follow specific software quality guidelines? If yes, which ones?**

FAIR4RS

**How will your software be packaged and distributed? Please provide a link to available packaging information (e.g. entry in a packaging registry, if available).**

The software will be made available as an R package on GitHub, available for direct installation from RStudio. As the package is s research compendium, it will not be made available in a packaging registry.

**How do you plan to procure long term maintenance of your software?**

### 2023-02-14 analysis/paper/paper.qmd
Remove duplicate references
Update script for loading packages

@@
 # Load packages ----
-## Managing paths
-library(here)
+## List packages to be loaded
+packages <- c(
+  "here",                         # Managing paths
+  "tidyverse",                    # Data manipulation and visualisation
+  "tidytext", "SnowballC", "tm",  # Text processing and analysis
+  "tidygeocoder"                  # Working with spatial data
+)
 
-## Data manipulation and visualisation
-library(tidyverse)
-
-## Text processing and analysis
-library(tidytext)
-library(SnowballC)
-library(tm)
-
-## Working with spatial data
-library(tidygeocoder)
+## Load required packages, install them if needed
+for (package in packages) {
+  if (!require(package, character.only = TRUE)) {
+    install.packages(package)
+  }
+  library(package, character.only = TRUE)
+}
 
@@
 [@hvitfeldt2021a]
-
-[@silge2017a] [@hvitfeldt2021a]
 

### 2023-02-14 analysis/paper/paper.qmd
Update comments in data reading code

@@
 # Read thesis metadata ----
-## All theses
-theses_all <-
-  readxl::read_xlsx(path = here("analysis", "data", "raw_data", "theses-all.xlsx"))
-
-## Theses with PDF available
 data_path <- here("analysis", "data", "raw_data")
@@
 
+## Read metadata for all theses
+all_theses <-
+  readxl::read_xlsx(path = here(data_path, "theses-all.xlsx"))
+
+## Read only metadata for theses with PDF available
 emu_theses <- 
@@
 #| echo: false
+
+# Extract text from PDFs
+# On the first run, this function might take extra time as a Python environment needs to be set up
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(theses_all)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(theses_all$grad_year, na.rm = TRUE), "-", max(theses_all$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
 The first year of the program was not represented, as PDF files were only available from 2008 onward.

### 2023-02-14 R/geocode.R
Add function to geocode thesis locations

@@ -1,0 +1,12 @@
+#' Geocode thesis locations
+#'
+#' The function creates two new columns, latitude and longitude,
+#' based on a column containing the address of the thesis.
+#'
+#' @param data Data frame with thesis data
+#' @return Data frame with latitude and longitude columns
+#'
+#' @export
+geocode_thesis_locations <- function(data) {
+  tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
+}

### 2023-02-14 analysis/paper/paper.qmd
Add function to geocode thesis locations
Create function to visualise thesis locations

--- 2023-02-13 analysis/paper/paper.qmd
@@
 
-# Geocode thesis locations
-emu_theses <- emu_theses %>%
-  geocode(location, method = 'osm', lat = latitude , long = longitude)
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses)
 
 # Show thesis locations on world map
-world <- map_data("world")
-ggplot() +
-  geom_map(
-    data = world, map = world,
-    aes(long, lat, map_id = region),
-    color = "black", fill = "lightgray", size = 0.1
-  ) +
-  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
-  coord_fixed() +
-  theme_void()
+visualize_thesis_locations(emu_theses)
 ```

### 2023-02-14 R/geocode.R

@@ -10,3 +10,25 @@
 geocode_thesis_locations <- function(data) {
   tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
 }
+
+
+#' Visualise thesis locations
+#'
+#' Show thesis locations on the map of the world.
+#'
+#' @param data Data frame with thesis data.
+#'
+#' @return A plot with thesis locations
+#' @export
+visualize_thesis_locations <- function(data) {
+  world <- ggplot2::map_data("world")
+  ggplot2::ggplot() +
+    ggplot2::geom_map(
+      data = world, map = world,
+      aes(long, lat, map_id = region),
+      color = "black", fill = "lightgray", size = 0.1
+    ) +
+    ggplot2::geom_point(data = data, aes(longitude, latitude), color = "red") +
+    ggplot2::coord_fixed() +
+    ggplot2::theme_void()
+}

### 2023-02-14 R/geocode.R
Update geocode function
Do not run the geocoder if latitude and longitude columns are already present in the input data frame.

@@ -8,7 +8,10 @@
 #'
 #' @export
 geocode_thesis_locations <- function(data) {
-  tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
+  if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
+    tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
+  }
+  return(data)
 }
 
 

### 2023-02-15 analysis/paper/paper.qmd
Updarte preview code chunk

@@
 
-# Preview theses
+# Preview theses in increasing order of graduation year
 emu_theses %>% 
@@
   pull(text_clean) %>% 
-  str_sub(1, 500)
+  head_text(head_n = 5)
 
-# Preview theses
+# Show distribution of the length of theses
 hist(str_length(emu_theses$text_clean))
-
-Encoding(emu_theses$text_clean)
 ```
@@
 
-# Add latitude and longitude information to theses data frame
-emu_theses <- geocode_thesis_locations(emu_theses)
+# Geocode thesis locations
+emu_theses <- emu_theses %>%
+  geocode(location, method = 'osm', lat = latitude , long = longitude)
 
 # Show thesis locations on world map
-visualize_thesis_locations(emu_theses)
+world <- map_data("world")
+ggplot() +
+  geom_map(
+    data = world, map = world,
+    aes(long, lat, map_id = region),
+    color = "black", fill = "lightgray", size = 0.1
+  ) +
+  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
+  coord_fixed() +
+  theme_void()
 ```

### 2023-02-15 analysis/paper/paper.qmd
Add analysis chunk from old script

@@
 
-# Preview theses in increasing order of graduation year
+# Preview theses
 emu_theses %>% 
@@
   pull(text_clean) %>% 
-  head_text(head_n = 5)
+  str_sub(1, 500)
 
-# Show distribution of the length of theses
+# Preview theses
 hist(str_length(emu_theses$text_clean))
+
+Encoding(emu_theses$text_clean)
 ```
@@
 
+emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
+  count(bigram, sort = TRUE) %>%
+  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+  # filter(n > 50)                ### filter words with a minimum count
+  mutate(bigram = reorder(bigram, n))
+
 emu_theses_bigrams_top %>%

### 2023-02-15 analysis/paper/paper.qmd
Add functions to calculate top n words

@@
 
+Counting frequencies of words per document
 ```{r}
-# Which are the top 20 most frequently used words in each thesis?
-emu_theses_top_20 <- emu_theses_words %>%
-  group_by(title) %>%
-  count(word, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50) %>%              ### filter words with a minimum count
-  ungroup() %>% 
-  mutate(word = reorder(word, n))
+# Which are the top n most frequently used words in each thesis?
+top_n_word_count_per_document(data = emu_theses_words, 
+                              top_n = 3, 
+                              title_col = 'title', 
+                              word_col = 'word')
+```
 
+```{r}
+# Which are the top n most frequently used words in all theses`?
+top_n_word_count_per_corpus(emu_theses_words, 5, "word")
+```
+
+```{r}
 # Which are first 5 most frequently used words in each thesis?
@@
   slice(1:5)
+```
 
+```{r}
 # Which theses contain "fabric" in the top 20 most frequently used words?

### 2023-02-15 R/analyse.R

@@ -1,0 +1,46 @@
+#' Find top n words with the highest count per document
+#'
+#' Given a data frame with a column storing the words found in a document
+#' identified with a title column and a positive number top_n, this function
+#' returns the top n words with the highest count in each document.
+#'
+#' @param data Data frame with theses, containing a column with words.
+#' @param top_n The number of top words to be returned by the function.
+#' @param title_col The name of the column from the input data frame containing the title of the document.
+#' @param word_col Name of the column from the input data frame containing all words.
+#' @param min_count Minimum count to be considered. Words below this value will not be included in the returned data frame.
+#'
+#' @return A data frame with the top n words with a count above the given minimum.
+#' @export
+top_n_word_count_per_document <- function(data, top_n, title_col = NULL, word_col = NULL, min_count = 1) {
+  data_top_n <- data |>
+    dplyr::group_by(!!sym(title_col)) |>
+    dplyr::count(!!sym(word_col), sort = TRUE) |>
+    dplyr::slice_max(n, n = top_n) |>
+    dplyr::filter(n > min_count) |>
+    dplyr::ungroup()
+
+  data_top_n
+}
+
+#' Find top n words with the highest count across all documents
+#'
+#' Given a data frame with a column storing the words found in a document
+#' and a positive number top_n, this function returns the top n words with the
+#' highest count across all documents.
+#'
+#' @param data Data frame with theses, containing a column with words.
+#' @param top_n The number of top words to be returned by the function.
+#' @param word_col Name of the column from the input data frame containing all words.
+#'
+#' @return A data frame with the top n words across all documents.
+#' @export
+top_n_word_count_per_corpus <- function(data, top_n, word_col = NULL) {
+  data_top_n <- data |>
+    dplyr::group_by(!!sym(word_col)) |>
+    dplyr::count(!!sym(word_col), sort = TRUE) |>
+    dplyr::ungroup() |>
+    dplyr::slice_max(n, n = top_n)
+
+  data_top_n
+}

### 2023-02-15 analysis/paper/paper.qmd
Update top n per corpus code chunk
Update top words code chunks

@@
 
-Counting frequencies of words per document
 ```{r}
@@
                               top_n = 3, 
-                              title_col = 'title', 
-                              word_col = 'word')
+                              title_col = "title", 
+                              word_col = "word")
 ```
@@
 # Which are the top n most frequently used words in all theses`?
-top_n_word_count_per_corpus(emu_theses_words, 5, "word")
-```
-
-```{r}
-# Which are first 5 most frequently used words in each thesis?
-emu_theses_first_5 <- emu_theses_words %>%
-  count(title, word) %>% 
-  group_by(title) %>% 
-  arrange(desc(n)) %>% 
-  slice(1:5)
+top_n_word_count_per_corpus(data = emu_theses_words, 
+                            top_n = 5, 
+                            word_col = "word")
 ```

### 2023-02-15 analysis/paper/paper.qmd
Update function names for top words

@@
 # Which are the top n most frequently used words in each thesis?
-top_n_word_count_per_document(data = emu_theses_words, 
+get_top_words_per_document(data = emu_theses_words, 
                            top_n = 3, 
@@
 # Which are the top n most frequently used words in all theses`?
-top_n_word_count_per_corpus(data = emu_theses_words, 
+get_top_words_per_corpus(data = emu_theses_words, 
                          top_n = 5, 

### 2023-02-15 R/analyse.R

@@ -1,4 +1,4 @@
-#' Find top n words with the highest count per document
+#' Extract top n words with the highest count per document
 #'
 #' Given a data frame with a column storing the words found in a document
 #' identified with a title column and a positive number top_n, this function
@@ -12,7 +12,7 @@
 #'
 #' @return A data frame with the top n words with a count above the given minimum.
 #' @export
-top_n_word_count_per_document <- function(data, top_n, title_col = NULL, word_col = NULL, min_count = 1) {
+get_top_words_per_document <- function(data, top_n, title_col = NULL, word_col = NULL, min_count = 1) {
   data_top_n <- data |>
     dplyr::group_by(!!sym(title_col)) |>
     dplyr::count(!!sym(word_col), sort = TRUE) |>
@@ -23,7 +23,7 @@ top_n_word_count_per_document <- function(data, top_n, title_col = NULL, word_co
   data_top_n
 }
 
-#' Find top n words with the highest count across all documents
+#' Extract top n words with the highest count across all documents
 #'
 #' Given a data frame with a column storing the words found in a document
 #' and a positive number top_n, this function returns the top n words with the
@@ -35,7 +35,7 @@ top_n_word_count_per_document <- function(data, top_n, title_col = NULL, word_co
 #'
 #' @return A data frame with the top n words across all documents.
 #' @export
-top_n_word_count_per_corpus <- function(data, top_n, word_col = NULL) {
+get_top_words_per_corpus <- function(data, top_n, word_col = NULL) {
   data_top_n <- data |>
     dplyr::group_by(!!sym(word_col)) |>
     dplyr::count(!!sym(word_col), sort = TRUE) |>

### 2023-02-16 analysis/paper/paper.qmd
Specify the language of the corpus

@@
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
+With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 
@@
 Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
-This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools.
+This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.
 
@@
 ```{r}
-# Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_theses_words, 
-                           top_n = 3, 
-                           title_col = "title", 
-                           word_col = "word")
-```
+# Which are the top 20 most frequently used words in each thesis?
+emu_theses_top_20 <- emu_theses_words %>%
+  group_by(title) %>%
+  count(word, sort = TRUE) %>%
+  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+  # filter(n > 50) %>%              ### filter words with a minimum count
+  ungroup() %>% 
+  mutate(word = reorder(word, n))
 
-```{r}
-# Which are the top n most frequently used words in all theses`?
-get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 5, 
-                         word_col = "word")
-```
+# Which are first 5 most frequently used words in each thesis?
+emu_theses_first_5 <- emu_theses_words %>%
+  count(title, word) %>% 
+  group_by(title) %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5)
 
-```{r}
 # Which theses contain "fabric" in the top 20 most frequently used words?

### 2023-02-16 analysis/paper/paper.qmd
Update description of stop words used

@@
 
--   Three types of stop words: global, subject-specific and document stop words.
--   Global stop words:
-    -   All theses were written in English, so for common English stop words the "snowball" lexicon was used
-    -   Optionally, try creating a custom stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words
--   Subject-specific stop words:
-    -   Manually constructed list, needs domain knowledge
-    -   Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words
--   Document stop words:
-    -   Stop words from titles and author names should only be removed from corresponding theses
+For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage.
+Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
+As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
+<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result.
+--> Subject-specific stop words, we manually constructed a list based on our domain knowledge.
+Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well.
+Finally, titles and author names were removed as document stop words on document-level.
 

### 2023-02-16 analysis/paper/paper.qmd
Update stopwords

--- 2023-02-16 analysis/paper/paper.qmd
@@
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
-With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 
@@
 Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
-This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.
+This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools.
 
@@
 #| label: stop-words
+
+emu_theses |> 
+  find_meta_stopwords(convert_to_regex = FALSE)
+
+emu_theses |> 
+  urbanism_stopwords()
 

### 2023-02-16 R/clean.R

@@ -99,6 +99,7 @@ urbanism_stopwords <- function(add_stopwords = NULL) {
   stop_words <-
     c(add_stopwords, "city", "urban", "urbanism", "hab", "km")
 
+  stop_words
 }
 
 #' Adjustable list of stopwords associated with master theses

### 2023-02-16 R/clean.R
Return stopwords in data frames

@@ -97,7 +97,7 @@ find_meta_stopwords <- function(metadata,
 #' @export
 urbanism_stopwords <- function(add_stopwords = NULL) {
   stop_words <-
-    c(add_stopwords, "city", "urban", "urbanism", "hab", "km")
+    tibble(word = c(add_stopwords, "city", "urban", "urbanism", "hab", "km"))
 
   stop_words
 }
@@ -110,8 +110,9 @@ urbanism_stopwords <- function(add_stopwords = NULL) {
 #' @export
 thesis_stopwords <- function(add_stopwords = NULL) {
   stop_words <-
-    c(add_stopwords, "preface", "foreword", "introduction", "conclusion", "thesis")
+    tibble(word = c(add_stopwords, "preface", "foreword", "introduction", "conclusion", "thesis"))
 
+  stop_words
 }
 
 

### 2023-02-16 analysis/paper/paper.qmd
Update script with stop words functions

@@
 
-emu_theses |> 
-  find_meta_stopwords(convert_to_regex = FALSE)
+meta_stopwords <- find_meta_stopwords(emu_theses, convert_to_regex = FALSE)
+urbanism_stopwords <- urbanism_stopwords() 
+thesis_stopwords <- thesis_stopwords()
 
-emu_theses |> 
-  urbanism_stopwords()
-
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
-                               "source", "author"))) %>% 
-  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
-  rbind(., data.frame(word = c("hab", "km")))
-
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
-  anti_join(stop_words_custom, by = "word")
+emu_theses_words <- emu_theses |> 
+  select(title, text) |>  
+  unnest_tokens(output = word, input = text) |>   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") |> 
+  anti_join(urbanism_stopwords, by = "word") |> 
+  anti_join(thesis_stopwords, by = "word")
 ```

### 2023-02-17 R/clean.R
Update author col names in meta stopwords function

@@ -48,8 +48,8 @@ head_text <- function(text_vector, head_n = 10, sub_start = 1 , sub_end = 500 ){
 #' @export
 find_meta_stopwords <- function(metadata,
                                 stop_cols =  list(
-                                  'author1' = c('author_firstname', 'author_surname'),
-                                  'author2' = c('author_firstname2', 'author_surname2'),
+                                  'author1' = c('first_name', 'last_name'),
+                                  'author2' = c('first_name_2', 'last_name_2'),
                                   'title'  = 'title'
                                 ),
                                 convert_to_regex = TRUE) {

### 2023-02-17 analysis/paper/paper.qmd
Remove stop words from text_clean column

@@
 
-meta_stopwords <- find_meta_stopwords(emu_theses, convert_to_regex = FALSE)
 urbanism_stopwords <- urbanism_stopwords() 
@@
 emu_theses_words <- emu_theses |> 
-  select(title, text) |>  
-  unnest_tokens(output = word, input = text) |>   # remove punctuation, convert to lowercase, seperate all words
+  select(title, text_clean) |> 
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, seperate all words
   anti_join(stop_words, by = "word") |> 

### 2023-02-17 analysis/paper/paper.qmd
Cache text extracted from pdf
Remove unnecessary comments

--- 2023-02-16 analysis/paper/paper.qmd
@@
 #| echo: false
+#| cache:true
+
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
 
-#head_text(emu_theses$text)
-
-# clean text 
 emu_theses$text_clean <- emu_theses$text %>%
@@
   str_remove_all(find_meta_stopwords(emu_theses) )  
-
 ```
@@
 #| label: stop-words
-
-emu_theses |> 
-  find_meta_stopwords(convert_to_regex = FALSE)
-
-emu_theses |> 
-  urbanism_stopwords()
 

### 2023-02-17 analysis/paper/paper.qmd
Make small style corrections
Accept md changes made by the RStudio visual editor
Correct cache code chunk label

@@
 #| echo: false
-#| cache:true
+#| cache: true
 
@@
 
-emu_theses$text_clean <- emu_theses$text %>%
-  clean_basic() %>%
+emu_theses$text_clean <- emu_theses$text |> 
+  clean_basic() |> 
   str_remove_all(find_meta_stopwords(emu_theses) )  
@@
 
-[@silge2017a]
-[@hvitfeldt2021a]
+[@silge2017a] [@hvitfeldt2021a]
 

### 2023-02-17 analysis/paper/paper.qmd
Update text
Start methods section on topic modeling
Add topicmodels package

@@
 library(tm)
+library(topicmodels)
 
@@
 
-A total number of `r nrow(emu_theses)` were analysed.
+A total number of `r nrow(emu_theses)` out of `r nrow(theses_all)` were analysed.
 
@@
 ```
+
+## Topic model
+
+Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
 

### 2023-02-17 analysis/paper/paper.qmd
Add script for tf-idf statistics
The tf-idf is used to show high-frequency words that are document-specific in a larger collection of documents.

@@
 
+```{r}
+#| label: tf-idf
+
+emu_theses_words
+
+emu_theses_words_count <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  ungroup()
+
+emu_theses_words_total <- emu_theses_words_count |> 
+  group_by(title) |> 
+  summarise(total = sum(n))
+
+emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)
+
+# emu_theses_words_count |> 
+#   group_by(title) |> 
+#   summarise(total = unique(tot
+
+# Preview term frequency distributions in a subset of theses
+ggplot(emu_theses_words_count |> 
+         filter(title == unique(title)[1:6]),
+       aes(n/total, fill = title)) +
+  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
+  xlim(NA, 0.0009) +
+  facet_wrap(~ title, ncol = 3, scales = "free_y")
+
+# Examine Zipf's law for the EMU theses
+## Calculate rank and term frequency
+freq_by_rank <- emu_theses_words_count |> 
+  group_by(title) |> 
+  mutate(rank = row_number(),
+         `term frequency` = n/total)
+
+## Visualise Zipf's law
+freq_by_rank |> 
+  ggplot(aes(rank, `term frequency`, color = title)) +
+  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
+  scale_x_log10() +
+  scale_y_log10()
+
+rank_subset <- freq_by_rank |> 
+  filter(rank < 500,
+         rank > 10)
+
+lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)
+
+freq_by_rank |> 
+  ggplot(aes(rank, `term frequency`, color = title)) +
+  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
+  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
+  scale_x_log10() +
+  scale_y_log10()
+
+## Calculate tf-idf
+emu_theses_words_count <- emu_theses_words_count |> 
+  bind_tf_idf(word, title, n)
+emu_theses_words_count
+
+emu_theses_words_count |> 
+  select(-total) |> 
+  arrange(desc(tf_idf))
+
+## Visualise the top n words with the highest tf-idf values for each of a subset of theses
+emu_theses_words_count |> 
+  filter(title == unique(title)[1:6]) |> 
+  arrange(desc(tf_idf)) |> 
+  mutate(word = factor(word, levels = rev(unique(word)))) |> 
+  group_by(title) |> 
+  top_n(15) |> 
+  ungroup() |> 
+  ggplot(aes(word, tf_idf, fill = title)) +
+  geom_col(show.legend = FALSE) +
+  labs(x = NULL, y = "tf-idf") +
+  facet_wrap(~ title,  ncol = 3, scales = "free") +
+  coord_flip()
+```
+
+The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
+As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
+
 # Discussion

### 2023-02-17 analysis/paper/paper.qmd
Update ngrams code chunk

--- 2023-02-17 analysis/paper/paper.qmd
@@
 library(tm)
-library(topicmodels)
 
@@
   facet_wrap( ~ title, scales = "free", ncol = 2)
+```
+
+```{r}
+#| label: ngrams
 
@@
 
-# bigrams
-emu_theses_bigrams <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-  anti_join(stop_words, by = c("first" = "word")) %>%
-  anti_join(stop_words, by = c("second" = "word")) %>%
-  anti_join(stop_words_custom, by = c("first" = "word")) %>%
-  anti_join(stop_words_custom, by = c("second" = "word")) %>%
-  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-  filter(first != second) %>% 
-  filter(first != "source")
+## Bigrams
+emu_theses_bigrams <- emu_theses |> 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2)
+
+emu_theses_bigrams_separated <- emu_theses_bigrams |>  
+  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE)
+
+emu_theses_bigrams_filtered <- emu_theses_bigrams_separated |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word"))
+
+bigram_counts <- emu_theses_bigrams_filtered |> 
+  count(first, second, sort = TRUE)
+
+bigram_counts
+
+  # # filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+  # filter(first != second) %>% 
+  # filter(first != "source")
 

### 2023-02-17 analysis/paper/paper.qmd
Add packages related to networks in the setup code chunk

--- 2023-02-17 analysis/paper/paper.qmd
@@
 # Load packages ----
-## Managing paths
-library(here)
+## List packages to be loaded
+packages <- c(
+  "here",                         # Managing paths
+  "tidyverse",                    # Data manipulation and visualisation
+  "tidytext", "SnowballC", "tm",  # Text processing and analysis
+  "tidygeocoder",                 # Working with spatial data
+  "igraph", "ggraph"              # Network analysis and visualsation
+)
 
-## Data manipulation and visualisation
-library(tidyverse)
-
-## Text processing and analysis
-library(tidytext)
-library(SnowballC)
-library(tm)
-library(topicmodels)
-
-## Working with spatial data
-library(tidygeocoder)
+## Load required packages, install them if needed
+for (package in packages) {
+  if (!require(package, character.only = TRUE)) {
+    install.packages(package)
+  }
+  library(package, character.only = TRUE)
+}
 
@@
 #| echo: false
-#| cache: true
 
+# Extract text from PDFs
+# On the first run, this function might take extra time as a Python environment needs to be set up
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
 
-emu_theses$text_clean <- emu_theses$text |> 
-  clean_basic() |> 
+#head_text(emu_theses$text)
+
+# clean text 
+emu_theses$text_clean <- emu_theses$text %>%
+  clean_basic() %>%
   str_remove_all(find_meta_stopwords(emu_theses) )
+
 ```
@@
 
-A total number of `r nrow(emu_theses)` out of `r nrow(theses_all)` were analysed.
+A total number of `r nrow(emu_theses)` were analysed.
 
@@
 
-[@silge2017a] [@hvitfeldt2021a]
+[@silge2017a]
+[@hvitfeldt2021a]
 
@@
 ```
-
-## Topic model
-
-Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
 
@@
 
-emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
-  mutate(bigram = reorder(bigram, n))
-
 emu_theses_bigrams_top %>%

### 2023-02-17 analysis/paper/paper.qmd
Add script to calculate bigrams and trigrams

--- 2023-02-17 analysis/paper/paper.qmd
@@
 
-# # wip function ngrams
-# get_ngrams <- function(df, n) {
-#   if (!is.numeric(n)) {
-#     stop("n is not numeric")
-#   }
-#   emu_theses %>%
-#     mutate(author = paste(first_name, last_name)) %>% 
-#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
-#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-#     anti_join(stop_words, by = c("first" = "word")) %>%
-#     anti_join(stop_words, by = c("second" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
-#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-#     filter(first != second) %>% 
-#     filter(first != "source")
-# }
-# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
+## TO DO: generalise bigrams and trigrams into an ngrams function
 
@@
 emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2)
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
 
-emu_theses_bigrams_separated <- emu_theses_bigrams |>  
-  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE)
-
-emu_theses_bigrams_filtered <- emu_theses_bigrams_separated |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word"))
-
-bigram_counts <- emu_theses_bigrams_filtered |> 
+bigram_counts <- emu_theses_bigrams |> 
   count(first, second, sort = TRUE)
@@
 
-  # # filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-  # filter(first != second) %>% 
-  # filter(first != "source")
+## Trigrams
+emu_theses_trigrams <- emu_theses |> 
+  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
+  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
+  filter(!first %in% stop_words$word,
+         !second %in% stop_words$word,
+         !third %in% stop_words$word) |>  
+  filter(str_detect(first, "[a-z]"), 
+         str_detect(second, "[a-z]"),
+         str_detect(third, "[a-z]"))
+
+trigram_counts <- emu_theses_trigrams |> 
+  count(first, second, sort = TRUE)
+
+trigram_counts
+
+# Bigram tf-idf
+bigram_tf_idf <- emu_theses_bigrams |> 
+  count(title, bigram) |> 
+  bind_tf_idf(bigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+bigram_tf_idf
+
+# Trigram tf-idf
+trigram_tf_idf <- emu_theses_trigrams |> 
+  count(title, trigram) |> 
+  bind_tf_idf(trigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+trigram_tf_idf
 

### 2023-02-17 analysis/paper/paper.qmd
Add bigram and trigram tf-idf

@@
 
-# Bigram tf-idf
+## Bigram tf-idf
 bigram_tf_idf <- emu_theses_bigrams |> 
@@
 
-# Trigram tf-idf
+## Trigram tf-idf
 trigram_tf_idf <- emu_theses_trigrams |> 
@@
 trigram_tf_idf
+
+## Show network of bigrams
+bigram_graph <- bigram_counts |> 
+  filter(n > 20) |> 
+  igraph::graph_from_data_frame()
 

### 2023-02-17 analysis/paper/paper.qmd
Update ngrams code chunk

@@
 
-## TO DO: generalise bigrams and trigrams into an ngrams function
+## TO DO: generalise bigrams and trigrams into ngrams function
 
@@
 bigram_graph <- bigram_counts |> 
-  filter(n > 20) |> 
-  igraph::graph_from_data_frame()
+  filter(n > 100) |> 
+  graph_from_data_frame()
+
+bigram_graph
+
+set.seed(2023)
+
+a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
+
+ggraph(bigram_graph, layout = "fr") +
+  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
+  geom_node_point(color = "lightblue", size = 5) +
+  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
+  theme_void()
+
+## Higgh frequency bigrams ----
 
@@
   count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 10) %>%        ### top_n is superseded; better use slice_max instead
+  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
@@
 emu_theses_bigrams_top %>%
+  filter(title == unique(title)[1:6]) |> 
   ggplot(aes(x = bigram, y = n, fill = title)) +
@@
        subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
-
-# find the most frequently used bigrams
-emu_thesis_bigrams_top <- emu_theses_bigrams %>%
-  group_by(title) %>% 
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 10) %>%                   ### change the number of top words to show
-  mutate(bigram = reorder(bigram, n)) %>% 
-  print()
+  facet_wrap( ~ title, scales = "free", ncol = 2) +
+  theme(legend.position = "none") 
 ```

### 2023-02-18 analysis/paper/paper.qmd
Update renv with reshape2 package for topic modeling

--- 2023-02-17 analysis/paper/paper.qmd
@@
 library(tidyverse)
+library(reshape2)
 
@@
   facet_wrap( ~ title, scales = "free", ncol = 2)
-```
 
-```{r}
-#| label: ngrams
+# # wip function ngrams
+# get_ngrams <- function(df, n) {
+#   if (!is.numeric(n)) {
+#     stop("n is not numeric")
+#   }
+#   emu_theses %>%
+#     mutate(author = paste(first_name, last_name)) %>% 
+#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
+#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+#     anti_join(stop_words, by = c("first" = "word")) %>%
+#     anti_join(stop_words, by = c("second" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
+#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+#     filter(first != second) %>% 
+#     filter(first != "source")
+# }
+# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
 
-## TO DO: generalise bigrams and trigrams into an ngrams function
-
-## Bigrams
-emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
-
-bigram_counts <- emu_theses_bigrams |> 
-  count(first, second, sort = TRUE)
-
-bigram_counts
-
-## Trigrams
-emu_theses_trigrams <- emu_theses |> 
-  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
-  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
-  filter(!first %in% stop_words$word,
-         !second %in% stop_words$word,
-         !third %in% stop_words$word) |>  
-  filter(str_detect(first, "[a-z]"), 
-         str_detect(second, "[a-z]"),
-         str_detect(third, "[a-z]"))
-
-trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, sort = TRUE)
-
-trigram_counts
-
-# Bigram tf-idf
-bigram_tf_idf <- emu_theses_bigrams |> 
-  count(title, bigram) |> 
-  bind_tf_idf(bigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-bigram_tf_idf
-
-# Trigram tf-idf
-trigram_tf_idf <- emu_theses_trigrams |> 
-  count(title, trigram) |> 
-  bind_tf_idf(trigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-trigram_tf_idf
+# bigrams
+emu_theses_bigrams <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+  anti_join(stop_words, by = c("first" = "word")) %>%
+  anti_join(stop_words, by = c("second" = "word")) %>%
+  anti_join(stop_words_custom, by = c("first" = "word")) %>%
+  anti_join(stop_words_custom, by = c("second" = "word")) %>%
+  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+  filter(first != second) %>% 
+  filter(first != "source")
 

### 2023-02-18 analysis/paper/paper.qmd
Add script for topic modeling
At this stage the output does not make much sense as the input text data is not cleaned properly. We will get back to this as soon as the data is ready for modeling.

@@
 
+```{r}
+emu_theses_dtm <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  cast_dtm(title, word, n)
+
+emu_theses_dtm
+
+emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
+emu_theses_lda
+
+emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
+emu_theses_topics
+
+# The top 10 words that are the most coomon in each topic
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(10, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
+
+emu_theses_top_terms |> 
+  mutate(term = reorder(term, beta)) |> 
+  ggplot(aes(term, beta, fill = factor(topic))) +
+  geom_col(show.legend = FALSE) +
+  facet_wrap(~ topic, scales = "free") +
+  coord_flip()
+
+# 
+beta_spread <- emu_theses_topics |> 
+  mutate(topic = paste0("topic", topic)) |> 
+  spread(topic, beta) |> 
+  filter(topic1 > 0.001, topic2 > .001) |> 
+  mutate(log_ratio = log2(topic2/topic1))
+
+beta_spread
+
+beta_spread %>%
+  group_by(direction = log_ratio > 0) %>%
+  slice_max(abs(log_ratio), n = 10) %>% 
+  ungroup() %>%
+  mutate(term = reorder(term, log_ratio)) %>%
+  ggplot(aes(log_ratio, term)) +
+  geom_col() +
+  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
+
+emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
+emu_theses_documents
+```
+
 # Results

### 2023-02-19 analysis/paper/paper.qmd
Add comments and adjust topic model text

@@
 
-Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
 
 ```{r}
+# The topicmodels package requires a document-term matrix as input
 emu_theses_dtm <- emu_theses_words |> 
@@
 
+# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+# In LDA, each document is a mixture of topics and each topic is a mixture of words
 emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
@@
 
+# 
 emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")

### 2023-02-19 analysis/paper/paper.qmd
Update topic model text

@@
 
-# 
+# Tidy model with beta statistic
 emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
@@
 
-# The top 10 words that are the most coomon in each topic
+# The top 10 words that are the most common in each topic
 emu_theses_top_terms <- emu_theses_topics |> 

### 2023-02-20 analysis/paper/paper.qmd

--- 2023-02-18 analysis/paper/paper.qmd
@@
 # Load packages ----
-## Managing paths
-library(here)
+## List packages to be loaded
+packages <- c(
+  "here",                         # Managing paths
+  "tidyverse",                    # Data manipulation and visualisation
+  "tidytext", "SnowballC", "tm",  # Text processing and analysis
+  "tidygeocoder",                 # Working with spatial data
+  "igraph", "ggraph"              # Network analysis and visualsation
+)
 
-## Data manipulation and visualisation
-library(tidyverse)
-library(reshape2)
-
-## Text processing and analysis
-library(tidytext)
-library(SnowballC)
-library(tm)
-library(topicmodels)
-
-## Working with spatial data
-library(tidygeocoder)
+## Load required packages, install them if needed
+for (package in packages) {
+  if (!require(package, character.only = TRUE)) {
+    install.packages(package)
+  }
+  library(package, character.only = TRUE)
+}
 
@@
 #| echo: false
-#| cache: true
+
+# Extract text from PDFs
+# On the first run, this function might take extra time as a Python environment needs to be set up
 
@@
 Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-
-```{r}
-emu_theses_dtm <- emu_theses_words |> 
-  count(title, word, sort = TRUE) |> 
-  cast_dtm(title, word, n)
-
-emu_theses_dtm
-
-emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
-emu_theses_lda
-
-emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
-emu_theses_topics
-
-# The top 10 words that are the most coomon in each topic
-emu_theses_top_terms <- emu_theses_topics |> 
-  group_by(topic) |> 
-  top_n(10, beta) |> 
-  ungroup() |> 
-  arrange(topic, -beta)
-
-emu_theses_top_terms |> 
-  mutate(term = reorder(term, beta)) |> 
-  ggplot(aes(term, beta, fill = factor(topic))) +
-  geom_col(show.legend = FALSE) +
-  facet_wrap(~ topic, scales = "free") +
-  coord_flip()
-
-# 
-beta_spread <- emu_theses_topics |> 
-  mutate(topic = paste0("topic", topic)) |> 
-  spread(topic, beta) |> 
-  filter(topic1 > 0.001, topic2 > .001) |> 
-  mutate(log_ratio = log2(topic2/topic1))
-
-beta_spread
-
-beta_spread %>%
-  group_by(direction = log_ratio > 0) %>%
-  slice_max(abs(log_ratio), n = 10) %>% 
-  ungroup() %>%
-  mutate(term = reorder(term, log_ratio)) %>%
-  ggplot(aes(log_ratio, term)) +
-  geom_col() +
-  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
-
-emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
-emu_theses_documents
-```
 

### 2023-02-20 analysis/paper/paper.qmd

@@
 ```{r}
-# Which are the top 20 most frequently used words in each thesis?
-emu_theses_top_20 <- emu_theses_words %>%
-  group_by(title) %>%
-  count(word, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50) %>%              ### filter words with a minimum count
-  ungroup() %>% 
-  mutate(word = reorder(word, n))
+# Which are the top n most frequently used words in each thesis?
+get_top_words_per_document(data = emu_theses_words, 
+                           top_n = 3, 
+                           title_col = "title", 
+                           word_col = "word")
+```
 
-# Which are first 5 most frequently used words in each thesis?
-emu_theses_first_5 <- emu_theses_words %>%
-  count(title, word) %>% 
-  group_by(title) %>% 
-  arrange(desc(n)) %>% 
-  slice(1:5)
+```{r}
+# Which are the top n most frequently used words in all theses`?
+get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 5, 
+                         word_col = "word")
+```
 
+```{r}
 # Which theses contain "fabric" in the top 20 most frequently used words?

### 2023-02-20 analysis/paper/paper.qmd

@@
   facet_wrap( ~ title, scales = "free", ncol = 2)
+```
 
-# # wip function ngrams
-# get_ngrams <- function(df, n) {
-#   if (!is.numeric(n)) {
-#     stop("n is not numeric")
-#   }
-#   emu_theses %>%
-#     mutate(author = paste(first_name, last_name)) %>% 
-#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
-#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-#     anti_join(stop_words, by = c("first" = "word")) %>%
-#     anti_join(stop_words, by = c("second" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
-#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-#     filter(first != second) %>% 
-#     filter(first != "source")
-# }
-# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
+```{r}
+#| label: ngrams
 
-# bigrams
-emu_theses_bigrams <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-  anti_join(stop_words, by = c("first" = "word")) %>%
-  anti_join(stop_words, by = c("second" = "word")) %>%
-  anti_join(stop_words_custom, by = c("first" = "word")) %>%
-  anti_join(stop_words_custom, by = c("second" = "word")) %>%
-  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-  filter(first != second) %>% 
-  filter(first != "source")
+## TO DO: generalise bigrams and trigrams into ngrams function
+
+## Bigrams
+emu_theses_bigrams <- emu_theses |> 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
+
+bigram_counts <- emu_theses_bigrams |> 
+  count(first, second, sort = TRUE)
+
+bigram_counts
+
+## Trigrams
+emu_theses_trigrams <- emu_theses |> 
+  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
+  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
+  filter(!first %in% stop_words$word,
+         !second %in% stop_words$word,
+         !third %in% stop_words$word) |>  
+  filter(str_detect(first, "[a-z]"), 
+         str_detect(second, "[a-z]"),
+         str_detect(third, "[a-z]"))
+
+trigram_counts <- emu_theses_trigrams |> 
+  count(first, second, sort = TRUE)
+
+trigram_counts
+
+## Bigram tf-idf
+bigram_tf_idf <- emu_theses_bigrams |> 
+  count(title, bigram) |> 
+  bind_tf_idf(bigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+bigram_tf_idf
+
+## Trigram tf-idf
+trigram_tf_idf <- emu_theses_trigrams |> 
+  count(title, trigram) |> 
+  bind_tf_idf(trigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+trigram_tf_idf
+
+## Show network of bigrams
+bigram_graph <- bigram_counts |> 
+  filter(n > 100) |> 
+  graph_from_data_frame()
+
+bigram_graph
+
+set.seed(2023)
+
+a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
+
+ggraph(bigram_graph, layout = "fr") +
+  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
+  geom_node_point(color = "lightblue", size = 5) +
+  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
+  theme_void()
+
+## Higgh frequency bigrams ----
 

### 2023-02-20 analysis/paper/paper.qmd

@@
   "tidyverse",                                 # Data manipulation and visualisation
-  "tidytext", "SnowballC", "tm",  # Text processing and analysis
+  "tidytext", "SnowballC", "tm","topicmodels"  # Text processing and analysis
   "tidygeocoder",                              # Working with spatial data

### 2023-02-20 R/geocode.R
Remove return from geocode function

@@ -11,7 +11,6 @@ geocode_thesis_locations <- function(data) {
   if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
     tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
   }
-  return(data)
 }
 
 

### 2023-02-20 analysis/paper/paper.qmd
Add locaiton as argument to geocode function

--- 2023-02-20 analysis/paper/paper.qmd
@@
 # Load packages ----
-## List packages to be loaded
-packages <- c(
-  "here",                         # Managing paths
-  "tidyverse",                    # Data manipulation and visualisation
-  "tidytext", "SnowballC", "tm",  # Text processing and analysis
-  "tidygeocoder",                 # Working with spatial data
-  "igraph", "ggraph"              # Network analysis and visualsation
-)
+## Managing paths
+library(here)
 
-## Load required packages, install them if needed
-for (package in packages) {
-  if (!require(package, character.only = TRUE)) {
-    install.packages(package)
-  }
-  library(package, character.only = TRUE)
-}
+## Data manipulation and visualisation
+library(tidyverse)
+
+## Text processing and analysis
+library(tidytext)
+library(SnowballC)
+library(tm)
+
+## Working with spatial data
+library(tidygeocoder)
 
@@
 # Read thesis metadata ----
+## All theses
+theses_all <-
+  readxl::read_xlsx(path = here("analysis", "data", "raw_data", "theses-all.xlsx"))
+
+## Theses with PDF available
 data_path <- here("analysis", "data", "raw_data")
@@
 
-## Read metadata for all theses
-all_theses <-
-  readxl::read_xlsx(path = here(data_path, "theses-all.xlsx"))
-
-## Read only metadata for theses with PDF available
 emu_theses <- 
@@
 #| echo: false
-
-# Extract text from PDFs
-# On the first run, this function might take extra time as a Python environment needs to be set up
-
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(theses_all)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(theses_all$grad_year, na.rm = TRUE), "-", max(theses_all$grad_year, na.rm = TRUE))`.
 The first year of the program was not represented, as PDF files were only available from 2008 onward.
@@
 
-# Geocode thesis locations
-emu_theses <- emu_theses %>%
-  geocode(location, method = 'osm', lat = latitude , long = longitude)
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses, "location")
 
 # Show thesis locations on world map
-world <- map_data("world")
-ggplot() +
-  geom_map(
-    data = world, map = world,
-    aes(long, lat, map_id = region),
-    color = "black", fill = "lightgray", size = 0.1
-  ) +
-  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
-  coord_fixed() +
-  theme_void()
+visualize_thesis_locations(emu_theses)
 ```
@@
 
-## Topic model
-
-Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-
 # Results
@@
   ungroup() %>% 
-  mutate(bigram = reorder(bigram, n))
-
-emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
   mutate(bigram = reorder(bigram, n))

### 2023-02-20 R/geocode.R

@@ -7,9 +7,9 @@
 #' @return Data frame with latitude and longitude columns
 #'
 #' @export
-geocode_thesis_locations <- function(data) {
+geocode_thesis_locations <- function(data, location = NULL) {
   if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
-    tidygeocoder::geocode(data, "location", method = 'osm', lat = latitude , long = longitude)
+    tidygeocoder::geocode(data, !!sym(location), method = 'osm', lat = latitude , long = longitude)
   }
 }
 

### 2023-02-20 analysis/paper/paper.qmd
Generalize geocode function with location

@@
 # Add latitude and longitude information to theses data frame
-emu_theses <- geocode_thesis_locations(emu_theses, "location")
+emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
+
+ geocoded_data <- tidygeocoder::geocode(emu_theses,
+      address = location,
+      method = 'osm',
+      lat = latitude ,
+      long = longitude
+ )
 

### 2023-02-20 R/geocode.R

@@ -7,9 +7,15 @@
 #' @return Data frame with latitude and longitude columns
 #'
 #' @export
-geocode_thesis_locations <- function(data, location = NULL) {
+geocode_thesis_locations <- function(data, location = "location") {
   if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
-    tidygeocoder::geocode(data, !!sym(location), method = 'osm', lat = latitude , long = longitude)
+    geocoded_data <- tidygeocoder::geocode(
+      data,
+      address = !!sym(location),
+      method = 'osm',
+      lat = latitude ,
+      long = longitude
+    )
   }
 }
 

### 2023-02-20 R/geocode.R
Update geocode functions

@@ -17,6 +17,7 @@ geocode_thesis_locations <- function(data, location = "location") {
       long = longitude
     )
   }
+  geocoded_data
 }
 
 
@@ -28,15 +29,15 @@ geocode_thesis_locations <- function(data, location = "location") {
 #'
 #' @return A plot with thesis locations
 #' @export
-visualize_thesis_locations <- function(data) {
+visualize_thesis_locations <- function(data, cols = c(map_col = 'black' , map_fill = 'lightgray', point_col = 'red' )) {
   world <- ggplot2::map_data("world")
   ggplot2::ggplot() +
     ggplot2::geom_map(
       data = world, map = world,
       aes(long, lat, map_id = region),
-      color = "black", fill = "lightgray", size = 0.1
+      color = cols['map_col'], fill = cols['map_fill'], size = 0.1
     ) +
-    ggplot2::geom_point(data = data, aes(longitude, latitude), color = "red") +
+    ggplot2::geom_point(data = data, aes(longitude, latitude), color = cols['point_col']) +
     ggplot2::coord_fixed() +
     ggplot2::theme_void()
 }

### 2023-02-20 README.Rmd
Update README.Rmd

@@
 
-# minTEMU - The Legacy of the European Postgraduate Master in Urbanism at TU Delft: A Text Mining Approach
+# mintEMU - The Legacy of the European Postgraduate Master in Urbanism at TU Delft: A Text Mining Approach
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/UD3-Lab/minTEMU/master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/UD3-Lab/minTEMU/main?urlpath=rstudio)
 

### 2023-02-21 analysis/paper/paper.qmd
Add missing comma in list of packages

@@
   "tidyverse",                                   # Data manipulation and visualisation
-  "tidytext", "SnowballC", "tm","topicmodels"  # Text processing and analysis
+  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
   "tidygeocoder",                                # Working with spatial data
@@
 
-For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage.
-Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
-As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
-<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result.
---> Subject-specific stop words, we manually constructed a list based on our domain knowledge.
-Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well.
-Finally, titles and author names were removed as document stop words on document-level.
+-   Three types of stop words: global, subject-specific and document stop words.
+-   Global stop words:
+    -   All theses were written in English, so for common English stop words the "snowball" lexicon was used
+    -   Optionally, try creating a custom stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words
+-   Subject-specific stop words:
+    -   Manually constructed list, needs domain knowledge
+    -   Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words
+-   Document stop words:
+    -   Stop words from titles and author names should only be removed from corresponding theses
 

### 2023-02-21 analysis/paper/paper.qmd
Increase dot size on map if multiple theses in the same location

@@
   ) +
-  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
+  geom_count(data = emu_theses, aes(longitude, latitude), color = "red", alpha = 0.7) +
+  scale_size_area(max_size = 3) +
   coord_fixed() +

### 2023-02-21 R/geocode.R
Replace geom_point in world map function with geom_count

@@ -37,7 +37,8 @@ visualize_thesis_locations <- function(data, cols = c(map_col = 'black' , map_fi
       aes(long, lat, map_id = region),
       color = cols['map_col'], fill = cols['map_fill'], size = 0.1
     ) +
-    ggplot2::geom_point(data = data, aes(longitude, latitude), color = cols['point_col']) +
+    ggplot2::geom_count(data = data, aes(longitude, latitude), color = cols['point_col']) +
+    ggplot2::scale_size_area(max_size = 5) +
     ggplot2::coord_fixed() +
     ggplot2::theme_void()
 }

### 2023-03-01 analysis/paper/paper.qmd

@@
   if (!require(package, character.only = TRUE)) {
-    install.packages(package)
+    install.packages(package,  "http://cran.us.r-project.org")
   }
@@
 }
+
+# library(here)
+# library(tidyverse)
+# library(tidytext)
+# library(SnowballC)
+# library(tm)
+# library(topicmodels)
+# library(tidygeocoder)
+# library(igraph)
+# library(ggraph)
 
@@
 
-A total number of `r nrow(emu_theses)` out of `r nrow(theses_all)` were analysed.
+A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.
 
@@
 # Show thesis locations on world map
-world <- map_data("world")
+world <- ggplot2::map_data("world")
 ggplot() +
@@
 ```{r}
-# Which theses contain "fabric" in the top 20 most frequently used words?
-emu_theses_top_20 %>% 
-  filter(word == "fabric")
+# Which are the top n most frequently used words in all theses`?
+emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 20, 
+                         word_col = "word")
+```
 
-emu_theses_top_20 %>%
-  slice(1:40) %>% 
-  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
-  geom_col() +
-  xlab(NULL) +
-  coord_flip() +
-  labs(y = "Count",
-       x = "Unique words",
-       title = "Count of unique words found in thesis",
-       subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
-```
+<!-- ```{r} -->
+<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
+<!-- emu_theses_top_20 %>%  -->
+<!--   filter(word == "fabric") -->
+
+<!-- emu_theses_top_20 %>% -->
+<!--   slice(1:40) %>%  -->
+<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
+<!--   geom_col() + -->
+<!--   xlab(NULL) + -->
+<!--   coord_flip() + -->
+<!--   labs(y = "Count", -->
+<!--        x = "Unique words", -->
+<!--        title = "Count of unique words found in thesis", -->
+<!--        subtitle = "Stop words were removed from the list") + -->
+<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
+<!-- ``` -->
 

### 2023-03-01 analysis/paper/paper.qmd

@@
 
-# Geocode thesis locations
-emu_theses <- emu_theses %>%
-  geocode(location, method = 'osm', lat = latitude , long = longitude)
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
+
+# emu_theses <- tidygeocoder::geocode(
+#   emu_theses,
+#   address = location,
+#   method = 'osm',
+#   lat = latitude ,
+#   long = longitude
+# )
 
 # Show thesis locations on world map
-world <- ggplot2::map_data("world")
-ggplot() +
-  geom_map(
-    data = world, map = world,
-    aes(long, lat, map_id = region),
-    color = "black", fill = "lightgray", size = 0.1
-  ) +
-  geom_count(data = emu_theses, aes(longitude, latitude), color = "red", alpha = 0.7) +
-  scale_size_area(max_size = 3) +
-  coord_fixed() +
-  theme_void()
+visualize_thesis_locations(emu_theses)
 ```

### 2023-03-01 analysis/paper/paper.qmd
Update geocode function and function call in paper

--- 2023-02-21 analysis/paper/paper.qmd
@@
 # Load packages ----
-## List packages to be loaded
-packages <- c(
-  "here",                                        # Managing paths
-  "tidyverse",                                   # Data manipulation and visualisation
-  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
-  "tidygeocoder",                                # Working with spatial data
-  "igraph", "ggraph"                             # Network analysis and visualsation
-)
+## Managing paths
+library(here)
 
-## Load required packages, install them if needed
-for (package in packages) {
-  if (!require(package, character.only = TRUE)) {
-    install.packages(package)
-  }
-  library(package, character.only = TRUE)
-}
+## Data manipulation and visualisation
+library(tidyverse)
+
+## Text processing and analysis
+library(tidytext)
+library(SnowballC)
+library(tm)
+
+## Working with spatial data
+library(tidygeocoder)
 
@@
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
-With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 
-A total number of `r nrow(emu_theses)` out of `r nrow(theses_all)` were analysed.
+A total number of `r nrow(emu_theses)` were analysed.
 
@@
 
-# Geocode thesis locations
-emu_theses <- emu_theses %>%
-  geocode(location, method = 'osm', lat = latitude , long = longitude)
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
 
 # Show thesis locations on world map
-world <- map_data("world")
-ggplot() +
-  geom_map(
-    data = world, map = world,
-    aes(long, lat, map_id = region),
-    color = "black", fill = "lightgray", size = 0.1
-  ) +
-  geom_count(data = emu_theses, aes(longitude, latitude), color = "red", alpha = 0.7) +
-  scale_size_area(max_size = 3) +
-  coord_fixed() +
-  theme_void()
+visualize_thesis_locations(emu_theses)
 ```
@@
   facet_wrap( ~ title, scales = "free", ncol = 2)
-```
 
-```{r}
-#| label: ngrams
+# # wip function ngrams
+# get_ngrams <- function(df, n) {
+#   if (!is.numeric(n)) {
+#     stop("n is not numeric")
+#   }
+#   emu_theses %>%
+#     mutate(author = paste(first_name, last_name)) %>% 
+#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
+#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+#     anti_join(stop_words, by = c("first" = "word")) %>%
+#     anti_join(stop_words, by = c("second" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
+#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+#     filter(first != second) %>% 
+#     filter(first != "source")
+# }
+# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
 
-## TO DO: generalise bigrams and trigrams into ngrams function
-
-## Bigrams
-emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
-
-bigram_counts <- emu_theses_bigrams |> 
-  count(first, second, sort = TRUE)
-
-bigram_counts
-
-## Trigrams
-emu_theses_trigrams <- emu_theses |> 
-  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
-  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
-  filter(!first %in% stop_words$word,
-         !second %in% stop_words$word,
-         !third %in% stop_words$word) |>  
-  filter(str_detect(first, "[a-z]"), 
-         str_detect(second, "[a-z]"),
-         str_detect(third, "[a-z]"))
-
-trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, sort = TRUE)
-
-trigram_counts
-
-## Bigram tf-idf
-bigram_tf_idf <- emu_theses_bigrams |> 
-  count(title, bigram) |> 
-  bind_tf_idf(bigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-bigram_tf_idf
-
-## Trigram tf-idf
-trigram_tf_idf <- emu_theses_trigrams |> 
-  count(title, trigram) |> 
-  bind_tf_idf(trigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-trigram_tf_idf
-
-## Show network of bigrams
-bigram_graph <- bigram_counts |> 
-  filter(n > 100) |> 
-  graph_from_data_frame()
-
-bigram_graph
-
-set.seed(2023)
-
-a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
-
-ggraph(bigram_graph, layout = "fr") +
-  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
-  geom_node_point(color = "lightblue", size = 5) +
-  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
-  theme_void()
-
-## Higgh frequency bigrams ----
+# bigrams
+emu_theses_bigrams <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+  anti_join(stop_words, by = c("first" = "word")) %>%
+  anti_join(stop_words, by = c("second" = "word")) %>%
+  anti_join(stop_words_custom, by = c("first" = "word")) %>%
+  anti_join(stop_words_custom, by = c("second" = "word")) %>%
+  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+  filter(first != second) %>% 
+  filter(first != "source")
 
@@
   count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+  slice_max(n, n = 10) %>%        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
@@
 
-emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
-  mutate(bigram = reorder(bigram, n))
-
 emu_theses_bigrams_top %>%
-  filter(title == unique(title)[1:6]) |> 
   ggplot(aes(x = bigram, y = n, fill = title)) +
@@
        subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2) +
-  theme(legend.position = "none") 
+  facet_wrap( ~ title, scales = "free", ncol = 2)
+
+# find the most frequently used bigrams
+emu_thesis_bigrams_top <- emu_theses_bigrams %>%
+  group_by(title) %>% 
+  count(bigram, sort = TRUE) %>%
+  slice_max(n, n = 10) %>%                   ### change the number of top words to show
+  mutate(bigram = reorder(bigram, n)) %>% 
+  print()
 ```
-
-```{r}
-#| label: tf-idf
-
-emu_theses_words
-
-emu_theses_words_count <- emu_theses_words |> 
-  count(title, word, sort = TRUE) |> 
-  ungroup()
-
-emu_theses_words_total <- emu_theses_words_count |> 
-  group_by(title) |> 
-  summarise(total = sum(n))
-
-emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)
-
-# emu_theses_words_count |> 
-#   group_by(title) |> 
-#   summarise(total = unique(tot
-
-# Preview term frequency distributions in a subset of theses
-ggplot(emu_theses_words_count |> 
-         filter(title == unique(title)[1:6]),
-       aes(n/total, fill = title)) +
-  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
-  xlim(NA, 0.0009) +
-  facet_wrap(~ title, ncol = 3, scales = "free_y")
-
-# Examine Zipf's law for the EMU theses
-## Calculate rank and term frequency
-freq_by_rank <- emu_theses_words_count |> 
-  group_by(title) |> 
-  mutate(rank = row_number(),
-         `term frequency` = n/total)
-
-## Visualise Zipf's law
-freq_by_rank |> 
-  ggplot(aes(rank, `term frequency`, color = title)) +
-  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
-  scale_x_log10() +
-  scale_y_log10()
-
-rank_subset <- freq_by_rank |> 
-  filter(rank < 500,
-         rank > 10)
-
-lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)
-
-freq_by_rank |> 
-  ggplot(aes(rank, `term frequency`, color = title)) +
-  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
-  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
-  scale_x_log10() +
-  scale_y_log10()
-
-## Calculate tf-idf
-emu_theses_words_count <- emu_theses_words_count |> 
-  bind_tf_idf(word, title, n)
-emu_theses_words_count
-
-emu_theses_words_count |> 
-  select(-total) |> 
-  arrange(desc(tf_idf))
-
-## Visualise the top n words with the highest tf-idf values for each of a subset of theses
-emu_theses_words_count |> 
-  filter(title == unique(title)[1:6]) |> 
-  arrange(desc(tf_idf)) |> 
-  mutate(word = factor(word, levels = rev(unique(word)))) |> 
-  group_by(title) |> 
-  top_n(15) |> 
-  ungroup() |> 
-  ggplot(aes(word, tf_idf, fill = title)) +
-  geom_col(show.legend = FALSE) +
-  labs(x = NULL, y = "tf-idf") +
-  facet_wrap(~ title,  ncol = 3, scales = "free") +
-  coord_flip()
-```
-
-The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
-As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
 

### 2023-03-01 R/geocode.R

@@ -11,7 +11,7 @@ geocode_thesis_locations <- function(data, location = "location") {
   if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
     geocoded_data <- tidygeocoder::geocode(
       data,
-      address = !!sym(location),
+      address = location,
       method = 'osm',
       lat = latitude ,
       long = longitude

### 2023-03-13 analysis/paper/paper.qmd
Cache text extracted from pdfs
Clean up comments in setup code chunk

--- 2023-03-01 analysis/paper/paper.qmd
@@
   "tidygeocoder",                                # Working with spatial data
-  "igraph", "ggraph"                             # Network analysis and visualsation
+  "igraph", "ggraph"                             # Network analysis and visualization
 )
@@
 }
-
-# library(here)
-# library(tidyverse)
-# library(tidytext)
-# library(SnowballC)
-# library(tm)
-# library(topicmodels)
-# library(tidygeocoder)
-# library(igraph)
-# library(ggraph)
 

### 2023-03-13 analysis/paper/paper.qmd
Update subject and document stopwords functions

--- 2023-03-01 analysis/paper/paper.qmd
@@
 
-#head_text(emu_theses$text)
-
-# clean text 
-emu_theses$text_clean <- emu_theses$text %>%
-  clean_basic() %>%
-  str_remove_all(find_meta_stopwords(emu_theses) )
-
+emu_theses$text_clean <- emu_theses$text |>
+  clean_basic() |>
+  str_remove_all(find_meta_stopwords(emu_theses)) |> 
+  str_remove_all(subject_stopwords()) |> 
+  str_remove_all(document_stopwords())
 ```
@@
 
-# Add latitude and longitude information to theses data frame
-emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
+# Geocode thesis locations
+emu_theses <- emu_theses %>%
+  geocode(location, method = 'osm', lat = latitude , long = longitude)
 
 # Show thesis locations on world map
-visualize_thesis_locations(emu_theses)
+world <- map_data("world")
+ggplot() +
+  geom_map(
+    data = world, map = world,
+    aes(long, lat, map_id = region),
+    color = "black", fill = "lightgray", size = 0.1
+  ) +
+  geom_point(data = emu_theses, aes(longitude, latitude), color = "red") +
+  coord_fixed() +
+  theme_void()
 ```
@@
 
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
-                               "source", "author"))) %>% 
-  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
-  rbind(., data.frame(word = c("hab", "km")))
+urbanism_stopwords <- urbanism_stopwords() 
+thesis_stopwords <- thesis_stopwords()
 
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
-  anti_join(stop_words_custom, by = "word")
+emu_theses_words <- emu_theses |> 
+  select(title, text_clean) |> 
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") |> 
+  anti_join(urbanism_stopwords, by = "word") |> 
+  anti_join(thesis_stopwords, by = "word")
 ```

### 2023-03-13 R/clean.R

@@ -95,11 +95,13 @@ find_meta_stopwords <- function(metadata,
 #' @param add_stopwords A vector of words to be added to the urbanism stopwords
 #' @return A vector of urbanism-related stopwords
 #' @export
-subject_stopwords <- function(add_stopwords = NULL) {
-  stop_words <-
-    tibble(word = c(add_stopwords, "city", "urban", "urbanism", "hab", "km"))
+subject_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
+  stop_words <- c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km",
+                  "emu", "european postgraduate master in urbanism", "advisor",
+                  "prof")
 
-  stop_words
+  if (convert_to_regex) return(paste(stop_words, collapse = "|"))
+  else return(stop_words)
 }
 
 #' Adjustable list of stopwords associated with master theses
@@ -108,11 +110,12 @@ subject_stopwords <- function(add_stopwords = NULL) {
 #' @param add_stopwords A vector of words to be added to the master theses-related stopwords
 #' @return A vector of thesis-related stopwords
 #' @export
-document_stopwords <- function(add_stopwords = NULL) {
-  stop_words <-
-    tibble(word = c(add_stopwords, "preface", "foreword", "introduction", "conclusion", "thesis"))
+document_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
+  stop_words <- c(add_stopwords, "preface", "foreword", "introduction", "conclusion",
+                  "thesis", "colophon", "acknowledgements", "references")
 
-  stop_words
+  if (convert_to_regex) return(paste(stop_words, collapse = "|"))
+  else return(stop_words)
 }
 
 

### 2023-03-13 analysis/paper/paper.qmd
Update stopword removal in paper

@@
 emu_theses$text_clean <- emu_theses$text |>
-  clean_basic() |>
-  str_remove_all(find_meta_stopwords(emu_theses)) |> 
-  str_remove_all(subject_stopwords()) |> 
-  str_remove_all(document_stopwords())
-```
-
-```{r}
-#| label: preview
-#| include: false
-
-# Preview theses
-emu_theses %>% 
-  arrange(graduation_year) %>% 
-  pull(text_clean) %>% 
-  str_sub(1, 500)
-
-# Preview theses
-hist(str_length(emu_theses$text_clean))
-
-Encoding(emu_theses$text_clean)
+  clean_basic() 
 ```
@@
 
-urbanism_stopwords <- urbanism_stopwords() 
-thesis_stopwords <- thesis_stopwords()
+emu_theses$text_clean <- emu_theses$text_clean |>
+  str_remove_all(find_meta_stopwords(emu_theses)) |> 
+  str_remove_all(subject_stopwords()) |> 
+  str_remove_all(document_stopwords())
 
@@
   select(title, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") |> 
-  anti_join(urbanism_stopwords, by = "word") |> 
-  anti_join(thesis_stopwords, by = "word")
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
+  anti_join(tidytext::stop_words, by = "word")
 ```

### 2023-03-13 analysis/paper/paper.qmd
Explore topic model

--- 2023-03-13 analysis/paper/paper.qmd
@@
 #| echo: false
+#| cache: true
+
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
   clean_basic() |> 
-  str_remove_all(find_meta_stopwords(emu_theses)) |> 
-  str_remove_all(subject_stopwords()) |> 
-  str_remove_all(document_stopwords())
+  str_remove_all(find_meta_stopwords(emu_theses) )  
 ```
@@
 
-urbanism_stopwords <- urbanism_stopwords() 
-thesis_stopwords <- thesis_stopwords()
+stop_words_custom <- emu_theses %>% 
+  unnest_tokens(output = word, input = title) %>%   ### add words from the title
+  select(word) %>% 
+  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
+                               "source", "author"))) %>% 
+  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
+  rbind(., data.frame(word = c("hab", "km")))
 
-emu_theses_words <- emu_theses |> 
-  select(title, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") |> 
-  anti_join(urbanism_stopwords, by = "word") |> 
-  anti_join(thesis_stopwords, by = "word")
+emu_theses_words <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") %>%  # remove stop words
+  anti_join(stop_words_custom, by = "word")
 ```
@@
   count(stem, sort = TRUE)
+```
+
+## Topic model
+
+Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+
+```{r}
+# The topicmodels package requires a document-term matrix as input
+emu_theses_dtm <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  cast_dtm(title, word, n)
+
+emu_theses_dtm
+
+# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+# In LDA, each document is a mixture of topics and each topic is a mixture of words
+emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
+emu_theses_lda
+
+# Tidy model with beta statistic
+emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
+emu_theses_topics
+
+# The top 10 words that are the most common in each topic
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
+
+emu_theses_top_terms |> 
+  mutate(term = reorder(term, beta)) |> 
+  ggplot(aes(term, beta, fill = factor(topic))) +
+  geom_col(show.legend = FALSE) +
+  facet_wrap(~ topic, scales = "free") +
+  coord_flip()
+
+# 
+beta_spread <- emu_theses_topics |> 
+  mutate(topic = paste0("topic", topic)) |> 
+  spread(topic, beta) |> 
+  filter(topic1 > 0.001, topic2 > .001) |> 
+  mutate(log_ratio = log2(topic2/topic1))
+
+beta_spread
+
+beta_spread %>%
+  group_by(direction = log_ratio > 0) %>%
+  slice_max(abs(log_ratio), n = 20) %>% 
+  ungroup() %>%
+  mutate(term = reorder(term, log_ratio)) %>%
+  ggplot(aes(log_ratio, term)) +
+  geom_col() +
+  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
+
+emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
+emu_theses_documents
+
+emu_theses_documents |> 
+  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
+  geom_bar(stat = "identity") + 
+  coord_flip()
+
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
+  geom_bar(stat = "identity") + 
+  coord_flip() + 
+  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)
+
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
+  geom_col()
 ```

### 2023-03-13 analysis/paper/paper.qmd
Explore topic model

@@
   coord_flip()
-
-emu_theses_documents |> 
-  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
-  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
-  geom_bar(stat = "identity") + 
-  coord_flip() + 
-  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)
-
-emu_theses_documents |> 
-  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
-  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
-  geom_col()
 ```

### 2023-03-13 analysis/paper/paper.qmd
Add notes to the introduction

--- 2023-03-13 analysis/paper/paper.qmd
@@
 # Load packages ----
-## Managing paths
-library(here)
+## List packages to be loaded
+packages <- c(
+  "here",                                        # Managing paths
+  "tidyverse",                                   # Data manipulation and visualisation
+  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
+  "tidygeocoder",                                # Working with spatial data
+  "igraph", "ggraph"                             # Network analysis and visualsation
+)
 
-## Data manipulation and visualisation
-library(tidyverse)
+## Load required packages, install them if needed
+for (package in packages) {
+  if (!require(package, character.only = TRUE)) {
+    install.packages(package,  "http://cran.us.r-project.org")
+  }
+  library(package, character.only = TRUE)
+}
 
-## Text processing and analysis
-library(tidytext)
-library(SnowballC)
-library(tm)
-
-## Working with spatial data
-library(tidygeocoder)
+# library(here)
+# library(tidyverse)
+# library(tidytext)
+# library(SnowballC)
+# library(tm)
+# library(topicmodels)
+# library(tidygeocoder)
+# library(igraph)
+# library(ggraph)
 
@@
 #| echo: false
+
+# Extract text from PDFs
+# On the first run, this function might take extra time as a Python environment needs to be set up
+
 emu_theses$text <- convert_pdf_text(pdf_paths)
@@
 emu_theses$text_clean <- emu_theses$text |> 
-  clean_basic() 
+  clean_basic() |> 
+  str_remove_all(find_meta_stopwords(emu_theses) )  
+```
+
+```{r}
+#| label: preview
+#| include: false
+
+# Preview theses in increasing order of graduation year
+emu_theses %>% 
+  arrange(graduation_year) %>% 
+  pull(text_clean) %>% 
+  head_text(head_n = 5)
+
+# Show distribution of the length of theses
+hist(str_length(emu_theses$text_clean))
 ```
@@
 
+Distinctive features of the program:  
+- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice
+- provides a bridge towards PhD-level research
+- all students had an exchange semester at one of the other three program universities
+
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(theses_all)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(theses_all$grad_year, na.rm = TRUE), "-", max(theses_all$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
 The first year of the program was not represented, as PDF files were only available from 2008 onward.
@@
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
+With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 
-A total number of `r nrow(emu_theses)` were analysed.
+A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.
 
@@
 
-emu_theses$text_clean <- emu_theses$text_clean |>
-  str_remove_all(find_meta_stopwords(emu_theses)) |> 
-  str_remove_all(subject_stopwords()) |> 
-  str_remove_all(document_stopwords())
+stop_words_custom <- emu_theses %>% 
+  unnest_tokens(output = word, input = title) %>%   ### add words from the title
+  select(word) %>% 
+  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
+                               "source", "author"))) %>% 
+  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
+  rbind(., data.frame(word = c("hab", "km")))
 
-emu_theses_words <- emu_theses |> 
-  select(title, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
-  anti_join(tidytext::stop_words, by = "word")
+emu_theses_words <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") %>%  # remove stop words
+  anti_join(stop_words_custom, by = "word")
 ```
@@
 
+## Topic model
+
+Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+
 # Results
@@
 ```{r}
-# Which are the top 20 most frequently used words in each thesis?
-emu_theses_top_20 <- emu_theses_words %>%
-  group_by(title) %>%
-  count(word, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50) %>%              ### filter words with a minimum count
-  ungroup() %>% 
-  mutate(word = reorder(word, n))
+# Which are the top n most frequently used words in each thesis?
+get_top_words_per_document(data = emu_theses_words, 
+                           top_n = 3, 
+                           title_col = "title", 
+                           word_col = "word")
+```
 
-# Which are first 5 most frequently used words in each thesis?
-emu_theses_first_5 <- emu_theses_words %>%
-  count(title, word) %>% 
-  group_by(title) %>% 
-  arrange(desc(n)) %>% 
-  slice(1:5)
+```{r}
+# Which are the top n most frequently used words in all theses`?
+get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 5, 
+                         word_col = "word")
+```
 
-# Which theses contain "fabric" in the top 20 most frequently used words?
-emu_theses_top_20 %>% 
-  filter(word == "fabric")
+```{r}
+# Which are the top n most frequently used words in all theses`?
+emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 20, 
+                         word_col = "word")
+```
 
-emu_theses_top_20 %>%
-  slice(1:40) %>% 
-  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
-  geom_col() +
-  xlab(NULL) +
-  coord_flip() +
-  labs(y = "Count",
-       x = "Unique words",
-       title = "Count of unique words found in thesis",
-       subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
+<!-- ```{r} -->
 
-# # wip function ngrams
-# get_ngrams <- function(df, n) {
-#   if (!is.numeric(n)) {
-#     stop("n is not numeric")
-#   }
-#   emu_theses %>%
-#     mutate(author = paste(first_name, last_name)) %>% 
-#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
-#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-#     anti_join(stop_words, by = c("first" = "word")) %>%
-#     anti_join(stop_words, by = c("second" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
-#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
-#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-#     filter(first != second) %>% 
-#     filter(first != "source")
-# }
-# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
+<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
 
-# bigrams
-emu_theses_bigrams <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
-  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
-  anti_join(stop_words, by = c("first" = "word")) %>%
-  anti_join(stop_words, by = c("second" = "word")) %>%
-  anti_join(stop_words_custom, by = c("first" = "word")) %>%
-  anti_join(stop_words_custom, by = c("second" = "word")) %>%
-  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
-  filter(first != second) %>% 
-  filter(first != "source")
+<!-- emu_theses_top_20 %>%  -->
+
+<!--   filter(word == "fabric") -->
+
+<!-- emu_theses_top_20 %>% -->
+
+<!--   slice(1:40) %>%  -->
+
+<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
+
+<!--   geom_col() + -->
+
+<!--   xlab(NULL) + -->
+
+<!--   coord_flip() + -->
+
+<!--   labs(y = "Count", -->
+
+<!--        x = "Unique words", -->
+
+<!--        title = "Count of unique words found in thesis", -->
+
+<!--        subtitle = "Stop words were removed from the list") + -->
+
+<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
+
+<!-- ``` -->
+
+```{r}
+#| label: ngrams
+
+## TO DO: generalise bigrams and trigrams into ngrams function
+
+## Bigrams
+emu_theses_bigrams <- emu_theses |> 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
+
+bigram_counts <- emu_theses_bigrams |> 
+  count(first, second, sort = TRUE)
+
+bigram_counts
+
+## Trigrams
+emu_theses_trigrams <- emu_theses |> 
+  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
+  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
+  filter(!first %in% stop_words$word,
+         !second %in% stop_words$word,
+         !third %in% stop_words$word) |>  
+  filter(str_detect(first, "[a-z]"), 
+         str_detect(second, "[a-z]"),
+         str_detect(third, "[a-z]"))
+
+trigram_counts <- emu_theses_trigrams |> 
+  count(first, second, sort = TRUE)
+
+trigram_counts
+
+## Bigram tf-idf
+bigram_tf_idf <- emu_theses_bigrams |> 
+  count(title, bigram) |> 
+  bind_tf_idf(bigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+bigram_tf_idf
+
+## Trigram tf-idf
+trigram_tf_idf <- emu_theses_trigrams |> 
+  count(title, trigram) |> 
+  bind_tf_idf(trigram, title, n) |> 
+  arrange(desc(tf_idf))
+
+trigram_tf_idf
+
+## Show network of bigrams
+bigram_graph <- bigram_counts |> 
+  filter(n > 100) |> 
+  graph_from_data_frame()
+
+bigram_graph
+
+set.seed(2023)
+
+a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
+
+ggraph(bigram_graph, layout = "fr") +
+  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
+  geom_node_point(color = "lightblue", size = 5) +
+  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
+  theme_void()
+
+## Higgh frequency bigrams ----
 
@@
        subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2)
+  facet_wrap( ~ title, scales = "free", ncol = 2) +
+  theme(legend.position = "none") 
+```
 
-# find the most frequently used bigrams
-emu_thesis_bigrams_top <- emu_theses_bigrams %>%
-  group_by(title) %>% 
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 10) %>%                   ### change the number of top words to show
-  mutate(bigram = reorder(bigram, n)) %>% 
-  print()
+```{r}
+#| label: tf-idf
+
+emu_theses_words
+
+emu_theses_words_count <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  ungroup()
+
+emu_theses_words_total <- emu_theses_words_count |> 
+  group_by(title) |> 
+  summarise(total = sum(n))
+
+emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)
+
+# emu_theses_words_count |> 
+#   group_by(title) |> 
+#   summarise(total = unique(tot
+
+# Preview term frequency distributions in a subset of theses
+ggplot(emu_theses_words_count |> 
+         filter(title == unique(title)[1:6]),
+       aes(n/total, fill = title)) +
+  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
+  xlim(NA, 0.0009) +
+  facet_wrap(~ title, ncol = 3, scales = "free_y")
+
+# Examine Zipf's law for the EMU theses
+## Calculate rank and term frequency
+freq_by_rank <- emu_theses_words_count |> 
+  group_by(title) |> 
+  mutate(rank = row_number(),
+         `term frequency` = n/total)
+
+## Visualise Zipf's law
+freq_by_rank |> 
+  ggplot(aes(rank, `term frequency`, color = title)) +
+  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
+  scale_x_log10() +
+  scale_y_log10()
+
+rank_subset <- freq_by_rank |> 
+  filter(rank < 500,
+         rank > 10)
+
+lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)
+
+freq_by_rank |> 
+  ggplot(aes(rank, `term frequency`, color = title)) +
+  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
+  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
+  scale_x_log10() +
+  scale_y_log10()
+
+## Calculate tf-idf
+emu_theses_words_count <- emu_theses_words_count |> 
+  bind_tf_idf(word, title, n)
+emu_theses_words_count
+
+emu_theses_words_count |> 
+  select(-total) |> 
+  arrange(desc(tf_idf))
+
+## Visualise the top n words with the highest tf-idf values for each of a subset of theses
+emu_theses_words_count |> 
+  filter(title == unique(title)[1:6]) |> 
+  arrange(desc(tf_idf)) |> 
+  mutate(word = factor(word, levels = rev(unique(word)))) |> 
+  group_by(title) |> 
+  top_n(15) |> 
+  ungroup() |> 
+  ggplot(aes(word, tf_idf, fill = title)) +
+  geom_col(show.legend = FALSE) +
+  labs(x = NULL, y = "tf-idf") +
+  facet_wrap(~ title,  ncol = 3, scales = "free") +
+  coord_flip()
 ```
+
+The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
+As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
 

### 2023-03-13 analysis/paper/paper.qmd
Only extract text from PDF if extracted text is missing

@@
 #| echo: false
+#| cache: true
 
-# Extract text from PDFs
+data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
+
+if (!file.exists(data_path)) {
   # On the first run, this function might take extra time as a Python environment needs to be set up
+  emu_theses$text <- convert_pdf_text(pdf_paths)
+  write_csv(emu_theses, data_path)
+} else {
+  emu_theses <- read_csv(data_path)
+}
 
-emu_theses$text <- convert_pdf_text(pdf_paths)
 ```
@@
 <!-- Complete paragraph about the program -->
-
-Distinctive features of the program:  
-- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice
-- provides a bridge towards PhD-level research
-- all students had an exchange semester at one of the other three program universities
 
@@
 <!-- ```{r} -->
-
 <!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
-
 <!-- emu_theses_top_20 %>%  -->
-
 <!--   filter(word == "fabric") -->
@@
 <!-- emu_theses_top_20 %>% -->
-
 <!--   slice(1:40) %>%  -->
-
 <!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
-
 <!--   geom_col() + -->
-
 <!--   xlab(NULL) + -->
-
 <!--   coord_flip() + -->
-
 <!--   labs(y = "Count", -->
-
 <!--        x = "Unique words", -->
-
 <!--        title = "Count of unique words found in thesis", -->
-
 <!--        subtitle = "Stop words were removed from the list") + -->
-
 <!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
-
 <!-- ``` -->

### 2023-03-17 analysis/paper/paper.qmd
Add notes on how to answer SRQ4

--- 2023-03-13 analysis/paper/paper.qmd
@@
 
-# Geocode thesis locations
-emu_theses <- emu_theses %>%
-  geocode(location, method = 'osm', lat = latitude , long = longitude)
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
 
 # Show thesis locations on world map
-world <- ggplot2::map_data("world")
-ggplot() +
-  geom_map(
-    data = world, map = world,
-    aes(long, lat, map_id = region),
-    color = "black", fill = "lightgray", size = 0.1
-  ) +
-  geom_count(data = emu_theses, aes(longitude, latitude), color = "red", alpha = 0.7) +
-  scale_size_area(max_size = 3) +
-  coord_fixed() +
-  theme_void()
+visualize_thesis_locations(emu_theses)
 ```

### 2023-03-17 analysis/paper/paper.qmd

@@
 emu_theses$text_clean <- emu_theses$text |>
-  clean_basic() |> 
-  str_remove_all(find_meta_stopwords(emu_theses) )  
-```
+  clean_basic()
 
-```{r}
-#| label: preview
-#| include: false
-
-# Preview theses in increasing order of graduation year
-emu_theses %>% 
-  arrange(graduation_year) %>% 
-  pull(text_clean) %>% 
-  head_text(head_n = 5)
-
-# Show distribution of the length of theses
-hist(str_length(emu_theses$text_clean))
 ```
@@
 
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
-                               "source", "author"))) %>% 
-  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
-  rbind(., data.frame(word = c("hab", "km")))
+emu_theses$text_clean <- emu_theses$text_clean |>
+  str_remove_all(find_meta_stopwords(emu_theses)) |> 
+  str_remove_all(subject_stopwords()) |> 
+  str_remove_all(document_stopwords())
 
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
-  anti_join(stop_words_custom, by = "word")
+emu_theses_words <- emu_theses |> 
+  select(title, text_clean) |> 
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
+  anti_join(tidytext::stop_words, by = "word")
 ```

### 2023-03-17 analysis/smp.qmd

@@
 The research compendium will be documented as an R package, that is, all functions used in data processing and analysis will have associated documentation available from the help interface of R.
-Once loaded as a package all documentation is available to the user.
+Once loaded as a package, all documentation is available to the user.
 The analysis will be fully documented in a Quarto computational notebook using R Markdown with the knitr engine.
@@
 
-Installation requirements are described in the README file of the repository on GitHub:
+Installation requirements are described in the [README file](https://github.com/UD3-Lab/mintEMU#how-to-run-in-your-browser-or-download-and-run-locally) of the repository on GitHub.
 
@@
 
+Citation guidelines are described in the [README file](https://github.com/UD3-Lab/mintEMU#how-to-cite) of the repository on GitHub.
+
 **How will your software be documented for future developers?**
 
+All the functions are well documented and the following R package structure so that they can be reused in other context.
+The repository README file has also a [Contributions](https://github.com/UD3-Lab/mintEMU#contributions) section, where [contributor guidelines](https://github.com/UD3-Lab/mintEMU/blob/main/CONTRIBUTING.md) and [Contributor code of Conduct](https://github.com/UD3-Lab/mintEMU/blob/main/CONDUCT.md) can be easily accessed.
+
 **How will your software be tested? Please provide a link to the (automated) testing results.**
+
+TBD
 
@@
 
+The snapshot of the GitHub repository (e.g. release associated with the publication of the paper) will be published on Zenodo.
+
 **How do you plan to procure long term maintenance of your software?**
+The long-term maintenance of the software is not foreseen in the project.The software is making use of the package management software, `renv`, which help user retrieve the required version of dependencies needed to use the functionalities of the software.

### 2023-03-17 analysis/smp.tex
explicitly stated package name in clean.R

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={mintEMU Software Management Plan},
  pdfauthor={Claudiu Forgaci},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{mintEMU Software Management Plan}
\author{Claudiu Forgaci}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, sharp corners, interior hidden, breakable]}{\end{tcolorbox}}\fi

This Software Management Plan uses the guidelines for medium management
level as described in the Practical guide to Software Management Plans
(\textbf{martinez-ortiz2022?}).

\textbf{Please provide a brief description of your software, stating its
purpose and intended audience.}

\emph{mintEMU} is a research compendium created as an R package with
\texttt{rrtools} in preparation of the scientific paper \emph{The Legacy
of the European Post-Master in Urbanism at TU Delft: A Text Mining
Approach}. The compendum contains the data, code for data processing and
analysis, as well as the paper presenting the results of the analysis.
The intended audience is the scientific community interested in the
research results, methods employed or data generated in the project.

The project was prompted by the closure of the European post-Master of
Urbanism (EMU) of the Department of Urbanism at the Faculty of
Architecture and the Built Environment, TU Delft. In order to describe
the legacy of the EMU program, including the distinctive features of its
didactic approach, this research aims to reveal the main topics taught
in it and how those topics had evolved through the years of the program.
To that end, we employ a text mining approach in which we analyse its
output: 90+ theses each with 100+ pages produced over the years for the
duration of the program between 2007-2021.

The findings will be presented in a journal article, accompanied by a
fully reproducible research compendium comprising an open dataset and
corresponding software, as well as a dashboard to enable users to
interact with the data and explore relations and patterns in the data
beyond the findings presented in the paper. The project is carried out
with in-kind support from the TU Delft Digital Competence Center.

\textbf{How will you manage versioning of your software?}

The software will be versioned with Git and hosted on GitHub to allow
for collaborative development. Although the research compendium is not a
software that will be versioned after publication, semantic versioning
will be used in the process leading up to publication, as follows:
development version 0.0.0.9000 in the research stage; patch version
0.0.1 with the publication of the preprint; minor version 0.1.0 and
subsequent increments of minor version at the time of submission for
review, and responses to reviews, respectively; and 1.0.0 for the
published paper.

\textbf{How will your software be documented for users? Please provide a
link to the documentation if available. How will you document your
software's contribution guidelines and governance structure?}

The research compendium will be documented as an R package, that is, all
functions used in data processing and analysis will have associated
documentation available from the help interface of R. Once loaded as a
package, all documentation is available to the user. The analysis will
be fully documented in a Quarto computational notebook using R Markdown
with the knitr engine.

\textbf{What licence will you give your software? How will you check
that it respects the licences of libraries and dependencies it uses?}

The software will use an MIT license.

\textbf{How will the installation requirements of your software be
documented? Please provide a link to the installation documentation if
available.}

Installation requirements are described in the
\href{https://github.com/UD3-Lab/mintEMU\#how-to-run-in-your-browser-or-download-and-run-locally}{README
file} of the repository on GitHub.

\textbf{How will users of your software be able to cite your software?
Please provide a link to your software citation file (CFF) if
available.}

Citation guidelines are described in the
\href{https://github.com/UD3-Lab/mintEMU\#how-to-cite}{README file} of
the repository on GitHub.

\textbf{How will your software be documented for future developers?}

All the functions are well documented and the following R package
structure so that they can be reused in other context. The repository
README file has also a
\href{https://github.com/UD3-Lab/mintEMU\#contributions}{Contributions}
section, where
\href{https://github.com/UD3-Lab/mintEMU/blob/main/CONTRIBUTING.md}{contributor
guidelines} and
\href{https://github.com/UD3-Lab/mintEMU/blob/main/CONDUCT.md}{Contributor
code of Conduct} can be easily accessed.

\textbf{How will your software be tested? Please provide a link to the
(automated) testing results.}

TBD

\textbf{Do you follow specific software quality guidelines? If yes,
which ones?}

FAIR4RS

\textbf{How will your software be packaged and distributed? Please
provide a link to available packaging information (e.g.~entry in a
packaging registry, if available).}

The software will be made available as an R package on GitHub, available
for direct installation from RStudio. As the package is s research
compendium, it will not be made available in a packaging registry.

The snapshot of the GitHub repository (e.g.~release associated with the
publication of the paper) will be published on Zenodo.

\textbf{How do you plan to procure long term maintenance of your
software?} The long-term maintenance of the software is not foreseen in
the project.The software is making use of the package management
software, \texttt{renv}, which help user retrieve the required version
of dependencies needed to use the functionalities of the software.

\end{document}

### 2023-03-17 R/clean.R

@@ -10,9 +10,9 @@ clean_basic <- function(text) {
 
   out_text <- text |>
     tolower() |>
-    str_replace_all("\\s{1,}", " ") |>
-    str_remove_all("[^0-9a-z ]") |>
-    str_squish()
+    stringr::str_replace_all("\\s{1,}", " ") |>
+    stringr::str_remove_all("[^0-9a-z ]") |>
+    stringr::str_squish()
 
   out_text
 
@@ -112,5 +112,7 @@ thesis_stopwords <- function(add_stopwords = NULL) {
     c(add_stopwords, "preface", "foreword", "introduction", "conclusion", "thesis")
 
 }
+
+
 
 

### 2023-03-17 analysis/paper/paper.qmd
minor changes

@@
 emu_theses <- 
-  read_csv(here(data_path, "theses-metadata.csv")) %>%
-  mutate(text = "") %>%
+  read_csv(here(data_path, "theses-metadata.csv")) |>
+  mutate(text = "") |>
   filter(!is.na(pdf_via))
@@
 # Preview theses in increasing order of graduation year
-emu_theses %>% 
-  arrange(graduation_year) %>% 
-  pull(text_clean) %>% 
+emu_theses |> 
+  arrange(graduation_year) |> 
+  pull(text_clean) |> 
   head_text(head_n = 5)
@@
 
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
+stop_words_custom <- emu_theses |> 
+  unnest_tokens(output = word, input = title) |>   ### add words from the title
+  select(word) |>
   rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
-                               "source", "author"))) %>% 
+                               "source", "author"))) |>
   # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
@@
 
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
+emu_theses_words <- emu_theses |>
+  mutate(author = paste(first_name, last_name)) |> 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) |> 
+  unnest_tokens(output = word, input = text) |>   # remove punctuation, convert to lowercase, seperate all words
+  anti_join(stop_words, by = "word") |>  # remove stop words
   anti_join(stop_words_custom, by = "word")
@@
   mutate(stem = wordStem(words = word, 
-                         language = "porter")) %>%  # Implement the Porter stemming algorithm for English provided by the SnowballC package
+                         language = "porter")) |>  # Implement the Porter stemming algorithm for English provided by the SnowballC package
   count(stem, sort = TRUE)
@@
 
-emu_theses_bigrams_top <- emu_theses_bigrams %>%
-  group_by(title) %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+emu_theses_bigrams_top <- emu_theses_bigrams |>
+  group_by(title) |>
+  count(bigram, sort = TRUE) |>
+  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
-  ungroup() %>% 
+  ungroup() |>
   mutate(bigram = reorder(bigram, n))
 
-emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
-  count(bigram, sort = TRUE) %>%
-  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+emu_theses_bigrams_top_all <- emu_theses_bigrams |>
+  count(bigram, sort = TRUE) |>
+  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
@@
 
-emu_theses_bigrams_top %>%
+emu_theses_bigrams_top |>
   filter(title == unique(title)[1:6]) |> 

### 2023-03-17 R/clean.R

@@ -11,7 +11,8 @@ clean_basic <- function(text) {
   out_text <- text |>
     tolower() |>
     stringr::str_replace_all("\\s{1,}", " ") |>
-    stringr::str_remove_all("[^0-9a-z ]") |>
+    stringr::str_remove_all("[[:punct:]]") |>
+    stringr::str_remove_all("\\b[0-9]+\\b") |>
     stringr::str_squish()
 
   out_text
@@ -97,8 +98,9 @@ find_meta_stopwords <- function(metadata,
 #' @export
 urbanism_stopwords <- function(add_stopwords = NULL) {
   stop_words <-
-    c(add_stopwords, "city", "urban", "urbanism", "hab", "km")
+    tibble::tibble(word = c(add_stopwords, "city", "urban", "urbanism", "hab", "km"))
 
+  stop_words
 }
 
 #' Adjustable list of stopwords associated with master theses
@@ -108,8 +110,20 @@ urbanism_stopwords <- function(add_stopwords = NULL) {
 #' @return A vector of thesis-related stopwords
 #' @export
 thesis_stopwords <- function(add_stopwords = NULL) {
-  stop_words <-
-    c(add_stopwords, "preface", "foreword", "introduction", "conclusion", "thesis")
+  stop_words <- tibble::tibble(
+    word =
+      c(
+        add_stopwords,
+        "preface",
+        "foreword",
+        "introduction",
+        "conclusion",
+        "thesis",
+        "source",
+        "author"
+      )
+  )
+  stop_words
 
 }
 

### 2023-03-20 analysis/paper/paper.qmd
Add note on how to answer SQ4
Update stopwords functions

--- 2023-03-17 analysis/paper/paper.qmd
@@
 # Load packages ----
-## List packages to be loaded
-packages <- c(
-  "here",                                        # Managing paths
-  "tidyverse",                                   # Data manipulation and visualisation
-  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
-  "tidygeocoder",                                # Working with spatial data
-  "igraph", "ggraph"                             # Network analysis and visualization
-)
+## Managing paths
+library(here)
 
-## Load required packages, install them if needed
-for (package in packages) {
-  if (!require(package, character.only = TRUE)) {
-    install.packages(package,  "http://cran.us.r-project.org")
-  }
-  library(package, character.only = TRUE)
-}
+## Data manipulation and visualisation
+library(tidyverse)
+
+## Text processing and analysis
+library(tidytext)
+library(SnowballC)
+library(tm)
+
+## Working with spatial data
+library(tidygeocoder)
 
@@
 emu_theses$text_clean <- emu_theses$text |>
-  clean_basic() |> 
-  str_remove_all(find_meta_stopwords(emu_theses) )  
-```
-
-```{r}
-#| label: preview
-#| include: false
-
-# Preview theses in increasing order of graduation year
-emu_theses %>% 
-  arrange(graduation_year) %>% 
-  pull(text_clean) %>% 
-  head_text(head_n = 5)
-
-# Show distribution of the length of theses
-hist(str_length(emu_theses$text_clean))
+  clean_basic() 
 ```
@@
 
-Distinctive features of the program:  
-- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice
-- provides a bridge towards PhD-level research
-- all students had an exchange semester at one of the other three program universities
-
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(theses_all)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(theses_all$grad_year, na.rm = TRUE), "-", max(theses_all$grad_year, na.rm = TRUE))`.
 The first year of the program was not represented, as PDF files were only available from 2008 onward.
@@
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
-With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 
-A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.
+A total number of `r nrow(emu_theses)` were analysed.
 
@@
 
-stop_words_custom <- emu_theses %>% 
-  unnest_tokens(output = word, input = title) %>%   ### add words from the title
-  select(word) %>% 
-  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
-                               "source", "author"))) %>% 
-  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
-  rbind(., data.frame(word = c("hab", "km")))
+emu_theses$text_clean <- emu_theses$text_clean |>
+  str_remove_all(find_meta_stopwords(emu_theses)) |> 
+  str_remove_all(urbanism_stopwords(
+    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
+  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))
 
-emu_theses_words <- emu_theses %>%
-  mutate(author = paste(first_name, last_name)) %>% 
-  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
-  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
-  anti_join(stop_words, by = "word") %>%  # remove stop words
-  anti_join(stop_words_custom, by = "word")
+emu_theses_words <- emu_theses |> 
+  select(title, text_clean) |> 
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
+  anti_join(tidytext::stop_words, by = "word")
 ```
@@
 ```{r}
-# Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_theses_words, 
-                           top_n = 3, 
-                           title_col = "title", 
-                           word_col = "word")
-```
+# Which are the top 20 most frequently used words in each thesis?
+emu_theses_top_20 <- emu_theses_words %>%
+  group_by(title) %>%
+  count(word, sort = TRUE) %>%
+  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
+  # filter(n > 50) %>%              ### filter words with a minimum count
+  ungroup() %>% 
+  mutate(word = reorder(word, n))
 
-```{r}
-# Which are the top n most frequently used words in all theses`?
-get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 5, 
-                         word_col = "word")
-```
+# Which are first 5 most frequently used words in each thesis?
+emu_theses_first_5 <- emu_theses_words %>%
+  count(title, word) %>% 
+  group_by(title) %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5)
 
-```{r}
-# Which are the top n most frequently used words in all theses`?
-emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 20, 
-                         word_col = "word")
-```
+# Which theses contain "fabric" in the top 20 most frequently used words?
+emu_theses_top_20 %>% 
+  filter(word == "fabric")
 
-<!-- ```{r} -->
+emu_theses_top_20 %>%
+  slice(1:40) %>% 
+  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
+  geom_col() +
+  xlab(NULL) +
+  coord_flip() +
+  labs(y = "Count",
+       x = "Unique words",
+       title = "Count of unique words found in thesis",
+       subtitle = "Stop words were removed from the list") +
+  facet_wrap( ~ title, scales = "free", ncol = 2)
 
-<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
+# # wip function ngrams
+# get_ngrams <- function(df, n) {
+#   if (!is.numeric(n)) {
+#     stop("n is not numeric")
+#   }
+#   emu_theses %>%
+#     mutate(author = paste(first_name, last_name)) %>% 
+#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
+#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+#     anti_join(stop_words, by = c("first" = "word")) %>%
+#     anti_join(stop_words, by = c("second" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
+#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
+#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+#     filter(first != second) %>% 
+#     filter(first != "source")
+# }
+# emu_theses_ngrams <- get_ngrams(emu_theses, 2)
 
-<!-- emu_theses_top_20 %>%  -->
-
-<!--   filter(word == "fabric") -->
-
-<!-- emu_theses_top_20 %>% -->
-
-<!--   slice(1:40) %>%  -->
-
-<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
-
-<!--   geom_col() + -->
-
-<!--   xlab(NULL) + -->
-
-<!--   coord_flip() + -->
-
-<!--   labs(y = "Count", -->
-
-<!--        x = "Unique words", -->
-
-<!--        title = "Count of unique words found in thesis", -->
-
-<!--        subtitle = "Stop words were removed from the list") + -->
-
-<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
-
-<!-- ``` -->
-
-```{r}
-#| label: ngrams
-
-## TO DO: generalise bigrams and trigrams into ngrams function
-
-## Bigrams
-emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
-
-bigram_counts <- emu_theses_bigrams |> 
-  count(first, second, sort = TRUE)
-
-bigram_counts
-
-## Trigrams
-emu_theses_trigrams <- emu_theses |> 
-  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
-  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
-  filter(!first %in% stop_words$word,
-         !second %in% stop_words$word,
-         !third %in% stop_words$word) |>  
-  filter(str_detect(first, "[a-z]"), 
-         str_detect(second, "[a-z]"),
-         str_detect(third, "[a-z]"))
-
-trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, sort = TRUE)
-
-trigram_counts
-
-## Bigram tf-idf
-bigram_tf_idf <- emu_theses_bigrams |> 
-  count(title, bigram) |> 
-  bind_tf_idf(bigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-bigram_tf_idf
-
-## Trigram tf-idf
-trigram_tf_idf <- emu_theses_trigrams |> 
-  count(title, trigram) |> 
-  bind_tf_idf(trigram, title, n) |> 
-  arrange(desc(tf_idf))
-
-trigram_tf_idf
-
-## Show network of bigrams
-bigram_graph <- bigram_counts |> 
-  filter(n > 100) |> 
-  graph_from_data_frame()
-
-bigram_graph
-
-set.seed(2023)
-
-a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
-
-ggraph(bigram_graph, layout = "fr") +
-  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
-  geom_node_point(color = "lightblue", size = 5) +
-  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
-  theme_void()
-
-## Higgh frequency bigrams ----
+# bigrams
+emu_theses_bigrams <- emu_theses %>%
+  mutate(author = paste(first_name, last_name)) %>% 
+  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
+  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
+  anti_join(stop_words, by = c("first" = "word")) %>%
+  anti_join(stop_words, by = c("second" = "word")) %>%
+  anti_join(stop_words_custom, by = c("first" = "word")) %>%
+  anti_join(stop_words_custom, by = c("second" = "word")) %>%
+  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
+  filter(first != second) %>% 
+  filter(first != "source")
 
@@
 emu_theses_bigrams_top %>%
-  filter(title == unique(title)[1:6]) |> 
   ggplot(aes(x = bigram, y = n, fill = title)) +

### 2023-03-20 R/clean.R

@@ -95,10 +95,8 @@ find_meta_stopwords <- function(metadata,
 #' @param add_stopwords A vector of words to be added to the urbanism stopwords
 #' @return A vector of urbanism-related stopwords
 #' @export
-subject_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
-  stop_words <- c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km",
-                  "emu", "european postgraduate master in urbanism", "advisor",
-                  "prof")
+urbanism_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
+  stop_words <- c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km")
 
   if (convert_to_regex) return(paste(stop_words, collapse = "|"))
   else return(stop_words)
@@ -110,7 +108,7 @@ subject_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
 #' @param add_stopwords A vector of words to be added to the master theses-related stopwords
 #' @return A vector of thesis-related stopwords
 #' @export
-document_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
+thesis_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
   stop_words <- c(add_stopwords, "preface", "foreword", "introduction", "conclusion",
                   "thesis", "colophon", "acknowledgements", "references")
 

### 2023-03-20 analysis/paper/paper.qmd

--- 2023-03-17 analysis/paper/paper.qmd
@@
   "tidygeocoder",                                # Working with spatial data
-  "igraph", "ggraph"                             # Network analysis and visualization
+  "igraph", "ggraph"                             # Network analysis and visualsation
 )
@@
 #| echo: false
-#| cache: true
 
-data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
+# Extract text from PDFs
+# On the first run, this function might take extra time as a Python environment needs to be set up
 
-if (!file.exists(data_path)) {
-  # On the first run, this function might take extra time as a Python environment needs to be set up
 emu_theses$text <- convert_pdf_text(pdf_paths)
-  write_csv(emu_theses, data_path)
-} else {
-  emu_theses <- read_csv(data_path)
-}
-
 ```

### 2023-03-20 analysis/paper/paper.qmd

@@
 -   To what extent were the topics and evolution thereof influenced by the exchange semester followed by the students?
+    - Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
+Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
 

### 2023-03-20 analysis/paper/paper.qmd

@@
   "here",                                        # Managing paths
-  "tidyverse",                                   # Data manipulation and visualisation
+  "tidyverse", "reshape2",                       # Data manipulation and visualisation
   "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
@@
 
-Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+
+```{r}
+# The topicmodels package requires a document-term matrix as input
+emu_theses_dtm <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  cast_dtm(title, word, n)
+
+emu_theses_dtm
+
+# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+# In LDA, each document is a mixture of topics and each topic is a mixture of words
+emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
+emu_theses_lda
+
+# Tidy model with beta statistic
+emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
+emu_theses_topics
+
+# The top 10 words that are the most common in each topic
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
+
+emu_theses_top_terms |> 
+  mutate(term = reorder(term, beta)) |> 
+  ggplot(aes(term, beta, fill = factor(topic))) +
+  geom_col(show.legend = FALSE) +
+  facet_wrap(~ topic, scales = "free") +
+  coord_flip()
+
+# 
+beta_spread <- emu_theses_topics |> 
+  mutate(topic = paste0("topic", topic)) |> 
+  spread(topic, beta) |> 
+  filter(topic1 > 0.001, topic2 > .001) |> 
+  mutate(log_ratio = log2(topic2/topic1))
+
+beta_spread
+
+beta_spread %>%
+  group_by(direction = log_ratio > 0) %>%
+  slice_max(abs(log_ratio), n = 20) %>% 
+  ungroup() %>%
+  mutate(term = reorder(term, log_ratio)) %>%
+  ggplot(aes(log_ratio, term)) +
+  geom_col() +
+  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
+
+emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
+emu_theses_documents
+
+emu_theses_documents |> 
+  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
+  geom_bar(stat = "identity") + 
+  coord_flip()
+
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
+  geom_bar(stat = "identity") + 
+  coord_flip() + 
+  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)
+
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
+  geom_col()
+```
 

### 2023-03-20 analysis/paper/paper.qmd

@@
   anti_join(tidytext::stop_words, by = "word")
+
 ```

### 2023-03-20 R/clean.R
Fix typo

@@ -52,7 +52,7 @@ find_meta_stopwords <- function(metadata,
                                   'author2' = c('first_name_2', 'last_name_2'),
                                   'title'  = 'title',
                                   'subtitle' = 'subtitle',
-                                  'full_title' = 'full_title                       
+                                  'full_title' = 'full_title'
                                 ),
                                 convert_to_regex = TRUE) {
   cols_meta <- names(metadata)
@@ -123,7 +123,7 @@ thesis_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
         "source",
         "author",
         "colophon",
-        "acknowledgements", 
+        "acknowledgements",
         "references"
         )
 

### 2023-03-20 R/geocode.R
Update geocode function

@@ -9,7 +9,7 @@
 #' @export
 geocode_thesis_locations <- function(data, location = "location") {
   if (!("latitude" %in% names(data)) | !("longitude" %in% names(data))) {
-    geocoded_data <- tidygeocoder::geocode(
+    data <- tidygeocoder::geocode(
       data,
       address = location,
       method = 'osm',
@@ -17,7 +17,7 @@ geocode_thesis_locations <- function(data, location = "location") {
       long = longitude
     )
   }
-  geocoded_data
+  data
 }
 
 

### 2023-03-20 tests/testthat/test-analyse.R
Add test test

@@ -1,0 +1,33 @@
+test_that("get_top_words_per_document returns top word for each title", {
+  df_input <-
+    tibble::tibble(
+      title = c(rep("t1", 6), rep("t2", 6), rep("t3", 7)),
+      word = c(
+        rep("w1", 3),
+        rep("w2", 2),
+        rep("w3", 1),
+        rep("w1", 2),
+        rep("w2", 3),
+        rep("w3", 1),
+        rep("w1", 4),
+        rep("w2", 2),
+        rep("w3", 1)
+      ))
+
+  expected_output <- tibble::tibble(
+    title = c("t1", "t2", "t3"),
+    word = c("w1", "w2", "w1"),
+    n = c(3, 3, 4))
+
+  df_output <-
+    get_top_words_per_document(
+      df_input,
+      top_n = 1,
+      title_col = "title",
+      word_col = "word",
+      min_count = 1
+    )
+
+  expect_equal(df_output, expected_output)
+
+})

### 2023-03-21 analysis/paper/paper.qmd
Update geocode function
Add function to convert data frame to document term matrix

--- 2023-03-20 analysis/paper/paper.qmd
@@
 ```{r}
-# The topicmodels package requires a document-term matrix as input
-emu_theses_dtm <- emu_theses_words |> 
-  count(title, word, sort = TRUE) |> 
-  cast_dtm(title, word, n)
-
+# The {topicmodels} package requires a document-term matrix as input
+emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
 emu_theses_dtm

### 2023-03-21 R/analyse.R

@@ -44,3 +44,19 @@ get_top_words_per_corpus <- function(data, top_n, word_col = NULL) {
 
   data_top_n
 }
+
+#' Convert data frame with word frequencies into document term matrix
+#'
+#' @param data Data frame containing colomns with titles, words and word counts
+#' @param title_col Name of the title column of the input data frame
+#' @param word_col Name of the word column of the input data frame
+#'
+#' @return A document term matrix
+#' @export
+convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
+  data_dtm <- data |>
+    dplyr::count(!!rlang::sym(title_col), !!rlang::sym(word_col), sort = TRUE) |>
+    tidytext::cast_dtm(!!rlang::sym(title_col), !!rlang::sym(word_col), n)
+
+  data_dtm
+}

### 2023-03-21 analysis/paper/paper.qmd
Add function to get top words per topic

--- 2023-03-20 analysis/paper/paper.qmd
@@
 # Add latitude and longitude information to theses data frame
-emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)
+emu_theses <- geocode_thesis_locations(emu_theses)
 
@@
 # The top 10 words that are the most common in each topic
-emu_theses_top_terms <- emu_theses_topics |> 
-  group_by(topic) |> 
-  top_n(20, beta) |> 
-  ungroup() |> 
-  arrange(topic, -beta)
+emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20, "topic", "beta")
 

### 2023-03-21 R/analyse.R

@@ -44,3 +44,22 @@ get_top_words_per_corpus <- function(data, top_n, word_col = NULL) {
 
   data_top_n
 }
+
+#' Extract top n terms with the highest count across all topics
+#'
+#' @param data Data frame with topics, terms and beta statistic.
+#' @param top_n The number of top words to be returned by the function.
+#' @param topic_col Name of the column containing topics.
+#' @param beta_col Name of the column containing the beta statistic for all terms.
+#'
+#' @return A data frame with the top n terms across all topics.
+#' @export
+get_top_words_per_topic <- function(data, top_n, topic_col = NULL, beta_col = NULL) {
+  data_top_terms <- data |>
+    dplyr::group_by(!!rlang::sym(topic_col)) |>
+    dplyr::top_n(top_n, !!rlang::sym(beta_col)) |>
+    dplyr::ungroup() |>
+    dplyr::arrange(!!rlang::sym(topic_col), -!!rlang::sym(beta_col))
+
+  data_top_terms
+}

### 2023-03-21 analysis/paper/paper.qmd
Update other top words functions

@@
 # The top 10 words that are the most common in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20, "topic", "beta")
+emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)
 
 emu_theses_top_terms |> 
-  mutate(term = reorder(term, beta)) |> 
-  ggplot(aes(term, beta, fill = factor(topic))) +
+  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
   geom_col(show.legend = FALSE) +
@@
 get_top_words_per_document(data = emu_theses_words, 
-                           top_n = 3, 
-                           title_col = "title", 
-                           word_col = "word")
+                           top_n = 3)
 ```
@@
 get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 5, 
-                         word_col = "word")
+                         top_n = 5)
 ```
@@
 # Which are the top n most frequently used words in all theses`?
-emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 20, 
-                         word_col = "word")
+get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 20)
 ```

### 2023-03-21 R/analyse.R

@@ -12,10 +12,17 @@
 #'
 #' @return A data frame with the top n words with a count above the given minimum.
 #' @export
-get_top_words_per_document <- function(data, top_n, title_col = NULL, word_col = NULL, min_count = 1) {
+get_top_words_per_document <- function(data,
+                                       top_n,
+                                       title_col = "title",
+                                       word_col = "word",
+                                       min_count = 1) {
+  title <- sym(title_col)
+  word <- sym(word_col)
+
   data_top_n <- data |>
-    dplyr::group_by(!!sym(title_col)) |>
-    dplyr::count(!!sym(word_col), sort = TRUE) |>
+    dplyr::group_by(!!title) |>
+    dplyr::count(!!word, sort = TRUE) |>
     dplyr::slice_max(n, n = top_n) |>
     dplyr::filter(n > min_count) |>
     dplyr::ungroup()
@@ -35,10 +42,12 @@ get_top_words_per_document <- function(data, top_n, title_col = NULL, word_col =
 #'
 #' @return A data frame with the top n words across all documents.
 #' @export
-get_top_words_per_corpus <- function(data, top_n, word_col = NULL) {
+get_top_words_per_corpus <- function(data, top_n, word_col = "word") {
+  word <- sym(word_col)
+
   data_top_n <- data |>
-    dplyr::group_by(!!sym(word_col)) |>
-    dplyr::count(!!sym(word_col), sort = TRUE) |>
+    dplyr::group_by(!!word) |>
+    dplyr::count(!!word, sort = TRUE) |>
     dplyr::ungroup() |>
     dplyr::slice_max(n, n = top_n)
 
@@ -54,12 +63,18 @@ get_top_words_per_corpus <- function(data, top_n, word_col = NULL) {
 #'
 #' @return A data frame with the top n terms across all topics.
 #' @export
-get_top_words_per_topic <- function(data, top_n, topic_col = NULL, beta_col = NULL) {
+get_top_words_per_topic <- function(data,
+                                    top_n,
+                                    topic_col = "topic",
+                                    beta_col = "beta") {
+  topic <- rlang::sym(topic_col)
+  beta <- rlang::sym(beta_col)
+
   data_top_terms <- data |>
-    dplyr::group_by(!!rlang::sym(topic_col)) |>
-    dplyr::top_n(top_n, !!rlang::sym(beta_col)) |>
[…]

### 2023-03-21 analysis/paper/paper.qmd
Remove duplicate analysis

--- 2023-03-21 analysis/paper/paper.qmd
@@
 # The top 10 words that are the most common in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20, "topic", "beta")
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
 

### 2023-04-17 analysis/paper/paper.qmd
add short words function
Update topic model functions

--- 2023-03-21 analysis/paper/paper.qmd
@@
 
+Additional checks were made to the document in search for previously undetected stop words.
+The checks included identifying words consisting of four or less characters that extensively used in the text corpus.
+
+```{r}
+#| label: short-stop-words
+
+#short_words(emu_theses$text_clean, scope = 'corpus')
+short_words(emu_theses_words$word)
+
+```
+
 Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
@@
 ```{r}
-# The topicmodels package requires a document-term matrix as input
-emu_theses_dtm <- emu_theses_words |> 
-  count(title, word, sort = TRUE) |> 
-  cast_dtm(title, word, n)
+emu_theses_topics <- get_topics(emu_theses_words)
 
-emu_theses_dtm
-
-# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
-# In LDA, each document is a mixture of topics and each topic is a mixture of words
-emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
-emu_theses_lda
-
-# Tidy model with beta statistic
-emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
-emu_theses_topics
+emu_theses_topics_terms <- tidy(emu_theses_topics, matrix = "beta")
+emu_theses_topics_documents <- tidy(emu_theses_topics, matrix = "gamma")
 
 # The top 10 words that are the most common in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20, "topic", "beta")
+emu_theses_top_terms <- emu_theses_topics_terms |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
 
@@
 # 
-beta_spread <- emu_theses_topics |> 
+beta_spread <- emu_theses_topics_terms |> 
   mutate(topic = paste0("topic", topic)) |> 
@@
 
-emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
-emu_theses_documents
-
-emu_theses_documents |> 
+emu_theses_topics_documents |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 
-emu_theses_documents |> 
+emu_theses_topics_documents |> 
   inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 

### 2023-04-17 R/analyse.R

@@ -60,3 +60,34 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 
   data_dtm
 }
+
+#' Get topics for a given data frame of documents and words
+#'
+#' @param data Data frame with theses, containing a column with words.
+#' @param k Number of desired clusters.
+#' @param matrix Statistics to be computed. Can be "beta" or "gamma"
+#' @param title_col The name of the column from the input data frame containing the title of the document.
+#' @param word_col Name of the column from the input data frame containing all words.
+#' @param seed Seed to control pseudorandomisation.
+#'
+#' @return An LDA model object.
+#' @export
+get_topics <-
+  function(data,
+           k = 2,
+           seed = 2023,
+           title_col = "title",
+           word_col = "word") {
+    stopifnot("k must be an integer of at least 2" = k >= 2)
+
+    # The {topicmodels} package requires a document-term matrix as input
+    dtm <- convert_to_dtm(data,
+                          title_col = title_col,
+                          word_col = word_col)
+
+    # We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+    # In LDA, each document is a mixture of topics and each topic is a mixture of words
+    lda <- topicmodels::LDA(dtm,
+                            k = k,
+                            control = list(seed = seed))
+}

### 2023-04-17 tests/testthat/test-analyse.R

@@ -31,3 +31,7 @@ test_that("get_top_words_per_document returns top word for each title", {
   expect_equal(df_output, expected_output)
 
 })
+
+test_that("get_topics throws an error if k is less than 2", {
+  expect_error(get_topics(mtcars, k = 1))
+})

### 2023-04-17 R/convert.R
added page numbers

@@ -6,7 +6,7 @@
 #' @param pdf_filenames  A vector of paths to input (pdf) files
 #' @return A vector of body of text; one string per pdf file
 #' @export
-convert_pdf_text <-function(pdf_filenames){
+convert_pdf_text <-function(pdf_filenames, by_page = TRUE){
 
   pdf_list<- as.list(pdf_filenames)
 
@@ -24,7 +24,10 @@ convert_pdf_text <-function(pdf_filenames){
   # Load python script with functions to extract texts
   reticulate::source_python(here::here("inst/python","extract-text-pdf-python.py"))
 
-  text_list <- convert_pdf(pdf_list) %>% unlist()
+  if (by_page == TRUE)
+    text_list <- convert_pdf(pdf_list) %>% unlist()
+  else
+    text_list <- convert_pdf(pdf_list, by_page = reticulate::r_to_py(TRUE)) # %>% unlist()
 
   text_list
 }

### 2023-04-17 R/extract-text-pdf-python.py

@@ -48,7 +48,7 @@ def check_file_path(file_path):
 
         return(file_list)
         
-def convert_pdf(file_list):
+def convert_pdf(file_list, by_page = True):
     '''
     Function converts the pdf files to text. It allow unusual pdf layout (multiple columns)
 
@@ -71,7 +71,8 @@ def convert_pdf(file_list):
             for page in doc:
                 text = page.get_text()
                 l.append(text)
-        l = ' '.join(l)
+        if not by_page
+          l = ' '.join(l)
         text_list.append(l)
     
     return(text_list)

### 2023-04-17 inst/python/extract-text-pdf-python.py

@@ -48,7 +48,7 @@ def check_file_path(file_path):
 
         return(file_list)
         
-def convert_pdf(file_list):
+def convert_pdf(file_list, by_page = True):
     '''
     Function converts the pdf files to text. It allow unusual pdf layout (multiple columns)
 
@@ -71,11 +71,11 @@ def convert_pdf(file_list):
             for page in doc:
                 text = page.get_text()
                 l.append(text)
-        l = ' '.join(l)
+        if not by_page:
+          l = ' '.join(l)
         text_list.append(l)
     
     return(text_list)
-
 
 
 

### 2023-04-21 analysis/paper/paper.qmd
Add description of topic modeling as a specific type of clustering method

@@
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+An important distinction from other clustering methods like k-means, is that topic models allow for mixed membership.
+This means that a topic is a probability distribution over a vocabulary and that all words have a probability of being part of a topic.
+Typically, words with highest probablities within a topic will be examined to interpret and describe the meaning of a topic.
 

### 2023-04-24 analysis/paper/paper.qmd
Update code chunk labels and fix references

--- 2023-03-21 analysis/paper/paper.qmd
@@
 
-```{r}
-#| label: setup
+```{r setup}
 #| echo: false
@@
 
-```{r}
-#| label: extract-pdf-text
+```{r extract-pdf-text}
 #| echo: false
@@
 
-```{r}
-#| label: clean-text
+```{r clean-text}
 #| echo: false
@@
 
-```{r}
-#| label: fig-map-theses
+```{r fig-map-theses}
 #| warning: false 
@@
 
-Figure @fig-map-theses shows the geographic spread of the theses included in the analysis.
+@fig-map-theses shows the geographic spread of the theses included in the analysis.
 
 ## Data analysis
-
-[@silge2017a] [@hvitfeldt2021a]
 
@@
 
-```{r}
-#| label: stop-words
+```{r stop-words}
 
@@
 
+Additional checks were made to the document in search for previously undetected stop words.
+The checks included identifying words consisting of four or less characters that extensively used in the text corpus.
+
+```{r}
+#| label: short-stop-words
+
+#short_words(emu_theses$text_clean, scope = 'corpus')
+short_words(emu_theses_words$word)
+
+```
+
 Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
 
-```{r}
-#| label: stemming
+```{r stemming}
 

### 2023-04-24 data-raw/emu_text.R
Update package data and script

@@ -1,9 +1,17 @@
-# Get dataset URL ---------------------------------------------------------
-url_emu_text <- ""
+# Download and import the dataset -----------------------------------------
+emu_text_url <- here::here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
+emu_text_raw_df <- readr::read_csv(emu_text_url)
 
-# Download the dataset ----------------------------------------------------
-emu_text <- readr::read_csv(url_emu_text)
+# Clean and geocode the dataset -------------------------------------------
 
-readr::write_csv(emu_text, here::here("inst", "extdata", "emu_text.csv"))
+emu_text_df <- mintEMU::geocode_thesis_locations(emu_text_raw_df)
 
-usethis::use_data(emu_text, overwrite = TRUE)
+emu_text_df$text_clean <- emu_text_df$text |>
+  mintEMU::clean_basic() |>
+  stringr::str_remove_all(mintEMU::find_meta_stopwords(emu_theses)) |>
+  stringr::str_remove_all(mintEMU::urbanism_stopwords(
+    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
+  stringr::str_remove_all(mintEMU::thesis_stopwords(
+    add_stopwords = c("advisor", "prof")))
+
+usethis::use_data(emu_text_df, overwrite = TRUE)

### 2023-04-24 analysis/paper/paper.qmd
Update package data and script
Update paper author

@@
 author:
-  - Jane Doe:
+  - Claudiu Forgaci:
       correspondence: "yes"
       email: janedoe@fosg.org
-      orcid: 0000-0003-1689-0557
-      institute: tud
-  - John Q. Doe:
-      email: johndoe@fosg.org
-      orcid: 0000-0003-1689-0558
+      orcid: 0000-0003-3218-5102
       institute: tud

### 2023-06-14 analysis/paper/paper.qmd
Update author email address

@@
       correspondence: "yes"
-      email: janedoe@fosg.org
+      email: C.Forgaci@tudelft.nl
       orcid: 0000-0003-3218-5102

### 2023-06-14 analysis/paper/paper.qmd
Update cache chunk
If the raw data or metadata changes, it should not be taken from cache.

@@
 } else {
-  emu_theses <- read_csv(data_path)
+  emu_theses$text <- read_csv(data_path)$text
 }

### 2023-06-22 analysis/paper/paper.qmd
Remove statement on the absence of theses from 2007
Create ngrams function

@@
 To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
-The first year of the program was not represented, as PDF files were only available from 2008 onward.
 
@@
 
-## TO DO: generalise bigrams and trigrams into ngrams function
+## Add lemmas as optional column to ngram function
 
@@
 emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
+  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
+
+emu_theses_bigrams |>
+  dplyr::mutate(first_lemma =  |> tidytext::unnest_tokens())
+
+spacyr::spacy_parse(emu_theses_bigrams$first)
 
@@
 
+bigram_stem_counts <- emu_theses_bigrams |> 
+  count(first_stem, second_stem, sort = TRUE)
+
 bigram_counts
+bigram_stem_counts
 
@@
 emu_theses_trigrams <- emu_theses |> 
-  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
-  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
-  filter(!first %in% stop_words$word,
-         !second %in% stop_words$word,
-         !third %in% stop_words$word) |>  
-  filter(str_detect(first, "[a-z]"), 
-         str_detect(second, "[a-z]"),
-         str_detect(third, "[a-z]"))
+  get_ngrams(n = 3, title_col = "title", text_col = "text")
 
 trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, sort = TRUE)
+  count(first, second, third, sort = TRUE)
 
@@
 bigram_tf_idf <- emu_theses_bigrams |> 
-  count(title, bigram) |> 
-  bind_tf_idf(bigram, title, n) |> 
+  count(title, ngram) |> 
+  bind_tf_idf(ngram, title, n) |> 
   arrange(desc(tf_idf))
@@
 trigram_tf_idf <- emu_theses_trigrams |> 
-  count(title, trigram) |> 
-  bind_tf_idf(trigram, title, n) |> 
+  count(title, ngram) |> 
+  bind_tf_idf(ngram, title, n) |> 
   arrange(desc(tf_idf))
@@
   geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
-  geom_node_point(color = "lightblue", size = 5) +
+  geom_node_point(color = "lightblue", linewidth = 5) +
   geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
@@
   group_by(title) |>
-  count(bigram, sort = TRUE) |>
+  count(ngram, sort = TRUE) |>
   slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
@@
   ungroup() |>
-  mutate(bigram = reorder(bigram, n))
+  mutate(bigram = reorder(ngram, n))
 
 emu_theses_bigrams_top_all <- emu_theses_bigrams |>
-  count(bigram, sort = TRUE) |>
+  count(ngram, sort = TRUE) |>
   slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
-  mutate(bigram = reorder(bigram, n))
+  mutate(ngram = reorder(ngram, n))
 
 emu_theses_bigrams_top |>
-  filter(title == unique(title)[1:6]) |> 
-  ggplot(aes(x = bigram, y = n, fill = title)) +
+  filter(title == unique(title)[1:10]) |> 
+  ggplot(aes(x = ngram, y = n, fill = title)) +
   geom_col() +

### 2023-06-22 R/analyse.R

@@ -71,8 +71,13 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 #'
 #' @return A data frame with an "ngram" column and columns for its constituent words
 #' @export
-get_ngrams <- function(data, n, title_col = NULL, text_col = NULL, stem = FALSE) {
-  data_ngrams <- data |>
+get_ngrams <-
+  function(data,
+           n,
+           title_col = NULL,
+           text_col = NULL,
+           stem = FALSE) {
+    data_ngrams <- data |>
     select(title_col, text_col) |>
     tidytext::unnest_tokens(ngram, text_col, token = "ngrams", n = n) |>
     tidyr::separate(ngram, into = c("first", "second", "third"),

### 2023-06-22 analysis/paper/paper.qmd
Add custom urbanism stopwords

--- 2023-06-14 analysis/paper/paper.qmd
@@
   str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
+    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
   str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

### 2023-07-03 analysis/paper/paper.qmd
Create function to get tf-idf statistic

@@
 
+The `tf-idf` statistic shows words that are the most important to one document in a collection of documents [@silge2017].
+In this case, it shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
+
+```{r tf-idf}
+## Calculate tf-idf statistic
+emu_theses_tf_idf <- emu_theses_words |> 
+  get_tf_idf(title_col = "title", word_col = "word")
+
+## Get top n words most specific to each thesis 
+emu_theses_tf_idf |> 
+  group_by(title) |> 
+  top_n(5, tf_idf) |> 
+  arrange(desc(tf_idf), .by_group = TRUE)
+```
+
 ```{r}
-#| label: tf-idf
-
-emu_theses_words
-
-emu_theses_words_count <- emu_theses_words |> 
-  count(title, word, sort = TRUE) |> 
-  ungroup()
-
-emu_theses_words_total <- emu_theses_words_count |> 
-  group_by(title) |> 
-  summarise(total = sum(n))
-
-emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)
-
-# emu_theses_words_count |> 
-#   group_by(title) |> 
-#   summarise(total = unique(tot
-
 # Preview term frequency distributions in a subset of theses
@@
   facet_wrap(~ title, ncol = 3, scales = "free_y")
+```
 
+```{r}
 # Examine Zipf's law for the EMU theses
@@
   scale_y_log10()
-
-## Calculate tf-idf
-emu_theses_words_count <- emu_theses_words_count |> 
-  bind_tf_idf(word, title, n)
-emu_theses_words_count
-
-emu_theses_words_count |> 
-  select(-total) |> 
-  arrange(desc(tf_idf))
 

### 2023-07-03 R/analyse.R

@@ -60,3 +60,29 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 
   data_dtm
 }
+
+#' Get `tf-idf` statistic
+#'
+#' @param data Data frame containing colomns with titles, words and word counts
+#' @param title_col Name of the title column of the input data frame
+#' @param word_col Name of the word column of the input data frame
+#'
+#' @return A data frame including the `tf-idf` statistic for each word
+#' @export
+get_tf_idf <- function(data, title_col = NULL, word_col = NULL) {
+  words_counts <- data |>
+    count(!!rlang::sym(title_col), !!rlang::sym(word_col), sort = TRUE) |>
+    ungroup()
+
+  words_total <- words_counts |>
+    group_by(!!rlang::sym(title_col)) |>
+    summarise(total = sum(n))
+
+  words_counts <- left_join(words_counts, words_total)
+
+  words_counts <- words_counts |>
+    bind_tf_idf(!!rlang::sym(word_col), !!rlang::sym(title_col), n) |>
+    select(-total)
+
+  words_counts
+}

### 2023-07-03 analysis/paper/paper.qmd
Add script to compute word embeddings

--- 2023-06-22 analysis/paper/paper.qmd
@@
 
-The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
-As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
+```{r word-embeddings}
+
+tidy_emu_theses <- emu_theses |> 
+  select(title, text_clean) |> 
+  unnest_tokens(word, text_clean) |> 
+  add_count(word) |> 
+  filter(n >= 50) |> 
+  select(-n)
+
+nested_words <- tidy_emu_theses |> 
+  nest(words = c(word))
+
+nested_words
+
+slide_windows <- function(tbl, window_size) {
+  skipgrams <- slider::slide(
+    tbl,
+    ~.x,
+    .after = window_size - 1,
+    .step = 1,
+    .complete = TRUE
+  )
+
+  safe_mutate <- safely(mutate)
+
+  out <- map2(skipgrams,
+              1:length(skipgrams),
+              ~ safe_mutate(.x, window_id = .y))
+
+  out |> 
+    transpose() |> 
+    pluck("result") |> 
+    compact() |> 
+    bind_rows()
+}
+
+library(widyr)
+library(furrr)
+
+plan(multisession)
+
+tidy_pmi <- nested_words |> 
+  mutate(words = future_map(words, slide_windows, 10L)) |> 
+  unnest(words) |> 
+  unite(window_id, title, window_id) |> 
+  pairwise_pmi(word, window_id)
+
+tidy_pmi
+
+tidy_word_vectors <- tidy_pmi |> 
+  widely_svd(
+    item1, item2, pmi,
+    nv = 100, maxit = 1000
+  )
+
+tidy_word_vectors
+
+nearest_neighbors <- function(df, token) {
+  df |> 
+    widely(
+      ~ {
+        y <- .[rep(token, nrow(.)), ]
+        res <- rowSums(. * y) /
+          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))
+
+        matrix(res, ncol = 1, dimnames = list(x = names(res)))
+      },
+      sort = TRUE
+    )(item1, dimension, value) |> 
+    select(-item2)
+}
+
+tidy_word_vectors |> 
+  nearest_neighbors("resilience")
+
+tidy_word_vectors |> 
+  filter(dimension <= 24) |> 
+  group_by(dimension) |> 
+  top_n(12, abs(value)) |> 
+  ungroup() |> 
+  mutate(item1 = reorder_within(item1, value, dimension)) |> 
+  ggplot(aes(item1, value, fill = dimension)) +
+  geom_col(alpha = 0.8, show.legend = FALSE) +
+  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
+  scale_x_reordered() +
+  coord_flip() +
+  labs(
+    x = NULL,
+    y = "Value",
+    title = "First 24 principal components for text of theses",
+    subtitle = "Top words contributing to the components that explain the most variation"
+  )
+
+word_matrix <- tidy_emu_theses |> 
+  count(title, word) |> 
+  cast_sparse(title, word, n)
+
+embedding_matrix <- tidy_word_vectors |> 
+  cast_sparse(item1, dimension, value)
+
+doc_matrix <- word_matrix %*% embedding_matrix
+
+dim(doc_matrix)
+
+doc_matrix
+```
 

### 2023-07-03 analysis/paper/paper.qmd
Write geocoded data frame to file

--- 2023-06-22 analysis/paper/paper.qmd
@@
 
-```{r extract-pdf-text}
+```{r get-pdf-text}
 #| echo: false
@@
 
-data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
+data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")
 
@@
   emu_theses$text <- convert_pdf_text(pdf_paths)
+
+  # Add latitude and longitude information to theses data frame
+  emu_theses <- geocode_thesis_locations(emu_theses)
+
+  # Write thesis dataframe to file
   write_csv(emu_theses, data_path)
@@
   emu_theses$text <- read_csv(data_path)$text
+  emu_theses$longitude <- read_csv(data_path)$longitude
+  emu_theses$latitude <- read_csv(data_path)$latitude
 }
@@
 #| fig-cap: "Location of EMU theses included in the analysis"
-
-# Add latitude and longitude information to theses data frame
-emu_theses <- geocode_thesis_locations(emu_theses)
 

### 2023-07-03 R/clean.R
created anonymysation checking function

@@ -272,7 +272,7 @@ normalise_words <-function(string_vec) {
 
   words <- mapply(c, words, normalised_words, SIMPLIFY=FALSE)
 
-  words <- lapply(words, unique) |> unlist()
+  words <- lapply(words, function(x) paste(unique(x), collapse = " ") ) |> unlist()
 
   words
 

### 2023-07-03 analysis/supplementary-materials/prepare-for-publication.R

@@ -11,18 +11,100 @@ data("emu_theses")
 emu_theses <- emu_theses |>
   arrange(title) |>
   mutate(ID  = seq_len(n()),
+# remove email addresses
          text_raw = str_replace_all(text,
-                                word_vec_combos(normalise_words(emu_theses$first_name),
-                                                normalise_words(emu_theses$last_name)),
-                                "AUTHOR_REMOVED"
-                                )
-         )
+                                    "\\b[A-Za-z0-9._%+-]+@[A-Za-z0-9.-]+\\.[A-Za-z]{2,}\\b" ),
+                                    "EMAIL_REMOVED")
 
-head_text(emu_theses$text)
+# Check if email addresses removed:
+str_extract_all(emu_theses$text_raw, "@")
 
-# Anonymyse raw text -----------
+anonymisation_checker <- function(text_vector, check_pattern ){
 
-emu_theses
+  occur <- str_extract_all(text_vector, check_pattern)
+  not_removed <- which(lengths(occur)>0)
+  loc <- str_locate_all(text_vector, check_pattern )
+
+  if(length(check_pattern) == 1 )
+    check_pattern <- rep(check_pattern, length(text_vector) )
+
+  for (nr in not_removed) {
+    n_occs <- nrow(loc[[nr]])
+
+    print(paste("PATTERN:", check_pattern[[nr]]))
+
+    for (occ in 1:n_occs) {
+      print(paste(
+        "NON_REMOVED,  NR",
+        nr,
+        ",",
+        occ ,
+        ":",
+        str_sub(text_vector[nr], loc[[nr]][occ] - 25, loc[[nr]][occ] +
+                  25)
+      ))
+    }
+  }
+
+}
+
+anonymisation_checker(emu_theses$text_raw, "@")
+
+# Anonymise raw text -----------
+text_raw <- mapply(str_replace_all, emu_theses$text_raw,
+                  word_vec_combos(normalise_words(emu_theses$first_name),
+                                  normalise_words(emu_theses$last_name)),
+                  "AUTHOR_REMOVED")
+
+# Check the Anonymisation
+head_text(text_raw)
+
+# Checking if still some first names prevailed
+str_extract_all(emu_theses$text, emu_theses$first_name)
+str_extract_all(text_raw, emu_theses$first_name)
+
+
[…]

### 2023-07-03 analysis/paper/paper.qmd
Update description and comments in topic model code chunk

@@
 author:
-  - Claudiu Forgaci:
+  - Jane Doe:
       correspondence: "yes"
-      email: C.Forgaci@tudelft.nl
-      orcid: 0000-0003-3218-5102
+      email: janedoe@fosg.org
+      orcid: 0000-0003-1689-0557
+      institute: tud
+  - John Q. Doe:
+      email: johndoe@fosg.org
+      orcid: 0000-0003-1689-0558
       institute: tud
@@
 
-```{r setup}
+```{r}
+#| label: setup
 #| echo: false
@@
 
-```{r get-pdf-text}
+```{r}
+#| label: extract-pdf-text
 #| echo: false
@@
 
-data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")
+data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
 
@@
   emu_theses$text <- convert_pdf_text(pdf_paths)
-
-  # Add latitude and longitude information to theses data frame
-  emu_theses <- geocode_thesis_locations(emu_theses)
-
-  # Write thesis dataframe to file
   write_csv(emu_theses, data_path)
 } else {
-  emu_theses$text <- read_csv(data_path)$text
-  emu_theses$longitude <- read_csv(data_path)$longitude
-  emu_theses$latitude <- read_csv(data_path)$latitude
+  emu_theses <- read_csv(data_path)
 }
@@
 
-```{r clean-text}
+```{r}
+#| label: clean-text
 #| echo: false
@@
 To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+The first year of the program was not represented, as PDF files were only available from 2008 onward.
 
@@
 
-```{r fig-map-theses}
+```{r}
+#| label: fig-map-theses
 #| warning: false 
@@
 #| fig-cap: "Location of EMU theses included in the analysis"
+
+# Add latitude and longitude information to theses data frame
+emu_theses <- geocode_thesis_locations(emu_theses)
 
@@
 
-@fig-map-theses shows the geographic spread of the theses included in the analysis.
+Figure @fig-map-theses shows the geographic spread of the theses included in the analysis.
 
 ## Data analysis
+
+[@silge2017a] [@hvitfeldt2021a]
 
@@
 
-```{r stop-words}
+```{r}
+#| label: stop-words
 
@@
   str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
+    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
   str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))
@@
 
-Additional checks were made to the document in search for previously undetected stop words.
-The checks included identifying words consisting of four or less characters that extensively used in the text corpus.
+Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
 
 ```{r}
-#| label: short-stop-words
-
-#short_words(emu_theses$text_clean, scope = 'corpus')
-short_words(emu_theses_words$word)
-
-```
-
-Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
-
-```{r stemming}
+#| label: stemming
 
@@
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+In LDA, each document is a mixture of topics and each topic is a mixture of words.
+The probability of a term being part of a topic is given by the beta statistic.
+The probability of a document being part of a topic is given by the gamma statistic.
 
 ```{r}
-# The {topicmodels} package requires a document-term matrix as input
-emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
+# The topicmodels package requires a document-term matrix as input
+emu_theses_dtm <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  cast_dtm(title, word, n)
+
 emu_theses_dtm
 
-# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
-# In LDA, each document is a mixture of topics and each topic is a mixture of words
+# Fit LDA model
 emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
@@
 
-# The top 10 words that are the most common in each topic
-emu_theses_top_terms <- emu_theses_topics |> 
-  group_by(topic) |> 
-  top_n(20, beta) |> 
-  ungroup() |> 
-  arrange(topic, -beta)
+# The top n words that are the most common in each topic
+emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)
 
 emu_theses_top_terms |> 
-  mutate(term = reorder(term, beta)) |> 
-  ggplot(aes(term, beta, fill = factor(topic))) +
+  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
   geom_col(show.legend = FALSE) +
@@
 ```{r}
-# Which are the top n most frequently used words in all theses`?
+# Which are the top n most frequently used words in all theses?
 get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 5, 
-                         word_col = "word")
-```
-
-```{r}
-# Which are the top n most frequently used words in all theses`?
-emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 20, 
-                         word_col = "word")
+                         top_n = 20)
 ```

### 2023-07-03 analysis/paper/paper.qmd

--- 2023-07-03 analysis/paper/paper.qmd
@@
 
-```{r get-pdf-text}
+```{r extract-pdf-text}
 #| echo: false
@@
   emu_theses$text <- convert_pdf_text(pdf_paths)
-
-  # Add latitude and longitude information to theses data frame
-  emu_theses <- geocode_thesis_locations(emu_theses)
-
-  # Write thesis dataframe to file
   write_csv(emu_theses, data_path)
@@
   emu_theses$text <- read_csv(data_path)$text
-  emu_theses$longitude <- read_csv(data_path)$longitude
-  emu_theses$latitude <- read_csv(data_path)$latitude
 }
@@
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+In LDA, each document is a mixture of topics and each topic is a mixture of words.
+The probability of a term being part of a topic is given by the beta statistic.
+The probability of a document being part of a topic is given by the gamma statistic.
 
@@
 
-# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
-# In LDA, each document is a mixture of topics and each topic is a mixture of words
+# Fit LDA model
 emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))

### 2023-07-03 R/clean.R
changes to replcements checker

@@ -277,3 +277,49 @@ normalise_words <-function(string_vec) {
   words
 
 }
+
+
+#' Check the result of string replacement
+#'
+#' Function checks if a specific check pattern is included in the string vector  ;
+#' if so, it would print the pattern and the instances of it's occurance in the string
+#'
+#' @param text_vector A vector string to be checked
+#' @param check_pattern A pattern used to check whether the replacement worked. Can be single value or a vector of the same length as text_vector
+#' @param char_before Number of characters to be shown before the start of the check_pattern. Default is 25
+#' @param char_after  Number of characters to be shown after the start of the check_pattern. Default is 25
+#'
+#' @export
+#'
+replacement_checker <- function(text_vector, check_pattern, char_before = 25, char_after = 25 ){
+
+  occur <- str_extract_all(text_vector, check_pattern)
+  not_removed <- which(lengths(occur)>0)
+  loc <- str_locate_all(text_vector, check_pattern )
+
+  if (!(length(check_pattern) == 1 | length(check_pattern) == length(text_vector) ))
+    stop("The `check_pattern` needs to be a string vector of length 1 or of the same length as `text_vector`")
+
+
+  if(length(check_pattern) == 1 )
+    check_pattern <- rep(check_pattern, length(text_vector) )
+
+  for (nr in not_removed) {
+    n_occs <- nrow(loc[[nr]])
+
+    print(paste("PATTERN:", check_pattern[[nr]]))
+
+    for (occ in 1:n_occs) {
+      print(paste(
+        "NON_REMOVED,  NR",
+        nr,
+        ",",
+        occ ,
+        ":",
+        str_sub(text_vector[nr], loc[[nr]][occ] - char_before, loc[[nr]][occ] +
+                  char_after)
+      ))
+    }
+  }
+
+}

### 2023-07-03 analysis/supplementary-materials/Package_data_workflow.R

@@ -1,5 +1,5 @@
 # Create data-raw folder to compute package datasets ----------------------
 usethis::use_data_raw()
 
-# Create data folder with data to be made available asa part of the pacakge
+# Create data folder with data to be made available as a part of the pacakge
 usethis::use_data(emu_theses, overwrite = TRUE)

### 2023-07-03 analysis/supplementary-materials/prepare-for-publication.R

@@ -3,54 +3,30 @@ library(tidyverse)
 library(here)
 library(mintEMU)
 
+# TODO Extract text from PDFs (move this step from paper.qmd)
+
 # Load data -----------
+
 data("emu_theses")
 
 
 # Add ID ( order based on title) ------------
+
+# TODO  remove special Latin characters before anonymisation
 emu_theses <- emu_theses |>
   arrange(title) |>
   mutate(ID  = seq_len(n()),
 # remove email addresses
          text_raw = str_replace_all(text,
-                                    "\\b[A-Za-z0-9._%+-]+@[A-Za-z0-9.-]+\\.[A-Za-z]{2,}\\b" ),
-                                    "EMAIL_REMOVED")
+                                    "\\b[A-Za-z0-9._%+-]+@[A-Za-z0-9.-]+\\.[A-Za-z]{2,}\\b" ,
+                                    "EMAIL_REMOVED"))
 
 # Check if email addresses removed:
-str_extract_all(emu_theses$text_raw, "@")
-
-anonymisation_checker <- function(text_vector, check_pattern ){
-
-  occur <- str_extract_all(text_vector, check_pattern)
-  not_removed <- which(lengths(occur)>0)
-  loc <- str_locate_all(text_vector, check_pattern )
-
-  if(length(check_pattern) == 1 )
-    check_pattern <- rep(check_pattern, length(text_vector) )
-
-  for (nr in not_removed) {
-    n_occs <- nrow(loc[[nr]])
-
-    print(paste("PATTERN:", check_pattern[[nr]]))
-
-    for (occ in 1:n_occs) {
-      print(paste(
-        "NON_REMOVED,  NR",
-        nr,
-        ",",
-        occ ,
-        ":",
-        str_sub(text_vector[nr], loc[[nr]][occ] - 25, loc[[nr]][occ] +
-                  25)
-      ))
-    }
-  }
-
-}
-
-anonymisation_checker(emu_theses$text_raw, "@")
+replacement_checker(emu_theses$text_raw, "@")
 
 # Anonymise raw text -----------
+
+# TODO add combinations of lower and uppercase in word combos
 text_raw <- mapply(str_replace_all, emu_theses$text_raw,
                   word_vec_combos(normalise_words(emu_theses$first_name),
                                   normalise_words(emu_theses$last_name)),
@@ -60,46 +36,17 @@ text_raw <- mapply(str_replace_all, emu_theses$text_raw,
[…]

### 2023-07-03 R/analyse.R
Update R/analyse.R
Co-authored-by: alwil <alexandra.wilczynska@gmail.com>

@@ -80,7 +80,7 @@ get_ngrams <-
     data_ngrams <- data |>
     select(title_col, text_col) |>
     tidytext::unnest_tokens(ngram, text_col, token = "ngrams", n = n) |>
-    tidyr::separate(ngram, into = c("first", "second", "third"),
+    tidyr::separate(ngram, into = paste("w" , 1:n, sep = "_")
                     sep = " ", remove = FALSE) |>
     dplyr::anti_join(tidytext::stop_words, by = c("first" = "word")) |>
     dplyr::anti_join(tidytext::stop_words, by = c("second" = "word")) |>

### 2023-07-03 R/analyse.R
Replace repeated mutate calles with mutate across

@@ -90,12 +90,11 @@ get_ngrams <-
 
   if (stem) {
     data_ngrams <- data_ngrams |>
-      dplyr::mutate(first_stem = SnowballC::wordStem(words = first,
-                                                     language = "porter")) |>
-      dplyr::mutate(second_stem = SnowballC::wordStem(words = second,
-                                                      language = "porter")) |>
-      dplyr::mutate(third_stem = SnowballC::wordStem(words = third,
-                                                     language = "porter"))
+      dplyr::mutate(dplyr::across(
+        tidyselect::starts_with("w_"),
+        ~ SnowballC::wordStem(words = .x, language = "porter"),
+        .names = "{.col}_stem"
+      ))
   }
 
   data_ngrams

### 2023-07-03 analysis/paper/paper.qmd
Generalise column-creation in ngrams function

--- 2023-07-03 analysis/paper/paper.qmd
@@
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
-In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the beta statistic.
-The probability of a document being part of a topic is given by the gamma statistic.
 
@@
 
-# Fit LDA model
+# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+# In LDA, each document is a mixture of topics and each topic is a mixture of words
 emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
@@
 
-# The top n words that are the most common in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)
+# The top 10 words that are the most common in each topic
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
 
 emu_theses_top_terms |> 
-  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
+  mutate(term = reorder(term, beta)) |> 
+  ggplot(aes(term, beta, fill = factor(topic))) +
   geom_col(show.legend = FALSE) +
@@
 get_top_words_per_document(data = emu_theses_words, 
-                           top_n = 3)
+                           top_n = 3, 
+                           title_col = "title", 
+                           word_col = "word")
 ```
@@
 ```{r}
-# Which are the top n most frequently used words in all theses?
+# Which are the top n most frequently used words in all theses`?
 get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 20)
+                         top_n = 5, 
+                         word_col = "word")
+```
+
+```{r}
+# Which are the top n most frequently used words in all theses`?
+emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 20, 
+                         word_col = "word")
 ```
@@
 emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
+  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
+
+spacyr::spacy_parse(emu_theses_bigrams$first)
 

### 2023-07-03 R/analyse.R

@@ -71,23 +71,28 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 #'
 #' @return A data frame with an "ngram" column and columns for its constituent words
 #' @export
-get_ngrams <-
-  function(data,
-           n,
-           title_col = NULL,
-           text_col = NULL,
-           stem = FALSE) {
-    data_ngrams <- data |>
+get_ngrams <- function(data,
+                       n,
+                       title_col = NULL,
+                       text_col = NULL,
+                       stem = FALSE) {
+
+  # Create ngrams column and separate columns for each of the n terms
+  data_ngrams <- data |>
     select(title_col, text_col) |>
     tidytext::unnest_tokens(ngram, text_col, token = "ngrams", n = n) |>
-    tidyr::separate(ngram, into = paste("w" , 1:n, sep = "_")
-                    sep = " ", remove = FALSE) |>
-    dplyr::anti_join(tidytext::stop_words, by = c("first" = "word")) |>
-    dplyr::anti_join(tidytext::stop_words, by = c("second" = "word")) |>
-    dplyr::anti_join(tidytext::stop_words, by = c("third" = "word")) |>
-    dplyr::filter(str_detect(first, "[a-z]") &
-             stringr::str_detect(second, "[a-z]"))
+    tidyr::separate(ngram, into = paste("w" , 1:n, sep = "_"),
+                    sep = " ", remove = FALSE)
 
+  # Remove stop words from each column starting with "w_"
+  w_cols <- names(data_ngrams)[grepl("w_", names(data_ngrams))]
+
+  for (i in w_cols) {
+    data_ngrams <- data_ngrams |>
+      anti_join(stop_words, by = set_names("word", i))
+  }
+
+  # Add columns with stemmed versions of individual terms if `stem = TRUE`
   if (stem) {
     data_ngrams <- data_ngrams |>
       dplyr::mutate(dplyr::across(

### 2023-07-03 analysis/paper/paper.qmd
Update usage of ngram column names

@@
 
-## Add lemmas as optional column to ngram function
-
 ## Bigrams
@@
 
-spacyr::spacy_parse(emu_theses_bigrams$first)
+spacyr::spacy_parse(emu_theses_bigrams$w_1)
 
 bigram_counts <- emu_theses_bigrams |> 
-  count(first, second, sort = TRUE)
+  count(w_1, w_2, sort = TRUE)
 
 bigram_stem_counts <- emu_theses_bigrams |> 
-  count(first_stem, second_stem, sort = TRUE)
+  count(w_1_stem, w_2_stem, sort = TRUE)
 
@@
 trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, third, sort = TRUE)
+  count(w_1, w_2, w_3, sort = TRUE)
 

### 2023-07-03 analysis/paper/paper.qmd

--- 2023-07-03 analysis/paper/paper.qmd
@@
 
-## TO DO: generalise bigrams and trigrams into ngrams function
-
 ## Bigrams
 emu_theses_bigrams <- emu_theses |> 
-  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
-  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
-  anti_join(stop_words, by = c("first" = "word")) |> 
-  anti_join(stop_words, by = c("second" = "word")) |> 
-  filter(str_detect(first, "[a-z]") & 
-         str_detect(second, "[a-z]"))
+  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
+
+spacyr::spacy_parse(emu_theses_bigrams$w_1)
 
 bigram_counts <- emu_theses_bigrams |> 
-  count(first, second, sort = TRUE)
+  count(w_1, w_2, sort = TRUE)
+
+bigram_stem_counts <- emu_theses_bigrams |> 
+  count(w_1_stem, w_2_stem, sort = TRUE)
 
 bigram_counts
+bigram_stem_counts
 
@@
 emu_theses_trigrams <- emu_theses |> 
-  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
-  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
-  filter(!first %in% stop_words$word,
-         !second %in% stop_words$word,
-         !third %in% stop_words$word) |>  
-  filter(str_detect(first, "[a-z]"), 
-         str_detect(second, "[a-z]"),
-         str_detect(third, "[a-z]"))
+  get_ngrams(n = 3, title_col = "title", text_col = "text")
 
 trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, sort = TRUE)
+  count(w_1, w_2, w_3, sort = TRUE)
 

### 2023-07-03 analysis/paper/paper.qmd
Document word embeddings script

--- 2023-07-03 analysis/paper/paper.qmd
@@
   "here",                                        # Managing paths
-  "tidyverse", "reshape2",                       # Data manipulation and visualisation
+  "tidyverse", "reshape2", "widyr",              # Data manipulation and visualisation
+  "furrr",                                       # Parallel use of purrr::map() functions
   "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
@@
 
-## Add lemmas as optional column to ngram function
+## TO DO: generalise bigrams and trigrams into ngrams function
 
@@
 emu_theses_bigrams <- emu_theses |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
-
-spacyr::spacy_parse(emu_theses_bigrams$first)
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
 
@@
 
-bigram_stem_counts <- emu_theses_bigrams |> 
-  count(first_stem, second_stem, sort = TRUE)
-
 bigram_counts
-bigram_stem_counts
 
@@
 emu_theses_trigrams <- emu_theses |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text")
+  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
+  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
+  filter(!first %in% stop_words$word,
+         !second %in% stop_words$word,
+         !third %in% stop_words$word) |>  
+  filter(str_detect(first, "[a-z]"), 
+         str_detect(second, "[a-z]"),
+         str_detect(third, "[a-z]"))
 
 trigram_counts <- emu_theses_trigrams |> 
-  count(first, second, third, sort = TRUE)
+  count(first, second, sort = TRUE)
 
@@
 bigram_tf_idf <- emu_theses_bigrams |> 
-  count(title, ngram) |> 
-  bind_tf_idf(ngram, title, n) |> 
+  count(title, bigram) |> 
+  bind_tf_idf(bigram, title, n) |> 
   arrange(desc(tf_idf))
@@
 trigram_tf_idf <- emu_theses_trigrams |> 
-  count(title, ngram) |> 
-  bind_tf_idf(ngram, title, n) |> 
+  count(title, trigram) |> 
+  bind_tf_idf(trigram, title, n) |> 
   arrange(desc(tf_idf))
@@
   geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
-  geom_node_point(color = "lightblue", linewidth = 5) +
+  geom_node_point(color = "lightblue", size = 5) +
   geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
@@
   group_by(title) |>
-  count(ngram, sort = TRUE) |>
+  count(bigram, sort = TRUE) |>
   slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
@@
   ungroup() |>
-  mutate(bigram = reorder(ngram, n))
+  mutate(bigram = reorder(bigram, n))
 
 emu_theses_bigrams_top_all <- emu_theses_bigrams |>
-  count(ngram, sort = TRUE) |>
+  count(bigram, sort = TRUE) |>
   slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
   # filter(n > 50)                ### filter words with a minimum count
-  mutate(ngram = reorder(ngram, n))
+  mutate(bigram = reorder(bigram, n))
 
 emu_theses_bigrams_top |>
-  filter(title == unique(title)[1:10]) |> 
-  ggplot(aes(x = ngram, y = n, fill = title)) +
+  filter(title == unique(title)[1:6]) |> 
+  ggplot(aes(x = bigram, y = n, fill = title)) +
   geom_col() +
@@
 
-The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
-As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
+Describe word embeddings ...
+
+```{r word-embeddings}
+#| include: false
+
+# Prepare tidy data frame for word embeddings
+tidy_emu_theses <- emu_theses |> 
+  select(title, text_clean) |> 
+  unnest_tokens(word, text_clean) |> 
+  add_count(word) |> 
+  filter(n >= 50) |> 
+  select(-n)
+
+# Create a nested data frame with one row per thesis
+nested_words <- tidy_emu_theses |> 
+  nest(words = c(word))
+
+```
+
+```{r word-embeddings}
+
+# Create sliding window function to calculate skipgram probabilities,
+# that is, how often each word occurs near each other word.
+# A small window size of 3-4 shows how a word is used, while a larger  
+# window size of, say, 10 gives more information about the domain topic.
+slide_windows <- function(tbl, window_size) {
+  # Iterate through an input vector with a fixed window size
+  skipgrams <- slider::slide(
+    tbl,
+    ~.x,
+    .after = window_size - 1,
+    .step = 1L,
+    .complete = TRUE  # Only evaluate function on complete windows
+  )
+
+  # Run even when unsuccessful
+  safe_mutate <- safely(mutate)
+
+  # Map safe_mutate to skipgrams 
+  out <- map2(skipgrams,
+              1:length(skipgrams),
+              ~ safe_mutate(.x, window_id = .y))
+
+  out |> 
+    transpose() |>  # Transpose `result` and `error` into columns,
+    pluck("result") |>  # only keep `result`,
+    compact() |>  # remove any NULL values [is this for the last windows?],
+    bind_rows()  # and bind all rows into a single data frame.
+}
+```
+
+```{r word-embeddings}
+
+# To reduce computation time, we use the {furrr} 
+# and underlying {future} packages for parallel processing
+future::plan(multisession)
+
+# To calculate how often words occur on their own or tigether with other words
+# we use point-wise mutual information (PMI)
+tidy_pmi <- nested_words |> 
+  mutate(words = future_map(words, slide_windows, 4L)) |> 
+  unnest(words) |> 
+  unite(window_id, title, window_id) |> 
+  pairwise_pmi(word, window_id)
+
+```
+
+The high-dimensional sparse matrix of word features is projected into reduced, 100-dimensional set of features.
+
+```{r word-embeddings}
+
+# Determine word vectors using SVD, a method for dimensionality reduction
+# The resulting matrix will be of size n_voc * n_dim instead of n_voc * n_voc
+tidy_word_vectors <- tidy_pmi |> 
+  widely_svd(
+    item1, item2, pmi,
+    nv = 100, maxit = 1000
+  )
+
+```
+
+```{r word-embeddings}
+
+# Calculate nearest neighbours of a given word
+nearest_neighbors <- function(df, token) {
+  df |> 
+    widely(
+      ~ {
+        y <- .[rep(token, nrow(.)), ]
+        res <- rowSums(. * y) /
+          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))
+
+        matrix(res, ncol = 1, dimnames = list(x = names(res)))
+      },
+      sort = TRUE
+    )(item1, dimension, value) |> 
+    select(-item2)
+}
+
+tidy_word_vectors |> 
+  nearest_neighbors("price")
+```
+
+```{r word-embeddings}
+
+# The first n components with top words found in the corpus of theses
+tidy_word_vectors |> 
+  filter(dimension <= 24) |> 
+  group_by(dimension) |> 
+  top_n(12, abs(value)) |> 
+  ungroup() |> 
+  mutate(item1 = reorder_within(item1, value, dimension)) |> 
+  ggplot(aes(item1, value, fill = dimension)) +
+  geom_col(alpha = 0.8, show.legend = FALSE) +
+  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
+  scale_x_reordered() +
+  coord_flip() +
+  labs(
+    x = NULL,
+    y = "Value",
+    title = "First 24 principal components for text of theses",
+    subtitle = "Top words contributing to the components that explain the most variation"
+  )
+
+```
+
+```{r word-embeddings}
+
+# Summarise word embeddings into document embeddings for modeling purposes
+word_matrix <- tidy_emu_theses |> 
+  count(title, word) |> 
+  cast_sparse(title, word, n)
+
+embedding_matrix <- tidy_word_vectors |> 
+  cast_sparse(item1, dimension, value)
+
+doc_matrix <- word_matrix %*% embedding_matrix
+
+dim(doc_matrix)
+
+doc_matrix
+
+```
 

### 2023-07-03 analysis/paper/paper.qmd
Create word embeddings functions

@@
 
-Describe word embeddings ...
+### Word embeddings
 
-```{r word-embeddings}
+[Describe word embeddings ...]
+
+```{r word-embeddings-pmi}
 #| include: false
 
-# Prepare tidy data frame for word embeddings
-tidy_emu_theses <- emu_theses |> 
-  select(title, text_clean) |> 
-  unnest_tokens(word, text_clean) |> 
-  add_count(word) |> 
-  filter(n >= 50) |> 
-  select(-n)
-
-# Create a nested data frame with one row per thesis
-nested_words <- tidy_emu_theses |> 
-  nest(words = c(word))
-
-```
-
-```{r word-embeddings}
-
-# Create sliding window function to calculate skipgram probabilities,
-# that is, how often each word occurs near each other word.
-# A small window size of 3-4 shows how a word is used, while a larger  
-# window size of, say, 10 gives more information about the domain topic.
-slide_windows <- function(tbl, window_size) {
-  # Iterate through an input vector with a fixed window size
-  skipgrams <- slider::slide(
-    tbl,
-    ~.x,
-    .after = window_size - 1,
-    .step = 1L,
-    .complete = TRUE  # Only evaluate function on complete windows
-  )
-
-  # Run even when unsuccessful
-  safe_mutate <- safely(mutate)
-
-  # Map safe_mutate to skipgrams 
-  out <- map2(skipgrams,
-              1:length(skipgrams),
-              ~ safe_mutate(.x, window_id = .y))
-
-  out |> 
-    transpose() |>  # Transpose `result` and `error` into columns,
-    pluck("result") |>  # only keep `result`,
-    compact() |>  # remove any NULL values [is this for the last windows?],
-    bind_rows()  # and bind all rows into a single data frame.
-}
-```
-
-```{r word-embeddings}
-
-# To reduce computation time, we use the {furrr} 
-# and underlying {future} packages for parallel processing
-future::plan(multisession)
-
-# To calculate how often words occur on their own or tigether with other words
-# we use point-wise mutual information (PMI)
-tidy_pmi <- nested_words |> 
-  mutate(words = future_map(words, slide_windows, 4L)) |> 
-  unnest(words) |> 
-  unite(window_id, title, window_id) |> 
-  pairwise_pmi(word, window_id)
+# To what extent do words co-occur with other words?
+tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")
 
@@
 
-```{r word-embeddings}
+```{r word-embeddings-svd}
 
 # Determine word vectors using SVD, a method for dimensionality reduction
-# The resulting matrix will be of size n_voc * n_dim instead of n_voc * n_voc
 tidy_word_vectors <- tidy_pmi |> 
@@
 
-```{r word-embeddings}
+[What terms should be checked with the function below?]
+
+```{r word-embeddings-nn}
 
 # Calculate nearest neighbours of a given word
-nearest_neighbors <- function(df, token) {
-  df |> 
-    widely(
-      ~ {
-        y <- .[rep(token, nrow(.)), ]
-        res <- rowSums(. * y) /
-          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))
+tidy_word_vectors |> 
+  nearest_neighbors("resilience")
 
-        matrix(res, ncol = 1, dimnames = list(x = names(res)))
-      },
-      sort = TRUE
-    )(item1, dimension, value) |> 
-    select(-item2)
-}
+```
 
-tidy_word_vectors |> 
-  nearest_neighbors("price")
-```
+The first [?] components... [what do they show?]
 
@@
 # The first n components with top words found in the corpus of theses
-tidy_word_vectors |> 
-  filter(dimension <= 24) |> 
-  group_by(dimension) |> 
-  top_n(12, abs(value)) |> 
-  ungroup() |> 
-  mutate(item1 = reorder_within(item1, value, dimension)) |> 
-  ggplot(aes(item1, value, fill = dimension)) +
-  geom_col(alpha = 0.8, show.legend = FALSE) +
-  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
-  scale_x_reordered() +
-  coord_flip() +
-  labs(
-    x = NULL,
-    y = "Value",
-    title = "First 24 principal components for text of theses",
-    subtitle = "Top words contributing to the components that explain the most variation"
-  )
+visualise_first_n_components(tidy_word_vectors, 24, 12)
 

### 2023-07-03 R/analyse.R

@@ -60,3 +60,139 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 
   data_dtm
 }
+
+#' Sliding window
+#'
+#' Sliding window function used to calculate skipgram probabilities,
+#' that is, how often each word occurs near each other word.
+#'
+#' @param tbl A vector of tokens.
+#' @param window_size A small window size of 3-4 shows how a word is used, while a larger window size of, say, 10 gives more information about the domain topic.
+#'
+#' @return
+#' @export
+#'
+slide_windows <- function(tbl, window_size) {
+  # Iterate through an input vector with a fixed window size
+  skipgrams <- slider::slide(
+    tbl,
+    ~.x,
+    .after = window_size - 1,
+    .step = 1L,
+    .complete = TRUE  # Only evaluate function on complete windows
+  )
+
+  # Run even when unsuccessful
+  safe_mutate <- purrr::safely(mutate)
+
+  # Map safe_mutate to skipgrams
+  out <- map2(skipgrams,
+              1:length(skipgrams),
+              ~ safe_mutate(.x, window_id = .y))
+
+  out |>
+    purrr::transpose() |>  # Transpose `result` and `error` into columns,
+    purrr::pluck("result") |>  # only keep `result`,
+    plyr::compact() |>  # remove any NULL values [is this for the last windows?],
+    dplyr::bind_rows()  # and bind all rows into a single data frame [?].
+}
+
+#' Pointwise mutual information of pairs of items
+#'
+#' @param data Data frame of theses
+#' @param title_col Name of the title column
+#' @param text_col Name of the text column to be tokenised
+#' @param parallel If `parallel = TRUE`, compute in parallel
+#'
+#' @return Data frame with two items and their co-occurrence
+#' @export
+#'
+get_pmi <- function(data,
+                    title_col = NULL,
+                    text_col = NULL,
+                    parallel = TRUE,
+                    window_size = 4L) {
+
+  # To reduce computation time, we use the {furrr}
+  # and underlying {future} packages for parallel processing
+  if (parallel){
+    future::plan(multisession)
+  }
+
+  # Prepare tidy data frame for word embeddings
+  tidy_emu_theses <- data |>
+    dplyr::select(!!rlang::sym(title_col), !!rlang::sym(text_col)) |>
[…]

### 2023-07-04 analysis/paper/paper.qmd
Update paper

--- 2023-07-03 analysis/paper/paper.qmd
@@
 As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
-<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result.
---> Subject-specific stop words, we manually constructed a list based on our domain knowledge.
+<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant improvement was observed in the result.
+--> We manually constructed a list of subject-specific stop words based on our domain knowledge.
 Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well.
-Finally, titles and author names were removed as document stop words on document-level.
+Finally, titles and author names were removed as document stop words.
 
@@
   anti_join(tidytext::stop_words, by = "word")
+
+# TODO can we use this instead of the short_words function?
+emu_theses_words <- emu_theses_words |> 
+  filter(nchar(word) > 4)
 
@@
   count(stem, sort = TRUE)
+
 ```
@@
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the beta statistic.
-The probability of a document being part of a topic is given by the gamma statistic.
+The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.
 
-```{r}
-# The {topicmodels} package requires a document-term matrix as input
-emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
-emu_theses_dtm
+```{r topic-model-lda}
 
-# Fit LDA model
-emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
-emu_theses_lda
+# Prepare document-term matrix as input for LDA model
+emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
 
-# Tidy model with beta statistic
-emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
-emu_theses_topics
+# Create LDA model
+emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
 
-# The top n words that are the most common in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)
+```
 
+```{r topic-model-beta}
+
+# Extract topics as mixtures of terms from LDA model
+emu_topics <- tidy(emu_lda, matrix = "beta")
+
+# The top n most common words in each topic
+emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)
+
+# TODO create visualisation function
 emu_theses_top_terms |> 
@@
 
-# 
+# TODO can this be generalised to more than two topics?
 beta_spread <- emu_theses_topics |> 
@@
   mutate(log_ratio = log2(topic2/topic1))
-
-beta_spread
 
@@
   labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
+```
 
-emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
-emu_theses_documents
+```{r topic-model-gamma}
 
-emu_theses_documents |> 
+# Extract documents as mixtures of topics from LDA model
+emu_documents <- tidy(emu_lda, matrix = "gamma")
+
+emu_documents |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 
-emu_theses_documents |> 
-  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+emu_documents |> 
+  inner_join(emu_theses |> 
+               select(title, graduation_year), 
+             by = c("document" = "title")) |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 
-emu_theses_documents |> 
-  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
+emu_documents |> 
+  inner_join(emu_theses |> 
+               select(title, graduation_year), 
+             by = c("document" = "title")) |> 
   ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
@@
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_theses_words, 
-                           top_n = 3)
+get_top_words_per_document(data = emu_theses_words, top_n = 3)
 ```
@@
 # Which are the top n most frequently used words in all theses?
-get_top_words_per_corpus(data = emu_theses_words, 
-                         top_n = 20)
+get_top_words_per_corpus(data = emu_theses_words, top_n = 20)
 ```
@@
 
+## TO DO: generalise bigrams and trigrams into ngrams function
+
 ## Bigrams
 emu_theses_bigrams <- emu_theses |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
-
-spacyr::spacy_parse(emu_theses_bigrams$w_1)
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
 
 bigram_counts <- emu_theses_bigrams |> 
-  count(w_1, w_2, sort = TRUE)
-
-bigram_stem_counts <- emu_theses_bigrams |> 
-  count(w_1_stem, w_2_stem, sort = TRUE)
+  count(first, second, sort = TRUE)
 
 bigram_counts
-bigram_stem_counts
 
@@
 emu_theses_trigrams <- emu_theses |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text")
+  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
+  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
+  filter(!first %in% stop_words$word,
+         !second %in% stop_words$word,
+         !third %in% stop_words$word) |>  
+  filter(str_detect(first, "[a-z]"), 
+         str_detect(second, "[a-z]"),
+         str_detect(third, "[a-z]"))
 
 trigram_counts <- emu_theses_trigrams |> 
-  count(w_1, w_2, w_3, sort = TRUE)
+  count(first, second, sort = TRUE)
 

### 2023-07-04 analysis/paper/paper.qmd

@@
 
-The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
-As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
+### Word embeddings
+
+[Describe word embeddings ...]
+
+```{r word-embeddings-pmi}
+#| include: false
+
+# To what extent do words co-occur with other words?
+tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")
+
+```
+
+The high-dimensional sparse matrix of word features is projected into reduced, 100-dimensional set of features.
+
+```{r word-embeddings-svd}
+
+# Determine word vectors using SVD, a method for dimensionality reduction
+tidy_word_vectors <- tidy_pmi |> 
+  widely_svd(
+    item1, item2, pmi,
+    nv = 100, maxit = 1000
+  )
+
+```
+
+[What terms should be checked with the function below?]
+
+```{r word-embeddings-nn}
+
+# Calculate nearest neighbours of a given word
+tidy_word_vectors |> 
+  nearest_neighbors("resilience")
+
+```
+
+The first [?] components... [what do they show?]
+
+```{r word-embeddings}
+
+# The first n components with top words found in the corpus of theses
+visualise_first_n_components(tidy_word_vectors, 24, 12)
+
+```
+
+```{r word-embeddings}
+
+# Summarise word embeddings into document embeddings for modeling purposes
+word_matrix <- tidy_emu_theses |> 
+  count(title, word) |> 
+  cast_sparse(title, word, n)
+
+embedding_matrix <- tidy_word_vectors |> 
+  cast_sparse(item1, dimension, value)
+
+doc_matrix <- word_matrix %*% embedding_matrix
+
+dim(doc_matrix)
+
+doc_matrix
+
+```
 

### 2023-07-04 analysis/paper/paper.qmd
Update top words and visualisation script in paper

@@
 # Which are the top n most frequently used words in all theses?
-get_top_words_per_corpus(data = emu_theses_words, top_n = 20)
+n = 20
+
+# TODO add topic to fill aesthetic
+get_top_words_per_corpus(data = emu_theses_words |> 
+                           mutate(word = textstem::lemmatize_words(emu_theses_words$word)), top_n = n) |> 
+  ggplot(aes(reorder(word, n), n)) +
+  geom_col(fill = "orange") +
+  geom_text(aes(label = n), size = 2, hjust = 1.1) +
+  coord_flip() +
+  xlab("Word") +
+  ylab("Frequency") +
+  labs(title = paste0("Top ", n, " most used words in the corpus of theses"))
+  theme_minimal() +
+  theme(panel.grid = element_blank())
+
 ```
-
-<!-- ```{r} -->
-
-<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
-
-<!-- emu_theses_top_20 %>%  -->
-
-<!--   filter(word == "fabric") -->
-
-<!-- emu_theses_top_20 %>% -->
-
-<!--   slice(1:40) %>%  -->
-
-<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
-
-<!--   geom_col() + -->
-
-<!--   xlab(NULL) + -->
-
-<!--   coord_flip() + -->
-
-<!--   labs(y = "Count", -->
-
-<!--        x = "Unique words", -->
-
-<!--        title = "Count of unique words found in thesis", -->
-
-<!--        subtitle = "Stop words were removed from the list") + -->
-
-<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
-
-<!-- ``` -->
 
@@
 spacyr::spacy_parse(emu_theses_bigrams$w_1)
+data.frame(emu_theses_bigrams$w_1, textstem::lemmatize_words(emu_theses_bigrams$w_1))
+
+set.seed(12345)
+
+# with spacy
+extract_rand_window <- function(x, p){
+    first_index = sample(seq(length(x) - p + 1), 1)
+    x[first_index:(first_index + p -1)]
+    }
+
+example_lemma <-list()
+for (i in 1:10) {
+  example_lemma[[i]] <- extract_rand_window(emu_theses_bigrams$w_1,10)
+}
+
+Original <- unlist(example_lemma)
+Lemmatized <- spacyr::spacy_parse(Original)
+
+# with textstem
+Lemmatized2 <- textstem::lemmatize_words(Original)
+
+lemma_example <- data.frame(Original, Lemmatized2)
+
+lemma_example
+
+data.frame(Lemmatized$token, Lemmatized$lemma, Lemmatized2)
 

### 2023-07-05 analysis/paper/paper.qmd
Update data collection description
Add LDAvis to topic modeling packages to be loaded

@@
   "furrr",                              # Parallel use of purrr::map() functions
-  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
+  "tidytext", "SnowballC", "tm",        # Text processing and analysis
+  "topicmodels", "LDAvis",              # Topic modeling
   "tidygeocoder",                       # Working with spatial data
@@
 With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
+
+The imported dataset contains the text data in the `text` column and related metadata in the other columns...
 

### 2023-07-05 analysis/paper/paper.qmd
Update pre-processing section

@@
 
-### Tokenisation
+### Pre-processing
 
@@
 
-### Stop words
-
-For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage.
+For analyses in which word-level tokens were used, stop words, i.e., common words or words that do not add much meaning, were removed in the pre-processing stage.
 Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
@@
 Finally, titles and author names were removed as document stop words.
+In the next pre-processing step, we normalised the resulting words, that is, reduced them to their canonical dictionary forms, the resulting words were lemmatised [@hvitfeldt2021a].
+Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.
 
@@
 
+# Add cleaned text column with removed stop words
 emu_theses$text_clean <- emu_theses$text_clean |>
@@
 
+# Unnest words and remove words of one and two letters
 emu_theses_words <- emu_theses |> 
   select(title, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
-  anti_join(tidytext::stop_words, by = "word")
-
-# TODO can we use this instead of the short_words function?
-emu_theses_words <- emu_theses_words |> 
-  filter(nchar(word) > 4)
+  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
+  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
+  mutate(word = textstem::lemmatize_words(word))  # lemmatise words
 
@@
 
-Additional checks were made to the document in search for previously undetected stop words.
-The checks included identifying words consisting of four or less characters that extensively used in the text corpus.
+Finally, additional checks were made to the document in search for previously undetected stop words.
+Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
 
-```{r}
-#| label: short-stop-words
+```{r short-stop-words}
+#| dependson: stop-words
 
-#short_words(emu_theses$text_clean, scope = 'corpus')
-short_words(emu_theses_words$word)
+#short_words(emu_theses_words$word, l_word = 3, l_list = 100)
 
-```
-
-Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
-
-```{r stemming}
-
-emu_theses_words %>% 
-  mutate(stem = wordStem(words = word, 
-                         language = "porter")) |>  # Implement the Porter stemming algorithm for English provided by the SnowballC package
-  count(stem, sort = TRUE)
+emu_theses_words <- emu_theses_words |>  
+  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
 

### 2023-07-05 analysis/paper/paper.qmd

@@
 ```{r topic-model-lda}
+emu_theses_topics <- get_topics(emu_theses_words)
+
+emu_theses_topics_terms <- tidy(emu_theses_topics, matrix = "beta")
+emu_theses_topics_documents <- tidy(emu_theses_topics, matrix = "gamma")
 
@@
 # TODO can this be generalised to more than two topics?
+# beta_spread <- emu_theses_topics_terms |> 
 beta_spread <- emu_theses_topics |> 

### 2023-07-05 analysis/paper/paper.qmd
Add wip topic modeling script

@@
 
-## Topic model
+### Topic model
 
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
@@
 
+The number of topics was empirically determined with the `ldatuning` package [@???].
+
 ```{r topic-model-lda}
-emu_theses_topics <- get_topics(emu_theses_words)
-
-emu_theses_topics_terms <- tidy(emu_theses_topics, matrix = "beta")
-emu_theses_topics_documents <- tidy(emu_theses_topics, matrix = "gamma")
+#| dependson: short-stop-words
 
 # Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
+# Reduce the size of the DTM by excluding words with a frequency lower than 5
+emu_dtm <- convert_to_dtm(emu_theses_words, 
+                          title_col = "title", 
+                          word_col = "word", 
+                          min_term_freq = 5)
 
-# Create LDA model
-emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
+# Fit LDA model with K clusters
+# # Determine K value empirically
+# K = get_number_of_topics(emu_dtm)
+# # This returns K = 23
+# TODO run the model with K = 23 and compare with results using qualitative value
+
+# Determine K value qualitatively
+# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
+emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
+               control = list(seed = 2023, iter = 500, verbose = 25))
+
+# Extract beta and theta statistics from LDA model
+beta <- posterior(emu_lda)$terms
+theta <- posterior(emu_lda)$topics
+
+# Preview 10 terms from each topic
+terms(emu_lda, 10)
+
+# Add topic pseudo-names based on the top 5 words in each topic
+topic_names <- name_topics(beta)
+
+```
+
+```{r topic-model-lda}
+
+# What is the dominant topic in each document?
+N = 20
+topic_prop_examples <- theta[1:N,]
+colnames(topic_prop_examples) <- topic_names
+vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
+                           document = factor(rownames(topic_prop_examples))), 
+                     variable.name = "topic", id.vars = "document")  
+ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
+  geom_bar(stat="identity") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
+  coord_flip() +
+  facet_wrap(~ document, ncol = 5) +
+  theme(legend.position = "none")
+```
+
+```{r}
+
+```
+
+```{r}
+
+# What are the most probable topics in the entire collection?
+topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
+names(topicProportions) <- topic_names     # assign the topic names we created before
+sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
+
+soP <- sort(topicProportions, decreasing = TRUE)
+paste(round(soP, 5), ":", names(soP))
+
+# How many documents are with the highest probability in each topic?
+K = 5
+countsOfPrimaryTopics <- rep(0, K)
+names(countsOfPrimaryTopics) <- topic_names
+for (i in 1:nDocs(emu_dtm)) {
+  topicsPerDoc <- theta[i, ] # select topic distribution for document i
+  # get first element position from ordered list
+  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
+  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
+}
+sort(countsOfPrimaryTopics, decreasing = TRUE)
+
+so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
+paste(so, ":", names(so))
+```
+
+```{r}
+topicToFilter <- 6  # you can set this manually ...
+# ... or have it selected by a term in the topic name (e.g. 'children')
+topicToFilter <- grep('network', topicNames)[1] 
+topicThreshold <- 0.2
+selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
+filteredCorpus <- corpus[selectedDocumentIndexes]
+# show length of filtered corpus
+filteredCorpus
+```
+
+```{r}
+
+N = 82
+topic_prop_per_year <- theta[1:N,]
+colnames(topic_prop_per_year) <- topic_names
+vizDataFrame <- melt(cbind(data.frame(topic_prop_per_year), 
+                           document = factor(rownames(topic_prop_per_year))), 
+                     variable.name = "topic", id.vars = "document")
+
+# plot topic proportions per decade as bar plot
+vizDataFrameWithYear <- vizDataFrame |> 
+  left_join(select(emu_theses, title, graduation_year), by = c("document" = "title")) |> 
+  ggplot(aes(x=factor(graduation_year), y=value, fill = topic)) + 
+  geom_bar(stat = "identity", position = "fill") +
+  scale_fill_viridis_d(name = "Topics") +
+  xlab("Graduation year") +
+  ylab("Proportion of topics") +
+  labs(title = "Proportion of summed topic probabilities over the years of the EMU program")
+
+```
+
+```{r lda-vis}
+
+topicmodels_json_ldavis <- function(fitted, corpus, doc_term) {
+
+  phi <- topicmodels::posterior(fitted)$terms |> as.matrix
+  theta <- topicmodels::posterior(fitted)$topics |> as.matrix
+  vocab <- colnames(phi)
+  doc_length <- vector()
+
+  for (i in 1:length(corpus)) {
+    temp <- paste(corpus[[i]]$content, collapse = " ")
+    doc_length <- c(doc_length, stringi::stri_count(temp, regex = '\\S+'))
+  }
+
+  temp_frequency <- doc_term |> as.matrix
+  freq_matrix <- data.frame(ST = colnames(temp_frequency),
+                            Freq = colSums(temp_frequency))
+  rm(temp_frequency)
+
+  json <- LDAvis::createJSON(phi = phi,
+                             theta = theta,
+                             doc.length = doc.length,
+                             vocab = vocab,
+                             term.frequency = freq_matrix$Freq)
+}
+
+LDAvis::serVis(emu_lda_json, out.dir = "vis", open.browser = FALSE)
 
@@
 ```{r topic-model-beta}
+#| dependson: topic-model-lda
 

### 2023-07-05 R/analyse.R
Add functions related to topic modeling

@@ -108,23 +108,86 @@ convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
 #' @export
 get_topics <-
   function(data,
-           k = 2,
-           seed = 2023,
            title_col = "title",
-           word_col = "word") {
+           word_col = "word",
+           k = 2,
+           min_term_freq = 1,
+           seed = 2023,
+           ...) {
     stopifnot("k must be an integer of at least 2" = k >= 2)
 
     # The {topicmodels} package requires a document-term matrix as input
     dtm <- convert_to_dtm(data,
                           title_col = title_col,
-                          word_col = word_col)
+                          word_col = word_col,
+                          min_term_freq = min_term_freq)
 
     # We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
     # In LDA, each document is a mixture of topics and each topic is a mixture of words
     lda <- topicmodels::LDA(dtm,
                             k = k,
-                            control = list(seed = seed))
+                            control = list(seed = seed, ...))
   }
+
+#' Add pseudo-names to topics in an LDA model
+#'
+#' Pseudo-names are obtained by pasting together the top n terms with
+#' the highest probability within a topic.
+#'
+#' @param beta Prior topic distribution over words
+#' @param n_words Number of words to include in the pseudo-names
+#'
+#' @return Character vector of topic pseudo-names
+#' @export
+#'
+name_topics <- function(beta, n_words = 5) {
+
+  topic_names <- c()
+
+  for (i in 1:nrow(beta)) {
+    name <- paste(names(head(sort(beta[i, ], decreasing = TRUE), n_words)), collapse = " ")
+    topic_names <- c(topic_names, name)
+  }
+
+  topic_names
+}
+
+#' Get number of topics for LDA model empirically
+#'
+#' @param dtm Document term matrix
+#' @param topics Range of number of topics
+#' @param metrics Metrics used to evaluate the number of topics
+#' @param method Method used for fitting the LDA model
+#' @param control Setting tuning parameters
+#' @param verbose Show progress in the console
+#'
+#' @return Plot and possibly a K value
+#' @export
+#'
+get_number_of_topics <- function(dtm,
+                                 topics = seq(from = 2, to = 30, by = 1),
+                                 metrics = c("CaoJuan2009",  "Deveaud2014"),
+                                 method = "Gibbs",
[…]

### 2023-07-05 R/analyse.R
Add minimum term frequency argument to dtm function

@@ -87,9 +87,10 @@ get_top_words_per_topic <- function(data,
 #'
 #' @return A document term matrix
 #' @export
-convert_to_dtm <- function(data, title_col = NULL, word_col = NULL) {
+convert_to_dtm <- function(data, title_col = NULL, word_col = NULL, min_term_freq = 1) {
   data_dtm <- data |>
     dplyr::count(!!rlang::sym(title_col), !!rlang::sym(word_col), sort = TRUE) |>
+    dplyr::filter(n > min_term_freq) |>
     tidytext::cast_dtm(!!rlang::sym(title_col), !!rlang::sym(word_col), n)
 
   data_dtm

### 2023-07-05 analysis/paper/paper.qmd
Add urbanism stop words found during modeling

@@
   str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))
+    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
+                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
+  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
 

### 2023-07-05 analysis/paper/paper.qmd
Add code chunk labels and update layout

--- 2023-07-05 analysis/paper/paper.qmd
@@
 ```{r topic-model-lda}
-emu_theses_topics <- get_topics(emu_theses_words)
-
-emu_theses_topics_terms <- tidy(emu_theses_topics, matrix = "beta")
-emu_theses_topics_documents <- tidy(emu_theses_topics, matrix = "gamma")
 
@@
 # TODO can this be generalised to more than two topics?
-# beta_spread <- emu_theses_topics_terms |> 
 beta_spread <- emu_theses_topics |> 
@@
 
-```{r}
+```{r top-words-document}
 # Which are the top n most frequently used words in each thesis?
@@
 
-```{r}
+```{r top-words-corpus}
 # Which are the top n most frequently used words in all theses?
-n = 20
-
-# TODO add topic to fill aesthetic
-get_top_words_per_corpus(data = emu_theses_words |> 
-                           mutate(word = textstem::lemmatize_words(emu_theses_words$word)), top_n = n) |> 
+# TODO add `topic` to `fill` aesthetic
+get_top_words_per_corpus(data = emu_theses_words, top_n = 50) |>
   ggplot(aes(reorder(word, n), n)) +
@@
   ylab("Frequency") +
-  labs(title = paste0("Top ", n, " most used words in the corpus of theses"))
+  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
   theme_minimal() +
   theme(panel.grid = element_blank())
-
 ```
@@
 ```{r tf-idf}
+
 ## Calculate tf-idf statistic
@@
   arrange(desc(tf_idf), .by_group = TRUE)
+
 ```
@@
 
-```{r word-embeddings}
+```{r word-embeddings-vis}
 
@@
 
-```{r word-embeddings}
+```{r doc-embeddings}
 

### 2023-07-05 analysis/paper/paper.qmd
Update analysis/paper/paper.qmd

@@
   "here",                                        # Managing paths
-  "tidyverse", "reshape2", "widyr",     # Data manipulation and visualisation
-  "furrr",                              # Parallel use of purrr::map() functions
-  "tidytext", "SnowballC", "tm",        # Text processing and analysis
-  "topicmodels", "LDAvis",              # Topic modeling
+  "tidyverse", "reshape2",                       # Data manipulation and visualisation
+  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
   "tidygeocoder",                                # Working with spatial data
@@
 
-```{r get-pdf-text}
+Description of text extraction; raw data available on 4TU.ResearchData at...
+
+```{r extract-pdf-text}
 #| echo: false
@@
 
-The imported dataset contains the text data in the `text` column and related metadata in the other columns...
-
 A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.
@@
 
-### Pre-processing
+### Tokenisation
 
@@
 
-For analyses in which word-level tokens were used, stop words, i.e., common words or words that do not add much meaning, were removed in the pre-processing stage.
+### Stop words
+
+For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage.
 Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
 As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
-<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant improvement was observed in the result.
---> We manually constructed a list of subject-specific stop words based on our domain knowledge.
+<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result.
+--> Subject-specific stop words, we manually constructed a list based on our domain knowledge.
 Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well.
-Finally, titles and author names were removed as document stop words.
-In the next pre-processing step, we normalised the resulting words, that is, reduced them to their canonical dictionary forms, the resulting words were lemmatised [@hvitfeldt2021a].
-Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.
+Finally, titles and author names were removed as document stop words on document-level.
 
@@
 
-# Add cleaned text column with removed stop words
 emu_theses$text_clean <- emu_theses$text_clean |>
@@
   str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
+    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
   str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))
 
-# Unnest words and remove words of one and two letters
 emu_theses_words <- emu_theses |> 
   select(title, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
-  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
-  mutate(word = textstem::lemmatize_words(word))  # lemmatise words
+  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
+  anti_join(tidytext::stop_words, by = "word")
 
@@
 
-Finally, additional checks were made to the document in search for previously undetected stop words.
-Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
+Additional checks were made to the document in search for previously undetected stop words.
+The checks included identifying words consisting of four or less characters that extensively used in the text corpus.
 
-```{r short-stop-words}
-#| dependson: stop-words
+```{r}
+#| label: short-stop-words
 
-#short_words(emu_theses_words$word, l_word = 3, l_list = 100)
+#short_words(emu_theses$text_clean, scope = 'corpus')
+short_words(emu_theses_words$word)
 
-emu_theses_words <- emu_theses_words |>  
-  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
+```
 
+Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].
+
+```{r stemming}
+
+emu_theses_words %>% 
+  mutate(stem = wordStem(words = word, 
+                         language = "porter")) |>  # Implement the Porter stemming algorithm for English provided by the SnowballC package
+  count(stem, sort = TRUE)
 ```
@@
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
-In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.
 
-```{r topic-model-lda}
+```{r}
+# The {topicmodels} package requires a document-term matrix as input
+emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
+emu_theses_dtm
 
-# Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
+# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
+# In LDA, each document is a mixture of topics and each topic is a mixture of words
+emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
+emu_theses_lda
 
-# Create LDA model
-emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
+# Tidy model with beta statistic
+emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
+emu_theses_topics
 
-```
+# The top 10 words that are the most common in each topic
+emu_theses_top_terms <- emu_theses_topics |> 
+  group_by(topic) |> 
+  top_n(20, beta) |> 
+  ungroup() |> 
+  arrange(topic, -beta)
 
-```{r topic-model-beta}
-
-# Extract topics as mixtures of terms from LDA model
-emu_topics <- tidy(emu_lda, matrix = "beta")
-
-# The top n most common words in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)
-
-# TODO create visualisation function
 emu_theses_top_terms |> 
-  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
+  mutate(term = reorder(term, beta)) |> 
+  ggplot(aes(term, beta, fill = factor(topic))) +
   geom_col(show.legend = FALSE) +
@@
 
-# TODO can this be generalised to more than two topics?
+# 
 beta_spread <- emu_theses_topics |> 
@@
   mutate(log_ratio = log2(topic2/topic1))
+
+beta_spread
 
@@
   labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
-```
 
-```{r topic-model-gamma}
+emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
+emu_theses_documents
 
-# Extract documents as mixtures of topics from LDA model
-emu_documents <- tidy(emu_lda, matrix = "gamma")
-
-emu_documents |> 
+emu_theses_documents |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 
-emu_documents |> 
-  inner_join(emu_theses |> 
-               select(title, graduation_year), 
-             by = c("document" = "title")) |> 
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 
-emu_documents |> 
-  inner_join(emu_theses |> 
-               select(title, graduation_year), 
-             by = c("document" = "title")) |> 
+emu_theses_documents |> 
+  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
   ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
@@
 
-```{r top-words-document}
+```{r}
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_theses_words, top_n = 3)
+get_top_words_per_document(data = emu_theses_words, 
+                           top_n = 3, 
+                           title_col = "title", 
+                           word_col = "word")
 ```
 
-```{r top-words-corpus}
-# Which are the top n most frequently used words in all theses?
-# TODO add `topic` to `fill` aesthetic
-get_top_words_per_corpus(data = emu_theses_words, top_n = 50) |>
-  ggplot(aes(reorder(word, n), n)) +
-  geom_col(fill = "orange") +
-  geom_text(aes(label = n), size = 2, hjust = 1.1) +
-  coord_flip() +
-  xlab("Word") +
-  ylab("Frequency") +
-  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
-  theme_minimal() +
-  theme(panel.grid = element_blank())
+```{r}
+# Which are the top n most frequently used words in all theses`?
+get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 5, 
+                         word_col = "word")
 ```
+
+```{r}
+# Which are the top n most frequently used words in all theses`?
+emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
+                         top_n = 20, 
+                         word_col = "word")
+```
+
+<!-- ```{r} -->
+
+<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->
+
+<!-- emu_theses_top_20 %>%  -->
+
+<!--   filter(word == "fabric") -->
+
+<!-- emu_theses_top_20 %>% -->
+
+<!--   slice(1:40) %>%  -->
+
+<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->
+
+<!--   geom_col() + -->
+
+<!--   xlab(NULL) + -->
+
+<!--   coord_flip() + -->
+
+<!--   labs(y = "Count", -->
+
+<!--        x = "Unique words", -->
+
+<!--        title = "Count of unique words found in thesis", -->
+
+<!--        subtitle = "Stop words were removed from the list") + -->
+
+<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->
+
+<!-- ``` -->
 
@@
 
+## TO DO: generalise bigrams and trigrams into ngrams function
+
 ## Bigrams
 emu_theses_bigrams <- emu_theses |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
-
-spacyr::spacy_parse(emu_theses_bigrams$w_1)
-data.frame(emu_theses_bigrams$w_1, textstem::lemmatize_words(emu_theses_bigrams$w_1))
-
-set.seed(12345)
-
-# with spacy
-extract_rand_window <- function(x, p){
-    first_index = sample(seq(length(x) - p + 1), 1)
-    x[first_index:(first_index + p -1)]
-    }
-
-example_lemma <-list()
-for (i in 1:10) {
-  example_lemma[[i]] <- extract_rand_window(emu_theses_bigrams$w_1,10)
-}
-
-Original <- unlist(example_lemma)
-Lemmatized <- spacyr::spacy_parse(Original)
-
-# with textstem
-Lemmatized2 <- textstem::lemmatize_words(Original)
-
-lemma_example <- data.frame(Original, Lemmatized2)
-
-lemma_example
-
-data.frame(Lemmatized$token, Lemmatized$lemma, Lemmatized2)
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
+  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
+  anti_join(stop_words, by = c("first" = "word")) |> 
+  anti_join(stop_words, by = c("second" = "word")) |> 
+  filter(str_detect(first, "[a-z]") & 
+         str_detect(second, "[a-z]"))
 
 bigram_counts <- emu_theses_bigrams |> 
-  count(w_1, w_2, sort = TRUE)
-
-bigram_stem_counts <- emu_theses_bigrams |> 
-  count(w_1_stem, w_2_stem, sort = TRUE)
+  count(first, second, sort = TRUE)
 
 bigram_counts
-bigram_stem_counts
 
@@
 
-The `tf-idf` statistic shows words that are the most important to one document in a collection of documents [@silge2017].
-In this case, it shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
+```{r}
+#| label: tf-idf
 
-```{r tf-idf}
+emu_theses_words
 
-## Calculate tf-idf statistic
-emu_theses_tf_idf <- emu_theses_words |> 
-  get_tf_idf(title_col = "title", word_col = "word")
+emu_theses_words_count <- emu_theses_words |> 
+  count(title, word, sort = TRUE) |> 
+  ungroup()
 
-## Get top n words most specific to each thesis 
-emu_theses_tf_idf |> 
+emu_theses_words_total <- emu_theses_words_count |> 
   group_by(title) |> 
-  top_n(5, tf_idf) |> 
-  arrange(desc(tf_idf), .by_group = TRUE)
+  summarise(total = sum(n))
 
-```
+emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)
 
-```{r}
+# emu_theses_words_count |> 
+#   group_by(title) |> 
+#   summarise(total = unique(tot
+
 # Preview term frequency distributions in a subset of theses
@@
   facet_wrap(~ title, ncol = 3, scales = "free_y")
-```
 
-```{r}
 # Examine Zipf's law for the EMU theses
@@
 
+## Calculate tf-idf
+emu_theses_words_count <- emu_theses_words_count |> 
+  bind_tf_idf(word, title, n)
+emu_theses_words_count
+
+emu_theses_words_count |> 
+  select(-total) |> 
+  arrange(desc(tf_idf))
+
 ## Visualise the top n words with the highest tf-idf values for each of a subset of theses
@@
 
-### Word embeddings
-
-[Describe word embeddings ...]
-
-```{r word-embeddings-pmi}
-#| include: false
-
-# To what extent do words co-occur with other words?
-tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")
-
-```
-
-The high-dimensional sparse matrix of word features is projected into reduced, 100-dimensional set of features.
-
-```{r word-embeddings-svd}
-
-# Determine word vectors using SVD, a method for dimensionality reduction
-tidy_word_vectors <- tidy_pmi |> 
-  widely_svd(
-    item1, item2, pmi,
-    nv = 100, maxit = 1000
-  )
-
-```
-
-[What terms should be checked with the function below?]
-
-```{r word-embeddings-nn}
-
-# Calculate nearest neighbours of a given word
-tidy_word_vectors |> 
-  nearest_neighbors("resilience")
-
-```
-
-The first [?] components... [what do they show?]
-
-```{r word-embeddings-vis}
-
-# The first n components with top words found in the corpus of theses
-visualise_first_n_components(tidy_word_vectors, 24, 12)
-
-```
-
-```{r doc-embeddings}
-
-# Summarise word embeddings into document embeddings for modeling purposes
-word_matrix <- tidy_emu_theses |> 
-  count(title, word) |> 
-  cast_sparse(title, word, n)
-
-embedding_matrix <- tidy_word_vectors |> 
-  cast_sparse(item1, dimension, value)
-
-doc_matrix <- word_matrix %*% embedding_matrix
-
-dim(doc_matrix)
-
-doc_matrix
-
-```
+The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
+As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
 

### 2023-07-05 analysis/paper/paper.qmd
Update topic model script

@@
   "tidytext", "SnowballC", "tm",        # Text processing and analysis
-  "topicmodels", "LDAvis",              # Topic modeling
+  "topicmodels", "LDAvis", "tsne",      # Topic modeling
   "tidygeocoder",                       # Working with spatial data
@@
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.
+The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
 
-The number of topics was empirically determined with the `ldatuning` package [@???].
+The number of topics was qualitatively determined
 
@@
 
-# Add topic pseudo-names based on the top 5 words in each topic
+# Add pseudo-names to topics based on the top 5 words in each topic
 topic_names <- name_topics(beta)
@@
   theme(legend.position = "none")
+
 ```
@@
 
-topicmodels_json_ldavis <- function(fitted, corpus, doc_term) {
+# TODO check if this can be included in the Shiny dashboard
+# Function to approximate the distance between topics
+svd_tsne <- function(x) tsne(svd(x)$u)
 
-  phi <- topicmodels::posterior(fitted)$terms |> as.matrix
-  theta <- topicmodels::posterior(fitted)$topics |> as.matrix
-  vocab <- colnames(phi)
-  doc_length <- vector()
+# Convert DTM into JSON required by the LDAvis package
+json <- LDAvis::createJSON(
+  phi = beta, 
+  theta = theta, 
+  doc.length = rowSums(as.matrix(emu_dtm)), 
+  vocab = colnames(emu_dtm), 
+  term.frequency = colSums(as.matrix(emu_dtm)),
+  mds.method = svd_tsne,
+  plot.opts = list(xlab="", ylab="")
+)
 
-  for (i in 1:length(corpus)) {
-    temp <- paste(corpus[[i]]$content, collapse = " ")
-    doc_length <- c(doc_length, stringi::stri_count(temp, regex = '\\S+'))
-  }
-
-  temp_frequency <- doc_term |> as.matrix
-  freq_matrix <- data.frame(ST = colnames(temp_frequency),
-                            Freq = colSums(temp_frequency))
-  rm(temp_frequency)
-
-  json <- LDAvis::createJSON(phi = phi,
-                             theta = theta,
-                             doc.length = doc.length,
-                             vocab = vocab,
-                             term.frequency = freq_matrix$Freq)
-}
-
-LDAvis::serVis(emu_lda_json, out.dir = "vis", open.browser = FALSE)
+# Visualise topics model with LDAvis
+LDAvis::serVis(json)
 
@@
 # Extract topics as mixtures of terms from LDA model
-emu_topics <- tidy(emu_lda, matrix = "beta")
+beta_tidy <- tidy(emu_lda, matrix = "beta")
 
 # The top n most common words in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)
+emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)
 

### 2023-07-05 analysis/paper/paper.qmd
Update top words per corpus

--- 2023-07-05 analysis/paper/paper.qmd
@@
   str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
-                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
+    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
+  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))
 
@@
 
-### Topic model
+## Topic model
 
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
@@
 
-The number of topics was empirically determined with the `ldatuning` package [@???].
-
 ```{r topic-model-lda}
-#| dependson: short-stop-words
 
 # Prepare document-term matrix as input for LDA model
-# Reduce the size of the DTM by excluding words with a frequency lower than 5
-emu_dtm <- convert_to_dtm(emu_theses_words, 
-                          title_col = "title", 
-                          word_col = "word", 
-                          min_term_freq = 5)
+emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
 
-# Fit LDA model with K clusters
-# # Determine K value empirically
-# K = get_number_of_topics(emu_dtm)
-# # This returns K = 23
-# TODO run the model with K = 23 and compare with results using qualitative value
-
-# Determine K value qualitatively
-# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
-emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
-               control = list(seed = 2023, iter = 500, verbose = 25))
-
-# Extract beta and theta statistics from LDA model
-beta <- posterior(emu_lda)$terms
-theta <- posterior(emu_lda)$topics
-
-# Preview 10 terms from each topic
-terms(emu_lda, 10)
-
-# Add topic pseudo-names based on the top 5 words in each topic
-topic_names <- name_topics(beta)
-
-```
-
-```{r topic-model-lda}
-
-# What is the dominant topic in each document?
-N = 20
-topic_prop_examples <- theta[1:N,]
-colnames(topic_prop_examples) <- topic_names
-vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
-                           document = factor(rownames(topic_prop_examples))), 
-                     variable.name = "topic", id.vars = "document")  
-ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
-  geom_bar(stat="identity") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
-  coord_flip() +
-  facet_wrap(~ document, ncol = 5) +
-  theme(legend.position = "none")
-```
-
-```{r}
-
-```
-
-```{r}
-
-# What are the most probable topics in the entire collection?
-topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
-names(topicProportions) <- topic_names     # assign the topic names we created before
-sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
-
-soP <- sort(topicProportions, decreasing = TRUE)
-paste(round(soP, 5), ":", names(soP))
-
-# How many documents are with the highest probability in each topic?
-K = 5
-countsOfPrimaryTopics <- rep(0, K)
-names(countsOfPrimaryTopics) <- topic_names
-for (i in 1:nDocs(emu_dtm)) {
-  topicsPerDoc <- theta[i, ] # select topic distribution for document i
-  # get first element position from ordered list
-  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
-  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
-}
-sort(countsOfPrimaryTopics, decreasing = TRUE)
-
-so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
-paste(so, ":", names(so))
-```
-
-```{r}
-topicToFilter <- 6  # you can set this manually ...
-# ... or have it selected by a term in the topic name (e.g. 'children')
-topicToFilter <- grep('network', topicNames)[1] 
-topicThreshold <- 0.2
-selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
-filteredCorpus <- corpus[selectedDocumentIndexes]
-# show length of filtered corpus
-filteredCorpus
-```
-
-```{r}
-
-N = 82
-topic_prop_per_year <- theta[1:N,]
-colnames(topic_prop_per_year) <- topic_names
-vizDataFrame <- melt(cbind(data.frame(topic_prop_per_year), 
-                           document = factor(rownames(topic_prop_per_year))), 
-                     variable.name = "topic", id.vars = "document")
-
-# plot topic proportions per decade as bar plot
-vizDataFrameWithYear <- vizDataFrame |> 
-  left_join(select(emu_theses, title, graduation_year), by = c("document" = "title")) |> 
-  ggplot(aes(x=factor(graduation_year), y=value, fill = topic)) + 
-  geom_bar(stat = "identity", position = "fill") +
-  scale_fill_viridis_d(name = "Topics") +
-  xlab("Graduation year") +
-  ylab("Proportion of topics") +
-  labs(title = "Proportion of summed topic probabilities over the years of the EMU program")
-
-```
-
-```{r lda-vis}
-
-topicmodels_json_ldavis <- function(fitted, corpus, doc_term) {
-
-  phi <- topicmodels::posterior(fitted)$terms |> as.matrix
-  theta <- topicmodels::posterior(fitted)$topics |> as.matrix
-  vocab <- colnames(phi)
-  doc_length <- vector()
-
-  for (i in 1:length(corpus)) {
-    temp <- paste(corpus[[i]]$content, collapse = " ")
-    doc_length <- c(doc_length, stringi::stri_count(temp, regex = '\\S+'))
-  }
-
-  temp_frequency <- doc_term |> as.matrix
-  freq_matrix <- data.frame(ST = colnames(temp_frequency),
-                            Freq = colSums(temp_frequency))
-  rm(temp_frequency)
-
-  json <- LDAvis::createJSON(phi = phi,
-                             theta = theta,
-                             doc.length = doc.length,
-                             vocab = vocab,
-                             term.frequency = freq_matrix$Freq)
-}
-
-LDAvis::serVis(emu_lda_json, out.dir = "vis", open.browser = FALSE)
+# Create LDA model
+emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
 
@@
 ```{r topic-model-beta}
-#| dependson: topic-model-lda
 
@@
 # TODO add `topic` to `fill` aesthetic
-get_top_words_per_corpus(data = emu_theses_words, top_n = 50) |>
+n = 50
+get_top_words_per_corpus(data = emu_theses_words, top_n = n) |>
   ggplot(aes(reorder(word, n), n)) +

### 2023-07-06 analysis/paper/paper.qmd
Update topic model code

--- 2023-07-05 analysis/paper/paper.qmd
@@
 
-```
-
-```{r}
-
 # What are the most probable topics in the entire collection?
@@
 # ... or have it selected by a term in the topic name (e.g. 'children')
-topicToFilter <- grep('network', topicNames)[1] 
+topicToFilter <- grep('river', topic_names)[1] 
 topicThreshold <- 0.2
 selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
-filteredCorpus <- corpus[selectedDocumentIndexes]
+filteredCorpus <- emu_theses[selectedDocumentIndexes,]
 # show length of filtered corpus
@@
 colnames(topic_prop_per_year) <- topic_names
-vizDataFrame <- melt(cbind(data.frame(topic_prop_per_year), 
+
+vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
                            document = factor(rownames(topic_prop_per_year))), 
-                     variable.name = "topic", id.vars = "document")
+                     variable.name = "topic", id.vars = "document") |> 
+    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))
 
-# plot topic proportions per decade as bar plot
-vizDataFrameWithYear <- vizDataFrame |> 
-  left_join(select(emu_theses, title, graduation_year), by = c("document" = "title")) |> 
-  ggplot(aes(x=factor(graduation_year), y=value, fill = topic)) + 
+# Plot topic proportions per year
+vizDataFrameWithYear |> 
+  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
   geom_bar(stat = "identity", position = "fill") +
@@
   ylab("Proportion of topics") +
-  labs(title = "Proportion of summed topic probabilities over the years of the EMU program")
+  labs(title = "Topic proportions per year")
 

### 2023-07-06 analysis/paper/paper.qmd
Update object names

--- 2023-07-05 analysis/paper/paper.qmd
@@
 ## Bigrams
-emu_theses_bigrams <- emu_theses |> 
+emu_bigrams <- emu_theses |> 
   get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
 
-spacyr::spacy_parse(emu_theses_bigrams$w_1)
-data.frame(emu_theses_bigrams$w_1, textstem::lemmatize_words(emu_theses_bigrams$w_1))
+spacyr::spacy_parse(emu_bigrams$w_1)
+data.frame(emu_bigrams$w_1, textstem::lemmatize_words(emu_bigrams$w_1))
 
@@
 for (i in 1:10) {
-  example_lemma[[i]] <- extract_rand_window(emu_theses_bigrams$w_1,10)
+  example_lemma[[i]] <- extract_rand_window(emu_bigrams$w_1,10)
 }
@@
 
-bigram_counts <- emu_theses_bigrams |> 
+bigram_counts <- emu_bigrams |> 
   count(w_1, w_2, sort = TRUE)
 
-bigram_stem_counts <- emu_theses_bigrams |> 
+bigram_stem_counts <- emu_bigrams |> 
   count(w_1_stem, w_2_stem, sort = TRUE)
@@
 ## Bigram tf-idf
-bigram_tf_idf <- emu_theses_bigrams |> 
+bigram_tf_idf <- emu_bigrams |> 
   count(title, ngram) |> 
@@
 
-emu_theses_bigrams_top <- emu_theses_bigrams |>
+emu_bigrams_top <- emu_bigrams |>
   group_by(title) |>
@@
 
-emu_theses_bigrams_top_all <- emu_theses_bigrams |>
+emu_bigrams_top_all <- emu_bigrams |>
   count(ngram, sort = TRUE) |>
@@
 
-emu_theses_bigrams_top |>
+emu_bigrams_top |>
   filter(title == unique(title)[1:10]) |> 

### 2023-07-06 analysis/paper/paper.qmd
Shorten `emu_theses` to `emu` in all object names

@@
 ## Read only metadata for theses with PDF available
-emu_theses <- 
+emu <- 
   read_csv(here(data_path, "theses-metadata.csv")) |>
@@
 
-pdf_paths <- here(data_path, emu_theses$file_name)
+pdf_paths <- here(data_path, emu$file_name)
 ```
@@
   # On the first run, this function might take extra time as a Python environment needs to be set up
-  emu_theses$text <- convert_pdf_text(pdf_paths)
+  emu$text <- convert_pdf_text(pdf_paths)
 
   # Add latitude and longitude information to theses data frame
-  emu_theses <- geocode_thesis_locations(emu_theses)
+  emu <- geocode_thesis_locations(emu)
 
   # Write thesis dataframe to file
-  write_csv(emu_theses, data_path)
+  write_csv(emu, data_path)
 } else {
-  emu_theses$text <- read_csv(data_path)$text
-  emu_theses$longitude <- read_csv(data_path)$longitude
-  emu_theses$latitude <- read_csv(data_path)$latitude
+  emu$text <- read_csv(data_path)$text
+  emu$longitude <- read_csv(data_path)$longitude
+  emu$latitude <- read_csv(data_path)$latitude
 }
@@
 
-emu_theses$text_clean <- emu_theses$text |>
+emu$text_clean <- emu$text |>
   clean_basic()
@@
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
 
@@
 
-A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.
+A total number of `r nrow(emu)` out of `r nrow(all_theses)` were analysed.
 
@@
 # Show thesis locations on world map
-visualize_thesis_locations(emu_theses)
+visualize_thesis_locations(emu)
 ```
@@
 # Add cleaned text column with removed stop words
-emu_theses$text_clean <- emu_theses$text_clean |>
-  str_remove_all(find_meta_stopwords(emu_theses)) |> 
+emu$text_clean <- emu$text_clean |>
+  str_remove_all(find_meta_stopwords(emu)) |> 
   str_remove_all(urbanism_stopwords(
@@
 # Unnest words and remove words of one and two letters
-emu_theses_words <- emu_theses |> 
+emu_words <- emu |> 
   select(title, text_clean) |> 
@@
 
-#short_words(emu_theses_words$word, l_word = 3, l_list = 100)
+#short_words(emu_words$word, l_word = 3, l_list = 100)
 
-emu_theses_words <- emu_theses_words |>  
+emu_words <- emu_words |>  
   filter(nchar(word) >= 3)  # keep only words that are at least three letters long
@@
 # Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
+emu_dtm <- convert_to_dtm(emu_words, "title", "word")
 
@@
 # The top n most common words in each topic
-emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)
+emu_top_terms <- get_top_words_per_topic(emu_topics, 20)
 
 # TODO create visualisation function
-emu_theses_top_terms |> 
+emu_top_terms |> 
   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
@@
 # TODO can this be generalised to more than two topics?
-beta_spread <- emu_theses_topics |> 
+beta_spread <- emu_topics |> 
   mutate(topic = paste0("topic", topic)) |> 
@@
 emu_documents |> 
-  inner_join(emu_theses |> 
+  inner_join(emu |> 
                select(title, graduation_year), 
@@
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_theses_words, top_n = 3)
+get_top_words_per_document(data = emu_words, top_n = 3)
 ```
@@
 n = 50
-get_top_words_per_corpus(data = emu_theses_words, top_n = n) |>
+get_top_words_per_corpus(data = emu_words, top_n = n) |>
   ggplot(aes(reorder(word, n), n)) +
@@
 ## Bigrams
-emu_bigrams <- emu_theses |> 
+emu_bigrams <- emu |> 
   get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
@@
 ## Trigrams
-emu_theses_trigrams <- emu_theses |> 
+emu_trigrams <- emu |> 
   get_ngrams(n = 3, title_col = "title", text_col = "text")
 
-trigram_counts <- emu_theses_trigrams |> 
+trigram_counts <- emu_trigrams |> 
   count(w_1, w_2, w_3, sort = TRUE)
@@
 ## Trigram tf-idf
-trigram_tf_idf <- emu_theses_trigrams |> 
+trigram_tf_idf <- emu_trigrams |> 
   count(title, ngram) |> 
@@
 ## Calculate tf-idf statistic
-emu_theses_tf_idf <- emu_theses_words |> 
+emu_tf_idf <- emu_words |> 
   get_tf_idf(title_col = "title", word_col = "word")
@@
 ## Get top n words most specific to each thesis 
-emu_theses_tf_idf |> 
+emu_tf_idf |> 
   group_by(title) |> 
@@
 # Preview term frequency distributions in a subset of theses
-ggplot(emu_theses_words_count |> 
+ggplot(emu_words_count |> 
          filter(title == unique(title)[1:6]),
@@
 ## Calculate rank and term frequency
-freq_by_rank <- emu_theses_words_count |> 
+freq_by_rank <- emu_words_count |> 
   group_by(title) |> 
@@
 ## Visualise the top n words with the highest tf-idf values for each of a subset of theses
-emu_theses_words_count |> 
+emu_words_count |> 
   filter(title == unique(title)[1:6]) |> 
@@
 # To what extent do words co-occur with other words?
-tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")
+tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")
 
@@
 # Summarise word embeddings into document embeddings for modeling purposes
-word_matrix <- tidy_emu_theses |> 
+word_matrix <- tidy_emu |> 
   count(title, word) |> 

### 2023-07-06 analysis/paper/paper.qmd
Format code
Remove lemmatization attempts from bigrams code chunk

@@
 ## Read metadata for all theses
-all_theses <-
-  readxl::read_xlsx(path = here(data_path, "theses-all.xlsx"))
+all_theses <- read_xlsx(path = here(data_path, "theses-all.xlsx"))
 
 ## Read only metadata for theses with PDF available
-emu <- 
-  read_csv(here(data_path, "theses-metadata.csv")) |>
+emu <- read_csv(here(data_path, "theses-metadata.csv")) |>
   mutate(text = "") |>
@@
   get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
-
-spacyr::spacy_parse(emu_bigrams$w_1)
-data.frame(emu_bigrams$w_1, textstem::lemmatize_words(emu_bigrams$w_1))
-
-set.seed(12345)
-
-# with spacy
-extract_rand_window <- function(x, p){
-    first_index = sample(seq(length(x) - p + 1), 1)
-    x[first_index:(first_index + p -1)]
-    }
-
-example_lemma <-list()
-for (i in 1:10) {
-  example_lemma[[i]] <- extract_rand_window(emu_bigrams$w_1,10)
-}
-
-Original <- unlist(example_lemma)
-Lemmatized <- spacyr::spacy_parse(Original)
-
-# with textstem
-Lemmatized2 <- textstem::lemmatize_words(Original)
-
-lemma_example <- data.frame(Original, Lemmatized2)
-
-lemma_example
-
-data.frame(Lemmatized$token, Lemmatized$lemma, Lemmatized2)
 

### 2023-07-06 analysis/paper/paper.qmd
Update package list

@@
   "here",                                       # Managing paths
-  "tidyverse", "reshape2", "widyr",     # Data manipulation and visualisation
+  "tidyverse", "reshape2", "widyr", "readxl",  # Data manipulation and visualisation
   "furrr",                                      # Parallel use of purrr::map() functions

### 2023-07-06 analysis/paper/paper.qmd
Add dependencies between code chunks
Add line breaks

@@
 
+Description of text extraction; raw data available on 4TU.ResearchData at...
+
 ```{r get-pdf-text}
@@
 #| echo: false
+#| dependson: get-pdf-text
 
@@
 #| label: ngrams
+#| dependson: clean-text
 
@@
 bigram_stem_counts
+```
 
+```{r}
 ## Trigrams

### 2023-07-06 data-raw/emu_text.R
Update package data script

@@ -1,12 +1,9 @@
-# Download and import the dataset -----------------------------------------
-emu_text_url <- here::here("analysis", "data", "derived_data", "emu_theses_with_text.csv")
-emu_text_raw_df <- readr::read_csv(emu_text_url)
+# Download and import the dataset ----
+emu_url <- here::here("analysis", "data", "derived_data", "emu_raw.csv")
+emu <- readr::read_csv(emu_url)
 
-# Clean and geocode the dataset -------------------------------------------
-
-emu_text_df <- mintEMU::geocode_thesis_locations(emu_text_raw_df)
-
-emu_text_df$text_clean <- emu_text_df$text |>
+# Clean the dataset ----
+emu$text_clean <- emu$text_raw |>
   mintEMU::clean_basic() |>
   stringr::str_remove_all(mintEMU::find_meta_stopwords(emu_theses)) |>
   stringr::str_remove_all(mintEMU::urbanism_stopwords(
@@ -14,4 +11,4 @@ emu_text_df$text_clean <- emu_text_df$text |>
   stringr::str_remove_all(mintEMU::thesis_stopwords(
     add_stopwords = c("advisor", "prof")))
 
-usethis::use_data(emu_text_df, overwrite = TRUE)
+usethis::use_data(emu, overwrite = TRUE)

### 2023-07-06 analysis/paper/paper.qmd

@@
   "tidytext", "SnowballC", "tm",                # Text processing and analysis
-  "topicmodels", "LDAvis",                      # Topic modeling
+  "topicmodels", "LDAvis", "tsne",              # Topic modeling
   "tidygeocoder",                               # Working with spatial data
@@
 
-## Topic model
+### Topic model
 
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.
+The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
+
+The number of topics was qualitatively determined
+
+```{r topic-model-lda}
+#| dependson: short-stop-words
+
+# Prepare document-term matrix as input for LDA model
+
+# Reduce the size of the DTM by excluding words with a frequency lower than 5
+emu_dtm <- convert_to_dtm(emu_theses_words, 
+                          title_col = "title", 
+                          word_col = "word", 
+                          min_term_freq = 5)
+
+# Fit LDA model with K clusters
+# # Determine K value empirically
+# K = get_number_of_topics(emu_dtm)
+# # This returns K = 23
+# TODO run the model with K = 23 and compare with results using qualitative value
+
+# Determine K value qualitatively
+# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
+emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
+               control = list(seed = 2023, iter = 500, verbose = 25))
+
+# Extract beta and theta statistics from LDA model
+beta <- posterior(emu_lda)$terms
+theta <- posterior(emu_lda)$topics
+
+# Preview 10 terms from each topic
+terms(emu_lda, 10)
+
+# Add pseudo-names to topics based on the top 5 words in each topic
+topic_names <- name_topics(beta)
+
+```
 
@@
 
-# Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_words, "title", "word")
+# What is the dominant topic in each document?
+N = 20
+topic_prop_examples <- theta[1:N,]
+colnames(topic_prop_examples) <- topic_names
+vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
+                           document = factor(rownames(topic_prop_examples))), 
+                     variable.name = "topic", id.vars = "document")  
+ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
+  geom_bar(stat="identity") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
+  coord_flip() +
+  facet_wrap(~ document, ncol = 5) +
+  theme(legend.position = "none")
 
-# Create LDA model
-emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
+```
+
+```{r}
+
+# What are the most probable topics in the entire collection?
+topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
+names(topicProportions) <- topic_names     # assign the topic names we created before
+sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
+
+soP <- sort(topicProportions, decreasing = TRUE)
+paste(round(soP, 5), ":", names(soP))
+
+# How many documents are with the highest probability in each topic?
+K = 5
+countsOfPrimaryTopics <- rep(0, K)
+names(countsOfPrimaryTopics) <- topic_names
+for (i in 1:nDocs(emu_dtm)) {
+  topicsPerDoc <- theta[i, ] # select topic distribution for document i
+  # get first element position from ordered list
+  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
+  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
+}
+sort(countsOfPrimaryTopics, decreasing = TRUE)
+
+so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
+paste(so, ":", names(so))
+```
+
+```{r}
+topicToFilter <- 6  # you can set this manually ...
+# ... or have it selected by a term in the topic name (e.g. 'children')
+topicToFilter <- grep('river', topic_names)[1] 
+topicThreshold <- 0.2
+selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
+filteredCorpus <- emu_theses[selectedDocumentIndexes,]
+# show length of filtered corpus
+filteredCorpus
+```
+
+```{r}
+
+N = 82
+topic_prop_per_year <- theta[1:N,]
+colnames(topic_prop_per_year) <- topic_names
+
+vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
+                           document = factor(rownames(topic_prop_per_year))), 
+                     variable.name = "topic", id.vars = "document") |> 
+    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))
+
+# Plot topic proportions per year
+vizDataFrameWithYear |> 
+  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
+  geom_bar(stat = "identity", position = "fill") +
+  scale_fill_viridis_d(name = "Topics") +
+  xlab("Graduation year") +
+  ylab("Proportion of topics") +
+  labs(title = "Topic proportions per year")
+
+```
+
+```{r lda-vis}
+
+# TODO check if this can be included in the Shiny dashboard
+# Function to approximate the distance between topics
+svd_tsne <- function(x) tsne(svd(x)$u)
+
+# Convert DTM into JSON required by the LDAvis package
+json <- LDAvis::createJSON(
+  phi = beta, 
+  theta = theta, 
+  doc.length = rowSums(as.matrix(emu_dtm)), 
+  vocab = colnames(emu_dtm), 
+  term.frequency = colSums(as.matrix(emu_dtm)),
+  mds.method = svd_tsne,
+  plot.opts = list(xlab="", ylab="")
+)
+
+# Visualise topics model with LDAvis
+LDAvis::serVis(json)
 
@@
 ```{r topic-model-beta}
+#| dependson: topic-model-lda
 
 # Extract topics as mixtures of terms from LDA model
-emu_topics <- tidy(emu_lda, matrix = "beta")
+beta_tidy <- tidy(emu_lda, matrix = "beta")
 
 # The top n most common words in each topic
-emu_top_terms <- get_top_words_per_topic(emu_topics, 20)
+emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)
 
@@
 # TODO can this be generalised to more than two topics?
+
 beta_spread <- emu_topics |> 

### 2023-07-10 analysis/paper/paper.qmd
Update analysis/paper/paper.qmd

@@
 colnames(topic_prop_examples) <- topic_names
+# TODO change to pivot_longer
 vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 

### 2023-07-10 analysis/paper/paper.qmd
Update analysis/paper/paper.qmd

@@
 
+# TODO change to pivot_longer
 vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 

### 2023-07-10 R/analyse.R
Update `get_ngrams` function

@@ -237,6 +237,7 @@ visualise_first_n_components <- function(word_vectors, dim = 24, top_n = 12) {
 #' @param text_col The name of the column where the text to be tokenised is located
 #' @param title_col The name of the title column
 #' @param stem If `TRUE` create additional columns with stemmed version of tokens
+#' @param lemma If `TRUE` create additional columns with lemmatised version of tokens
 #'
 #' @return A data frame with an "ngram" column and columns for its constituent words
 #' @export
@@ -244,7 +245,9 @@ get_ngrams <- function(data,
                        n,
                        title_col = NULL,
                        text_col = NULL,
-                       stem = FALSE) {
+                       rm_n_short_words = 2,
+                       stem = FALSE,
+                       lemma = FALSE) {
 
   # Create ngrams column and separate columns for each of the n terms
   data_ngrams <- data |>
@@ -256,9 +259,24 @@ get_ngrams <- function(data,
   # Remove stop words from each column starting with "w_"
   w_cols <- names(data_ngrams)[grepl("w_", names(data_ngrams))]
 
+  # Create vector of short words to be removed
+  short_words <- data |>
+    select(text) |>
+    unnest_tokens(output = word, input = text) |>
+    anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
+    mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
+    filter(nchar(word) <= rm_n_short_words)
+
   for (i in w_cols) {
     data_ngrams <- data_ngrams |>
-      anti_join(stop_words, by = set_names("word", i))
+      anti_join(stop_words, by = set_names("word", i)) |>
+      anti_join(short_words, by = set_names("word", i)) |>
+      anti_join(urbanism_stopwords(add_stopwords = c("emu", "tu", "delft", "ku",
+                                                     "leuven", "upc", "barcelona",
+                                                     "iuav", "venice"),
+                                   convert_to_regex = FALSE) %>%
+                  data.frame(word = .),
+                by = set_names("word", i))
   }
 
   # Add columns with stemmed versions of individual terms if `stem = TRUE`
@@ -268,6 +286,16 @@ get_ngrams <- function(data,
         tidyselect::starts_with("w_"),
         ~ SnowballC::wordStem(words = .x, language = "porter"),
         .names = "{.col}_stem"
+      ))
+  }
+
[…]

### 2023-07-10 analysis/paper/paper.qmd
Update ngram script in paper

--- 2023-07-06 analysis/paper/paper.qmd
@@
   "tidytext", "SnowballC", "tm",                # Text processing and analysis
-  "topicmodels", "LDAvis", "tsne",              # Topic modeling
+  "topicmodels", "LDAvis",                      # Topic modeling
   "tidygeocoder",                               # Working with spatial data
@@
 
-### Topic model
+## Topic model
 
 Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
-We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
+We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
 In LDA, each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
-
-The number of topics was qualitatively determined
-
-```{r topic-model-lda}
-#| dependson: short-stop-words
-
-# Prepare document-term matrix as input for LDA model
-
-# Reduce the size of the DTM by excluding words with a frequency lower than 5
-emu_dtm <- convert_to_dtm(emu_theses_words, 
-                          title_col = "title", 
-                          word_col = "word", 
-                          min_term_freq = 5)
-
-# Fit LDA model with K clusters
-# # Determine K value empirically
-# K = get_number_of_topics(emu_dtm)
-# # This returns K = 23
-# TODO run the model with K = 23 and compare with results using qualitative value
-
-# Determine K value qualitatively
-# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
-emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
-               control = list(seed = 2023, iter = 500, verbose = 25))
-
-# Extract beta and theta statistics from LDA model
-beta <- posterior(emu_lda)$terms
-theta <- posterior(emu_lda)$topics
-
-# Preview 10 terms from each topic
-terms(emu_lda, 10)
-
-# Add pseudo-names to topics based on the top 5 words in each topic
-topic_names <- name_topics(beta)
-
-```
+The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.
 
@@
 
-# What is the dominant topic in each document?
-N = 20
-topic_prop_examples <- theta[1:N,]
-colnames(topic_prop_examples) <- topic_names
-vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
-                           document = factor(rownames(topic_prop_examples))), 
-                     variable.name = "topic", id.vars = "document")  
-ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
-  geom_bar(stat="identity") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
-  coord_flip() +
-  facet_wrap(~ document, ncol = 5) +
-  theme(legend.position = "none")
+# Prepare document-term matrix as input for LDA model
+emu_dtm <- convert_to_dtm(emu_words, "title", "word")
 
-```
-
-```{r}
-
-# What are the most probable topics in the entire collection?
-topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
-names(topicProportions) <- topic_names     # assign the topic names we created before
-sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
-
-soP <- sort(topicProportions, decreasing = TRUE)
-paste(round(soP, 5), ":", names(soP))
-
-# How many documents are with the highest probability in each topic?
-K = 5
-countsOfPrimaryTopics <- rep(0, K)
-names(countsOfPrimaryTopics) <- topic_names
-for (i in 1:nDocs(emu_dtm)) {
-  topicsPerDoc <- theta[i, ] # select topic distribution for document i
-  # get first element position from ordered list
-  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
-  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
-}
-sort(countsOfPrimaryTopics, decreasing = TRUE)
-
-so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
-paste(so, ":", names(so))
-```
-
-```{r}
-topicToFilter <- 6  # you can set this manually ...
-# ... or have it selected by a term in the topic name (e.g. 'children')
-topicToFilter <- grep('river', topic_names)[1] 
-topicThreshold <- 0.2
-selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
-filteredCorpus <- emu_theses[selectedDocumentIndexes,]
-# show length of filtered corpus
-filteredCorpus
-```
-
-```{r}
-
-N = 82
-topic_prop_per_year <- theta[1:N,]
-colnames(topic_prop_per_year) <- topic_names
-
-vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
-                           document = factor(rownames(topic_prop_per_year))), 
-                     variable.name = "topic", id.vars = "document") |> 
-    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))
-
-# Plot topic proportions per year
-vizDataFrameWithYear |> 
-  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
-  geom_bar(stat = "identity", position = "fill") +
-  scale_fill_viridis_d(name = "Topics") +
-  xlab("Graduation year") +
-  ylab("Proportion of topics") +
-  labs(title = "Topic proportions per year")
-
-```
-
-```{r lda-vis}
-
-# TODO check if this can be included in the Shiny dashboard
-# Function to approximate the distance between topics
-svd_tsne <- function(x) tsne(svd(x)$u)
-
-# Convert DTM into JSON required by the LDAvis package
-json <- LDAvis::createJSON(
-  phi = beta, 
-  theta = theta, 
-  doc.length = rowSums(as.matrix(emu_dtm)), 
-  vocab = colnames(emu_dtm), 
-  term.frequency = colSums(as.matrix(emu_dtm)),
-  mds.method = svd_tsne,
-  plot.opts = list(xlab="", ylab="")
-)
-
-# Visualise topics model with LDAvis
-LDAvis::serVis(json)
+# Create LDA model
+emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))
 
@@
 ```{r topic-model-beta}
-#| dependson: topic-model-lda
 
 # Extract topics as mixtures of terms from LDA model
-beta_tidy <- tidy(emu_lda, matrix = "beta")
+emu_topics <- tidy(emu_lda, matrix = "beta")
 
 # The top n most common words in each topic
-emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)
+emu_top_terms <- get_top_words_per_topic(emu_topics, 20)
 
@@
 # TODO can this be generalised to more than two topics?
-
 beta_spread <- emu_topics |> 
@@
 
-```{r}
-#| label: ngrams
+```{r ngrams}
 #| dependson: clean-text
@@
 emu_bigrams <- emu |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
+  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", lemma = TRUE) |> 
+  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = "_", remove = FALSE)
 
 bigram_counts <- emu_bigrams |> 
-  count(w_1, w_2, sort = TRUE)
+  count(ngram, sort = TRUE) |> 
+  filter(n > 20)
 
-bigram_stem_counts <- emu_bigrams |> 
-  count(w_1_stem, w_2_stem, sort = TRUE)
+emu_bigrams <- emu_bigrams |> 
+  filter(ngram %in% bigram_counts$ngram)
 
-bigram_counts
-bigram_stem_counts
+bigram_lemma_counts <- emu_bigrams |> 
+  count(w_1_lemma, w_2_lemma, sort = TRUE) |> 
+  filter(n > 5)
+
+# Concatenate bigrams
+
+# # First attempt using a named vector in `str_replace_all()`
+c_bigrams <- paste(emu_bigrams$ngram_lemma, sep = "_")
+names(c_bigrams) <- emu_bigrams$ngram
+
+system.time(
+  emu$text_clean <- emu$text_clean |>
+    str_replace_all(fixed(c_bigrams))
+)
+
+# Second attempt with `mgsub::mgsub()`
+# emu$text_clean <- mgsub(emu$text_clean,
+#                         pattern = emu_bigrams$ngram,
+#                         replacement = emu_bigrams$ngram_lemma)
+
 ```
@@
 ```{r}
+
 ## Trigrams
 emu_trigrams <- emu |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text")
+  get_ngrams(n = 3, title_col = "title", text_col = "text_clean", lemma = TRUE)
 
@@
 
-trigram_counts
+trigram_lemma_counts <- emu_trigrams |> 
+  count(w_1_lemma, w_2_lemma, w_3_lemma, sort = TRUE)
 
+```
+
+```{r}
 ## Bigram tf-idf
@@
 bigram_tf_idf
+```
 
+```{r}
 ## Trigram tf-idf
@@
 trigram_tf_idf
+```
 
+```{r}
 ## Show network of bigrams

### 2023-07-10 analysis/paper/paper.qmd
Add stashed changes

--- 2023-07-10 analysis/paper/paper.qmd
@@
 
+# TODO replace the following lines with the metadata included in the package
+# emu_metadata <- data(emu_metadata)
 ## Read metadata for all theses
@@
 
+# TODO replace this and next code chunk with reading `text_raw` from the package
+# emu <- data(emu_raw)
 pdf_paths <- here(data_path, emu$file_name)
@@
 #| dependson: topic-model-lda
+
+# TODO Explore including concatenated collocations in the topic model
 

### 2023-07-10 R/clean.R

@@ -102,7 +102,7 @@ find_meta_stopwords <- function(metadata,
 #' @return A vector of urbanism-related stopwords
 #' @export
 urbanism_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
-  stop_words <- c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km")
+  stop_words <- c(add_stopwords, "city", "cities", "urban\\b", "urbanism", "hab", "km")
 
   if (convert_to_regex) return(paste(stop_words, collapse = "|"))
   else return(stop_words)

### 2023-07-10 analysis/supplementary-materials/Package_data_workflow.R
separated metadata file from the text file

@@ -1,5 +1,5 @@
 # Create data-raw folder to compute package datasets ----------------------
 usethis::use_data_raw()
 
-# Create data folder with data to be made available as a part of the pacakge
+# Create data folder with data to be made available as a part of the package
 usethis::use_data(emu, overwrite = TRUE)

### 2023-07-10 analysis/supplementary-materials/prepare-for-publication.R

@@ -112,10 +112,22 @@ emu_theses <- emu_theses |>
          longitude, abstract,
          text_raw)
 
-# write CSV file ----------------
+# separate text file and metadata file ----------------
+emu_metadata <- emu_theses |>
+  select(-text_raw)
 
-emu_url <- here("analysis", "data", "derived_data", "emu_raw.csv")
+emu_raw <- emu_theses |>
+  select(ID, text_raw)
 
-readr::write_csv(emu_theses, emu_url)
+# write CSV files ----------------
 
+# emu raw text file
+emu_raw_path <- here("analysis", "data", "derived_data", "emu_raw.csv")
+
+readr::write_csv(emu_raw, emu_path)
+
+# emu metadata file
+emu_meta_path <- here("analysis", "data", "derived_data", "emu_metadata.csv")
+
+readr::write_csv(emu_metadata, emu_meta_path)
 

### 2023-07-10 analysis/paper/paper.qmd
Update code chunk lables, options, names and todos

@@
 #| warning: false
+#| message: false
 
@@
   "here",                                       # Managing paths
+  # TODO remove reshape2 if `melt()` is replaced by`pivot_longer()`
   "tidyverse", "reshape2", "widyr", "readxl",   # Data manipulation and visualisation
@@
 emu <- read_csv(here(data_path, "theses-metadata.csv")) |>
-  mutate(text = "") |>
+  mutate(text_raw = "") |>
   filter(!is.na(pdf_via))
@@
 #| cache: true
+#| message: false
 
@@
   # On the first run, this function might take extra time as a Python environment needs to be set up
-  emu$text <- convert_pdf_text(pdf_paths)
+  emu$text_raw <- convert_pdf_text(pdf_paths)
 
@@
 } else {
-  emu$text <- read_csv(data_path)$text
+  emu$text_raw <- read_csv(data_path)$text
   emu$longitude <- read_csv(data_path)$longitude
@@
 
-```{r topic-model-lda}
+```{r topic-model-lda-1}
 #| dependson: short-stop-words
@@
 
-```{r topic-model-lda}
+```{r topic-model-lda-2}
 

### 2023-07-10 analysis/supplementary-materials/prepare-for-publication.R
fixed a typo

@@ -124,7 +124,7 @@ emu_raw <- emu_theses |>
 # emu raw text file
 emu_raw_path <- here("analysis", "data", "derived_data", "emu_raw.csv")
 
-readr::write_csv(emu_raw, emu_path)
+readr::write_csv(emu_raw, emu_raw_path)
 
 # emu metadata file
 emu_meta_path <- here("analysis", "data", "derived_data", "emu_metadata.csv")

### 2023-07-10 analysis/paper/paper.qmd
Update code chunk names

@@
 
-```{r ngrams}
-#| dependson: clean-text
+```{r bigrams}
 
@@
 
-```{r}
+```{r trigrams}
 
@@
 
-```{r}
+```{r bigram-tfidf}
 ## Bigram tf-idf
@@
 
-```{r}
+```{r trigram-tfidf}
 ## Trigram tf-idf

### 2023-07-10 analysis/paper/paper.qmd
Change lemmatised ngram separator to space

@@
 
-## Bigrams
 emu_bigrams <- emu |> 
   get_ngrams(n = 2, title_col = "title", text_col = "text_clean", lemma = TRUE) |> 
-  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = "_", remove = FALSE)
+  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = " ", remove = FALSE)
 

### 2023-07-10 R/analyse.R
Update text_raw column name in get_ngrams function

@@ -356,8 +356,8 @@ get_ngrams <- function(data,
 
   # Create vector of short words to be removed
   short_words <- data |>
-    select(text) |>
-    unnest_tokens(output = word, input = text) |>
+    select(text_raw) |>
+    unnest_tokens(output = word, input = text_raw) |>
     anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
     mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
     filter(nchar(word) <= rm_n_short_words)

### 2023-07-10 analysis/supplementary-materials/Package_data_workflow.R
Update analysis/supplementary-materials/Package_data_workflow.R
Co-authored-by: alwil <alexandra.wilczynska@gmail.com>

@@ -1,5 +1,4 @@
 # Create data-raw folder to compute package datasets ----------------------
 usethis::use_data_raw()
 
-# Create data folder with data to be made available as a part of the package
 usethis::use_data(emu, overwrite = TRUE)

### 2023-07-10 analysis/supplementary-materials/Package_data_workflow.R
Update analysis/supplementary-materials/Package_data_workflow.R
Co-authored-by: alwil <alexandra.wilczynska@gmail.com>

@@ -1,4 +1,3 @@
 # Create data-raw folder to compute package datasets ----------------------
 usethis::use_data_raw()
 
-usethis::use_data(emu, overwrite = TRUE)

### 2023-07-11 analysis/paper/paper.qmd

--- 2023-07-10 analysis/paper/paper.qmd
@@
 
-# Read thesis metadata ----
-data_path <- here("analysis", "data", "raw_data")
-pdf_names <- dir(data_path, pattern = "*.pdf")
-
-# TODO replace the following lines with the metadata included in the package
-# emu_metadata <- data(emu_metadata)
-## Read metadata for all theses
-all_theses <- read_xlsx(path = here(data_path, "theses-all.xlsx"))
-
-## Read only metadata for theses with PDF available
-emu <- read_csv(here(data_path, "theses-metadata.csv")) |>
-  mutate(text_raw = "") |>
-  filter(!is.na(pdf_via))
-
-# TODO replace this and next code chunk with reading `text_raw` from the package
-# emu <- data(emu_raw)
-pdf_paths <- here(data_path, emu$file_name)
 ```
@@
 
-data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")
+# Read thesis metadata ----
+data(emu_metadata)
 
-if (!file.exists(data_path)) {
-  # On the first run, this function might take extra time as a Python environment needs to be set up
-  emu$text_raw <- convert_pdf_text(pdf_paths)
+# Load the raw text dataset 
+data(emu_raw)
 
-  # Add latitude and longitude information to theses data frame
-  emu <- geocode_thesis_locations(emu)
-
-  # Write thesis dataframe to file
-  write_csv(emu, data_path)
-} else {
-  emu$text_raw <- read_csv(data_path)$text
-  emu$longitude <- read_csv(data_path)$longitude
-  emu$latitude <- read_csv(data_path)$latitude
-}
+# Join  text with the metadata file  
+emu <- left_join(emu_raw, emu_metadata, by = "ID")
 
@@
 
-emu$text_clean <- emu$text |>
+emu$text_clean <- emu$text_raw |>
   clean_basic()
@@
 
-```{r ngrams}
+```{r}
+#| label: ngrams
 #| dependson: clean-text
@@
 emu_bigrams <- emu |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", lemma = TRUE) |> 
-  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = "_", remove = FALSE)
+  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
 
 bigram_counts <- emu_bigrams |> 
-  count(ngram, sort = TRUE) |> 
-  filter(n > 20)
+  count(w_1, w_2, sort = TRUE)
 
-emu_bigrams <- emu_bigrams |> 
-  filter(ngram %in% bigram_counts$ngram)
+bigram_stem_counts <- emu_bigrams |> 
+  count(w_1_stem, w_2_stem, sort = TRUE)
 
-bigram_lemma_counts <- emu_bigrams |> 
-  count(w_1_lemma, w_2_lemma, sort = TRUE) |> 
-  filter(n > 5)
-
-# Concatenate bigrams
-
-# # First attempt using a named vector in `str_replace_all()`
-c_bigrams <- paste(emu_bigrams$ngram_lemma, sep = "_")
-names(c_bigrams) <- emu_bigrams$ngram
-
-system.time(
-  emu$text_clean <- emu$text_clean |>
-    str_replace_all(fixed(c_bigrams))
-)
-
-# Second attempt with `mgsub::mgsub()`
-# emu$text_clean <- mgsub(emu$text_clean,
-#                         pattern = emu_bigrams$ngram,
-#                         replacement = emu_bigrams$ngram_lemma)
-
+bigram_counts
+bigram_stem_counts
 ```
@@
 ```{r}
-
 ## Trigrams
 emu_trigrams <- emu |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text_clean", lemma = TRUE)
+  get_ngrams(n = 3, title_col = "title", text_col = "text")
 
@@
 
-trigram_lemma_counts <- emu_trigrams |> 
-  count(w_1_lemma, w_2_lemma, w_3_lemma, sort = TRUE)
+trigram_counts
 
-```
-
-```{r}
 ## Bigram tf-idf
@@
 bigram_tf_idf
-```
 
-```{r}
 ## Trigram tf-idf
@@
 trigram_tf_idf
-```
 
-```{r}
 ## Show network of bigrams

### 2023-07-11 R/ngrams.R
Add missing namespace to get_ngrams function

@@ -20,7 +20,7 @@ get_ngrams <- function(data,
 
   # Create ngrams column and separate columns for each of the n terms
   data_ngrams <- data |>
-    select(title_col, text_col) |>
+    dplyr::select(title_col, text_col) |>
     tidytext::unnest_tokens(ngram, text_col, token = "ngrams", n = n) |>
     tidyr::separate(ngram, into = paste("w" , 1:n, sep = "_"),
                     sep = " ", remove = FALSE)

### 2023-07-11 analysis/paper/paper.qmd
Move ngrams to pre-processing

--- 2023-07-10 analysis/paper/paper.qmd
@@
 
-```{r clean-text}
-#| echo: false
-#| dependson: get-pdf-text
-
-emu$text_clean <- emu$text |>
-  clean_basic()
-
-```
-
 # Introduction
@@
 
-```{r stop-words}
+```{r pre-process-text}
 
-# Add cleaned text column with removed stop words
+meta_sw <- find_meta_stopwords(emu)
+urbanism_sw <- urbanism_stopwords(
+    add_stopwords = c("emu", "european postgraduate master in urbanism", 
+                      "european post master in urbanism",
+                      "european postmaster in urbanism",
+                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning"))
+custom_sw <- c("space", "spatial", "spaces", "plan", "public", "development") |>
+  (\(.) paste0("\\b", ., "\\b"))() |> 
+  paste(collapse = "|")
+
+# Add cleaned text column
+emu$text_clean <- emu$text_raw |>
+  # Remove white spaces and punctuation, and change text to lower-case
+  clean_basic() |>
+  # Remove title, subtitle and full title
+  str_remove_all(meta_sw) |> 
+  # Remove dataset-specific words
+  str_remove_all(urbanism_sw) |>
+  # Remove thesis-specific words
+  str_remove_all(thesis_sw)
+
+# Concatenate ngrams
+emu$text_clean <- emu |> 
+  c_ngrams(n_max = 3, text_col = "text_clean",
+           title_col = "title", min_freq = 20, lemma = TRUE)
+
+  # Words that were not included in ngrams are safe to remove now
 emu$text_clean <- emu$text_clean |> 
-  str_remove_all(find_meta_stopwords(emu)) |> 
-  str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
-                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
+  str_remove_all(custom_sw)
 
-# Unnest words and remove words of one and two letters
+```
+
+```{r tokenise}
+
+# TODO Replace title column with ID column
+
+# Unnest tokens and remove words of one and two letters
 emu_words <- emu |> 
@@
   anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
-  mutate(word = textstem::lemmatize_words(word))  # lemmatise words
+  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
+  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
 
@@
 Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
-
-```{r short-stop-words}
-#| dependson: stop-words
-
-#short_words(emu_words$word, l_word = 3, l_list = 100)
-
-emu_words <- emu_words |>  
-  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
-
-```
 
@@
 ```{r topic-model-lda-1}
-#| dependson: short-stop-words
 
@@
 # Reduce the size of the DTM by excluding words with a frequency lower than 5
-emu_dtm <- convert_to_dtm(emu_theses_words, 
+emu_dtm <- convert_to_dtm(emu_words, 
                           title_col = "title", 
@@
 
-```{r bigrams}
-
-emu_bigrams <- emu |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", lemma = TRUE) |> 
-  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = " ", remove = FALSE)
-
-bigram_counts <- emu_bigrams |> 
-  count(ngram, sort = TRUE) |> 
-  filter(n > 20)
-
-emu_bigrams <- emu_bigrams |> 
-  filter(ngram %in% bigram_counts$ngram)
-
-bigram_lemma_counts <- emu_bigrams |> 
-  count(w_1_lemma, w_2_lemma, sort = TRUE) |> 
-  filter(n > 5)
-
-# Concatenate bigrams
-
-# # First attempt using a named vector in `str_replace_all()`
-c_bigrams <- paste(emu_bigrams$ngram_lemma, sep = "_")
-names(c_bigrams) <- emu_bigrams$ngram
-
-system.time(
-  emu$text_clean <- emu$text_clean |>
-    str_replace_all(fixed(c_bigrams))
-)
-
-# Second attempt with `mgsub::mgsub()`
-# emu$text_clean <- mgsub(emu$text_clean,
-#                         pattern = emu_bigrams$ngram,
-#                         replacement = emu_bigrams$ngram_lemma)
-
-```
-
-```{r trigrams}
-
-## Trigrams
-emu_trigrams <- emu |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text_clean", lemma = TRUE)
-
-trigram_counts <- emu_trigrams |> 
-  count(w_1, w_2, w_3, sort = TRUE)
-
-trigram_lemma_counts <- emu_trigrams |> 
-  count(w_1_lemma, w_2_lemma, w_3_lemma, sort = TRUE)
-
-```
+A frequency count of ngrams reveals collocations like "public space" or "socio-spatial segregation" across the corpus.
+To include these in the topic model, we concatenated them with an underscore, so that they are considered one word during the word-level tokenisation.
+Only bigrams and trigrams, that is two- and three-word collocations, were extracted from the corpus.
 

### 2023-07-11 R/clean.R
Fix typos, add todos and update stopword functions

@@ -1,6 +1,6 @@
 #' Clean body of text
 #'
-#' Function removes white space, punctuation and  converts the strings to lowercase
+#' Function removes white space, punctuation and converts the strings to lowercase
 #'
 #' @param text A vector string to be converted
 #' @return A cleaned vector string
@@ -37,7 +37,7 @@ head_text <- function(text_vector, head_n = 10, sub_start = 1 , sub_end = 500 ){
   head_text
 }
 
-
+# TODO Should we remove the author names from this function as anonymisation has already been performed on the raw data?
 #' Find custom stop-phrases based on metadata file
 #'
 #' Function scans the dataset with metadata for fields that can be removed from the body fo text before the text analysis
@@ -102,7 +102,9 @@ find_meta_stopwords <- function(metadata,
 #' @return A vector of urbanism-related stopwords
 #' @export
 urbanism_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
-  stop_words <- c(add_stopwords, "city", "cities", "urban\\b", "urbanism", "hab", "km")
+  stop_words <- paste("\\b",
+                      c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km"),
+                      "\\b", sep = "")
 
   if (convert_to_regex) return(paste(stop_words, collapse = "|"))
   else return(stop_words)
@@ -116,19 +118,14 @@ urbanism_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
 #' @return A vector of thesis-related stopwords
 #' @export
 thesis_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
-  stop_words <- c(
-        add_stopwords,
-        "preface",
-        "foreword",
-        "introduction",
-        "conclusion",
-        "thesis",
-        "source",
-        "author",
-        "colophon",
-        "acknowledgements",
-        "references"
-        )
+  stop_words <- paste("\\b", c(add_stopwords,
+                               "preface", "foreword", "acknowledgements",
+                               "colophon", "introduction", "conclusion",
+                               "thesis",
+                               "source",
+                               "author", "advisor",
+                               "references", "bibliography"),
+                      "\\b", sep = "")
 
   if (convert_to_regex) return(paste(stop_words, collapse = "|"))
   else return(stop_words)

### 2023-07-11 analysis/paper/paper.qmd
Update script for word counts

--- 2023-07-11 analysis/paper/paper.qmd
@@
 
+# Read thesis metadata ----
+data_path <- here("analysis", "data", "raw_data")
+pdf_names <- dir(data_path, pattern = "*.pdf")
+
+# TODO replace the following lines with the metadata included in the package
+# emu_metadata <- data(emu_metadata)
+## Read metadata for all theses
+all_theses <- read_xlsx(path = here(data_path, "theses-all.xlsx"))
+
+## Read only metadata for theses with PDF available
+emu <- read_csv(here(data_path, "theses-metadata.csv")) |>
+  mutate(text_raw = "") |>
+  filter(!is.na(pdf_via))
+
+# TODO replace this and next code chunk with reading `text_raw` from the package
+# emu <- data(emu_raw)
+pdf_paths <- here(data_path, emu$file_name)
 ```
@@
 
-# Read thesis metadata ----
-data(emu_metadata)
+data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")
 
-# Load the raw text dataset 
-data(emu_raw)
+if (!file.exists(data_path)) {
+  # On the first run, this function might take extra time as a Python environment needs to be set up
+  emu$text_raw <- convert_pdf_text(pdf_paths)
 
-# Join  text with the metadata file  
-emu <- left_join(emu_raw, emu_metadata, by = "ID")
+  # Add latitude and longitude information to theses data frame
+  emu <- geocode_thesis_locations(emu)
+
+  # Write thesis dataframe to file
+  write_csv(emu, data_path)
+} else {
+  emu$text_raw <- read_csv(data_path)$text
+  emu$longitude <- read_csv(data_path)$longitude
+  emu$latitude <- read_csv(data_path)$latitude
+}
 
@@
 
-emu$text_clean <- emu$text_raw |>
+emu$text_clean <- emu$text |>
   clean_basic()
@@
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_words, top_n = 3)
+get_top_words_per_document(data = emu_words, top_n = 10)
 ```
@@
 # Preview term frequency distributions in a subset of theses
+
+emu_words_count <- emu_words |> 
+  count(title, word, sort = TRUE) |> 
+  ungroup()
+
+emu_words_total <- emu_words_count |> 
+  group_by(title) |> 
+  summarise(total = sum(n))
+
+emu_words_count <- left_join(emu_words_count, emu_words_total)
+
 ggplot(emu_words_count |> 

### 2023-07-11 analysis/paper/paper.qmd
Update layout

--- 2023-07-11 analysis/paper/paper.qmd
@@
 
-```{r get-pdf-text}
+```{r load-data}
 #| echo: false
-#| cache: true
 #| message: false
 
-data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")
+# Read thesis metadata
+data(emu_metadata)
 
-if (!file.exists(data_path)) {
-  # On the first run, this function might take extra time as a Python environment needs to be set up
-  emu$text_raw <- convert_pdf_text(pdf_paths)
+# Load raw text data
+data(emu_raw)
 
-  # Add latitude and longitude information to theses data frame
-  emu <- geocode_thesis_locations(emu)
-
-  # Write thesis dataframe to file
-  write_csv(emu, data_path)
-} else {
-  emu$text_raw <- read_csv(data_path)$text
-  emu$longitude <- read_csv(data_path)$longitude
-  emu$latitude <- read_csv(data_path)$latitude
-}
+# Join text with the metadata file  
+emu <- left_join(emu_raw, emu_metadata, by = "ID")
 

### 2023-07-11 analysis/paper/paper.qmd
Add anonymisation stop words

@@
 thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning"))
+ananymisiation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
 custom_sw <- c("space", "spatial", "spaces", "plan", "public", "development") |>
@@
   # Remove thesis-specific words
-  str_remove_all(thesis_sw)
+  str_remove_all(thesis_sw) |> 
+  # Remove text included with anonymisation
+  str_remove_all(ananymisiation_sw)
 

### 2023-07-12 analysis/paper/paper.qmd
Update text

--- 2023-07-11 analysis/paper/paper.qmd
@@
 
-```{r get-pdf-text}
+```{r load-data}
 #| echo: false
-#| cache: true
 #| message: false
 
-# Read thesis metadata ----
+# Read thesis metadata
 data(emu_metadata)
 
-# Load the raw text dataset 
+# Load raw text data
 data(emu_raw)
@@
 
-Distinctive features of the program:  
-- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice
-- provides a bridge towards PhD-level research
-- all students had an exchange semester at one of the other three program universities
+Distinctive features of the program:\
+- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice - provides a bridge towards PhD-level research - all students had an exchange semester at one of the other three program universities
 
-In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program.
+In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
 To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+
+<!-- If the focus of the paper shifts to topic modeling the paragraphs below can be swapped with the paragraphs describing the EMU program. -->
+
+This paper leverages topic modeling as an unsupervised machine learning method to reveal latent semantic structure and relationships between the latent topics in a large document collection [@muchene2021].
+
+NLP... and the focus on its subtopic topic modeling
+
+Topic modeling provides a way to generative
+
+-   bag of words
+
+-   decomposition into two low-dimensional matrices for doc-topic distribution and word-topic distribution, respectively
+
+-   LDA and extensions of it (CTM, HDP, GLDA, DTM) are broadly adopted in topic modeling [@muchene2021]
+
+<!-- Describe the research gaps -->
+
+Research gaps and opportunities:
+
+-   The scarcity of studies that use statistical methods for unstructured data in urbanism education
+
+-   Theses are representative outputs of educational programs that can be used as data for text mining.
+The growing availability of theses in digital repositories presents an opportunity to that end.
+The TU Delft repository was established in 2007, in the year when the EMU program started.
+
+-   Applications of topic modeling in the field of urbanism
+
+-   Patterns revealed in past education can potentially better inform future decisions in education in the field
+
+-   The EMU program...
+
+Challenges:
+
+Contribution:
+
+-   This paper showcases reproducible research workflow.
+It is part of a reproducible research compendium and accompanied by a published dataset.
+The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023].
+The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository \[ADD CITATION\] and it is available for exploration in a Shiny dashboard at \[ADD LINK\]
+
+-   Give an overview of th predominant topics and their evolution over the years of the EMU program.
+In addition to the natural document classification <!--# explain --> of documents provided by a repository, LDA allows documents to be part of multiple topics [@muchene2021].
+
+<!--# Outline -->
+
+The following section describes the case study data, pre-processing steps and the topic modeling workflow.
 
@@
 -   To what extent were the topics and evolution thereof influenced by the exchange semester followed by the students?
+
     -   Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
@@
 
-## Data collection
+## Data
 
+Of the total of `r nrow(all_theses)` theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->.
+Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
+The full-text documents were obtained either from the TU Delft Education Repository (an API for downloading a large set of full-text data is not available), when available, or from the authors via email.
+Even though repository entries are open access, it was unclear if the content of the theses can be mined.
+Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
+Only PDFs for which permission was obtained were included in the corpus.With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
 The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
-With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
+@tbl-data gives an overview of the data included in the analysis.
+
+```{r tbl-data}
+#| tbl-cap: Overview of theses included in the analysis
+#| tbl-cap-location: top
+
+```
 
@@
 
-A total number of `r nrow(emu)` out of `r nrow(all_theses)` were analysed.
+For each thesis, henceforth referred to as "document", the metadata consisted of the `ID`, `title`, `graduation_year`, `graduation_semester`, `exchange_semester`, `location`, `latitude`, `longitude`, `abstract`, `text_raw` and `text_clean` <!--# check column names or add them in an inline code chunk with epoxy -->.
+
+Both the raw and cleaned data are published in the 4TU.ResearchData repository \[ADD CITATION\].
 
@@
 This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.
+`unnest_tokens()` removes punctuation, converts to lower-case, and separate all words into individual tokens.
+
+Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified during preliminarey EDA were concatenated by replacing white spaces with underscores.
 
@@
 
-Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
 We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
@@
 
-The number of topics was qualitatively determined
+LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].
+
+Latent topics are measured as patterns of word co-occurrence
+
+Describe model parameters:
+
+-   collection of M = `r nrow(emu)` documents
+
+-   vocabulary of V = `r length(unique(emu_words$word))` words
+
+Determining tuning parameters:
+
+-   The number of topics was qualitatively determined
+
+LDA requires a document-term matrix as input.
 
@@
 
+## Overall patterns
+
+The tf-idf statistic shows that...
+
+## Topics
+
+## The dynamics of thesis topics
+
+Topics such as ... have gained importance over the years, while ... have lost prominence.
+
 ```{r top-words-document}
@@
 
-[Describe word embeddings ...]
+\[Describe word embeddings ...\]
 
@@
 
-[What terms should be checked with the function below?]
+\[What terms should be checked with the function below?\]
 
@@
 
-The first [?] components... [what do they show?]
+The first \[?\] components... \[what do they show?\]
 
@@
 # Discussion
+
+-   Although the focus of this study is on describing a fixed corpus though unsupervised machine learning, it can provide input into supervised predictive models.
 

### 2023-07-12 analysis/paper/paper.qmd
Update text and make small corrections in code chunks

@@
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.
+To that end, we employed a text mining approach in which we analysed its output: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE), "-", max(emu$graduation_year, na.rm = TRUE))`.
 
@@
 
-NLP... and the focus on its subtopic topic modeling
+In Natural Language Processing (NLP) ... and the focus on its subtopic topic modeling
 
@@
 
-Of the total of `r nrow(all_theses)` theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->.
+Of the total of 96 theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->.
 Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
@@
 
+-   corpus D
+
+-   N = `r length(emu_words$word)` words
+
 -   vocabulary of V = `r length(unique(emu_words$word))` words
@@
 ```{r topic-model-lda-1}
-#| dependson: short-stop-words
+#| include: false
 
@@
 
-```{r topic-model-lda-2}
+```{r fig-topics-per-doc}
 
@@
                      variable.name = "topic", id.vars = "document")  
-ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
+ggplot(data = vizDataFrame, aes(topic, value), ylab = "proportion") + 
   geom_bar(stat="identity") +
@@
 # Extract topics as mixtures of terms from LDA model
-beta_tidy <- tidy(emu_lda, matrix = "beta")
+emu_topics <- tidy(emu_lda, matrix = "beta")
 
 # The top n most common words in each topic
-emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)
+emu_top_terms <- get_top_words_per_topic(emu_topics, 20)
 
@@
 emu_trigrams <- emu |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text")
+  get_ngrams(n = 3, title_col = "title", text_col = "text_raw")
 
@@
   scale_y_log10()
+```
+
+```{r}
+#| eval: false
 
@@
 emu_words_count |> 
-  filter(title == unique(title)[1:6]) |> 
+  get_tf_idf(title_col = "title", word_col = "word") |> 
+  filter(title == unique(title)[1:2]) |>  # Keep only the first 6 theses
   arrange(desc(tf_idf)) |> 
@@
 
+tidy_words <- emu |> 
+  select(title, text_clean) |> 
+  unnest_tokens(word, text_clean) |> 
+  add_count(word) |> 
+  filter(n >= 50) |> 
+  select(-n)
+
 # Summarise word embeddings into document embeddings for modeling purposes
-word_matrix <- tidy_emu |> 
+word_matrix <- tidy_words |> 
   count(title, word) |> 

### 2023-07-12 analysis/paper/paper.qmd

@@
 emu <- left_join(emu_raw, emu_metadata, by = "ID")
-
-```
-
-```{r clean-text}
-#| echo: false
-#| dependson: get-pdf-text
-
-emu$text_clean <- emu$text_raw |>
-  clean_basic()
 
@@
 
-```{r stop-words}
+```{r pre-process-text}
 
-# Add cleaned text column with removed stop words
+meta_sw <- find_meta_stopwords(emu)
+urbanism_sw <- urbanism_stopwords(
+    add_stopwords = c("emu", "european postgraduate master in urbanism", 
+                      "european post master in urbanism",
+                      "european postmaster in urbanism",
+                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning"))
+ananymisiation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
+custom_sw <- c("space", "spatial", "spaces", "plan", "public", "development") |>
+  (\(.) paste0("\\b", ., "\\b"))() |> 
+  paste(collapse = "|")
+
+# Add cleaned text column
+emu$text_clean <- emu$text_raw |>
+  # Remove white spaces and punctuation, and change text to lower-case
+  clean_basic() |>
+  # Remove title, subtitle and full title
+  str_remove_all(meta_sw) |> 
+  # Remove dataset-specific words
+  str_remove_all(urbanism_sw) |>
+  # Remove thesis-specific words
+  str_remove_all(thesis_sw) |> 
+  # Remove text included with anonymisation
+  str_remove_all(ananymisiation_sw)
+
+# Concatenate ngrams
+emu$text_clean <- emu |> 
+  c_ngrams(n_max = 3, text_col = "text_clean",
+           title_col = "title", min_freq = 20, lemma = TRUE)
+
+  # Words that were not included in ngrams are safe to remove now
 emu$text_clean <- emu$text_clean |> 
-  str_remove_all(find_meta_stopwords(emu)) |> 
-  str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
-                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
+  str_remove_all(custom_sw)
 
-# Unnest words and remove words of one and two letters
+```
+
+```{r tokenise}
+
+# TODO Replace title column with ID column
+
+# Unnest tokens and remove words of one and two letters
 emu_words <- emu |> 
@@
 
-```{r}
-#| label: ngrams
-#| dependson: clean-text
+A frequency count of ngrams reveals collocations like "public space" or "socio-spatial segregation" across the corpus.
+To include these in the topic model, we concatenated them with an underscore, so that they are considered one word during the word-level tokenisation.
+Only bigrams and trigrams, that is two- and three-word collocations, were extracted from the corpus.
 
-## Bigrams
+```{r bigram-tfidf}
+
 emu_bigrams <- emu |> 
@@
 
-bigram_stem_counts <- emu_bigrams |> 
-  count(w_1_stem, w_2_stem, sort = TRUE)
+bigram_tf_idf <- emu_bigrams |> 
+  count(title, ngram) |> 
+  bind_tf_idf(ngram, title, n) |> 
+  arrange(desc(tf_idf))
 
-bigram_counts
-bigram_stem_counts
+bigram_tf_idf
 ```
 
-```{r}
-## Trigrams
+```{r trigram-tfidf}
+
 emu_trigrams <- emu |> 
@@
 
-trigram_counts
-
-## Bigram tf-idf
-bigram_tf_idf <- emu_bigrams |> 
-  count(title, ngram) |> 
-  bind_tf_idf(ngram, title, n) |> 
-  arrange(desc(tf_idf))
-
-bigram_tf_idf
-
-## Trigram tf-idf
 trigram_tf_idf <- emu_trigrams |> 

### 2023-07-12 analysis/paper/paper.qmd
Add PLOS ONE template

--- 2023-07-12 analysis/paper/paper.qmd
@@
 format: 
-  docx:
-    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
+  # docx:
+  #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
+  pdf:
+    template: "../templates/plos_latex_template.tex"
 execute:

### 2023-07-12 analysis/templates/plos_latex_template.tex

% Template for PLoS
% Version 3.6 Aug 2022
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage[nopatch=eqnum]{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS journals} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* correspondingauthor@institute.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\begin{eqnarray}
\label{eq:schemeP}
	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
\end{eqnarray}

\section*{Materials and methods}
\subsection*{Etiam eget sapien nibh}

% For figure citations, please use "Fig" instead of "Figure".
Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
\subsubsection*{3rd level heading} 
Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\begin{enumerate}
	\item{react}
	\item{diffuse free particles}
	\item{increment time by dt and go to 1}
\end{enumerate}

\subsection*{Sed ac quam id nisi malesuada congue}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{itemize}
	\item First bulleted item.
	\item Second bulleted item.
	\item Third bulleted item.
\end{itemize}

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.

\section*{Conclusion}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
% 
\begin{thebibliography}{10}

\bibitem{bib1}
Conant GC, Wolfe KH.
\newblock {{T}urning a hobby into a job: how duplicated genes find new
  functions}.
\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

\bibitem{bib2}
Ohno S.
\newblock Evolution by gene duplication.
\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
  Springer-Verlag.; 1970.

\bibitem{bib3}
Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
  infection through a transposon insertion followed by a {D}uplication}.
\newblock PLoS Genet. 2011 Oct;7(10):e1002337.

\end{thebibliography}

\end{document}

### 2023-07-12 analysis/paper/paper.qmd
Add pdf output

--- 2023-07-12 analysis/paper/paper.qmd
@@
 date-format: long
-format: 
+format: pdf
   # docx:
   #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
-  pdf:
-    template: "../templates/plos_latex_template.tex"
+  # pdf:
+  #   template: "../templates/plos_latex_template.tex"
 execute:

### 2023-07-13 analysis/paper/paper.qmd
Add abstract

--- 2023-07-12 analysis/paper/paper.qmd
@@
 abstract: |
-  Text of abstract
+  Prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft,
+  this paper aims to describe the legacy of the EMU. To that end, it reports on the
+  analysis of the graduation theses produced over 2007-2021 when the program was running.
+  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
+  and evolutions thereof, found in the full texts of the theses. Results reveal # topics
+  that are to a large extent consistent with the set-up of the program in two thematic 
+  semesters at TU Delft and three possible exchange semester locations, each characterised 
+  by a specific approach. The paper discusses the relevance of the findings for the
+  state of the profession and education in the field. Set up as a reproducible research
+  compendium accompanied by a published FAIR dataset of the theses and an interactive
+  dashboard, the analysis presented can be both reproduced and applied to other datasets. 
 keywords: |
-  urbanism; education; post-master; text mining
+  urbanism; education; post-master; text mining; topic 
 highlights: |

### 2023-07-13 R/data.R
added full title and subtitle to the metadata file

@@ -25,7 +25,9 @@
 #'   \item{ID}{ID of the thesis}
 #'   \item{graduation_year}{Raw text of the thesis.}
 #'   \item{graduation_semester}{Raw text of the thesis.}
+#'   \item{full_title}{Full title of the thesis.}
 #'   \item{title}{Title of the thesis.}
+#'   \item{subtitle}{Subtitle of the thesis.}
 #'   \item{location}{The location that the thesis is focused on.}
 #'   \item{latitude}{Latitude of the thesis location.}
 #'   \item{longitude}{Longitude of the thesis location.}

### 2023-07-13 analysis/supplementary-materials/prepare-for-publication.R

@@ -98,18 +98,22 @@ emu_theses$text_raw  <- text_raw
 emu_theses <- emu_theses |>
   select(ID, graduation_year,
          graduation_semester,
-         title, location, abstract,
-                text_raw)
+         title, full_title, subtitle,
+         location,
+         abstract,
+         text_raw)
 
 # Add longitude and latitude columns ---------
 emu_theses <- geocode_thesis_locations(emu_theses)
 
 # reorder the columns in the dataset ---------
 emu_theses <- emu_theses |>
-  select(ID, graduation_year,
+  select(ID,
+         graduation_year,
          graduation_semester,
-         title, location, latitude,
-         longitude, abstract,
+         full_title, title, subtitle,
+         location, latitude, longitude,
+         abstract,
          text_raw)
 
 # separate text file and metadata file ----------------

### 2023-07-13 analysis/paper/paper.qmd
Update code chunk labels and visibility options

--- 2023-07-13 analysis/paper/paper.qmd
@@
 execute:
-  echo: true
+  echo: false
   warning: false
@@
 ```{r setup}
-#| echo: false
-#| warning: false
-#| message: false
+#| include: false
 
@@
 ```{r load-data}
-#| echo: false
-#| message: false
+#| include: false
 
-# Read thesis metadata
+# Load thesis metadata
 data(emu_metadata)
@@
 ```{r clean-text}
-#| echo: false
-#| dependson: get-pdf-text
+#| include: false
 
@@
 ```{r fig-map-theses}
-#| warning: false 
-#| echo: false
 #| fig-cap: "Location of EMU theses included in the analysis"
@@
 ```{r short-stop-words}
-#| dependson: stop-words
+#| include: false
 
-#short_words(emu_words$word, l_word = 3, l_list = 100)
+# short_words(emu_words$word, l_word = 3, l_list = 100)
 
@@
 ```{r topic-model-lda-1}
-#| include: false
 
@@
 
-```{r}
+```{r topics-per-corpus}
 
@@
 ```{r word-embeddings-pmi}
-#| include: false
 

### 2023-07-13 analysis/paper/paper.qmd
Add process diagram

@@
   "tidygeocoder",                               # Working with spatial data
-  "igraph", "ggraph"                            # Network analysis and visualization
+  "igraph", "ggraph", "DiagrammeR"              # Network analysis and visualization
 )
@@
 Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
+
+```{r process-diagram}
+
+mermaid("
+graph LR
+    A[Data collection]-->B[Pre-processing]
+    B-->C[Information extraction]
+    C-->D[Visualisation]
+    C-->B
+    D-->E[Results]
+")
+
+```
 

### 2023-07-13 analysis/paper/paper.qmd
Add comment on how to visualise topics per docs

--- 2023-07-13 analysis/paper/paper.qmd
@@
 
+```{r process-diagram}
+
+mermaid("
+graph LR
+    A[Data collection]-->B[Pre-processing]
+    B-->C[Information extraction]
+    C-->D[Visualisation]
+    C-->B
+    D-->E[Results]
+")
+
+```
+
 ## Data
@@
 ```{r fig-topics-per-doc}
+#| fig-cap: Topic probability distribution per document
 
+# TODO Replace this with a table containing topics, documents, and theta
+#      sorted by theta in decreasing order
 # What is the dominant topic in each document?

### 2023-07-13 analysis/paper/paper.qmd
Remove white spaces

--- 2023-07-13 analysis/paper/paper.qmd
@@
 devtools::load_all(".")
-
 ```
@@
 emu <- left_join(emu_raw, emu_metadata, by = "ID")
-
 ```
@@
 ")
-
 ```
@@
   mutate(word = textstem::lemmatize_words(word))  # lemmatise words
-
 ```
@@
 ```{r}
+
 topicToFilter <- 6  # you can set this manually ...
@@
   labs(title = "Topic proportions per year")
-
 ```
@@
 ```{r top-words-document}
+
 # Which are the top n most frequently used words in each thesis?
@@
 ```{r top-words-corpus}
+
 # Which are the top n most frequently used words in all theses?
@@
 bigram_stem_counts
-```
 
-```{r}
 ## Trigrams
@@
 ```{r}
+
 # Preview term frequency distributions in a subset of theses
-
 emu_words_count <- emu_words |> 
@@
 tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")
-
 ```
@@
   )
-
 ```
@@
   nearest_neighbors("resilience")
-
 ```
@@
 visualise_first_n_components(tidy_word_vectors, 24, 12)
-
 ```
@@
 dim(doc_matrix)
-
-doc_matrix
-
 ```

### 2023-07-13 analysis/paper/paper.qmd

--- 2023-07-13 analysis/paper/paper.qmd
@@
   filter(nchar(word) >= 3)  # keep only words that are at least three letters long
-
 ```

### 2023-07-13 R/clean.R
remove punctuation and add word boundaries

@@ -284,13 +284,22 @@ normalise_words <- function(string_vec) {
   sentesized_words <- lapply(normalised_words, stringr::str_to_title)
 
 
+  # Final list of words
   final_words <- mapply(c, normalised_words, lowerised_words, sentesized_words, SIMPLIFY=FALSE)
 
+  # Remove punctuation
+  final_words   <- lapply(final_words,
+                          stringr::str_remove_all, '[[:punct:]]+')
+
+
+  # embed in the word boundaries
+  final_words   <- lapply(final_words, \(x) dplyr::if_else(is.na(unique(x)), NA_character_,
+                                                           unique(paste0("\\b", x, "\\b"))))
+
   final_words  <- lapply(final_words ,
-                         function(x) dplyr::if_else(is.na(unique(x)), NA_character_,
+                         \(x) dplyr::if_else(is.na(x), NA_character_,
                                                     paste(unique(x), collapse = " ")) |> unique()) |> unlist()
 
-  final_words
 
 }
 

### 2023-07-13 analysis/supplementary-materials/prepare-for-publication.R

@@ -77,6 +77,8 @@ pattern_fn <- lapply(pattern_fn, paste, collapse = "|") |>
 text_raw <- mapply(str_replace_all, text_raw, pattern_fn,
                    "FIRST_NAME_REMOVED")
 
+head_text(text_raw[55:60] )
+
 
 # Replace lastname with placeholder
 pattern_ln <- normalise_words(emu_theses$last_name)
@@ -86,6 +88,9 @@ pattern_ln <- lapply(pattern_ln, paste, collapse = "|") |>
 
 text_raw <- mapply(str_replace_all, text_raw, pattern_ln,
                    "LAST_NAME_REMOVED")
+
+head_text(text_raw[55:60] )
+
 
 # replacement_checker(text_raw, emu_theses$last_name, char_before = 0)
 

### 2023-07-13 analysis/supplementary-materials/prepare-for-publication.R
Update analysis/supplementary-materials/prepare-for-publication.R

@@ -77,7 +77,6 @@ pattern_fn <- lapply(pattern_fn, paste, collapse = "|") |>
 text_raw <- mapply(str_replace_all, text_raw, pattern_fn,
                    "FIRST_NAME_REMOVED")
 
-head_text(text_raw[55:60] )
 
 
 # Replace lastname with placeholder

### 2023-07-13 analysis/supplementary-materials/prepare-for-publication.R
Update analysis/supplementary-materials/prepare-for-publication.R

@@ -88,7 +88,6 @@ pattern_ln <- lapply(pattern_ln, paste, collapse = "|") |>
 text_raw <- mapply(str_replace_all, text_raw, pattern_ln,
                    "LAST_NAME_REMOVED")
 
-head_text(text_raw[55:60] )
 
 
 # replacement_checker(text_raw, emu_theses$last_name, char_before = 0)

### 2023-07-13 analysis/paper/paper.qmd
Add keyword
Update label and caption in process diagram code chunk

--- 2023-07-13 analysis/paper/paper.qmd
@@
 keywords: |
-  urbanism; education; post-master; text mining; topic 
+  urbanism; education; post-master; text mining; topic modeling
 highlights: |
@@
 
-```{r process-diagram}
+The research process (@fig-process-diagram)...
+
+```{r fig-process-diagram}
+#| fig-cap: Text mining process
 

### 2023-07-13 analysis/paper/paper.qmd
Add small corrections under data collection

@@
 ```
-
-Description of text extraction; raw data available on 4TU.ResearchData at...
 
@@
 
-## Data
+## Data collection
 
@@
 @tbl-data gives an overview of the data included in the analysis.
+
+Description of text extraction; raw data available on 4TU.ResearchData at...
 
@@
 
-Both the raw and cleaned data are published in the 4TU.ResearchData repository \[ADD CITATION\].
+Both the raw and cleaned data are published in the 4TU.ResearchData repository [ADD CITATION].
 
@@
 
-## Data analysis
-
-### Pre-processing
+## Pre-processing
 
@@
 
+# TODO replace with cleaning steps including ngram concatenation
 # Add cleaned text column with removed stop words

### 2023-07-13 analysis/paper/paper.qmd
Update topic modeling part

@@
 
+## Information extraction
+
 ### Topic model
@@
 # Prepare document-term matrix as input for LDA model
-
 # Reduce the size of the DTM by excluding words with a frequency lower than 5
@@
 # Fit LDA model with K clusters
-# # Determine K value empirically
-# K = get_number_of_topics(emu_dtm)
-# # This returns K = 23
-# TODO run the model with K = 23 and compare with results using qualitative value
+## Determine K value empirically
+k = get_number_of_topics(emu_dtm)
+# TODO run the model with the empirically determined k and compare (or combine) with results using qualitative value
+emu_lda_k <- LDA(emu_dtm, k = k, method = "Gibbs",
+               control = list(seed = 2023, iter = 500, verbose = 25))
 
-# Determine K value qualitatively
-# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
-emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
+## Determine K value qualitatively
+## K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
+emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
                control = list(seed = 2023, iter = 500, verbose = 25))
@@
 # What is the dominant topic in each document?
-N = 20
+N = 82
 topic_prop_examples <- theta[1:N,]
-colnames(topic_prop_examples) <- topic_names
+colnames(topic_prop_examples) <- str_sub(topic_names, 1, 40)
 # TODO change to pivot_longer
 vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
-                           document = factor(rownames(topic_prop_examples))), 
+                           document = factor(str_sub(rownames(topic_prop_examples), 1, 20))), 
                      variable.name = "topic", id.vars = "document")  
@@
   theme(legend.position = "none")
+
+vizDataFrame |> 
+  group_by(document) |> 
+  mutate(value_max = max(value),
+         topic_max = topic[value == value_max])
 

### 2023-07-13 analysis/paper/paper.qmd

@@
 emu <- left_join(emu_raw, emu_metadata, by = "ID")
-```
-
-```{r clean-text}
-#| include: false
-
-emu$text_clean <- emu$text_raw |>
-  clean_basic()
-
 ```
@@
 
-```{r stop-words}
+```{r pre-process-text}
 
-# TODO replace with cleaning steps including ngram concatenation
-# Add cleaned text column with removed stop words
+meta_sw <- find_meta_stopwords(emu)
+urbanism_sw <- urbanism_stopwords(
+    add_stopwords = c("emu", "european postgraduate master in urbanism", 
+                      "european post master in urbanism",
+                      "european postmaster in urbanism",
+                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning"))
+ananymisiation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
+custom_sw <- c("space", "spatial", "spaces", "plan", "public", "development") |>
+  (\(.) paste0("\\b", ., "\\b"))() |> 
+  paste(collapse = "|")
+
+# Add cleaned text column
+emu$text_clean <- emu$text_raw |>
+  # Remove white spaces and punctuation, and change text to lower-case
+  clean_basic() |>
+  # Remove title, subtitle and full title
+  str_remove_all(meta_sw) |> 
+  # Remove dataset-specific words
+  str_remove_all(urbanism_sw) |>
+  # Remove thesis-specific words
+  str_remove_all(thesis_sw) |> 
+  # Remove text included with anonymisation
+  str_remove_all(ananymisiation_sw)
+
+# Concatenate ngrams
+emu$text_clean <- emu |> 
+  c_ngrams(n_max = 3, text_col = "text_clean",
+           title_col = "title", min_freq = 20, lemma = TRUE)
+
+  # Words that were not included in ngrams are safe to remove now
 emu$text_clean <- emu$text_clean |> 
-  str_remove_all(find_meta_stopwords(emu)) |> 
-  str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
-                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
+  str_remove_all(custom_sw)
 
-# Unnest words and remove words of one and two letters
+```
+
+```{r tokenise}
+
+# TODO Replace title column with ID column
+
+# Unnest tokens and remove words of one and two letters
 emu_words <- emu |> 
@@
   anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
-  mutate(word = textstem::lemmatize_words(word))  # lemmatise words
+  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
+  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
 ```
@@
 Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
-
-```{r short-stop-words}
-#| include: false
-
-# short_words(emu_words$word, l_word = 3, l_list = 100)
-
-emu_words <- emu_words |>  
-  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
-
-```
 
@@
 
-bigram_stem_counts <- emu_bigrams |> 
-  count(w_1_stem, w_2_stem, sort = TRUE)
+bigram_tf_idf <- emu_bigrams |> 
+  count(title, ngram) |> 
+  bind_tf_idf(ngram, title, n) |> 
+  arrange(desc(tf_idf))
 
-bigram_counts
-bigram_stem_counts
+bigram_tf_idf
+```
 
-## Trigrams
+```{r trigram-tfidf}
+
 emu_trigrams <- emu |> 

### 2023-07-13 R/convert.R
changed tidyverse pipe to native pipe

@@ -24,7 +24,7 @@ convert_pdf_text <-function(pdf_filenames){
   # Load python script with functions to extract texts
   reticulate::source_python(here::here("inst/python","extract-text-pdf-python.py"))
 
-  text_list <- convert_pdf(pdf_list) %>% unlist()
+  text_list <- convert_pdf(pdf_list) |> unlist()
 
   text_list
 }

### 2023-07-13 analysis/paper/paper.qmd
Update topic model script with different Ks

--- 2023-07-13 analysis/paper/paper.qmd
@@
 # Extract beta and theta statistics from LDA model
-beta <- posterior(emu_lda)$terms
-theta <- posterior(emu_lda)$topics
+beta_5 <- posterior(emu_lda_5)$terms
+theta_5 <- posterior(emu_lda_5)$topics
 
-# Preview 10 terms from each topic
-terms(emu_lda, 10)
+beta_k <- posterior(emu_lda_k)$terms
+theta_k <- posterior(emu_lda_k)$topics
 
 # Add pseudo-names to topics based on the top 5 words in each topic
-topic_names <- name_topics(beta)
+topic_names_k <- name_topics(beta_k)
+topic_names_5 <- name_topics(beta_5)
 
@@
 # What is the dominant topic in each document?
-N = 82
-topic_prop_examples <- theta[1:N,]
-colnames(topic_prop_examples) <- str_sub(topic_names, 1, 40)
+N = 70
+topic_prop_examples <- theta_5[1:N,]
+colnames(topic_prop_examples) <- topic_names_5
 # TODO change to pivot_longer
@@
                            document = factor(str_sub(rownames(topic_prop_examples), 1, 20))), 
-                     variable.name = "topic", id.vars = "document")  
-ggplot(data = vizDataFrame, aes(topic, value), ylab = "proportion") + 
+                     variable.name = "topic", id.vars = "document") |> 
+  group_by(document) |> 
+  mutate(value_max = max(value),
+         topic_max = topic[value == value_max])
+ggplot(data = vizDataFrame, aes(topic, value, fill = topic_max), ylab = "proportion") + 
   geom_bar(stat="identity") +
@@
   coord_flip() +
-  facet_wrap(~ document, ncol = 5) +
-  theme(legend.position = "none")
-
-vizDataFrame |> 
-  group_by(document) |> 
-  mutate(value_max = max(value),
-         topic_max = topic[value == value_max])
+  facet_wrap(~ document, ncol = 10) +
+  theme(legend.position = "bottom") +
+  scale_fill_viridis_d(name = "Topic") +
+  labs(title = "Theses as distributions over topics",
+       subtitle = "Charts colored by the highest probability")
 

### 2023-07-13 analysis/paper/paper.qmd
Update preprocessing steps

@@
 
-# TODO replace with cleaning steps including ngram concatenation
-# Add cleaned text column with removed stop words
+meta_sw <- find_meta_stopwords(emu)
+emu_sw <- urbanism_stopwords(
+    add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
+                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "figure", "ning"))
+anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
+custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
+               "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
+               "create", "exist", "analys(i|e)s") |>
+  (\(.) paste0("\\b", ., "\\b"))() |> 
+  paste(collapse = "|")
+
+# Add cleaned text column
+emu$text_clean <- emu$text_raw |> 
+  # Remove text included with anonymisation
+  str_remove_all(anonymisation_sw) |>
+  # Remove white spaces and punctuation, and change text to lower-case
+  clean_basic() |>
+  # Remove title, subtitle and full title
+  str_remove_all(meta_sw) |> 
+  # Remove dataset-specific words
+  str_remove_all(emu_sw) |>
+  # Remove thesis-specific words
+  str_remove_all(thesis_sw)
+
+# Concatenate ngrams
+emu$text_clean <- emu |> 
+  c_ngrams(n_max = 3, text_col = "text_clean",
+           title_col = "title", min_freq = 20, lemma = TRUE)
+
+# Words that were not included in ngrams are safe to remove now
 emu$text_clean <- emu$text_clean |> 
-  str_remove_all(find_meta_stopwords(emu)) |> 
-  str_remove_all(urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
-                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
-  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))
+  str_remove_all(custom_sw)
 
-# Unnest words and remove words of one and two letters
+# # Get all concatenated ngrams
+# unique(unlist(str_match_all(emu$text_clean, "[A-Za-z]+(?:_[A-Za-z]+)+")[1]))
+
+# # Check if the word "space" was removed
+# unique(unlist(str_match_all(emu$text_clean, "\\bspace\\b")))
+# unique(unlist(str_match_all(emu$text_clean, "space(?:_[A-Za-z]+)+")))
+
+```
+
+```{r tokenise}
+
+# TODO Replace title column with ID column
+
+# Unnest tokens and remove words of one and two letters
 emu_words <- emu |> 
@@
   anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
-  mutate(word = textstem::lemmatize_words(word))  # lemmatise words
+  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
+  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
+
+# # Check if concatenated ngrams are retained
+# unlist(str_match_all(emu_words$word, "[A-Za-z]+(?:_[A-Za-z]+)+"))
 ```

### 2023-07-20 analysis/paper/paper.qmd

@@
 
-```{r stop-words}
+```{r pre-process-text}
 

### 2023-09-06 analysis/paper/paper.qmd

@@
   c_ngrams(n_max = 3, text_col = "text_clean",
-           title_col = "title", min_freq = 20, lemma = TRUE)
+           id_col = "ID", min_freq = 20, lemma = TRUE)
 
@@
 
-# TODO Replace title column with ID column
-
 # Unnest tokens and remove words of one and two letters
 emu_words <- emu |> 
-  select(title, text_clean) |> 
+  select(ID, text_clean) |> 
   unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
@@
 emu_dtm <- convert_to_dtm(emu_words, 
-                          title_col = "title", 
+                          id_col = "ID", 
                           word_col = "word", 
@@
                      variable.name = "topic", id.vars = "document") |> 
-    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))
+    left_join(select(emu_theses, ID, "year" = graduation_year), by = c("document" = "ID"))
 
@@
   inner_join(emu |> 
-               select(title, graduation_year), 
-             by = c("document" = "title")) |> 
+               select(ID, graduation_year), 
+             by = c("document" = "ID")) |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
   inner_join(emu |> 
-               select(title, graduation_year), 
-             by = c("document" = "title")) |> 
+               select(ID, graduation_year), 
+             by = c("document" = "ID")) |> 
   ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
@@
 emu_bigrams <- emu |> 
-  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)
+  get_ngrams(n = 2, id_col = "ID", text_col = "text_clean", stem = TRUE)
 
@@
 bigram_tf_idf <- emu_bigrams |> 
-  count(title, ngram) |> 
-  bind_tf_idf(ngram, title, n) |> 
+  count(ID, ngram) |> 
+  bind_tf_idf(ngram, ID, n) |> 
   arrange(desc(tf_idf))
@@
 emu_trigrams <- emu |> 
-  get_ngrams(n = 3, title_col = "title", text_col = "text_raw")
+  get_ngrams(n = 3, id_col = "ID", text_col = "text_raw")
 
@@
 trigram_tf_idf <- emu_trigrams |> 
-  count(title, ngram) |> 
-  bind_tf_idf(ngram, title, n) |> 
+  count(ID, ngram) |> 
+  bind_tf_idf(ngram, ID, n) |> 
   arrange(desc(tf_idf))
@@
 emu_bigrams_top <- emu_bigrams |>
-  group_by(title) |>
+  group_by(ID) |>
   count(ngram, sort = TRUE) |>
@@
 emu_bigrams_top |>
-  filter(title == unique(title)[1:10]) |> 
-  ggplot(aes(x = ngram, y = n, fill = title)) +
+  filter(ID == unique(ID)[1:10]) |> 
+  ggplot(aes(x = ngram, y = n, fill = ID)) +
   geom_col() +
@@
        subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ title, scales = "free", ncol = 2) +
+  facet_wrap( ~ ID, scales = "free", ncol = 2) +
   theme(legend.position = "none") 
@@
 emu_tf_idf <- emu_words |> 
-  get_tf_idf(title_col = "title", word_col = "word")
+  get_tf_idf(id_col = "ID", word_col = "word")
 
@@
 emu_tf_idf |> 
-  group_by(title) |> 
+  group_by(ID) |> 
   top_n(5, tf_idf) |> 
@@
 emu_words_count <- emu_words |> 
-  count(title, word, sort = TRUE) |> 
+  count(ID, word, sort = TRUE) |> 
   ungroup()
@@
 emu_words_total <- emu_words_count |> 
-  group_by(title) |> 
+  group_by(ID) |> 
   summarise(total = sum(n))
@@
 ggplot(emu_words_count |> 
-         filter(title == unique(title)[1:6]),
-       aes(n/total, fill = title)) +
+         filter(ID == unique(ID)[1:6]),
+       aes(n/total, fill = ID)) +
   geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
   xlim(NA, 0.0009) +
-  facet_wrap(~ title, ncol = 3, scales = "free_y")
+  facet_wrap(~ ID, ncol = 3, scales = "free_y")
 ```
@@
 freq_by_rank <- emu_words_count |> 
-  group_by(title) |> 
+  group_by(ID) |> 
   mutate(rank = row_number(),
@@
 freq_by_rank |> 
-  ggplot(aes(rank, `term frequency`, color = title)) +
+  ggplot(aes(rank, `term frequency`, color = ID)) +
   geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
@@
 emu_words_count |> 
-  get_tf_idf(title_col = "title", word_col = "word") |> 
-  filter(title == unique(title)[1:2]) |>  # Keep only the first 6 theses
+  get_tf_idf(id_col = "ID", word_col = "word") |> 
+  filter(ID == unique(ID)[1:2]) |>  # Keep only the first 6 theses
   arrange(desc(tf_idf)) |> 
   mutate(word = factor(word, levels = rev(unique(word)))) |> 
-  group_by(title) |> 
+  group_by(ID) |> 
   top_n(15) |> 
   ungroup() |> 
-  ggplot(aes(word, tf_idf, fill = title)) +
+  ggplot(aes(word, tf_idf, fill = ID)) +
   geom_col(show.legend = FALSE) +
   labs(x = NULL, y = "tf-idf") +
-  facet_wrap(~ title,  ncol = 3, scales = "free") +
+  facet_wrap(~ ID,  ncol = 3, scales = "free") +
   coord_flip()
@@
 # To what extent do words co-occur with other words?
-tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")
+tidy_pmi <- get_pmi(emu, id_col = "ID", text_col = "text_clean")
 ```
@@
 tidy_words <- emu |> 
-  select(title, text_clean) |> 
+  select(ID, text_clean) |> 
   unnest_tokens(word, text_clean) |> 
@@
 word_matrix <- tidy_words |> 
-  count(title, word) |> 
-  cast_sparse(title, word, n)
+  count(ID, word) |> 
+  cast_sparse(ID, word, n)
 

### 2023-09-06 analysis/paper/paper.qmd
Update code chunk options
Some code chunks do not run and prevent rendering. I temporarily made them inactive with eval: false.

--- 2023-07-20 analysis/paper/paper.qmd
@@
 #| fig-cap: Text mining process
+#| eval: false
 
@@
 ```{r topics-per-corpus}
+#| eval: false
 
@@
 ```{r}
+#| eval: false
 
@@
 ```{r lda-vis}
+#| eval: false
 
@@
 # Extract topics as mixtures of terms from LDA model
-emu_topics <- tidy(emu_lda, matrix = "beta")
+emu_topics <- tidy(emu_lda_5, matrix = "beta")
 
@@
 ```{r topic-model-gamma}
+#| eval: false
 
 # Extract documents as mixtures of topics from LDA model
-emu_documents <- tidy(emu_lda, matrix = "gamma")
+emu_documents <- tidy(emu_lda_5, matrix = "gamma")
 
@@
 emu_documents |>
+  mutate(document = as.numeric(document)) |> 
   inner_join(emu |> 
-               select(title, graduation_year), 
-             by = c("document" = "title")) |> 
+               select(ID, graduation_year), 
+             by = c("document" = "ID")) |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 ```{r top-words-document}
+#| eval: false
 
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_words, top_n = 10)
+get_top_words_per_document(data = emu_words, title_col = "ID", top_n = 10)
 ```
@@
 emu_tf_idf <- emu_words |> 
-  get_tf_idf(title_col = "title", word_col = "word")
+  get_tf_idf(title_col = "ID", word_col = "word")
 
@@
 emu_words_count |> 
-  get_tf_idf(title_col = "title", word_col = "word") |> 
-  filter(title == unique(title)[1:2]) |>  # Keep only the first 6 theses
+  get_tf_idf(title_col = "ID", word_col = "word") |> 
+  filter(ID == unique(ID)[1:2]) |>  # Keep only the first 6 theses
   arrange(desc(tf_idf)) |> 
   mutate(word = factor(word, levels = rev(unique(word)))) |> 
-  group_by(title) |> 
+  group_by(ID) |> 
   top_n(15) |> 
   ungroup() |> 
-  ggplot(aes(word, tf_idf, fill = title)) +
+  ggplot(aes(word, tf_idf, fill = ID)) +
   geom_col(show.legend = FALSE) +
   labs(x = NULL, y = "tf-idf") +
-  facet_wrap(~ title,  ncol = 3, scales = "free") +
+  facet_wrap(~ ID,  ncol = 3, scales = "free") +
   coord_flip()
@@
 
-\[Describe word embeddings ...\]
+Describe word embeddings ...
 
@@
 # To what extent do words co-occur with other words?
-tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")
+tidy_pmi <- get_pmi(emu, title_col = "ID", text_col = "text_clean")
 ```
@@
 
-\[What terms should be checked with the function below?\]
+What terms should be checked with the function below?
 
@@
 
-The first \[?\] components... \[what do they show?\]
+The first # components... what do they show?
 
@@
 
--   Although the focus of this study is on describing a fixed corpus though unsupervised machine learning, it can provide input into supervised predictive models.
+Although the focus of this study is on describing a fixed corpus though unsupervised machine learning, it can provide input into supervised predictive models.
 

### 2023-09-06 R/analyse.R
Fix bug with title_col use in tidy_pmi function

@@ -265,7 +265,7 @@ get_pmi <- function(data,
   tidy_pmi <- nested_words |>
     dplyr::mutate(words = furrr::future_map(words, slide_windows, window_size)) |>
     tidyr::unnest(words) |>
-    tidyr::unite(window_id, title, window_id) |>
+    tidyr::unite(window_id, !!rlang::sym(title_col), window_id) |>
     widyr::pairwise_pmi(word, window_id)
 
   tidy_pmi

### 2023-09-06 analysis/paper/paper.qmd

--- 2023-09-06 analysis/paper/paper.qmd
@@
 emu_documents |>
+  mutate(document = as.numeric(document)) |> 
   inner_join(emu |> 

### 2023-09-06 R/app_server.R
removed superfluous code

@@ -54,22 +54,6 @@ app_server <- function(session,input, output) {
    })
  # Render plots
 
-  output$distPlot <- shiny::renderPlot({
-    # generate bins based on input$bins from ui.R
-    x    <- faithful[, 2]
-    bins <- seq(min(x), max(x), length.out = 30)
-
-    # draw the histogram with the specified number of bins
-    hist(
-      x,
-      breaks = bins,
-      col = 'darkgray',
-      border = 'white',
-      xlab = 'Waiting time to next eruption (in mins)',
-      main = 'Histogram of waiting times'
-    )
-  })
-
   # ID of value clicked on the map
 
   clicked_id <- shiny::reactive({input$thesis_location_map_marker_click$id}) |>
@@ -168,5 +152,8 @@ app_server <- function(session,input, output) {
 
    }) |>
      shiny::bindEvent(input$thesis_location_map_marker_click)
+
+
+
 
   }

### 2023-09-06 R/analyse.R
Replace title_col with id_col

@@ -265,7 +265,7 @@ get_pmi <- function(data,
   tidy_pmi <- nested_words |>
     dplyr::mutate(words = furrr::future_map(words, slide_windows, window_size)) |>
     tidyr::unnest(words) |>
-    tidyr::unite(window_id, !!rlang::sym(title_col), window_id) |>
+    tidyr::unite(window_id, !!rlang::sym(id_col), window_id) |>
     widyr::pairwise_pmi(word, window_id)
 
   tidy_pmi

### 2023-09-12 analysis/paper/paper.qmd
Replace title_col with id_col
Update cleaning in data preparation script and package data

@@
 
+# Load cleaned text data
+data(emu_clean)
+
 # Join text with the metadata file  
-emu <- left_join(emu_raw, emu_metadata, by = "ID")
+emu <- emu_raw |> 
+  left_join(emu_metadata, by = "ID") |> 
+  left_join(emu_clean, by = "ID")
 ```
@@
 Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.
-
-```{r pre-process-text}
-
-meta_sw <- find_meta_stopwords(emu)
-emu_sw <- urbanism_stopwords(
-    add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
-                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
-thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "figure", "ning"))
-anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
-  paste(collapse = "|")
-custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
-               "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
-               "create", "exist", "analys(i|e)s") |>
-  (\(.) paste0("\\b", ., "\\b"))() |> 
-  paste(collapse = "|")
-
-# Add cleaned text column
-emu$text_clean <- emu$text_raw |> 
-  # Remove text included with anonymisation
-  str_remove_all(anonymisation_sw) |>
-  # Remove white spaces and punctuation, and change text to lower-case
-  clean_basic() |>
-  # Remove title, subtitle and full title
-  str_remove_all(meta_sw) |> 
-  # Remove dataset-specific words
-  str_remove_all(emu_sw) |>
-  # Remove thesis-specific words
-  str_remove_all(thesis_sw)
-
-# Concatenate ngrams
-emu$text_clean <- emu |> 
-  c_ngrams(n_max = 3, text_col = "text_clean",
-           id_col = "ID", min_freq = 20, lemma = TRUE)
-
-# Words that were not included in ngrams are safe to remove now
-emu$text_clean <- emu$text_clean |> 
-  str_remove_all(custom_sw)
-
-# # Get all concatenated ngrams
-# unique(unlist(str_match_all(emu$text_clean, "[A-Za-z]+(?:_[A-Za-z]+)+")[1]))
-
-# # Check if the word "space" was removed
-# unique(unlist(str_match_all(emu$text_clean, "\\bspace\\b")))
-# unique(unlist(str_match_all(emu$text_clean, "space(?:_[A-Za-z]+)+")))
-
-```
 
@@
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_words, title_col = "ID", top_n = 10)
+get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 10)
 ```

### 2023-09-12 data-raw/emu_text.R

@@ -14,20 +14,52 @@ emu_metadata <- readr::read_csv(emu_meta_path)
 usethis::use_data(emu_metadata, overwrite = TRUE)
 
 # Clean the dataset ----
-# TODO adapt cleaning procedure once all the steps are known (#105)
 
-emu_theses <- dplyr::left_join(emu_raw, emu_metadata, by  = "ID")
+emu <- dplyr::left_join(emu_raw, emu_metadata, by  = "ID")
 
-emu_theses$text_clean <- emu_theses$text_raw |>
-  mintEMU::clean_basic() |>
-  stringr::str_remove_all(mintEMU::find_meta_stopwords(emu_theses)) |>
-  stringr::str_remove_all(mintEMU::urbanism_stopwords(
-    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
-  stringr::str_remove_all(mintEMU::thesis_stopwords(
-    add_stopwords = c("advisor", "prof")))
+meta_sw <- find_meta_stopwords(emu)
+emu_sw <- urbanism_stopwords(
+  add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
+                    "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "figure", "ning"))
+anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
+custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
+               "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
+               "create", "exist", "analys(i|e)s") |>
+  (\(.) paste0("\\b", ., "\\b"))() |>
+  paste(collapse = "|")
 
-emu_clean <- emu_theses |>
+# Add cleaned text column
+emu$text_clean <- emu$text_raw |>
+  # Remove text included with anonymisation
+  str_remove_all(anonymisation_sw) |>
+  # Remove white spaces and punctuation, and change text to lower-case
+  clean_basic() |>
+  # Remove title, subtitle and full title
+  str_remove_all(meta_sw) |>
[…]

### 2023-09-13 analysis/paper/paper.qmd
Update topic model variable names

--- 2023-09-06 analysis/paper/paper.qmd
@@
 # What are the most probable topics in the entire collection?
-topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
-names(topicProportions) <- topic_names     # assign the topic names we created before
+topicProportions <- colSums(theta_5) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
+names(topicProportions) <- topic_names_5     # assign the topic names we created before
 sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
@@
 countsOfPrimaryTopics <- rep(0, K)
-names(countsOfPrimaryTopics) <- topic_names
+names(countsOfPrimaryTopics) <- topic_names_5
 for (i in 1:nDocs(emu_dtm)) {
-  topicsPerDoc <- theta[i, ] # select topic distribution for document i
+  topicsPerDoc <- theta_5[i, ] # select topic distribution for document i
   # get first element position from ordered list
@@
 # ... or have it selected by a term in the topic name (e.g. 'children')
-topicToFilter <- grep('river', topic_names)[1] 
+topicToFilter <- grep('river', topic_names_5)[1] 
 topicThreshold <- 0.2
-selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
-filteredCorpus <- emu_theses[selectedDocumentIndexes,]
+selectedDocumentIndexes <- which(theta_5[, topicToFilter] >= topicThreshold)
+filteredCorpus <- emu[selectedDocumentIndexes,]
 # show length of filtered corpus
@@
 N = 82
-topic_prop_per_year <- theta[1:N,]
-colnames(topic_prop_per_year) <- topic_names
+topic_prop_per_year <- theta_5[1:N,]
+colnames(topic_prop_per_year) <- topic_names_5
 
@@
                      variable.name = "topic", id.vars = "document") |> 
-    left_join(select(emu_theses, ID, "year" = graduation_year), by = c("document" = "ID"))
+    left_join(select(emu, ID, "year" = graduation_year), by = c("document" = "ID"))
 
@@
 json <- LDAvis::createJSON(
-  phi = beta, 
-  theta = theta, 
+  phi = beta_5, 
+  theta = theta_5, 
   doc.length = rowSums(as.matrix(emu_dtm)), 

### 2023-09-13 analysis/paper/paper.qmd
Update c_ngrams check
Update facet names in plot of theses topics

@@
 # # Check if concatenated ngrams are retained
-# unlist(str_match_all(emu_words$word, "[A-Za-z]+(?:_[A-Za-z]+)+"))
+# sort(table(unlist(str_match_all(emu_words$word, "[A-Za-z]+(?:_[A-Za-z]+)+"))), decreasing = TRUE)
 ```
@@
   coord_flip() +
-  facet_wrap(~ document, ncol = 10) +
+  facet_wrap(~ as.numeric(document), ncol = 10) +
   theme(legend.position = "bottom") +

### 2023-09-13 analysis/supplementary-materials/1-merge-metadata-files.R
Remove theses for which permission was not granted

@@ -61,6 +61,7 @@ theses_meta %>%
   arrange(last_name) %>%
   select(-title, -uuid, -count_na,
          -contains("last_name_split"), -author,
+         -contains("permission_re"), -contains("granted_2"),
          -`publication year`,
          -`publication type`,
          -publisher,

### 2023-09-13 analysis/supplementary-materials/3-prepare-for-publication.R

@@ -100,6 +100,7 @@ emu_theses$text_raw  <- text_raw
 
 # Make selection of columns ----------------
 emu_theses <- emu_theses |>
+  filter(as.logical(permission_granted) == TRUE) |>
   select(ID, graduation_year,
          graduation_semester,
          title, full_title, subtitle, link,

### 2023-10-29 data-raw/emu_text.R
Update namespaces and cleaning procedure in emu_text

@@ -7,21 +7,21 @@ usethis::use_data(emu_raw, overwrite = TRUE)
 
 # emu metadata file
 emu_meta_path <- here::here("analysis", "data", "derived_data", "emu_metadata.csv")
-
 emu_metadata <- readr::read_csv(emu_meta_path)
 
 # add metadata to the package
 usethis::use_data(emu_metadata, overwrite = TRUE)
 
 # Clean the dataset ----
-
 emu <- dplyr::left_join(emu_raw, emu_metadata, by  = "ID")
 
-meta_sw <- mintEMU::find_meta_stopwords(emu)
+# Prepare lists of stop words
+meta_sw <- mintEMU::find_meta_stopwords(emu, clean_meta = FALSE)
 emu_sw <- mintEMU::urbanism_stopwords(
   add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
                     "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
-thesis_sw <- mintEMU::thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "figure", "ning"))
+thesis_sw <- mintEMU::thesis_stopwords(
+  add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
 anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
   paste(collapse = "|")
 custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
@@ -30,18 +30,36 @@ custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "deve
   (\(.) paste0("\\b", ., "\\b"))() |>
   paste(collapse = "|")
 
+# # Add cleaned text column
+# emu$text_clean <- emu$text_raw |>
+#   # Remove hyphenation
+#   stringr::str_replace_all("-\\n", "") |>
+#   # Remove text included with anonymisation
+#   stringr::str_remove_all(anonymisation_sw) |>
+#   # Remove white spaces and punctuation, and change text to lower-case
+#   mintEMU::clean_basic() |>
+#   # Remove title, subtitle and full title
+#   stringr::str_remove_all(meta_sw) |>
+#   # Remove dataset-specific words
[…]

### 2023-10-29 analysis/supplementary-materials/1-prepare-metadata.R
Add table with total theses per year

@@ -91,4 +91,7 @@ theses_meta %>%
   separate(full_title, c("title", "subtitle"), sep = ":", remove = FALSE) %>%
   write_csv(file = here("analysis", "data", "raw_data", "theses-metadata.csv"))
 
-
+## Total theses per year
+readxl::read_xlsx(path = here("analysis", "data", "raw_data", "theses-all.xlsx")) |>
+  count(grad_year) |>
+  write_csv(file = here("analysis", "data", "derived_data", "theses-per-year.csv"))

### 2023-10-29 analysis/supplementary-materials/create-rrtools.R
Update analysis/supplementary-materials/create-rrtools.R

@@ -6,5 +6,5 @@ renv::init()
 rrtools::add_dependencies_to_description()
 devtools::document()
 
-# allows using magrittr pipe in the pacakge:
+# allows using magrittr pipe in the package:
 usethis::use_pipe()

### 2023-10-29 analysis/paper/paper.qmd
Add note about evaluation through topic coherence

@@
 abstract: |
-  Prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft,
+  Prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021,
   this paper aims to describe the legacy of the EMU. To that end, it reports on the
@@
 
+# TODO remove this section after all analysis has been translated into functions
 # Load packages ----
@@
 packages <- c(
-  "here",                                       # Managing paths
-  # TODO remove reshape2 if `melt()` is replaced by`pivot_longer()`
-  "tidyverse", "reshape2", "widyr", "readxl",   # Data manipulation and visualisation
+  "here",
+  "tidyverse", "widyr", "readxl",   # Data manipulation and visualisation
   "furrr",                          # Parallel use of purrr::map() functions
@@
 
+# TODO should this be replaced with a call to `library(mintEMU)`?
 # Load analysis functions ----
@@
 
-# Load thesis metadata
-data(emu_metadata)
+# TODO merge into setup code chunk
 
-# Load raw text data
-data(emu_raw)
+# Load data for analysis
+data(emu_metadata)  # Thesis metadata
+data(emu_raw)       # Raw text data
+data(emu_clean)     # Cleaned text data
 
-# Load cleaned text data
-data(emu_clean)
+# Join metadata and cleaned text data
+emu <- emu_metadata |> 
+  left_join(emu_raw, by = "ID") |> 
+  left_join(emu_clean, by = "ID")
 
-# Join text with the metadata file  
-emu <- emu_raw |> 
-  left_join(emu_metadata, by = "ID") |> 
-  left_join(emu_clean, by = "ID")
+theses_all_per_year <- read_csv(file = here("analysis", "data", "derived_data", "theses-per-year.csv"))
+
+# # Prepare data for analysis
+# ## Create lists of stop words
+# thesis_sw <- thesis_stopwords(
+#   add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
+# custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
+#                "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
+#                "create", "exist", "analys(i|e)s") |>
+#   (\(.) paste0("\\b", ., "\\b"))() |>
+#   paste(collapse = "|")
+# emu_sw <- urbanism_stopwords(
+#   add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
+#                     "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
+# 
+# emu$text_clean <- emu$text_clean |> 
+#   # Remove white spaces and punctuation, and change text to lower-case
+#   clean_basic() |>
+#   # Remove thesis-specific words
+#   str_remove_all(thesis_sw) |> 
+#   # Remove dataset-specific words
+#   str_remove_all(emu_sw)
+# 
+# # Concatenate ngrams
+# emu$text_clean <- emu |>
+#   mintEMU::c_ngrams(n_max = 3, text_col = "text_clean",
+#            id_col = "ID", min_freq = 20, lemma = TRUE)
+# 
+# # Words that were not included in ngrams are safe to remove now
+# emu$text_clean <- emu$text_clean |>
+#   # Words that were not included in ngrams are safe to remove now
+#   str_remove_all(custom_sw)
+
 ```
@@
 The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft.
-The EMU was an advanced master ran jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia.
-<!-- Complete paragraph about the program -->
+The EMU is an advanced master started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005.
+The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and experience in practice.
+With a combined research and design program, it provided a bridge towards advanced practice as well as research trajectories after graduation.
+All students followed an exchange semester at one of the other three program universities, allowing them to have a broader view on the field of urbanism.
+Although located in Europe and named "European", its geographic scope is global, especially at the level of graduation, considering that a large majority of students were of non-European origin.
+Influenced by the strong urbanism tradition of the four universities involved, the program yielded a group of almost 100 graduates.
 
-Distinctive features of the program:\
-- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice - provides a bridge towards PhD-level research - all students had an exchange semester at one of the other three program universities
+In 2021, TU Delft stepped out the program, but its approach has remained imprinted on the TU Delft approach to urbanism. <!--# paragraph to be elaborated -->
 
 In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
-To that end, we employed a text mining approach in which we analysed its output: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE), "-", max(emu$graduation_year, na.rm = TRUE))`.
+To that end, the research presented in this paper employed a text mining approach to analyse the output of the EMU: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.
 
@@
 
-This paper leverages topic modeling as an unsupervised machine learning method to reveal latent semantic structure and relationships between the latent topics in a large document collection [@muchene2021].
+**The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012].
+At the same time, urbanism research in general and urbanism education studies in particular that use statistical methods for unstructured data, such as topic modeling, are scarce.
+This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then.
+Patterns revealed in past education can potentially better inform future decisions in education in the field.
+Although the corpus analysed in this study is relatively small, it is large enough to make a quantitative analytical approach untenable.
 
-In Natural Language Processing (NLP) ... and the focus on its subtopic topic modeling
+Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of NLP, as an unsupervised machine learning method to reveal latent the semantic structure and relationships between latent topics in a large document collection [@muchene2021].
+Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a *soft clustering algorithm* in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a].
+LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposition into two low-dimensional matrices for doc-topic distribution and word-topic distribution, respectively.
+Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a].
+Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022]
 
-Topic modeling provides a way to generative
-
--   bag of words
-
--   decomposition into two low-dimensional matrices for doc-topic distribution and word-topic distribution, respectively
-
--   LDA and extensions of it (CTM, HDP, GLDA, DTM) are broadly adopted in topic modeling [@muchene2021]
-
-<!-- Describe the research gaps -->
-
-Research gaps and opportunities:
-
--   The scarcity of studies that use statistical methods for unstructured data in urbanism education
-
--   Theses are representative outputs of educational programs that can be used as data for text mining.
-The growing availability of theses in digital repositories presents an opportunity to that end.
-The TU Delft repository was established in 2007, in the year when the EMU program started.
-
--   Applications of topic modeling in the field of urbanism
-
--   Patterns revealed in past education can potentially better inform future decisions in education in the field
-
--   The EMU program...
-
-Challenges:
-
-Contribution:
-
--   This paper showcases reproducible research workflow.
+In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**.
 It is part of a reproducible research compendium and accompanied by a published dataset.
@@
 -   Give an overview of th predominant topics and their evolution over the years of the EMU program.
-In addition to the natural document classification <!--# explain --> of documents provided by a repository, LDA allows documents to be part of multiple topics [@muchene2021].
-
-<!--# Outline -->
 
 The following section describes the case study data, pre-processing steps and the topic modeling workflow.
+Then results of the analysis are summarised in...
+The paper concludes with a discussion on what the lessons learned from the EMU program mean for urbanism education and practice in general as well as on the broader challenges and opportunities of the text mining approach taken in the paper.
 
@@
 
-Research questions:
-
--   What were the main topics addressed in the EMU program?
-
--   How did those topics evolve throughout the duration of the program between 2007-2021?
-
--   To what extent were the topics and evolution thereof influenced by the assignments given to the students throughout their studies leading up to their thesis?
-
--   To what extent were the topics and evolution thereof influenced by the exchange semester followed by the students?
-
-    -   Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
-Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
-
-The research process (@fig-process-diagram)...
+To describe the legacy of the EMU program this research answers two questions: What were the main topics addressed in the EMU program?
+And how did those topics evolve throughout the duration of the program between 2007-2021?
+The paper then discusses to what extent those topics and evolution thereof were influenced by the assignments given to the students throughout their studies leading up to their thesis and by the exchange semester followed by the students.
+The research process is described in @fig-process-diagram.
 
@@
 
+1.
+
+2.  Pre-processing ...
+
+3.  Information extraction ... topic modeling:
+
+3.1. tokenisation with a window of 1-3 words
+
+3.2. DTM, discarding terms with a frequency under a specified threshold
+
+3.3. Assign a K value manually
+
+3.4. Calculate topic coherence for K values ranging from 2 to K
+
+3.5. Plot K values agianst coherence scores and select K with the highest coherence value
+
+3.6. Fit model with K determined in previous step
+
+3.7. Get top 20 words in each topic to understand what it is about
+
+3.8. Run a cluster dendrogram (Hellinger distance) to see which topics should be grouped together
+
+3.9. Visualise topics with word clouds
+
+OR Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
+Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
+
+4.  Visualisation ...
+
+5.  Results ...
+
 ## Data collection
 
-Of the total of 96 theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->.
-Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
-The full-text documents were obtained either from the TU Delft Education Repository (an API for downloading a large set of full-text data is not available), when available, or from the authors via email.
-Even though repository entries are open access, it was unclear if the content of the theses can be mined.
+The dataset consists of full texts and metadata of a subset of EMU theses written in English.
+The full texts are available as PDF documents with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
+The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository.
+In addition, metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
+Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published.
 Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
-Only PDFs for which permission was obtained were included in the corpus.With the exception of a bilingual English-Spanish thesis, all thesis were written in English.
-The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
-@tbl-data gives an overview of the data included in the analysis.
-
-Description of text extraction; raw data available on 4TU.ResearchData at...
+As a result, of the total of `r sum(theses_all_per_year$n)` theses presented in the EMU program, full-text documents were obtained for 83 of which permissions were granted for `r nrow(emu)`.
+Only PDFs for which permission was obtained were included in the corpus.
+The final corpus contains at least one thesis for each year of graduation between 2007-2021 as shown in @tbl-data.
 
@@
 
+emu |> 
+  count(graduation_year) |> 
+  left_join(theses_all_per_year, by = c("graduation_year" = "grad_year")) |> 
+  mutate(graduation_year = as.factor(graduation_year),
+         `% of all theses` = scales::label_percent()(round(n.x / n.y, 4))) |>
+  bind_rows(data.frame(graduation_year = "", 
+                       n.x = nrow(emu), 
+                       n.y = sum(theses_all_per_year$n),
+                       `% of all theses` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n)),
+                       check.names = FALSE)) |> 
+  rename(`Graduation year` = graduation_year,
+         `No. theses included` = n.x) |>
+  select(-n.y) |>
+  kableExtra::kable(align = "r")
 ```
 
-The imported dataset contains the text data in the `text` column and related metadata in the other columns...
+## Pre-processing
 
-For each thesis, henceforth referred to as "document", the metadata consisted of the `ID`, `title`, `graduation_year`, `graduation_semester`, `exchange_semester`, `location`, `latitude`, `longitude`, `abstract`, `text_raw` and `text_clean` <!--# check column names or add them in an inline code chunk with epoxy -->.
+The text was extracted from the original theses with . The text was
 
-Both the raw and cleaned data are published in the 4TU.ResearchData repository [ADD CITATION].
+Description of text extraction;
+
+The raw data available on 4TU.ResearchData at...The imported dataset contains the text data in the `text` column and related metadata in the other columns...
+
+For each thesis, henceforth referred to as "document", the following cleaning steps were performed:
+
+1.
+Text was extracted from the original PDF documents with the the PyMuPDF Python package.
+During extraction, the front matter (cover page, colophon, contents, acknowledgements) and back matter (references) were removed.
+The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction.
+2.  The extracted text was cleaned of line breaks, hyphenation, occurrences of the thesis title, author name, and page numbers.
+
+Only relevant metadata was kept
+
+The resulting dataset consisted of the `ID`, `title`, `graduation_year`, `graduation_semester`, `exchange_semester`, `location`, `latitude`, `longitude`, `abstract`, `text_raw` and `text_clean` <!--# check column names or add them in an inline code chunk with epoxy -->.
+
+Both the raw and cleaned data are published in the 4TU.ResearchData repository \[ADD CITATION\].
 
@@
 
-## Pre-processing
+## Information extraction
+
+### Tokenisation
 
@@
 
-Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified during preliminarey EDA were concatenated by replacing white spaces with underscores.
+Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified during preliminary EDA were concatenated by replacing white spaces and hyphens with underscores.
+This way, they could be included alongside words in the topic model.
 
@@
 
+```{r eval=FALSE}
+ent1 <- spacyr::entity_extract(spacyr::spacy_parse(emu$text_clean[1:13]))
+
+ent1 |> 
+  filter(entity_type == "LOC") |> 
+  pull(entity) |> 
+  unique()
+
+ent2 <- ent1 |> 
+  mutate(entity = tolower(entity)) |> 
+  group_by(entity) |> 
+  mutate(bign = length(unique(doc_id))) |> 
+  ungroup() |> 
+  count(doc_id, entity, bign) |> 
+  filter(bign > 3, nchar(entity) > 4)
+
+dtm <- tidytext::cast_sparse(data = ent2,
+                             row = doc_id,
+                             column = entity,
+                             value = n)
+
+lda <- text2vec::LDA$new(n_topics = 50) 
+fit <- lda$fit_transform(dtm, progressbar = F)
+tm_summary <- # incomplete code
+
+```
+
+```{r eval=FALSE}
+# Find locations
+entity::location_entity(head_text(emu$text_clean[15], sub_end = 50000))
+```
+
 ```{r tokenise}
+
+# TODO create tokenizer function
 
@@
 
-## Information extraction
-
 ### Topic model
 
-Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a group of documents [@silge2017].
+Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017].
 We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
@@
 -   The number of topics was qualitatively determined
+-   Evaluating the number of topics quantitatively requires can be done through the measure of **topic coherence**
 
-LDA requires a document-term matrix as input.
+LDA requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
 
-```{r topic-model-lda-1}
+```{r topic-model-dtm}
+
+# TODO should tf or tf-idf be used for weighting?
 
 # Prepare document-term matrix as input for LDA model
-# Reduce the size of the DTM by excluding words with a frequency lower than 5
 emu_dtm <- convert_to_dtm(emu_words, 
@@
                           min_term_freq = 5)
+```
 
+```{r topic-model-lda-1}
 # Fit LDA model with K clusters
@@
 k = get_number_of_topics(emu_dtm)
-# TODO run the model with the empirically determined k and compare (or combine) with results using qualitative value
+# TODO run the model with the empirically determined k and compare with results using qualitative value
 emu_lda_k <- LDA(emu_dtm, k = k, method = "Gibbs",
                control = list(seed = 2023, iter = 500, verbose = 25))
+
+beta_k <- posterior(emu_lda_k)$terms
+theta_k <- posterior(emu_lda_k)$topics
+
+# Add pseudo-names to topics based on the top 5 words in each topic
+topic_names_k <- name_topics(beta_k)
 
@@
 
-beta_k <- posterior(emu_lda_k)$terms
-theta_k <- posterior(emu_lda_k)$topics
-
 # Add pseudo-names to topics based on the top 5 words in each topic
-topic_names_k <- name_topics(beta_k)
 topic_names_5 <- name_topics(beta_5)
@@
 # What is the dominant topic in each document?
-N = 70
+N = nrow(theta_5)
 topic_prop_examples <- theta_5[1:N,]
 colnames(topic_prop_examples) <- topic_names_5
-# TODO change to pivot_longer
-vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
+
+vizDataFrame <- pivot_longer(
+  data = data.frame(topic_prop_examples, 
                     document = factor(str_sub(rownames(topic_prop_examples), 1, 20))),
-                     variable.name = "topic", id.vars = "document") |> 
+  cols = -document,
+  names_to = "topic",
+  values_to = "value") |> 
   group_by(document) |> 
@@
 ```{r topics-per-corpus}
-#| eval: false
 
 # What are the most probable topics in the entire collection?
-topicProportions <- colSums(theta_5) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
-names(topicProportions) <- topic_names_5     # assign the topic names we created before
-sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order
-
-soP <- sort(topicProportions, decreasing = TRUE)
-paste(round(soP, 5), ":", names(soP))
+topic_proportions <- get_topic_proportions(emu_lda_5)
+# paste0(round(topic_proportions$proportion, 2), ": ", topic_proportions$topic)
 
 # How many documents are with the highest probability in each topic?
-K = 5
-countsOfPrimaryTopics <- rep(0, K)
-names(countsOfPrimaryTopics) <- topic_names_5
-for (i in 1:nDocs(emu_dtm)) {
-  topicsPerDoc <- theta_5[i, ] # select topic distribution for document i
-  # get first element position from ordered list
-  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
-  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
-}
-sort(countsOfPrimaryTopics, decreasing = TRUE)
+counts_of_primary_topics <- count_docs_per_topic(emu_lda_5)
+# paste0(counts_of_primary_topics$count, ": ", topic_proportions$topic)
 
-so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
-paste(so, ":", names(so))
+topic_proportions |> 
+  left_join(counts_of_primary_topics, by = "topic")
 ```
@@
 
-topicToFilter <- 6  # you can set this manually ...
-# ... or have it selected by a term in the topic name (e.g. 'children')
-topicToFilter <- grep('river', topic_names_5)[1] 
-topicThreshold <- 0.2
-selectedDocumentIndexes <- which(theta_5[, topicToFilter] >= topicThreshold)
-filteredCorpus <- emu[selectedDocumentIndexes,]
+topic_to_filter <- 1  # set manually
+# or have it selected by a term in the topic name (e.g. 'airport')
+topic_to_filter <- grep('water', topic_names_5)[1] 
+topic_threshold <- 0.2
+selected_document_indices <- which(theta_5[, topic_to_filter] >= topic_threshold)
+filtered_corpus <- emu[selected_document_indices,]
 # show length of filtered corpus
-filteredCorpus
+nrow(filtered_corpus)
 ```
@@
 
-N = 82
+N = nrow(emu)
 topic_prop_per_year <- theta_5[1:N,]
@@
 
-# TODO change to pivot_longer
-vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
-                           document = factor(rownames(topic_prop_per_year))), 
-                     variable.name = "topic", id.vars = "document") |> 
-    left_join(select(emu, ID, "year" = graduation_year), by = c("document" = "ID"))
+viz_df_with_year <-
+  pivot_longer(
+    data = data.frame(topic_prop_examples,
+                      document = factor(str_sub(
+                        rownames(topic_prop_examples), 1, 20))),
+    cols = -document,
+    names_to = "topic",
+    values_to = "value") |> 
+  left_join(mutate(emu, ID = as.factor(ID), "year" = graduation_year), by = c("document" = "ID"))
 
+# TODO change to discrete colors
+# TODO try representing this as a line chart
 # Plot topic proportions per year
-vizDataFrameWithYear |> 
+viz_df_with_year |> 
   ggplot(aes(x = factor(year), y = value, fill = topic)) + 
@@
 # TODO check if this can be included in the Shiny dashboard
+# TODO remove from paper
+# TODO how to show distances between topics in the paper?
 # Function to approximate the distance between topics
@@
 ```{r topic-model-beta}
-
-# TODO Explore including concatenated collocations in the topic model
 
@@
 
+# TODO check if this is consistent with the results above
+# TODO is gamma the same as theta ?!
 emu_documents |>
@@
 
+# TODO check if this plot is useful and, if not, remove it
 emu_documents |> 
@@
 # Which are the top n most frequently used words in all theses?
-# TODO add `topic` to `fill` aesthetic
-n = 50
+terms <- as.data.frame(posterior(emu_lda_5)$terms)
+rownames(terms) <- topic_names_5
+terms <- terms |> 
+  mutate(topic = rownames(terms)) |> 
+  pivot_longer(-topic,
+               names_to = "term",
+               values_to = "prob") |> 
+  group_by(term) |> 
+  mutate(max_topic = topic[which.max(prob)]) |> 
+  filter(topic == max_topic) |> 
+  ungroup()
+
+emu_words_topics <- emu_words |> 
+  left_join(terms, by = c("word" = "term"))
+
+n = 20
 get_top_words_per_corpus(data = emu_words, top_n = n) |>
+  left_join(terms, by = c("word" = "term")) |> 
   ggplot(aes(reorder(word, n), n)) +
-  geom_col(fill = "orange") +
+  geom_col(aes(fill = max_topic)) +
   geom_text(aes(label = n), size = 2, hjust = 1.1) +
@@
 trigram_tf_idf
+
+trigram_tf_idf |> arrange(desc(n))
 ```
@@
 tidy_word_vectors |> 
-  nearest_neighbors("resilience")
+  nearest_neighbors("water")
 ```
 
-The first # components... what do they show?
+The first \# components... what do they show?
 

### 2023-10-31 analysis/paper/paper.qmd
Update introduction text

@@
 
-The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft.
-The EMU is an advanced master started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005.
-The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and experience in practice.
-With a combined research and design program, it provided a bridge towards advanced practice as well as research trajectories after graduation.
-All students followed an exchange semester at one of the other three program universities, allowing them to have a broader view on the field of urbanism.
-Although located in Europe and named "European", its geographic scope is global, especially at the level of graduation, considering that a large majority of students were of non-European origin.
-Influenced by the strong urbanism tradition of the four universities involved, the program yielded a group of almost 100 graduates.
+The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft.
+The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005.
+The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field.
+With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation.
+All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism.
+Although located in Europe and named "European", its geographic scope is global, especially at the stage of graduation, considering that a large majority of students were of non-European origin.
+Influenced by the strong urbanism tradition of the four universities involved, the program yielded a group of almost 100 graduates at TU Delft.
 
-In 2021, TU Delft stepped out the program, but its approach has remained imprinted on the TU Delft approach to urbanism. <!--# paragraph to be elaborated -->
+In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism.
+Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and to some extent MSc Landscape Architecture tracks.
+There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit.
+Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges.
+Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight on what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.
 
-In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
+To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
 To that end, the research presented in this paper employed a text mining approach to analyse the output of the EMU: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.
-
-<!-- If the focus of the paper shifts to topic modeling the paragraphs below can be swapped with the paragraphs describing the EMU program. -->
 

### 2023-10-31 analysis/paper/paper.qmd
Update data collection text and in-line code

@@
 
-## Data collection
+## Data collection {#datacol}
 
-The dataset consists of full texts and metadata of a subset of EMU theses written in English.
-The full texts are available as PDF documents with complex layouts typical to an urbanism project in which text and various types of graphics are combined.
+**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined.
 The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository.
-In addition, metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
+Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
 Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published.
 Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
-As a result, of the total of `r sum(theses_all_per_year$n)` theses presented in the EMU program, full-text documents were obtained for 83 of which permissions were granted for `r nrow(emu)`.
+As a result, of the total of `r sum(theses_all_per_year$n_total)` theses presented in the EMU program, full-text documents were obtained for `r sum(theses_all_per_year$n_pdfs)` of which permissions were granted for `r nrow(emu)`.
 Only PDFs for which permission was obtained were included in the corpus.
-The final corpus contains at least one thesis for each year of graduation between 2007-2021 as shown in @tbl-data.
+The final corpus contains at least one thesis for each year of graduation between `r min(emu$graduation_year)`-`r max(emu$graduation_year)` as shown in @tbl-data.
 
@@
 
-emu |> 
-  count(graduation_year) |> 
-  left_join(theses_all_per_year, by = c("graduation_year" = "grad_year")) |> 
-  mutate(graduation_year = as.factor(graduation_year),
-         `% of all theses` = scales::label_percent()(round(n.x / n.y, 4))) |>
-  bind_rows(data.frame(graduation_year = "", 
-                       n.x = nrow(emu), 
-                       n.y = sum(theses_all_per_year$n),
-                       `% of all theses` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n)),
+theses_all_per_year |> 
+  select(grad_year, n_total, n_permissions) |> 
+  mutate(grad_year = as.factor(grad_year),
+         `% of all theses` = scales::label_percent()(round(n_permissions / n_total, 4))) |> 
+  bind_rows(data.frame(grad_year = "", 
+                       n_permissions = nrow(emu), 
+                       n_total = sum(theses_all_per_year$n_total),
+                       `% of all theses` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
                        check.names = FALSE)) |> 
-  rename(`Graduation year` = graduation_year,
-         `No. theses included` = n.x) |>
-  select(-n.y) |>
-  kableExtra::kable(align = "r")
+  rename(`No. theses` = n_total,
+         `Graduation year` = grad_year,
+         `No. theses included` = n_permissions) |> 
+  kableExtra::kable(align = "lrrr")
+
 ```

### 2023-10-31 analysis/supplementary-materials/1-prepare-metadata.R
Create summary metadata table with included theses

@@ -93,5 +93,9 @@ theses_meta %>%
 
 ## Total theses per year
 readxl::read_xlsx(path = here("analysis", "data", "raw_data", "theses-all.xlsx")) |>
-  count(grad_year) |>
+  # count(grad_year, pdf_exists, permission_granted) |>
+  group_by(grad_year) |>
+  summarise(n_total = n(),
+            n_pdfs = sum(as.logical(pdf_exists), na.rm = TRUE),
+            n_permissions = sum(as.logical(permission_granted), na.rm = TRUE)) |>
   write_csv(file = here("analysis", "data", "derived_data", "theses-per-year.csv"))

### 2023-12-05 analysis/supplementary-materials/2-prepare-for-publication.R
Update preprocessing scripts

@@ -95,7 +95,7 @@ text_raw <- mapply(str_replace_all, text_raw, pattern_ln,
 # head_text(text_raw)
 
 # incorporate into the main dataset
-emu_theses$text_raw  <- text_raw
+emu_theses$text_raw <- text_raw
 
 
 # Make selection of columns -----------------------------------------------
@@ -103,10 +103,13 @@ emu_theses <- emu_theses |>
   filter(as.logical(permission_granted) == TRUE) |>
   select(ID, graduation_year,
          graduation_semester,
-         title, full_title, subtitle, link,
+         full_title, title, subtitle, link,
          location,
          abstract,
-         text_raw)
+         text_raw) |>
+  rename(grad_year = graduation_year,
+         grad_sem = graduation_semester,
+         loc = location)
 
 
 # Add longitude and latitude columns --------------------------------------
@@ -116,17 +119,35 @@ emu_theses <- geocode_thesis_locations(emu_theses)
 # Reorder the columns in the dataset --------------------------------------
 emu_theses <- emu_theses |>
   select(ID,
-         graduation_year,
-         graduation_semester,
+         grad_year,
+         grad_sem,
          full_title, title, subtitle, link,
-         location, latitude, longitude,
+         loc, lat, long,
          abstract,
          text_raw)
 
 
+# Clean raw data for publication ----------------------------------------
+meta_sw <- mintEMU::find_meta_stopwords(emu_theses, clean_meta = FALSE)
+anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
+  paste(collapse = "|")
+
+emu_theses$text_raw <- emu_theses$text_raw |>
+  # Remove hyphenation
+  stringr::str_replace_all("-\\n", "") |>
+  # Remove line breaks
+  stringr::str_replace_all(" \\n", " ") |>
+  # Remove page numbers
+  stringr::str_replace_all("\\\n{0,}\\s{0,}\\d+\\s{0,}\\.?\\\n", "\n") |>
+  # Remove text included with anonymisation
+  stringr::str_remove_all(anonymisation_sw) |>
+  # Remove occurrences of the full title
+  stringr::str_remove_all(meta_sw)
+
+
 # Separate metadata file and text file ------------------------------------
 emu_metadata <- emu_theses |>
[…]

### 2023-12-05 data-raw/emu_text.R

@@ -1,62 +1,49 @@
-# Download and import the dataset ----
-emu_raw_path <- here::here("analysis", "data", "derived_data", "emu_raw.csv")
+# Download and import the dataset -----------------------------------------
+
+# TODO Replace paths to emu_raw.csv and emu_metadata.csv with URL of published dataset
+
+## Raw text data ----
+emu_raw_path <- here::here("analysis", "data", "derived_data", "emu-raw.csv")
 emu_raw <- readr::read_csv(emu_raw_path)
 
-# add raw dataset to the package
+### Add raw dataset to the package
 usethis::use_data(emu_raw, overwrite = TRUE)
 
-# emu metadata file
-emu_meta_path <- here::here("analysis", "data", "derived_data", "emu_metadata.csv")
+## Metadata ----
+emu_meta_path <- here::here("analysis", "data", "derived_data", "emu-metadata.csv")
 emu_metadata <- readr::read_csv(emu_meta_path)
 
-# add metadata to the package
+### Add metadata to the package
 usethis::use_data(emu_metadata, overwrite = TRUE)
 
-# Clean the dataset ----
-emu <- dplyr::left_join(emu_raw, emu_metadata, by  = "ID")
+# Clean the dataset -------------------------------------------------------
+emu <- dplyr::left_join(emu_raw, emu_metadata, by = "ID")
 
-# Prepare lists of stop words
-meta_sw <- mintEMU::find_meta_stopwords(emu, clean_meta = FALSE)
+## Prepare lists of stop words ----
+# meta_sw <- mintEMU::find_meta_stopwords(emu, clean_meta = FALSE)
 emu_sw <- mintEMU::urbanism_stopwords(
   add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
                     "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
 thesis_sw <- mintEMU::thesis_stopwords(
   add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
-anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
-  paste(collapse = "|")
[…]

### 2023-12-05 R/analyse.R
Add ldavis function

@@ -407,6 +407,33 @@ count_docs_per_topic <- function(lda) {
              row.names = NULL)
 }
 
+#' Visualise LDA model
+#'
+#' @param phi   A matrix with each row representing the probability
+#'              distribution of a topic over terms
+#' @param theta A matrix with each row representing the probability
+#'              distribution of a documents over topics
+#' @param dtm   Document-term matrix used as input to the LDA model
+#'
+#' @return  Visualise LDA in a browser.
+#' @export
+visualise_lda <- function(phi, theta, dtm) {
 
+  # Function to approximate the distance between topics
+  svd_tsne <- function(x) tsne::tsne(svd(x)$u)
 
+  # Convert DTM into JSON required by the LDAvis package
+  json <- LDAvis::createJSON(
+    phi = phi,
+    theta = theta,
+    doc.length = rowSums(as.matrix(dtm)),
+    vocab = colnames(dtm),
+    term.frequency = colSums(as.matrix(dtm)),
+    mds.method = svd_tsne,
+    plot.opts = list(xlab="", ylab="")
+  )
+
+  # Visualise topics model with LDAvis
+  LDAvis::serVis(json)
+}
 

### 2023-12-05 analysis/paper/paper.qmd
Update comments on document structure in paper
Add various changes to paper

@@
   #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
-  # pdf:
-  #   template: "../templates/plos_latex_template.tex"
+  pdf:
+    template: "../templates/plos_latex_template.tex"
 execute:
@@
 
-<!-- With the following code you can access and display values from the yml header above. -->
+<!-- The following code display values from the YAML header above. -->
 
@@
 
-<!-- The actual document text starts here: -->
+<!-- Setup code, not shown in the rendered document. -->
 
@@
 
-theses_all_per_year <- read_csv(file = here("analysis", "data", "derived_data", "theses-per-year.csv"))
+theses_all_per_year <- read_csv(file = here("analysis", "data", "derived_data", "emu-theses-per-year.csv"))
 
-# # Prepare data for analysis
-# ## Create lists of stop words
-# thesis_sw <- thesis_stopwords(
-#   add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
-# custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
-#                "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
-#                "create", "exist", "analys(i|e)s") |>
-#   (\(.) paste0("\\b", ., "\\b"))() |>
-#   paste(collapse = "|")
-# emu_sw <- urbanism_stopwords(
-#   add_stopwords = c("emu", "european post[\\-]?(graduate)?[\\s]?master[s]? in urbanism",
-#                     "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
-# 
-# emu$text_clean <- emu$text_clean |> 
-#   # Remove white spaces and punctuation, and change text to lower-case
-#   clean_basic() |>
-#   # Remove thesis-specific words
-#   str_remove_all(thesis_sw) |> 
-#   # Remove dataset-specific words
-#   str_remove_all(emu_sw)
-# 
-# # Concatenate ngrams
-# emu$text_clean <- emu |>
-#   mintEMU::c_ngrams(n_max = 3, text_col = "text_clean",
-#            id_col = "ID", min_freq = 20, lemma = TRUE)
-# 
-# # Words that were not included in ngrams are safe to remove now
-# emu$text_clean <- emu$text_clean |>
-#   # Words that were not included in ngrams are safe to remove now
-#   str_remove_all(custom_sw)
+word_counts <- vector(mode = "numeric", length = nrow(emu))
+for (i in 1:nrow(emu)) {
+  word_counts[i] <- 
+    tidytext::unnest_tokens(tbl = emu[i,], output = word, input = text_clean) |> nrow()
+}
 
@@
 
-# Introduction
+<!-- The actual document text starts here: -->
+
+# Introduction {#intro}
 
@@
 All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism.
-Although located in Europe and named "European", its geographic scope is global, especially at the stage of graduation, considering that a large majority of students were of non-European origin.
-Influenced by the strong urbanism tradition of the four universities involved, the program yielded a group of almost 100 graduates at TU Delft.
+Although located in Europe and named "European", its geographic scope was global, especially at the stage of graduation, considering that a large majority of students were of non-European origin.
 
 In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism.
-Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and to some extent MSc Landscape Architecture tracks.
+Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and to some extent MSc Landscape Architecture tracks.
 There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit.
 Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges.
-Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight on what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.
+Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight into what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.
 
 To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
-To that end, the research presented in this paper employed a text mining approach to analyse the output of the EMU: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.
+To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an average of `r round(mean(word_counts, na.rm = TRUE), -3)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.
+
+### Topic modeling {#topmod}
 
 **The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012].
-At the same time, urbanism research in general and urbanism education studies in particular that use statistical methods for unstructured data, such as topic modeling, are scarce.
+At the same time, urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data.
 This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then.
 Patterns revealed in past education can potentially better inform future decisions in education in the field.
-Although the corpus analysed in this study is relatively small, it is large enough to make a quantitative analytical approach untenable.
 
-Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of NLP, as an unsupervised machine learning method to reveal latent the semantic structure and relationships between latent topics in a large document collection [@muchene2021].
+Although the corpus analysed in this study is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.
+
+Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
 Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a *soft clustering algorithm* in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a].
-LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposition into two low-dimensional matrices for doc-topic distribution and word-topic distribution, respectively.
+LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively.
 Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a].
-Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022]
+Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].
 
 In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**.
-It is part of a reproducible research compendium and accompanied by a published dataset.
+Together with a published FAIR dataset, it forms a reproducible research compendium.
 The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023].
-The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository \[ADD CITATION\] and it is available for exploration in a Shiny dashboard at \[ADD LINK\]
+The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository \[ADD CITATION\] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.
 
@@
 The following section describes the case study data, pre-processing steps and the topic modeling workflow.
-Then results of the analysis are summarised in...
+The results of the analysis are then summarised in response to two research questions.
 The paper concludes with a discussion on what the lessons learned from the EMU program mean for urbanism education and practice in general as well as on the broader challenges and opportunities of the text mining approach taken in the paper.
 
-# Methods
+# Methods {#methods}
 
 To describe the legacy of the EMU program this research answers two questions: What were the main topics addressed in the EMU program?
-And how did those topics evolve throughout the duration of the program between 2007-2021?
+And how did those topics evolve over the duration of the program?
 The paper then discusses to what extent those topics and evolution thereof were influenced by the assignments given to the students throughout their studies leading up to their thesis and by the exchange semester followed by the students.
@@
     B-->C[Information extraction]
+    C-->B
     C-->D[Visualisation]
-    C-->B
     D-->E[Results]
@@
 ```
-
-1.
-
-2.  Pre-processing ...
-
-3.  Information extraction ... topic modeling:
-
-3.1. tokenisation with a window of 1-3 words
-
-3.2. DTM, discarding terms with a frequency under a specified threshold
-
-3.3. Assign a K value manually
-
-3.4. Calculate topic coherence for K values ranging from 2 to K
-
-3.5. Plot K values agianst coherence scores and select K with the highest coherence value
-
-3.6. Fit model with K determined in previous step
-
-3.7. Get top 20 words in each topic to understand what it is about
-
-3.8. Run a cluster dendrogram (Hellinger distance) to see which topics should be grouped together
-
-3.9. Visualise topics with word clouds
-
-OR Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
-Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
-
-4.  Visualisation ...
-
-5.  Results ...
 
@@
 Only PDFs for which permission was obtained were included in the corpus.
-The final corpus contains at least one thesis for each year of graduation between `r min(emu$graduation_year)`-`r max(emu$graduation_year)` as shown in @tbl-data.
+The final corpus contains at least one thesis for each year of graduation between `r min(emu$graduation_year)`-`r max(emu$graduation_year)` as shown in @tbl-data-selection.
+@fig-map-theses shows the geographic spread of the theses included in the analysis.
 
-```{r tbl-data}
+```{r tbl-data-selection}
 #| tbl-cap: Overview of theses included in the analysis
@@
                        check.names = FALSE)) |> 
+  filter(!is.na(grad_year)) |> 
   rename(`No. theses` = n_total,
@@
 ```
-
-## Pre-processing
-
-The text was extracted from the original theses with . The text was
-
-Description of text extraction;
-
-The raw data available on 4TU.ResearchData at...The imported dataset contains the text data in the `text` column and related metadata in the other columns...
-
-For each thesis, henceforth referred to as "document", the following cleaning steps were performed:
-
-1.
-Text was extracted from the original PDF documents with the the PyMuPDF Python package.
-During extraction, the front matter (cover page, colophon, contents, acknowledgements) and back matter (references) were removed.
-The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction.
-2.  The extracted text was cleaned of line breaks, hyphenation, occurrences of the thesis title, author name, and page numbers.
-
-Only relevant metadata was kept
-
-The resulting dataset consisted of the `ID`, `title`, `graduation_year`, `graduation_semester`, `exchange_semester`, `location`, `latitude`, `longitude`, `abstract`, `text_raw` and `text_clean` <!--# check column names or add them in an inline code chunk with epoxy -->.
-
-Both the raw and cleaned data are published in the 4TU.ResearchData repository \[ADD CITATION\].
 
@@
 
-@fig-map-theses shows the geographic spread of the theses included in the analysis.
+## Pre-processing {#prepro}
+
+For each full text, henceforth referred to as "document", the following cleaning steps were performed:
+
+1.
+Text was extracted from the original PDF documents with the the PyMuPDF Python package.
+During extraction, the front matter (cover page, colophon, table of contents, acknowledgements) and back matter (references, acknowledgements, appendices) were excluded.
+The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction.
+The extracted text was stored in the `text_raw` column.
+2.
+The extracted text was cleaned in R of line breaks, hyphenation, occurrences of the thesis title, author name, and page numbers.
+The cleaned text data was added to each thesis in a new column `text_clean`.
+3.
+From the metadata only the title, graduation year, exchange semester and location of the case studied in the theses were kept.
+Column names were shortened and updated to follow the naming conventions used in the dataset.
+
+The resulting dataset consists of the id, `title`, `grad_year`, `exch_sem`, `loc`, `lat`, `lon`, `text_raw` and `text_clean`, as shown in @tbl-data-variables.
+
+```{r tbl-data-variables}
+
+# TODO make sure that the columns reported in the text match those included in the dataset
+
+tibble::tibble(Column = names(emu), 
+               Type = purrr::map(emu, class),
+               Completeness = 
+                 purrr::map(emu, is.na) |> 
+                 purrr::map(mean) |> 
+                 purrr::map(\(x){1 - as.numeric(x)}) |> 
+                 purrr::map(scales::label_percent())) |>
+  kableExtra::kable()
+
+```
+
+Both the raw and cleaned data are published in the 4TU.ResearchData repository [@forgaci2023].
 
@@
 
+<!-- Add intro to this section -->
+
+Information extraction as a process of extracting information from unstructured data.
+
 ### Tokenisation
 
-Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
+<!-- [Add to vignette:] Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
 This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.
 `unnest_tokens()` removes punctuation, converts to lower-case, and separate all words into individual tokens.
+-->
 
-Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified during preliminary EDA were concatenated by replacing white spaces and hyphens with underscores.
+Word-level tokenisation was applied on the corpus.
+Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified with high frequency (at least 20 on a corpus level) during EDA were concatenated by replacing white spaces and hyphens with underscores.
 This way, they could be included alongside words in the topic model.
 
-For analyses in which word-level tokens were used, stop words, i.e., common words or words that do not add much meaning, were removed in the pre-processing stage.
+S**top words, i.e., common words or words that do not add much meaning, were removed**.
 Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
@@
 
-```{r tokenise}
-
-# TODO create tokenizer function
-
-# Unnest tokens and remove words of one and two letters
-emu_words <- emu |> 
-  select(ID, text_clean) |> 
-  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
-  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
-  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
-  filter(nchar(word) >= 3)  # keep only words that are at least three letters long
-
-# # Check if concatenated ngrams are retained
-# sort(table(unlist(str_match_all(emu_words$word, "[A-Za-z]+(?:_[A-Za-z]+)+"))), decreasing = TRUE)
+```{r tokenize}
+# Unnest tokens (words, bigrams and trigrams) and remove words of one and two letters
+emu_words <- tokenize(emu)
 ```
@@
 Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
+
+### Document-term matrix
+
+The topic model requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
+
+```{r dtm}
+# Prepare document-term matrix as input for LDA model
+emu_dtm <- convert_to_dtm(emu_words, 
+                          id_col = "ID", 
+                          word_col = "word", 
+                          min_term_freq = 5)
+```
 
@@
 -   The number of topics was qualitatively determined
--   Evaluating the number of topics quantitatively requires can be done through the measure of **topic coherence**
+-   Evaluating the number of topics quantitatively can be done through the measure of **topic coherence**
+    -   Calculate topic coherence for K values ranging from 2 to K
+    -   Plot K values against coherence scores and select K with the highest coherence value
 
-LDA requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
+A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics.
+Hence, finding a suitable value entails a trade-off between interpretability and specificity.
+This is very much dependent on the data at hand and the research question.
 
-```{r topic-model-dtm}
+#### Determining the value of K empirically
 
-# TODO should tf or tf-idf be used for weighting?
-
-# Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_words, 
-                          id_col = "ID", 
-                          word_col = "word", 
-                          min_term_freq = 5)
+```{r topic-model-lda-1}
+# Determine K value empirically
+k = get_number_of_topics(emu_dtm)
 ```
 
+Fit model with K determined in previous step...
+
 ```{r topic-model-lda-1}
-# Fit LDA model with K clusters
-## Determine K value empirically
-k = get_number_of_topics(emu_dtm)
-# TODO run the model with the empirically determined k and compare with results using qualitative value
 emu_lda_k <- LDA(emu_dtm, k = k, method = "Gibbs",
@@
 topic_names_k <- name_topics(beta_k)
+```
 
+Get top 20 words in each topic to understand what it is about
+
+Run a cluster dendrogram (Hellinger distance) to see which topics should be grouped together
+
+```{r}
+# Perform hierarchical cluster analysis.
+
+rownames(beta_k) <- topic_names_k
+dend <- as.dist(beta_k) |> 
+  hclust() |> 
+  as.dendrogram()
+
+# Plot with Color in function of the cluster
+leafcolor <- colormap::colormap(colormap = colormap::colormaps$viridis, nshades = 5, format = "hex", alpha = 1, reverse = FALSE)
+dend |> 
+  # set("labels_col", value = leafcolor, k=5) %>%
+  # set("branches_k_color", value = leafcolor, k = 5) %>%
+  # plot(horiz=TRUE, axes=FALSE)
+  ggdendro::ggdendrogram(rotate = TRUE)
+
+dendr    <- ggdendro::dendro_data(dend, type="rectangle") # convert for ggplot
+clust    <- cutree(as.dist(beta_k) |> hclust(),k = 5)                    # find 2 clusters
+clust.df <- data.frame(label=names(clust), cluster=factor(clust))
+# dendr[["labels"]] has the labels, merge with clust.df based on label column
+dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
+# plot the dendrogram; note use of color=cluster in geom_text(...)
+ggplot() + 
+  geom_segment(data=ggdendro::segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
+  geom_text(data=ggdendro::label(dendr), aes(x, y - 0.002, label=label, hjust=0, color=cluster), 
+           size=3) +
+  coord_flip() + 
+  scale_y_reverse(expand=c(0.1, 0.1)) +
+  theme(axis.line.y=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.ticks.y=element_blank(),
+        axis.text.x=element_blank(),
+        axis.text.y=element_blank(),
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        panel.background=element_rect(fill="white"),
+        panel.grid=element_blank())
+```
+
+Visualise topics with word clouds
+
+#### Determining the value of K qualitatively
+
+OR Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
+Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
+
+```{r topic-model-lda-1}
 ## Determine K value qualitatively
@@
     values_to = "value") |> 
-  left_join(mutate(emu, ID = as.factor(ID), "year" = graduation_year), by = c("document" = "ID"))
+  left_join(mutate(emu, ID = as.factor(ID), "year" = grad_year), by = c("document" = "ID"))
 
@@
 # TODO how to show distances between topics in the paper?
-# Function to approximate the distance between topics
-svd_tsne <- function(x) tsne(svd(x)$u)
 
-# Convert DTM into JSON required by the LDAvis package
-json <- LDAvis::createJSON(
-  phi = beta_5, 
+visualise_lda(phi = beta_5, 
               theta = theta_5, 
-  doc.length = rowSums(as.matrix(emu_dtm)), 
-  vocab = colnames(emu_dtm), 
-  term.frequency = colSums(as.matrix(emu_dtm)),
-  mds.method = svd_tsne,
-  plot.opts = list(xlab="", ylab="")
-)
-
-# Visualise topics model with LDAvis
-LDAvis::serVis(json)
+              dtm = emu_dtm)
 
@@
 # The top n most common words in each topic
-emu_top_terms <- get_top_words_per_topic(emu_topics, 20)
+emu_top_terms <- get_top_words_per_topic(data = emu_topics, top_n = 20)
 
@@
   inner_join(emu |> 
-               select(ID, graduation_year), 
+               select(ID, grad_year), 
              by = c("document" = "ID")) |> 
@@
   coord_flip() + 
-  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)
+  facet_wrap(~ grad_year, ncol = 1, scales = "free", drop = TRUE)
 
@@
   inner_join(emu |> 
-               select(ID, graduation_year), 
+               select(ID, grad_year), 
              by = c("document" = "ID")) |> 
-  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
+  ggplot(aes(x = factor(gradyear), y = topic, fill = factor(topic))) +
   geom_col()
@@
 emu_bigrams_top |>
-  filter(ID == unique(ID)[1:10]) |> 
+  filter(ID == unique(ID)[1:20]) |> 
   ggplot(aes(x = ngram, y = n, fill = ID)) +
@@
 
-# Determine word vectors using SVD, a method for dimensionality reduction
+# Deter                                                                        mine word vectors using SVD, a method for dimensionality reduction
 tidy_word_vectors <- tidy_pmi |> 

### 2023-12-06 R/analyse.R
Add digit argument to get_topic_proportions function, default value 2

@@ -357,7 +357,7 @@ get_tf_idf <- function(data, id_col = NULL, word_col = NULL) {
 #'
 #' @return Data frame with topics and their proportions
 #' @export
-get_topic_proportions <- function(lda) {
+get_topic_proportions <- function(lda, digits = 2) {
 
   topic_prob <- posterior(lda)$topics
   term_prob <- posterior(lda)$terms
@@ -369,7 +369,8 @@ get_topic_proportions <- function(lda) {
   # assign topic names
   names(topic_proportions) <- mintEMU::name_topics(term_prob)
   # show summed proportions in decreased order
-  sorted_topic_proportions <- sort(topic_proportions, decreasing = TRUE)
+  sorted_topic_proportions <- sort(topic_proportions, decreasing = TRUE) |>
+    round(digits = digits)
   data.frame(topic = names(sorted_topic_proportions),
              proportion = sorted_topic_proportions,
              row.names = NULL)

### 2023-12-06 analysis/paper/paper.qmd
Remove pdf template

@@
   #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
-  pdf:
-    template: "../templates/plos_latex_template.tex"
+  # pdf:
+  #   template: "../templates/plos_latex_template.tex"
 execute:

### 2023-12-06 analysis/paper/paper.qmd
Update column names in topic summary table
Update topic model description and fix k to 5

@@
 
-The topic model requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
-
 ```{r dtm}
@@
 
+The topic model requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
+
 ### Topic model
@@
 We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
-In LDA, each document is a mixture of topics and each topic is a mixture of words.
+LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].
+This means that each document is a mixture of topics and each topic is a mixture of words.
 The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
 
-LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].
+The corpus (D) consists of a collection of M = `r nrow(emu)` documents with a total number of N = `r length(emu_words$word)` words and a vocabulary of V = `r length(unique(emu_words$word))` words.
+After the exclusion of words with a frequency lower than 5, the vocabulary was reduced to V = `r ncol(emu_dtm)`.
 
-Latent topics are measured as patterns of word co-occurrence
-
-Describe model parameters:
-
--   collection of M = `r nrow(emu)` documents
-
--   corpus D
-
--   N = `r length(emu_words$word)` words
-
--   vocabulary of V = `r length(unique(emu_words$word))` words
-
-Determining tuning parameters:
-
--   The number of topics was qualitatively determined
--   Evaluating the number of topics quantitatively can be done through the measure of **topic coherence**
-    -   Calculate topic coherence for K values ranging from 2 to K
-    -   Plot K values against coherence scores and select K with the highest coherence value
-
+A key tuning parameter in an LDA model is the number of clusters K.
 A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics.
@@
 This is very much dependent on the data at hand and the research question.
+Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach:
 
-#### Determining the value of K empirically
+- TU Delft, Fall semester: Urban region networks
+- TU Delft, Spring semester: Constructing the sustainable city
+- IUAV Venice, exchange semester:
+- KU Leuven, exchange semester:
+- UPC Barcelona, exchange semester:
 
-```{r topic-model-lda-1}
-# Determine K value empirically
-k = get_number_of_topics(emu_dtm)
-```
-
-Fit model with K determined in previous step...
-
-```{r topic-model-lda-1}
-emu_lda_k <- LDA(emu_dtm, k = k, method = "Gibbs",
-               control = list(seed = 2023, iter = 500, verbose = 25))
-
-beta_k <- posterior(emu_lda_k)$terms
-theta_k <- posterior(emu_lda_k)$topics
-
-# Add pseudo-names to topics based on the top 5 words in each topic
-topic_names_k <- name_topics(beta_k)
-```
+To check how well the clustering is aligned with where the student carried out their exchange semester, we used a confusion matrix to visualize the relation between real values and values assigned through clustering.
 
@@
 
-Run a cluster dendrogram (Hellinger distance) to see which topics should be grouped together
-
-```{r}
-# Perform hierarchical cluster analysis.
-
-rownames(beta_k) <- topic_names_k
-dend <- as.dist(beta_k) |> 
-  hclust() |> 
-  as.dendrogram()
-
-# Plot with Color in function of the cluster
-leafcolor <- colormap::colormap(colormap = colormap::colormaps$viridis, nshades = 5, format = "hex", alpha = 1, reverse = FALSE)
-dend |> 
-  # set("labels_col", value = leafcolor, k=5) %>%
-  # set("branches_k_color", value = leafcolor, k = 5) %>%
-  # plot(horiz=TRUE, axes=FALSE)
-  ggdendro::ggdendrogram(rotate = TRUE)
-
-dendr    <- ggdendro::dendro_data(dend, type="rectangle") # convert for ggplot
-clust    <- cutree(as.dist(beta_k) |> hclust(),k = 5)                    # find 2 clusters
-clust.df <- data.frame(label=names(clust), cluster=factor(clust))
-# dendr[["labels"]] has the labels, merge with clust.df based on label column
-dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
-# plot the dendrogram; note use of color=cluster in geom_text(...)
-ggplot() + 
-  geom_segment(data=ggdendro::segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
-  geom_text(data=ggdendro::label(dendr), aes(x, y - 0.002, label=label, hjust=0, color=cluster), 
-           size=3) +
-  coord_flip() + 
-  scale_y_reverse(expand=c(0.1, 0.1)) +
-  theme(axis.line.y=element_blank(),
-        axis.ticks.x=element_blank(),
-        axis.ticks.y=element_blank(),
-        axis.text.x=element_blank(),
-        axis.text.y=element_blank(),
-        axis.title.x=element_blank(),
-        axis.title.y=element_blank(),
-        panel.background=element_rect(fill="white"),
-        panel.grid=element_blank())
-```
-
 Visualise topics with word clouds
 
-#### Determining the value of K qualitatively
+```{r lda-k-5}
+#| message: false
 
-OR Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester.
-Use a confusion matrix to visualize the relation between real values and values assigned through clustering.
-
-```{r topic-model-lda-1}
-## Determine K value qualitatively
-## K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
+## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
 emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
@@
 topic_proportions |> 
-  left_join(counts_of_primary_topics, by = "topic")
+  left_join(counts_of_primary_topics, by = "topic") |> 
+  rename(Topic = topic,
+         `Proportion per corpus` = proportion,
+         `No. of documents with primary topic` = count)
 ```

### 2023-12-13 R/analyse.R
Remove unused word embeddings functions

@@ -190,141 +190,6 @@ get_number_of_topics <- function(dtm,
   # }
 }
 
-#' Sliding window
-#'
-#' Sliding window function used to calculate skipgram probabilities,
-#' that is, how often each word occurs near each other word.
-#'
-#' @param tbl A vector of tokens.
-#' @param window_size A small window size of 3-4 shows how a word is used, while a larger window size of, say, 10 gives more information about the domain topic.
-#'
-#' @return Data frame with skipgram probabilities
-#' @export
-#'
-slide_windows <- function(tbl, window_size) {
-  # Iterate through an input vector with a fixed window size
-  skipgrams <- slider::slide(
-    tbl,
-    ~.x,
-    .after = window_size - 1,
-    .step = 1L,
-    .complete = TRUE  # Only evaluate function on complete windows
-  )
-
-  # Run even when unsuccessful
-  safe_mutate <- purrr::safely(mutate)
-
-  # Map safe_mutate to skipgrams
-  out <- map2(skipgrams,
-              1:length(skipgrams),
-              ~ safe_mutate(.x, window_id = .y))
-
-  out |>
-    purrr::transpose() |>  # Transpose `result` and `error` into columns,
-    purrr::pluck("result") |>  # only keep `result`,
-    plyr::compact() |>  # remove any NULL values [is this for the last windows?],
-    dplyr::bind_rows()  # and bind all rows into a single data frame [?].
-}
-
-#' Pointwise mutual information of pairs of items
-#'
-#' @param data Data frame of theses
-#' @param id_col Name of the ID column
-#' @param text_col Name of the text column to be tokenised
-#' @param parallel If `parallel = TRUE`, compute in parallel
-#'
-#' @return Data frame with two items and their co-occurrence
-#' @export
-#'
-get_pmi <- function(data,
-                    id_col = NULL,
-                    text_col = NULL,
-                    parallel = TRUE,
-                    window_size = 4L) {
-
-  # To reduce computation time, we use the {furrr}
-  # and underlying {future} packages for parallel processing
-  if (parallel){
-    future::plan(multisession)
-  }
-
-  # Prepare tidy data frame for word embeddings
-  tidy_emu_theses <- data |>
-    dplyr::select(!!rlang::sym(id_col), !!rlang::sym(text_col)) |>
[…]

### 2023-12-13 R/tokenize.R
Keep year column in the tokenize() function

@@ -3,6 +3,7 @@
 #' @param data A data frame containing a text column
 #' @param id_col Unique document id
 #' @param text_col Name of the column containing the text to be tokenized
+#' @param year_col Name of the column containing the year of graduation
 #' @param min_length Minimum length of words to be kept
 #' @param lemma Lemmatise words
 #'
@@ -11,14 +12,16 @@
 tokenize <- function(data,
                      id_col = "ID",
                      text_col = "text_clean",
+                     year_col = "grad_year",
                      min_length = 3,
                      lemma = TRUE) {
 
   id <- rlang::sym(id_col)
   text <- rlang::sym(text_col)
+  year <- rlang::sym(year_col)
 
   words_df <- data |>
-    dplyr::select(!!id, !!text) |>
+    dplyr::select(!!id, !!text, !!year) |>
     tidytext::unnest_tokens(output = word, input = !!text) |>  # remove punctuation, convert to lower-case, separate all words
     dplyr::anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
     dplyr::mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words

### 2023-12-13 R/analyse.R
Add get_word_counts() function

@@ -1,3 +1,22 @@
+#' Get word counts for corpus
+#'
+#' @param data A data frame comprising a corpus of texts
+#' @param text_col Name of the column in which the text is stored
+#'
+#' @return A numeric vector of word counts, one for each document in the corpus
+#' @export
+get_word_counts <- function(data, text_col = NULL) {
+  text_col <- rlang::sym(text_col)
+
+  word_counts <- vector(mode = "numeric", length = nrow(data))
+  for (i in 1:nrow(data)) {
+    word_counts[i] <-
+      tidytext::unnest_tokens(tbl = data[i,], output = word, input = !!text_col) |> nrow()
+  }
+
+  word_counts
+}
+
 #' Extract top n words with the highest count per document
 #'
 #' Given a data frame with a column storing the words found in a document

### 2023-12-13 R/analyse.R
Update vis_lda function name

@@ -302,7 +302,7 @@ count_docs_per_topic <- function(lda) {
 #'
 #' @return  Visualise LDA in a browser.
 #' @export
-visualise_lda <- function(phi, theta, dtm) {
+vis_lda <- function(phi, theta, dtm) {
 
   # Function to approximate the distance between topics
   svd_tsne <- function(x) tsne::tsne(svd(x)$u)

### 2023-12-13 R/analyse.R
Add visualisation functions

@@ -322,3 +322,60 @@ vis_lda <- function(phi, theta, dtm) {
   LDAvis::serVis(json)
 }
 
+#' Visualise top words per topic
+#'
+#' @param ... Arguments to be passed to `get_top_words_per_topic()`.
+#'
+#' @return `ggplot` bar chart of top n words faceted by topic.
+#' @export
+vis_top_words_per_topic <- function(...) {
+
+  top_terms <- mintEMU::get_top_words_per_topic(...)
+
+  top_terms |>
+    ggplot2::ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
+    ggplot2::geom_col(show.legend = FALSE) +
+    ggplot2::facet_wrap(~ topic, scales = "free_y", ncol = 3) +
+    ggplot2::ylab("Beta") +
+    ggplot2::xlab("Top terms") +
+    ggplot2::labs(title = "Top n terms per topic") +
+    ggplot2::coord_flip()
+}
+
+#' Visualise top words per corpus
+#'
+#' @param ... Arguments to be passed to `get_top_words_per_corpus()`.
+#'
+#' @return `ggplot` bar chart of top n words colored by topic.
+#' @export
+vis_top_words_per_corpus <- function(lda, words, top_n = 20) {
+
+  terms <- as.data.frame(posterior(lda)$terms)
+  rownames(terms) <- mintEMU::name_topics(posterior(lda)$terms)
+
+  terms <- terms |>
+    dplyr::mutate(topic = rownames(terms)) |>
+    tidyr::pivot_longer(-topic,
+                        names_to = "term",
+                        values_to = "prob") |>
+    dplyr::group_by(term) |>
+    dplyr::mutate(max_topic = topic[which.max(prob)]) |>
+    dplyr::filter(topic == max_topic) |>
+    dplyr::ungroup()
+
+  emu_words_topics <- emu_words |>
+    left_join(terms, by = c("word" = "term"))
+
+  top_terms <- mintEMU::get_top_words_per_corpus(words, top_n)
+
+  top_terms |>
+    dplyr::left_join(terms, by = c("word" = "term")) |>
+    ggplot2::ggplot(aes(reorder(word, n), n)) +
+    ggplot2::geom_col(aes(fill = max_topic)) +
+    ggplot2::geom_text(ggplot2::aes(label = n), size = 2, hjust = 1.1) +
+    ggplot2::coord_flip() +
+    ggplot2::xlab("Word") +
[…]

### 2023-12-13 R/analyse.R
Update analysis functions

@@ -55,16 +55,16 @@ get_top_words_per_document <- function(data,
 #' and a positive number top_n, this function returns the top n words with the
 #' highest count across all documents.
 #'
-#' @param data Data frame with theses, containing a column with words.
+#' @param words Data frame with theses, containing a column with words.
 #' @param top_n The number of top words to be returned by the function.
 #' @param word_col Name of the column from the input data frame containing all words.
 #'
 #' @return A data frame with the top n words across all documents.
 #' @export
-get_top_words_per_corpus <- function(data, top_n, word_col = "word") {
+get_top_words_per_corpus <- function(words, top_n, word_col = "word") {
   word <- rlang::sym(word_col)
 
-  data_top_n <- data |>
+  data_top_n <- words |>
     dplyr::group_by(!!word) |>
     dplyr::count(!!word, sort = TRUE) |>
     dplyr::ungroup() |>
@@ -75,7 +75,7 @@ get_top_words_per_corpus <- function(data, top_n, word_col = "word") {
 
 #' Extract top n terms with the highest count across all topics
 #'
-#' @param data Data frame with topics, terms and beta statistic.
+#' @param data LDA model.
 #' @param top_n The number of top words to be returned by the function.
 #' @param topic_col Name of the column containing topics.
 #' @param beta_col Name of the column containing the beta statistic for all terms.
@@ -86,10 +86,14 @@ get_top_words_per_topic <- function(data,
                                     top_n,
                                     topic_col = "topic",
                                     beta_col = "beta") {
+
+  # Extract data frame with topics, terms and beta statistic from LDA model
+  topics <- tidytext::tidy(data, matrix = "beta")
+
   topic <- rlang::sym(topic_col)
   beta <- rlang::sym(beta_col)
 
-  data_top_terms <- data |>
+  data_top_terms <- topics |>
     dplyr::group_by(!!topic) |>
     dplyr::top_n(top_n,!!beta) |>
     dplyr::ungroup() |>
@@ -379,3 +383,4 @@ vis_top_words_per_corpus <- function(lda, words, top_n = 20) {
     ggplot2::labs(title = paste0("Top ", top_n, " most used words in the corpus of theses")) +
     ggplot2::theme_minimal() +
     ggplot2::theme(panel.grid = ggplot2::element_blank())
+}

### 2023-12-13 analysis/paper/paper.qmd
Apply various changes in paper

@@
 abstract: |
-  Prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021,
-  this paper aims to describe the legacy of the EMU. To that end, it reports on the
-  analysis of the graduation theses produced over 2007-2021 when the program was running.
+  This paper aims to describe the legacy of the European post-Master of Urbanism 
+  (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
+  analysis of the graduation theses produced over the duration of the program.
   The text mining approach allows for a close scrutiny of the latent thematic patterns, 
-  and evolutions thereof, found in the full texts of the theses. Results reveal # topics
+  and evolutions thereof, found in the full texts of the theses. Results reveal 5 topics
   that are to a large extent consistent with the set-up of the program in two thematic 
@@
   by a specific approach. The paper discusses the relevance of the findings for the
-  state of the profession and education in the field. Set up as a reproducible research
-  compendium accompanied by a published FAIR dataset of the theses and an interactive
-  dashboard, the analysis presented can be both reproduced and applied to other datasets. 
+  state of the profession and education in the field. Set up as a research
+  compendium accompanied by a published FAIR dataset and an interactive dashboard, 
+  the research is fully reproudcible and enables further analysis. 
 keywords: |
@@
 
-# TODO remove this section after all analysis has been translated into functions
+# TODO remove unused packages after all analysis has been translated into functions
 # Load packages ----
@@
 
-word_counts <- vector(mode = "numeric", length = nrow(emu))
-for (i in 1:nrow(emu)) {
-  word_counts[i] <- 
-    tidytext::unnest_tokens(tbl = emu[i,], output = word, input = text_clean) |> nrow()
-}
+word_counts <- get_word_counts(emu, text_col = "text_clean")
 
@@
 All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism.
-Although located in Europe and named "European", its geographic scope was global, especially at the stage of graduation, considering that a large majority of students were of non-European origin.
+Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.
 
 In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism.
-Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and to some extent MSc Landscape Architecture tracks.
+Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks.
 There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit.
@@
 To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
-To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an average of `r round(mean(word_counts, na.rm = TRUE), -3)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.
+To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format()` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`.
+Although the corpus of thesis is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.
 
-### Topic modeling {#topmod}
+## A text mining approach {#textmin}
 
 **The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012].
-At the same time, urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data.
+Often such data is available in unstructured textual form and formats that are not easily machine-readable, such as Adobe PDF or Microsoft Word documents [@benchimol2022].
+While quantitative text analysis have been widely adopted in computationla social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data.
 This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then.
@@
 
-Although the corpus analysed in this study is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.
-
 Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
-Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a *soft clustering algorithm* in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a].
+Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a].
 LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively.
@@
 The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023].
-The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository \[ADD CITATION\] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.
-
--   Give an overview of th predominant topics and their evolution over the years of the EMU program.
+The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository [@forgaci2023] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.
 
@@
 
-mermaid("
-graph LR
+mermaid("graph LR
     A[Data collection]-->B[Pre-processing]
@@
 
-**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined.
+**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined <!--add figure-->.
 The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository.
@@
 Only PDFs for which permission was obtained were included in the corpus.
-The final corpus contains at least one thesis for each year of graduation between `r min(emu$graduation_year)`-`r max(emu$graduation_year)` as shown in @tbl-data-selection.
+The final corpus contains at least one thesis for each year of graduation between `r min(emu$grad_year)`-`r max(emu$grad_year)` as shown in @tbl-data-selection.
 @fig-map-theses shows the geographic spread of the theses included in the analysis.
@@
 
-## Information extraction
+## Information extraction {#infoext}
 
@@
 
-Information extraction as a process of extracting information from unstructured data.
+Information extraction is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting the topic model
 
 ### Tokenisation
-
-<!-- [Add to vignette:] Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package.
-This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.
-`unnest_tokens()` removes punctuation, converts to lower-case, and separate all words into individual tokens.
--->
 
@@
 
-S**top words, i.e., common words or words that do not add much meaning, were removed**.
+**Stop words, i.e., common words or words that do not add much meaning, were removed**.
 Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
 As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
-<!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant improvement was observed in the result.
---> We manually constructed a list of subject-specific stop words based on our domain knowledge.
+We manually constructed a list of subject-specific stop words based on our domain knowledge.
 Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well.
@@
 Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.
-
-```{r eval=FALSE}
-ent1 <- spacyr::entity_extract(spacyr::spacy_parse(emu$text_clean[1:13]))
-
-ent1 |> 
-  filter(entity_type == "LOC") |> 
-  pull(entity) |> 
-  unique()
-
-ent2 <- ent1 |> 
-  mutate(entity = tolower(entity)) |> 
-  group_by(entity) |> 
-  mutate(bign = length(unique(doc_id))) |> 
-  ungroup() |> 
-  count(doc_id, entity, bign) |> 
-  filter(bign > 3, nchar(entity) > 4)
-
-dtm <- tidytext::cast_sparse(data = ent2,
-                             row = doc_id,
-                             column = entity,
-                             value = n)
-
-lda <- text2vec::LDA$new(n_topics = 50) 
-fit <- lda$fit_transform(dtm, progressbar = F)
-tm_summary <- # incomplete code
-
-```
-
-```{r eval=FALSE}
-# Find locations
-entity::location_entity(head_text(emu$text_clean[15], sub_end = 50000))
-```
 
@@
 
-Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017].
+Similar to clustering on numeric data [@rüdiger2022], topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017].
 We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
@@
 
-A key tuning parameter in an LDA model is the number of clusters K.
+A key tuning parameter (i.e., a manually specified parameter) in an LDA model is the number of clusters K.
 A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics.
@@
 This is very much dependent on the data at hand and the research question.
-Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach:
+Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach.
 
-- TU Delft, Fall semester: Urban region networks
-- TU Delft, Spring semester: Constructing the sustainable city
-- IUAV Venice, exchange semester:
-- KU Leuven, exchange semester:
-- UPC Barcelona, exchange semester:
+| University    | Semester        | Topic                                            |
+|------------------|------------------|------------------------------------|
+| TU Delft      | Fall semester   | Urban region networks                            |
+|               | Spring semester | Constructing the sustainable city                |
+| IUAV Venice   | Fall semester   | Territories of dispersion: Situations, Scenarios |
+| KU Leuven     | Fall semester   |                                                  |
+| UPC Barcelona | Fall semester   | Urban Transformations                            |
+
+Each topic is assigned a pseudo-name comprising the top 5 most frequent words of the topic.
+The top 20 words were used to interpret and describe the meaning of the topics.
+Overall results per topic are visualised with bar charts and the evolution of topics and top words is visualised with line charts.
+
+Using the $\theta$ statistic, the distribution of documents over topics is visualised to classify documents accoriding to their dominant topic and to show to what extent a document is specific to a topic or spread across topics.
+
+Visualise topics with word clouds
 
@@
 
-Get top 20 words in each topic to understand what it is about
-
-Visualise topics with word clouds
-
-```{r lda-k-5}
-#| message: false
+```{r lda}
+#| include: false
 
@@
 emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
-                 control = list(seed = 2023, iter = 500, verbose = 25))
+                 control = list(seed = 2023, iter = 500, verbose = 100))
 
@@
 
-```{r fig-topics-per-doc}
+# Results
+
+<!-- Add intro paragraph -->
+
+## Topics
+
+```{r fig-topics-doc}
 #| fig-cap: Topic probability distribution per document
@@
 # What is the dominant topic in each document?
-N = nrow(theta_5)
-topic_prop_examples <- theta_5[1:N,]
-colnames(topic_prop_examples) <- topic_names_5
+topic_prop <- as.data.frame(theta_5)
+colnames(topic_prop) <- paste0(paste0("Topic ", 1:5, ": "), topic_names_5)
 
-vizDataFrame <- pivot_longer(
-  data = data.frame(topic_prop_examples, 
-                    document = factor(str_sub(rownames(topic_prop_examples), 1, 20))),
+topic_prop_viz <- 
+  pivot_longer(data = data.frame(topic_prop,
+                                 document = factor(rownames(topic_prop)),
+                                 check.names = FALSE),
     cols = -document,
@@
          topic_max = topic[value == value_max])
-ggplot(data = vizDataFrame, aes(topic, value, fill = topic_max), ylab = "proportion") + 
+
+p <- ggplot(data = topic_prop_viz, 
+       aes(topic, value, fill = topic_max), 
+       ylab = "proportion") + 
   geom_bar(stat="identity") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
+  labs(title = "Theses as distributions over topics",
+       subtitle = "Charts colored by the topic with the highest probability") +
+  ylab("Probability") +
+  scale_x_discrete(labels = paste0("Topic ", 1:5)) +
   coord_flip() +
-  facet_wrap(~ as.numeric(document), ncol = 10) +
-  theme(legend.position = "bottom") +
-  scale_fill_viridis_d(name = "Topic") +
-  labs(title = "Theses as distributions over topics",
-       subtitle = "Charts colored by the highest probability")
+  facet_wrap(~ factor(document, levels = rownames(topic_prop)), ncol = 10) +
+  theme(legend.position = "bottom",
+        legend.justification = "left",
+        legend.direction = "vertical",
+        axis.title.y = element_blank()) +
+  scale_fill_viridis_d(name = "Dominant topic")
+
+p_tab <- ggplotGrob(p)
+# print(p_tab)
+
+gtable_filter_remove <- function (x, name, trim = TRUE){
+    matches <- !(x$layout$name %in% name)
+    x$layout <- x$layout[matches, , drop = FALSE]
+    x$grobs <- x$grobs[matches]
+    if (trim) 
+        x <- gtable_trim(x)
+    x
+}
+
+select_grobs <- function() {
+  x <- vector()
+  for (i in 2:10) {
+    for (j in 1:10) {
+      x <- c(x, paste0("axis-l-", i, "-", j))
+      x <- c(x, paste0("axis-b-", i, "-", j))
+    }
+  }
+  x
+}
+
+p_filtered <- gtable_filter_remove(p_tab, name = select_grobs(), trim = FALSE)
+
+library(grid)
+grid::grid.newpage()
+grid::grid.draw(p_filtered)
 
@@
 
-```{r topics-per-corpus}
+```{r topic-proportions-corpus}
 
-# What are the most probable topics in the entire collection?
+# What is the mean probability of the documents being part of each topic?
 topic_proportions <- get_topic_proportions(emu_lda_5)
-# paste0(round(topic_proportions$proportion, 2), ": ", topic_proportions$topic)
 
@@
 counts_of_primary_topics <- count_docs_per_topic(emu_lda_5)
-# paste0(counts_of_primary_topics$count, ": ", topic_proportions$topic)
 
+# Show summary table of mean probabilities and document counts per topic
 topic_proportions |> 
@@
 
-```{r}
-#| eval: false
+The top 20 most frequently used words in the corpus (@fig-top-words-corpus), do not reflect the mean proportion ...
 
-topic_to_filter <- 1  # set manually
-# or have it selected by a term in the topic name (e.g. 'airport')
-topic_to_filter <- grep('water', topic_names_5)[1] 
-topic_threshold <- 0.2
-selected_document_indices <- which(theta_5[, topic_to_filter] >= topic_threshold)
-filtered_corpus <- emu[selected_document_indices,]
-# show length of filtered corpus
-nrow(filtered_corpus)
+```{r fig-top-words-corpus}
+# Which are the top n most frequently used words in all theses?
+vis_top_words_per_corpus(lda = emu_lda_5, words = emu_words, top_n = 20)
 ```
 
-```{r}
-#| eval: false
+```{r top-words-document}
+# Which are the top n most frequently used words in each thesis?
+get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 3)
+```
 
-N = nrow(emu)
-topic_prop_per_year <- theta_5[1:N,]
-colnames(topic_prop_per_year) <- topic_names_5
-
-viz_df_with_year <-
-  pivot_longer(
-    data = data.frame(topic_prop_examples,
-                      document = factor(str_sub(
-                        rownames(topic_prop_examples), 1, 20))),
-    cols = -document,
-    names_to = "topic",
-    values_to = "value") |> 
-  left_join(mutate(emu, ID = as.factor(ID), "year" = grad_year), by = c("document" = "ID"))
-
-# TODO change to discrete colors
-# TODO try representing this as a line chart
-# Plot topic proportions per year
-viz_df_with_year |> 
-  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
-  geom_bar(stat = "identity", position = "fill") +
-  scale_fill_viridis_d(name = "Topics") +
-  xlab("Graduation year") +
-  ylab("Proportion of topics") +
-  labs(title = "Topic proportions per year")
+```{r top-words-topic}
+vis_top_words_per_topic(data = emu_lda_5, top_n = 20)
 ```
@@
 
-visualise_lda(phi = beta_5, 
+vis_lda(phi = beta_5,
         theta = theta_5,
@@
 
-```{r topic-model-beta}
+The `tf-idf` statistic shows words that are the most important to one document in a collection of documents [@silge2017].
+In this case, it shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
 
-# Extract topics as mixtures of terms from LDA model
-emu_topics <- tidy(emu_lda_5, matrix = "beta")
+```{r tf-idf}
 
-# The top n most common words in each topic
-emu_top_terms <- get_top_words_per_topic(data = emu_topics, top_n = 20)
+## Calculate tf-idf statistic
+emu_tf_idf <- emu_words |> 
+  get_tf_idf(id_col = "ID", word_col = "word")
 
-# TODO create visualisation function
-emu_top_terms |> 
-  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
-  geom_col(show.legend = FALSE) +
-  facet_wrap(~ topic, scales = "free") +
-  coord_flip()
+## Get top n words most specific to each thesis 
+emu_tf_idf |> 
+  group_by(ID) |> 
+  top_n(5, tf_idf) |> 
+  arrange(desc(tf_idf), .by_group = TRUE)
 
-# TODO can this be generalised to more than two topics?
+```
 
-beta_spread <- emu_topics |> 
-  mutate(topic = paste0("topic", topic)) |> 
-  spread(topic, beta) |> 
-  filter(topic1 > 0.001, topic2 > .001) |> 
-  mutate(log_ratio = log2(topic2/topic1))
+### Topic 1: `r topic_names_5[1]`
 
-beta_spread %>%
-  group_by(direction = log_ratio > 0) %>%
-  slice_max(abs(log_ratio), n = 20) %>% 
-  ungroup() %>%
-  mutate(term = reorder(term, log_ratio)) %>%
-  ggplot(aes(log_ratio, term)) +
-  geom_col() +
-  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
+### Topic 2: `r topic_names_5[2]`
+
+### Topic 3: `r topic_names_5[3]`
+
+### Topic 4: `r topic_names_5[4]`
+
+### Topic 5: `r topic_names_5[5]`
+
+## The dynamics of thesis topics
+
+```{r fig-topics-evolution}
+#| eval: false
+
+topic_prop_per_year <- theta_5[1:nrow(emu),]
+colnames(topic_prop_per_year) <- topic_names_5
+
+viz_df_with_year <-
+  pivot_longer(
+    data = data.frame(topic_prop_per_year,
+                      document = factor(str_sub(
+                        rownames(topic_prop_per_year), 1, 20)),
+                      check.names = FALSE),
+    cols = -document,
+    names_to = "topic",
+    values_to = "value") |> 
+  left_join(mutate(emu, ID = as.factor(ID), "year" = grad_year), by = c("document" = "ID"))
+
+# TODO change to discrete colors
+# TODO represent this as a line chart
+# Plot topic proportions per year
+viz_df_with_year |> 
+  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
+  geom_bar(stat = "identity", position = "fill") +
+  scale_fill_viridis_d(name = "Topics") +
+  xlab("Graduation year") +
+  ylab("Proportion of topics") +
+  labs(title = "Topic proportions per year")
 ```
@@
 emu_documents |> 
+  mutate(document = as.numeric(document)) |> 
   inner_join(emu |> 
@@
              by = c("document" = "ID")) |> 
-  ggplot(aes(x = factor(gradyear), y = topic, fill = factor(topic))) +
+  ggplot(aes(x = factor(grad_year), y = topic, fill = factor(topic))) +
   geom_col()
-```
-
-# Results
-
-## Overall patterns
-
-The tf-idf statistic shows that...
-
-## Topics
-
-## The dynamics of thesis topics
-
-Topics such as ... have gained importance over the years, while ... have lost prominence.
-
-```{r top-words-document}
-#| eval: false
-
-# Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 10)
-```
-
-```{r top-words-corpus}
-
-# Which are the top n most frequently used words in all theses?
-terms <- as.data.frame(posterior(emu_lda_5)$terms)
-rownames(terms) <- topic_names_5
-terms <- terms |> 
-  mutate(topic = rownames(terms)) |> 
-  pivot_longer(-topic,
-               names_to = "term",
-               values_to = "prob") |> 
-  group_by(term) |> 
-  mutate(max_topic = topic[which.max(prob)]) |> 
-  filter(topic == max_topic) |> 
-  ungroup()
-
-emu_words_topics <- emu_words |> 
-  left_join(terms, by = c("word" = "term"))
-
-n = 20
-get_top_words_per_corpus(data = emu_words, top_n = n) |>
-  left_join(terms, by = c("word" = "term")) |> 
-  ggplot(aes(reorder(word, n), n)) +
-  geom_col(aes(fill = max_topic)) +
-  geom_text(aes(label = n), size = 2, hjust = 1.1) +
-  coord_flip() +
-  xlab("Word") +
-  ylab("Frequency") +
-  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
-  theme_minimal() +
-  theme(panel.grid = element_blank())
-```
-
-A frequency count of ngrams reveals collocations like "public space" or "socio-spatial segregation" across the corpus.
-To include these in the topic model, we concatenated them with an underscore, so that they are considered one word during the word-level tokenisation.
-Only bigrams and trigrams, that is two- and three-word collocations, were extracted from the corpus.
-
-```{r bigram-tfidf}
-
-emu_bigrams <- emu |> 
-  get_ngrams(n = 2, id_col = "ID", text_col = "text_clean", stem = TRUE)
-
-bigram_counts <- emu_bigrams |> 
-  count(w_1, w_2, sort = TRUE)
-
-bigram_tf_idf <- emu_bigrams |> 
-  count(ID, ngram) |> 
-  bind_tf_idf(ngram, ID, n) |> 
-  arrange(desc(tf_idf))
-
-bigram_tf_idf
-```
-
-```{r trigram-tfidf}
-
-emu_trigrams <- emu |> 
-  get_ngrams(n = 3, id_col = "ID", text_col = "text_raw")
-
-trigram_counts <- emu_trigrams |> 
-  count(w_1, w_2, w_3, sort = TRUE)
-
-trigram_tf_idf <- emu_trigrams |> 
-  count(ID, ngram) |> 
-  bind_tf_idf(ngram, ID, n) |> 
-  arrange(desc(tf_idf))
-
-trigram_tf_idf
-
-trigram_tf_idf |> arrange(desc(n))
-```
-
-```{r}
-## Show network of bigrams
-bigram_graph <- bigram_counts |> 
-  filter(n > 100) |> 
-  graph_from_data_frame()
-
-bigram_graph
-
-set.seed(2023)
-
-a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
-
-ggraph(bigram_graph, layout = "fr") +
-  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
-  geom_node_point(color = "lightblue", linewidth = 5) +
-  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
-  theme_void()
-
-## Higgh frequency bigrams ----
-
-emu_bigrams_top <- emu_bigrams |>
-  group_by(ID) |>
-  count(ngram, sort = TRUE) |>
-  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
-  ungroup() |>
-  mutate(bigram = reorder(ngram, n))
-
-emu_bigrams_top_all <- emu_bigrams |>
-  count(ngram, sort = TRUE) |>
-  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
-  # filter(n > 50)                ### filter words with a minimum count
-  mutate(ngram = reorder(ngram, n))
-
-emu_bigrams_top |>
-  filter(ID == unique(ID)[1:20]) |> 
-  ggplot(aes(x = ngram, y = n, fill = ID)) +
-  geom_col() +
-  xlab(NULL) +
-  coord_flip() +
-  labs(y = "Count",
-       x = "Unique words",
-       title = "Count of unique words found in thesis",
-       subtitle = "Stop words were removed from the list") +
-  facet_wrap( ~ ID, scales = "free", ncol = 2) +
-  theme(legend.position = "none") 
-```
-
-The `tf-idf` statistic shows words that are the most important to one document in a collection of documents [@silge2017].
-In this case, it shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words.
-
-```{r tf-idf}
-
-## Calculate tf-idf statistic
-emu_tf_idf <- emu_words |> 
-  get_tf_idf(id_col = "ID", word_col = "word")
-
-## Get top n words most specific to each thesis 
-emu_tf_idf |> 
-  group_by(ID) |> 
-  top_n(5, tf_idf) |> 
-  arrange(desc(tf_idf), .by_group = TRUE)
-
 ```
@@
 
-# Preview term frequency distributions in a subset of theses
-emu_words_count <- emu_words |> 
-  count(ID, word, sort = TRUE) |> 
-  ungroup()
+top_words <- get_top_words_per_corpus(emu_words, top_n = 5, word_col = "word")
 
-emu_words_total <- emu_words_count |> 
-  group_by(ID) |> 
-  summarise(total = sum(n))
+word_order <- emu_words |>
+  group_by(word, grad_year) |>
+  count(word, sort = TRUE) |>
+  ungroup() |>
+  arrange(grad_year,-n) |>
+  group_by(grad_year) |>
+  mutate(relative_frequency = n / sum(n) * 100) |>
+  mutate(rank = round(rank(-n), 0)) |>
+  ungroup() |>
+  filter(word %in% top_words$word)
 
-emu_words_count <- left_join(emu_words_count, emu_words_total)
-
-ggplot(emu_words_count |> 
-         filter(ID == unique(ID)[1:6]),
-       aes(n/total, fill = ID)) +
-  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
-  xlim(NA, 0.0009) +
-  facet_wrap(~ ID, ncol = 3, scales = "free_y")
+ggplot(word_order,
+       aes(
+         x = grad_year,
+         y = relative_frequency,
+         color = word,
+         label = rank
+       )) +
+  # aes(fill = "white") +
+  geom_smooth(lwd = 0.5) +
+  # geom_point(size = 8, shape = 21, fill = "white", stroke = 1.2) +
+  # geom_text(ggplot2::aes(x = grad_year), color = "black") +
+  ggplot2::scale_color_viridis_d() +
+  ggplot2::xlab("Year") + 
+  ggplot2::ylab("Frequency (%)") +
+  facet_wrap(~ word) +
+  theme(legend.position = "none", 
+        title = element_blank(),
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 15))
 ```
 
-```{r}
-# Examine Zipf's law for the EMU theses
-## Calculate rank and term frequency
-freq_by_rank <- emu_words_count |> 
-  group_by(ID) |> 
-  mutate(rank = row_number(),
-         `term frequency` = n/total)
-
-## Visualise Zipf's law
-freq_by_rank |> 
-  ggplot(aes(rank, `term frequency`, color = ID)) +
-  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
-  scale_x_log10() +
-  scale_y_log10()
-
-rank_subset <- freq_by_rank |> 
-  filter(rank < 500,
-         rank > 10)
-
-lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)
-
-freq_by_rank |> 
-  ggplot(aes(rank, `term frequency`, color = ID)) +
-  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
-  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
-  scale_x_log10() +
-  scale_y_log10()
-```
-
-```{r}
-#| eval: false
-
-## Visualise the top n words with the highest tf-idf values for each of a subset of theses
-emu_words_count |> 
-  get_tf_idf(id_col = "ID", word_col = "word") |> 
-  filter(ID == unique(ID)[1:2]) |>  # Keep only the first 6 theses
-  arrange(desc(tf_idf)) |> 
-  mutate(word = factor(word, levels = rev(unique(word)))) |> 
-  group_by(ID) |> 
-  top_n(15) |> 
-  ungroup() |> 
-  ggplot(aes(word, tf_idf, fill = ID)) +
-  geom_col(show.legend = FALSE) +
-  labs(x = NULL, y = "tf-idf") +
-  facet_wrap(~ ID,  ncol = 3, scales = "free") +
-  coord_flip()
-```
-
-### Word embeddings
-
-Describe word embeddings ...
-
-```{r word-embeddings-pmi}
-
-# To what extent do words co-occur with other words?
-tidy_pmi <- get_pmi(emu, id_col = "ID", text_col = "text_clean")
-```
-
-The high-dimensional sparse matrix of word features is projected into reduced, 100-dimensional set of features.
-
-```{r word-embeddings-svd}
-
-# Deter                                                                        mine word vectors using SVD, a method for dimensionality reduction
-tidy_word_vectors <- tidy_pmi |> 
-  widely_svd(
-    item1, item2, pmi,
-    nv = 100, maxit = 1000
-  )
-```
-
-What terms should be checked with the function below?
-
-```{r word-embeddings-nn}
-
-# Calculate nearest neighbours of a given word
-tidy_word_vectors |> 
-  nearest_neighbors("water")
-```
-
-The first \# components... what do they show?
-
-```{r word-embeddings-vis}
-
-# The first n components with top words found in the corpus of theses
-visualise_first_n_components(tidy_word_vectors, 24, 12)
-```
-
-```{r doc-embeddings}
-
-tidy_words <- emu |> 
-  select(ID, text_clean) |> 
-  unnest_tokens(word, text_clean) |> 
-  add_count(word) |> 
-  filter(n >= 50) |> 
-  select(-n)
-
-# Summarise word embeddings into document embeddings for modeling purposes
-word_matrix <- tidy_words |> 
-  count(ID, word) |> 
-  cast_sparse(ID, word, n)
-
-embedding_matrix <- tidy_word_vectors |> 
-  cast_sparse(item1, dimension, value)
-
-doc_matrix <- word_matrix %*% embedding_matrix
-
-dim(doc_matrix)
-```
+As shown in @fig-topics-evolution, topics such as ... have gained importance over the years, while ... have lost prominence.
 
@@
 Although the focus of this study is on describing a fixed corpus though unsupervised machine learning, it can provide input into supervised predictive models.
+
+Research directions...
 

### 2024-02-27 R/analyse.R
Add missing namespaces in analysis functions

@@ -233,7 +233,7 @@ get_tf_idf <- function(data, id_col = NULL, word_col = NULL) {
   words_counts <- left_join(words_counts, words_total)
 
   words_counts <- words_counts |>
-    bind_tf_idf(!!rlang::sym(word_col), !!rlang::sym(id_col), n) |>
+    tidytext::bind_tf_idf(!!rlang::sym(word_col), !!rlang::sym(id_col), n) |>
     select(-total)
 
   words_counts
@@ -247,8 +247,8 @@ get_tf_idf <- function(data, id_col = NULL, word_col = NULL) {
 #' @export
 get_topic_proportions <- function(lda, digits = 2) {
 
-  topic_prob <- posterior(lda)$topics
-  term_prob <- posterior(lda)$terms
+  topic_prob <- modeltools::posterior(lda)$topics
+  term_prob <- modeltools::posterior(lda)$terms
 
   ndocs <- length(unique(rownames(topic_prob)))
 
@@ -272,8 +272,8 @@ get_topic_proportions <- function(lda, digits = 2) {
 #' @export
 count_docs_per_topic <- function(lda) {
 
-  topic_prob <- posterior(lda)$topics
-  term_prob <- posterior(lda)$terms
+  topic_prob <- modeltools::posterior(lda)$topics
+  term_prob <- modeltools::posterior(lda)$terms
 
   k <- ncol(topic_prob)
   ndocs <- length(unique(rownames(topic_prob)))
@@ -354,8 +354,8 @@ vis_top_words_per_topic <- function(...) {
 #' @export
 vis_top_words_per_corpus <- function(lda, words, top_n = 20) {
 
-  terms <- as.data.frame(posterior(lda)$terms)
-  rownames(terms) <- mintEMU::name_topics(posterior(lda)$terms)
+  terms <- as.data.frame(modeltools::posterior(lda)$terms)
+  rownames(terms) <- mintEMU::name_topics(modeltools::posterior(lda)$terms)
 
   terms <- terms |>
     dplyr::mutate(topic = rownames(terms)) |>

### 2024-02-27 analysis/templates/Sage_LaTeX_Guidelines.tex
Update Sage template files

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[Afour,sageh,times]{sagej}

\usepackage{moreverb,url}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2016}

\begin{document}

\runninghead{Smith and Wittkopf}

\title{A demonstration of the \LaTeXe\ class file for
\itshape{SAGE Publications}}

\author{Alistair Smith\affilnum{1} and Hendrik Wittkopf\affilnum{2}}

\affiliation{\affilnum{1}Sunrise Setting Ltd, UK\\
\affilnum{2}SAGE Publications Ltd, UK}

\corrauth{Alistair Smith, Sunrise Setting Ltd
Brixham Laboratory,
Freshwater Quarry,
Brixham, Devon,
TQ5~8BA, UK.}

\email{alistair.smith@sunrise-setting.co.uk}

\begin{abstract}
This paper describes the use of the \LaTeXe\
\textsf{\journalclass} class file for setting papers to be
submitted to a \textit{SAGE Publications} journal.
The template can be downloaded \href{http://www.uk.sagepub.com/repository/binaries/SAGE LaTeX template.zip}{here}.
\end{abstract}

\keywords{Class file, \LaTeXe, \textit{SAGE Publications}}

\maketitle

\section{Introduction}
Many authors submitting to research journals use \LaTeXe\ to
prepare their papers. This paper describes the
\textsf{\journalclass} class file which can be used to convert
articles produced with other \LaTeXe\ class files into the correct
form for submission to \textit{SAGE Publications}.

The \textsf{\journalclass} class file preserves much of the
standard \LaTeXe\ interface so that any document which was
produced using the standard \LaTeXe\ \textsf{article} style can
easily be converted to work with the \textsf{\journalclassshort}
style. However, the width of text and typesize will vary from that
of \textsf{article.cls}; therefore, \textit{line breaks will change}
and it is likely that displayed mathematics and tabular material
will need re-setting.

In the following sections we describe how to lay out your code to
use \textsf{\journalclass} to reproduce much of the typographical look of
the \textit{SAGE} journal that you wish to submit to. However, this paper is not a guide to
using \LaTeXe\ and we would refer you to any of the many books
available (see, for example, \cite{R1}, \cite{R2} and \cite{R3}).

\section{The three golden rules}
Before we proceed, we would like to stress \textit{three golden
rules} that need to be followed to enable the most efficient use
of your code at the typesetting stage:
\begin{enumerate}
\item[(i)] keep your own macros to an absolute minimum;

\item[(ii)] as \TeX\ is designed to make sensible spacing
decisions by itself, do \textit{not} use explicit horizontal or
vertical spacing commands, except in a few accepted (mostly
mathematical) situations, such as \verb"\," before a
differential~d, or \verb"\quad" to separate an equation from its
qualifier;

\item[(iii)] follow the journal reference style.
\end{enumerate}

\section{Getting started} The \textsf{\journalclassshort} class file should run
on any standard \LaTeXe\ installation. If any of the fonts, style
files or packages it requires are missing from your installation,
they can be found on the \textit{\TeX\ Collection} DVDs or downloaded from
CTAN.

\begin{figure*}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\documentclass[<options>]{sagej}

\begin{document}

\runninghead{<Author surnames>}

\title{<Initial capital only>}

\author{<An Author\affilnum{1},
Someone Else\affilnum{2} and
Perhaps Another\affilnum{1}>}

\affiliation{<\affilnum{1}First and third authors' affiliation\\
\affilnum{2}Second author affiliation>}

\corrauth{<Corresponding author's name and full postal address>}

\email{<Corresponding author's email address>}

\begin{abstract}
<Text>
\end{abstract}

\keywords{<List keywords>}

\maketitle

\section{Introduction}
.
.
.
\end{boxedverbatim}
\end{center}
\caption{Example header text.\label{F1}}
\end{figure*}

\section{The article header information}
The heading for any file using \textsf{\journalclass} is shown in
Figure~\ref{F1}. You must select options for the trim/text area and
the reference style of the journal you are submitting to.
The choice of \verb+options+ are listed

% in Table~\ref{T1}.
%
% \begin{table}[h]
% \small\sf\centering
% \caption{The choice of options.\label{T1}}
% \begin{tabular}{lll}
% \toprule
% Option&Trim and font size&Columns\\
% \midrule
% \texttt{shortAfour}& 210\times 280 mm, 10pt& Double column\\
% \texttt{Afour} &210\times 297 mm, 10pt& Double column\\
% \texttt{MCfour} &189\times 246 mm, 10pt& Double column\\
% \texttt{PCfour} &170\times 242 mm, 10pt& Double column\\
% \texttt{Royal} &156\times 234 mm, 10pt& Single column\\
% \texttt{Crown} &7.25\times 9.5 in, 10pt&Single column\\
% \texttt{Review} & 156\times 234 mm, 12pt & Single column\\
% \bottomrule
% \end{tabular}\\[10pt]
% \begin{tabular}{ll}
% \toprule
% Option&Reference style\\
% \midrule
% \texttt{sageh}&SAGE Harvard style (author-year)\\
% \texttt{sagev}&SAGE Vancouver style (superscript numbers)\\
% \texttt{sageapa}&APA style (author-year)\\
% \bottomrule
% \end{tabular}
% \end{table}

For example, if your journal is short A4 sized, uses Times fonts and has Harvard style references then you would need\\
{\small\verb+\documentclass[ShortAfour,times,sageh]{sagej}+}

Most \textit{SAGE} journals are published using Times fonts but if for any reason you have a problem using Times you can
easily resort to Computer Modern fonts by removing the
\verb"times" option.

\subsection{`Review' option}
Some journals (for example, \emph{Journal of the Society for Clinical Trials}) require that
papers are set single column and with a larger font size to help with the review process.
If this is a requirement for the journal that you are submitting to, just add the \verb+Review+ option to the \verb+\documenclass[]{sagej}+ line.

\subsection{Remarks}
\begin{enumerate}
\item[(i)] In \verb"\runninghead" use `\textit{et~al.}' if there
are three or more authors.

\item[(ii)] For multiple author papers please note the use of \verb"\affilnum" to
link names and affiliations. The corresponding author details need to be included using the
\verb+\corrauth+ and \verb+\email+ commands.

\item[(iii)] For submitting a double-spaced manuscript, add
\verb"doublespace" as an option to the documentclass line.

\item[(iv)] The abstract should be capable of standing by itself,
in the absence of the body of the article and of the bibliography.
Therefore, it must not contain any reference citations.

\item[(v)] Keywords are separated by commas.

\item[(vi)] If you are submitting to a \textit{SAGE} journal that requires numbered sections (for example, IJRR), please add the command
  \verb+\setcounter{secnumdepth}{3}+ just above the \verb+\begin{document}+ line.

\end{enumerate}

\section{The body of the article}

\subsection{Mathematics} \textsf{\journalclass} makes the full
functionality of \AmS\/\TeX\ available. We encourage the use of
the \verb"align", \verb"gather" and \verb"multline" environments
for displayed mathematics. \textsf{amsthm} is used for setting
theorem-like and proof environments. The usual \verb"\newtheorem"
command needs to be used to set up the environments for your
particular document.

\subsection{Figures and tables} \textsf{\journalclass} includes the
\textsf{graphicx} package for handling figures.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}

For further details on how to size figures, etc., with the
\textsf{graphicx} package see, for example, \cite{R1}
or \cite{R3}.

The standard coding for a table is shown in Figure~\ref{F2}.

\begin{figure}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\begin{table}
\small\sf\centering
\caption{<Table caption.>}
\begin{tabular}{<table alignment>}
\toprule
<column headings>\\
\midrule
<table entries
(separated by & as usual)>\\
<table entries>\\
.
.
.\\
\bottomrule
\end{tabular}
\end{table}
\end{boxedverbatim}
\end{center}
\caption{Example table layout.\label{F2}}
\end{figure}

\subsection{Cross-referencing}
The use of the \LaTeX\ cross-reference system
for figures, tables, equations, etc., is encouraged
(using \verb"\ref{<name>}" and \verb"\label{<name>}").

\subsection{End of paper special sections}
Depending on the requirements of the journal that you are submitting to,
there are macros defined to typeset various special sections.

The commands available are:
\begin{verbatim}
\begin{acks}
To typeset an
  "Acknowledgements" section.
\end{acks}
\end{verbatim}

\begin{verbatim}
\begin{biog}
To typeset an
  "Author biography" section.
\end{biog}
\end{verbatim}

\begin{verbatim}
\begin{biogs}
To typeset an
  "Author Biographies" section.
\end{biogs}
\end{verbatim}

%\newpage

\begin{verbatim}
\begin{dci}
To typeset a "Declaration of
  conflicting interests" section.
\end{dci}
\end{verbatim}

\begin{verbatim}
\begin{funding}
To typeset a "Funding" section.
\end{funding}
\end{verbatim}

\begin{verbatim}
\begin{sm}
To typeset a
  "Supplemental material" section.
\end{sm}
\end{verbatim}

\subsection{Endnotes}
Most \textit{SAGE} journals use endnotes rather than footnotes, so any notes should be coded as \verb+\endnote{<Text>}+.
Place the command \verb+\theendnotes+ just above the Reference section to typeset the endnotes.

To avoid any confusion for papers that use Vancouver style references,  footnotes/endnotes should be edited into the text.

\subsection{References}
Please note that the files \textsf{SageH.bst} and \textsf{SageV.bst} are included with the class file
for those authors using \BibTeX.
The files work in a completely standard way, and you just need to uncomment one of the lines in the below example depending on what style you require:
\begin{verbatim}
%%Harvard (name/date)
%\bibliographystyle{SageH}
%%Vancouver (numbered)
%\bibliographystyle{SageV}
\bibliography{<YourBibfile.bib>}
\end{verbatim}
and remember to add the relevant option to the \verb+\documentclass[]{sagej}+ line as listed in Table~\ref{T1}.

%\section{Support for \textsf{\journalclass}}
%We offer on-line support to participating authors. Please contact
%us via e-mail at \dots
%
%We would welcome any feedback, positive or otherwise, on your
%experiences of using \textsf{\journalclass}.

\section{Copyright statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ SAGE Publications Ltd,
1 Oliver's Yard, 55 City Road, London, EC1Y~1SP, UK. All
rights reserved.

\subsection{Rules of use}
This class file is made available for use by authors who wish to
prepare an article for publication in a \textit{SAGE Publications} journal.
The user may not exploit any
part of the class file commercially.

This class file is provided on an \textit{as is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantablility or fitness for a particular purpose. There will
be no duty on the author[s] of the software or SAGE Publications Ltd
to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\begin{acks}
This class file was developed by Sunrise Setting Ltd,
Brixham, Devon, UK.\\
Website: \url{http://www.sunrise-setting.co.uk}
\end{acks}

\begin{thebibliography}{99}
\bibitem[Kopka and Daly(2003)]{R1}
Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
Addison-Wesley.

\bibitem[Lamport(1994)]{R2}
Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
2nd~edn. Addison-Wesley.

\bibitem[Mittelbach and Goossens(2004)]{R3}
Mittelbach~F and Goossens~M (2004) \textit{The \LaTeX\ Companion},
2nd~edn. Addison-Wesley.

\end{thebibliography}

\end{document}

### 2024-02-29 analysis/paper/paper.qmd
Update author information in YAML header

@@
 author:
-  - Claudiu Forgaci:
+  - name: Claudiu Forgaci
     correspondence: "yes"
@@
     orcid: 0000-0003-3218-5102
-      institute: tud
-institute:
-  - tud:
+    affiliations: 
+      - ref: tud
+affiliations:
+  - id: tud
     name: Delft University of Technology

### 2024-02-29 analysis/paper/paper.qmd
Change output to html

--- 2023-12-13 analysis/paper/paper.qmd
@@
 date-format: long
-format: pdf
+format:
+  html: default
   # docx:
-  #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
-  # pdf:
-  #   template: "../templates/plos_latex_template.tex"
+  #   reference-doc: ../templates/template.docx
 execute:

### 2024-02-29 analysis/paper/paper.qmd
Fix typo in abstract

@@
   compendium accompanied by a published FAIR dataset and an interactive dashboard, 
-  the research is fully reproudcible and enables further analysis. 
+  the research is fully reproducible and enables further analysis. 
 keywords: |

### 2024-02-29 analysis/paper/paper.qmd
Disable keywords and highlights
Update setup code chunk

@@
 
-<!-- The following code display values from the YAML header above. -->
+<!-- The following code can be used to show values from the YAML header in DOCX output. -->
 
-Keywords: `r rmarkdown::metadata$keywords`
+<!-- Keywords: `r rmarkdown::metadata$keywords` -->
 
-Highlights: `r rmarkdown::metadata$highlights`
+<!-- Highlights: `r rmarkdown::metadata$highlights` -->
 
@@
 
-```{r setup}
+```{r}
+#| label: setup
 #| include: false
 
-# TODO remove unused packages after all analysis has been translated into functions
 # Load packages ----
-## List packages to be loaded
-packages <- c(
-  "here",
-  "tidyverse", "widyr", "readxl",   # Data manipulation and visualisation
-  "furrr",                          # Parallel use of purrr::map() functions
-  "tidytext", "SnowballC", "tm",    # Text processing and analysis
-  "topicmodels", "LDAvis", "tsne",  # Topic modeling
-  "tidygeocoder",                   # Working with spatial data
-  "igraph", "ggraph", "DiagrammeR"  # Network analysis and visualization
-)
+library(tidyverse)
 
-## Load required packages, install them if needed
-for (package in packages) {
-  if (!require(package, character.only = TRUE)) {
-    install.packages(package,  "http://cran.us.r-project.org")
-  }
-  library(package, character.only = TRUE)
-}
+# Load analysis functions ----
+## TODO should this be replaced with a call to `library(mintEMU)`?
+devtools::load_all(".")
 
-# TODO should this be replaced with a call to `library(mintEMU)`?
-# Load analysis functions ----
-devtools::load_all(".")
-```
+# Load data ----
+data(emu_metadata)  # Thesis metadata
+data(emu_raw)       # Raw full text data
+data(emu_clean)     # Cleaned full text data
 
-```{r load-data}
-#| include: false
-
-# TODO merge into setup code chunk
-
-# Load data for analysis
-data(emu_metadata)  # Thesis metadata
-data(emu_raw)       # Raw text data
-data(emu_clean)     # Cleaned text data
-
-# Join metadata and cleaned text data
+# Join metadata and cleaned text data ----
 emu <- emu_metadata |> 
   left_join(emu_raw, by = "ID") |> 
-  left_join(emu_clean, by = "ID")
+  left_join(emu_clean, by = "ID") |> 
+  select(-c(grad_sem, title, subtitle, abstract)) |> 
+  rename(title = full_title) |> 
+  select("ID", "title", everything())
 
-theses_all_per_year <- read_csv(file = here("analysis", "data", "derived_data", "emu-theses-per-year.csv"))
-
+# Get counts per corpus ----
+theses_all_per_year <- read_csv(file = here::here("analysis", "data", "derived_data", "emu-theses-per-year.csv"))
 word_counts <- get_word_counts(emu, text_col = "text_clean")
-
 ```

### 2024-02-29 analysis/paper/paper.qmd
Update intro text

@@
 
-# Introduction {#intro}
+## Introduction {#intro}
 
-The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft.
+The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021.
 The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005.
-The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field.
+The program distinguished itself internationally as an advanced post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field.
 With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation.
@@
 
-In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism.
-Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks.
+In 2021, TU Delft left the program, but its approach has remained imprinted on the TU Delft approach to urbanism.
+Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities—mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters—are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks.
 There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit.
@@
 To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program.
-To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format()` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`.
-Although the corpus of thesis is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.
+To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU program: 96 theses with a rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format(big.mark=",")` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`.
+Although the corpus of theses is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.
 
@@
 
-**The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012].
+The rise of computational social sciences [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012].
 Often such data is available in unstructured textual form and formats that are not easily machine-readable, such as Adobe PDF or Microsoft Word documents [@benchimol2022].
-While quantitative text analysis have been widely adopted in computationla social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data.
-This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then.
+While quantitative text analysis have been widely adopted in computational social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data.
+This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories—the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then.
 Patterns revealed in past education can potentially better inform future decisions in education in the field.
 
-Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
-Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a].
+Acknowledging that opportunity, this paper leverages topic modeling, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
+Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022].
 LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively.
-Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a].
+Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022].
 Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].
 
-In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**.
+In addition to describing the legacy of the EMU, this paper showcases a reproducible research workflow.
 Together with a published FAIR dataset, it forms a reproducible research compendium.

### 2024-02-29 analysis/paper/paper.qmd
Apply text and code revisions throughout the paper

@@
 
-```{r fig-process-diagram}
-#| fig-cap: Text mining process
-#| eval: false
+::: {#fig-process-diagram}
+```{mermaid}
+graph LR
+    A[Data collection]-->B[Pre-processing]
+    B-->C[Text mining]
+    C-->D[Visualisation]
+    D-->B
+    D-->E[Results]
+```
 
-mermaid("graph LR
-    A[Data collection]-->B[Pre-processing]
-    B-->C[Information extraction]
-    C-->B
-    C-->D[Visualisation]
-    D-->E[Results]
-    ")
-```
+Text mining process
+:::
 
@@
 
-**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined <!--add figure-->.
-The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository.
-Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository.
-Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published.
-Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
-As a result, of the total of `r sum(theses_all_per_year$n_total)` theses presented in the EMU program, full-text documents were obtained for `r sum(theses_all_per_year$n_pdfs)` of which permissions were granted for `r nrow(emu)`.
-Only PDFs for which permission was obtained were included in the corpus.
-The final corpus contains at least one thesis for each year of graduation between `r min(emu$grad_year)`-`r max(emu$grad_year)` as shown in @tbl-data-selection.
-@fig-map-theses shows the geographic spread of the theses included in the analysis.
+```{r}
+#| label: data-collection
+#| include: false
 
-```{r tbl-data-selection}
-#| tbl-cap: Overview of theses included in the analysis
-#| tbl-cap-location: top
+n_theses_all <- sum(theses_all_per_year$n_total)
+n_theses_pdf <- sum(theses_all_per_year$n_pdfs)
+n_theses <- nrow(emu)
 
-theses_all_per_year |> 
+start_year <- min(emu$grad_year)
+end_year <- max(emu$grad_year)
+
+theses_overview <- theses_all_per_year |> 
   select(grad_year, n_total, n_permissions) |> 
   mutate(grad_year = as.factor(grad_year),
-         `% of all theses` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
-  bind_rows(data.frame(grad_year = "", 
+         `% of all theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
+  bind_rows(data.frame(grad_year = "2007-2021", 
                        n_permissions = nrow(emu), 
                        n_total = sum(theses_all_per_year$n_total),
-                       `% of all theses` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
+                       `% of all theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
                        check.names = FALSE)) |> 
@@
          `Graduation year` = grad_year,
-         `No. theses included` = n_permissions) |>
-  kableExtra::kable(align = "lrrr")
+         `No. theses included` = n_permissions)
 
+perc_theses_included <- tail(theses_overview, 1)[ncol(theses_overview)] |> pull()
 ```
 
-```{r fig-map-theses}
+The dataset consists of full texts and metadata of a subset of EMU theses written in English.
+The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined @fig-pdf-example.
+The full-text documents were obtained either from the TU Delft Education Repository upon request, or from the authors via email if not available on the repository.
+Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records if unavailable on the repository.
+Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published.
+Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
+As a result, of the total of `r n_theses_all` theses presented in the EMU program, full-text documents were obtained for `r n_theses_pdf` of which permissions were granted for `r n_theses`, that is, `r perc_theses_included` of all theses.
+Only PDFs for which permission was obtained were included in the corpus.
+The final corpus contains at least one thesis for each year of graduation between `r start_year`-`r end_year` as shown in @tbl-data-selection.
+@fig-map-theses shows the geographic spread of the theses included in the analysis.
+
+::: {#fig-pdf-example}
+![](pdf-example.png)
+
+Example of facing pages from one of the theses in the corpus.
+:::
+
+```{r}
+#| label: tbl-data-selection
+#| tbl-cap: Overview of theses included in the analysis
+#| tbl-cap-location: top
+
+theses_overview |>
+  kableExtra::kable(align = "lrrr")
+```
+
+```{r}
+#| label: fig-map-theses
 #| fig-cap: "Location of EMU theses included in the analysis"
+#| warning: false
 
@@
 1.
-Text was extracted from the original PDF documents with the the PyMuPDF Python package.
+Text was extracted from the original PDF documents with the the PyMuPDF Python package [@pymupdf/2024].
 During extraction, the front matter (cover page, colophon, table of contents, acknowledgements) and back matter (references, acknowledgements, appendices) were excluded.
 The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction.
-The extracted text was stored in the `text_raw` column.
+The extracted text was stored in the `text_raw` column of the dataset.
 2.
@@
 
-The resulting dataset consists of the id, `title`, `grad_year`, `exch_sem`, `loc`, `lat`, `lon`, `text_raw` and `text_clean`, as shown in @tbl-data-variables.
+The resulting dataset consists of `ID`, `title`, `grad_year`, `loc`, `lat`, `lon`, `text_raw` and `text_clean`, as shown in @tbl-data-variables.
 
-```{r tbl-data-variables}
-
-# TODO make sure that the columns reported in the text match those included in the dataset
+```{r}
+#| label: tbl-data-variables
+#| tbl-cap: Variables included in the dataset.
+#| tbl-cap-location: top
 
@@
                  purrr::map(scales::label_percent())) |>
-  kableExtra::kable()
+  kableExtra::kable(align = "llr")
 
@@
 
-## Information extraction {#infoext}
+## Text mining {#infoext}
 
-<!-- Add intro to this section -->
+<!-- TODO Add intro to this section -->
 
-Information extraction is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting the topic model
+Text mining is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting a topic model and visualisation of the results.
 
-### Tokenisation
+### Tokenization
 
-Word-level tokenisation was applied on the corpus.
-Bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified with high frequency (at least 20 on a corpus level) during EDA were concatenated by replacing white spaces and hyphens with underscores.
+Word-level tokenization was applied on the corpus.
+In addition, bigrams such as "public space" and trigrams such as "socio-spatial segregation" identified with high frequency (at least 20 on a corpus level) during exploratory data analysis (EDA) were concatenated by replacing white spaces and hyphens with underscores.
 This way, they could be included alongside words in the topic model.
 
-**Stop words, i.e., common words or words that do not add much meaning, were removed**.
+Stop words, i.e., common words or words that do not add much meaning, were removed.
 Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated.
-As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words.
+As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words [@porter2001].
 We manually constructed a list of subject-specific stop words based on our domain knowledge.
@@
 
-```{r tokenize}
+```{r}
+#| label: tokens
+
 # Unnest tokens (words, bigrams and trigrams) and remove words of one and two letters
-emu_words <- tokenize(emu)
+words <- tokenize(emu)
 ```
@@
 Finally, additional checks were made to the document in search for previously undetected stop words.
-Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, including "map" and "low", were meaningful and occurring in high frequency.
+Short words of one and two letters could be safely removed, while three-letter words were kept, as many of those, such as "map", were meaningful and occurring in high frequency.
 
@@
 
-```{r dtm}
+```{r}
+#| label: dtm
 # Prepare document-term matrix as input for LDA model
-emu_dtm <- convert_to_dtm(emu_words, 
+dtm <- convert_to_dtm(words, 
                           id_col = "ID", 
@@
 
-The topic model requires a document-term matrix (DTM) as input. To reduce the size of the DTM, words with a frequency lower than 5 were excluded.
+<!-- TODO Elaborate or merge with previous section -->
+
+The tokenized corpus was then converted into a document-term matrix (DTM), which is the format needed as input for the topic model.
+To reduce the size of the DTM, words with a corpus-level frequency lower than 5 were excluded.
 
@@
 Similar to clustering on numeric data [@rüdiger2022], topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017].
-We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling.
-LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].
+We used Latent Dirichlet Allocation (LDA), a generative probabilist topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].
 This means that each document is a mixture of topics and each topic is a mixture of words.
-The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
+The probability of a word being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.
 
-The corpus (D) consists of a collection of M = `r nrow(emu)` documents with a total number of N = `r length(emu_words$word)` words and a vocabulary of V = `r length(unique(emu_words$word))` words.
-After the exclusion of words with a frequency lower than 5, the vocabulary was reduced to V = `r ncol(emu_dtm)`.
+The corpus ($D$) consists of a collection of M = `r nrow(emu)` documents with a total number of N = `r length(words$word) |> format(big.mark=",")` words and a vocabulary of V = `r length(unique(words$word)) |> format(big.mark=",")` words.
+After the exclusion of words with a frequency lower than 5, the vocabulary was reduced to V = `r ncol(dtm) |> format(big.mark=",")`.
 
-A key tuning parameter (i.e., a manually specified parameter) in an LDA model is the number of clusters K.
-A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics.
-Hence, finding a suitable value entails a trade-off between interpretability and specificity.
+A key tuning parameter (i.e., a manually specified parameter) in an LDA model is the number of clusters $K$.
+A $K$ with a smaller value will yield more general topics whereas a $K$ with a high value will result in more specific topics.
+Hence, finding a suitable value entails a trade-off between interpretability and specificity<!-- add reference here -->.
 This is very much dependent on the data at hand and the research question.
-Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach.
+Given the small size of the corpus and the way the program was structured, the value of $K = 5$ was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach, as described in @tbl-semesters.
 
 | University    | Semester        | Topic                                            |
-|------------------|------------------|------------------------------------|
+|-----------------|-----------------|--------------------------------------|
 | TU Delft      | Fall semester   | Urban region networks                            |
@@
 
-Each topic is assigned a pseudo-name comprising the top 5 most frequent words of the topic.
-The top 20 words were used to interpret and describe the meaning of the topics.
-Overall results per topic are visualised with bar charts and the evolution of topics and top words is visualised with line charts.
+: Topics covered by each EMU semester at TU Delft and in the exchange locations. {#tbl-semesters}
 
-Using the $\theta$ statistic, the distribution of documents over topics is visualised to classify documents accoriding to their dominant topic and to show to what extent a document is specific to a topic or spread across topics.
+Each topic was assigned a pseudo-name comprising the top 5 most frequent words of the topic.
+The top 20 words were used to interpret and describe the meaning of the topics qualitatively.
+Overall results per topic were visualised with bar charts and the evolution of topics and top words was visualised with line charts.
+
+Using the $\theta$ statistic, the distribution of documents over topics is visualised to classify documents according to their dominant topic and to show to what extent a document is specific to a topic or spread across topics.
 
@@
 
-```{r lda}
+```{r}
+#| label: lda
 #| include: false
 
+# TODO transform this into a function
+
 ## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
-emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
+lda <- topicmodels::LDA(dtm, k = 3, method="Gibbs",
                         control = list(seed = 2023, iter = 500, verbose = 100))
@@
 # Extract beta and theta statistics from LDA model
-beta_5 <- posterior(emu_lda_5)$terms
-theta_5 <- posterior(emu_lda_5)$topics
+beta <- modeltools::posterior(lda)$terms
+theta <- modeltools::posterior(lda)$topics
 
 # Add pseudo-names to topics based on the top 5 words in each topic
-topic_names_5 <- name_topics(beta_5)
+topic_names_5 <- name_topics(beta)
 
@@
 
-```{r fig-topics-doc}
+```{r}
+#| label: fig-topics-doc
 #| fig-cap: Topic probability distribution per document
@@
 # What is the dominant topic in each document?
-topic_prop <- as.data.frame(theta_5)
-colnames(topic_prop) <- paste0(paste0("Topic ", 1:5, ": "), topic_names_5)
+topic_prop <- as.data.frame(theta)
+colnames(topic_prop) <- paste0(paste0("Topic ", 1:length(topic_names_5), ": "), topic_names_5)
 
@@
         legend.direction = "vertical",
-        axis.title.y = element_blank()) +
+        axis.title.y = element_blank(),
+        strip.text = element_text(size = 5),
+        strip.background = element_rect(size = 1)) +
   scale_fill_viridis_d(name = "Dominant topic")
@@
 
-```{r topic-proportions-corpus}
+```{r}
+#| label: topic-proportions-corpus
 
 # What is the mean probability of the documents being part of each topic?
-topic_proportions <- get_topic_proportions(emu_lda_5)
+topic_proportions <- get_topic_proportions(lda)
 
 # How many documents are with the highest probability in each topic?
-counts_of_primary_topics <- count_docs_per_topic(emu_lda_5)
+counts_of_primary_topics <- count_docs_per_topic(lda)
 
@@
 
-```{r fig-top-words-corpus}
+```{r}
+#| label: fig-top-words-corpus
+
 # Which are the top n most frequently used words in all theses?
-vis_top_words_per_corpus(lda = emu_lda_5, words = emu_words, top_n = 20)
+n <- 20
+vis_top_words_per_corpus(lda = lda, words = words, top_n = n)
 ```
 
-```{r top-words-document}
+```{r}
+#| label: top-words-document
+
 # Which are the top n most frequently used words in each thesis?
-get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 3)
+n <- 3
+get_top_words_per_document(data = words, id_col = "ID", top_n = n)
 ```
 
-```{r top-words-topic}
-vis_top_words_per_topic(data = emu_lda_5, top_n = 20)
+```{r}
+#| label: top-words-topic
+
+# Which are the top n most frequently used words in each thesis?
+n <- 20
+vis_top_words_per_topic(data = lda, top_n = n)
 ```
 
-```{r lda-vis}
-#| eval: false
+```{r}
+theta |> 
+  as_tibble() |> 
+  mutate(ID = as.numeric(rownames(theta))) |> 
+  pivot_longer(cols = `1`:ncol(theta),
+               names_to = "topic",
+               values_to = "prob") |> 
+  group_by(ID) |> 
+  mutate(max_prob = max(prob, na.rm = FALSE),
+         max_topic = if_else(round(prob, digits = 7) == round(max_prob, digits = 7), topic, NA)) |> 
+  ungroup() |> 
+  select(ID, max_prob, max_topic) |> 
+  drop_na() |> 
+  left_join(emu) |> 
+  select(ID, max_prob, max_topic, title, text_clean) |> 
+  # group_by(max_topic) |> 
+  # filter(max_prob == max(max_prob)) |> 
+  arrange(max_topic, desc(max_prob)) |> 
+  select(max_topic, max_prob, ID, title) |> 
+  kableExtra::kable()
 
-# TODO check if this can be included in the Shiny dashboard
-# TODO remove from paper
-# TODO how to show distances between topics in the paper?
-
-vis_lda(phi = beta_5,
-        theta = theta_5,
-        dtm = emu_dtm)
-
+# words_repr <- words |> 
+#   left_join(repr_theses) |> 
+#   drop_na()
+# 
+# set.seed(2023)
+# wordcloud::wordcloud(words_repr |> filter(ID == unique(words_repr$ID)[3]) |> pull(word), min.freq = 60)
 ```
@@
 
-```{r tf-idf}
+```{r}
+#| label: tf-idf
 
 ## Calculate tf-idf statistic
-emu_tf_idf <- emu_words |> 
+tf_idf <- words |> 
   get_tf_idf(id_col = "ID", word_col = "word")
@@
 ## Get top n words most specific to each thesis 
-emu_tf_idf |> 
+tf_idf |> 
   group_by(ID) |> 
@@
 
+**People and places:** *community land house develop region*
+
+Community, social, ... local scale
+
 ### Topic 2: `r topic_names_5[2]`
+
+**Landscapes:** *water landscape system river land*
+
+The most dominant topic highlights the systemic, landscape-driven component of the program. This topic was central in the Spring semester at TU Delft.
 
@@
 
+**Spatial morphology:** *block street build athens house*
+
+This topic is predominantly morphological... density... plot...
+
+The theses with the highest probabilities in this cluster...
+
 ### Topic 4: `r topic_names_5[4]`
+
+**Integrated projects:** *park build main house community*
 
@@
 
+**Regional networks:** *territory land rural potential region*
+
+This topic represents territorial challenges that go beyond the urban to consider
+
+### Integrated projects
+
+Theses that have a broader distribution across topics have an integrative character.
+Do those theses have a relatively higher frequency of the term "integrate"?
+Also, Topics \# and \# share the dominant term "system" as they both represent systemic approaches
+
 ## The dynamics of thesis topics
 
-```{r fig-topics-evolution}
+```{r}
+#| label: fig-topics-evolution
 #| eval: false
 
-topic_prop_per_year <- theta_5[1:nrow(emu),]
+topic_prop_per_year <- theta[1:nrow(emu),]
 colnames(topic_prop_per_year) <- topic_names_5
@@
 
-# TODO change to discrete colors
 # TODO represent this as a line chart
@@
 
-```{r topic-model-gamma}
+```{r}
+#| label: topic-model-gamma
 #| eval: false
@@
 # Extract documents as mixtures of topics from LDA model
-emu_documents <- tidy(emu_lda_5, matrix = "gamma")
+documents <- tidy(lda, matrix = "gamma")
 
-emu_documents |> 
+documents |> 
   ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
@@
 # TODO is gamma the same as theta ?!
-emu_documents |>
+documents |>
   mutate(document = as.numeric(document)) |> 
@@
 # TODO check if this plot is useful and, if not, remove it
-emu_documents |> 
+documents |> 
   mutate(document = as.numeric(document)) |> 
@@
 
-top_words <- get_top_words_per_corpus(emu_words, top_n = 5, word_col = "word")
+top_words <- get_top_words_per_corpus(words, top_n = 5, word_col = "word")
 
-word_order <- emu_words |>
+word_order <- words |>
   group_by(word, grad_year) |>
@@
 
+-   show mean length of theses over time
+
+-   discuss why theses tend to be longer over time: more frictionless workflows, pre-made layouts, more focus on research
+
 # Discussion
+
+The relevance of the findings for the state of the profession and urbanism education...
+
+What are the core, stable elements of urbanism practice and how does the field need to adapt to changing societal challenges?
+
+The topics can be useful for searching...
+
+The number of topics...
 
@@
 
-# Conclusion
+-   what is generalisable to urbanism theses in general
+
+    -   increasing length is probably the case for all theses
+
+# Conclusion {#conclusion}
 
 # Acknowledgements
+
+I would like to acknowledge the invaluable contribution of Aleksandra Wilczynska to the development of this paper as a reproducible research compendium as part of DCC support provided by the TU Delft library.
+I would also like to thank former EMU coordinators Birgit Hausleitner and Luiz De Carvalho Filho for their support in collecting the metadata and data used in this paper, as well as to Bjarne van der Drift and Greta Samulionytė for their assistance in completing the metadata.
 
@@
 
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+This paper was generated on `r Sys.time()` using the following computational environment and dependencies:
 
@@
 ```{r}
+#| label: last-commit
 # what commit is this file at? 

### 2024-02-29 data-raw/emu_text.R
Update list of custom stopwords

@@ -28,9 +28,9 @@ thesis_sw <- mintEMU::thesis_stopwords(
   add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
 # anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
 #   paste(collapse = "|")
-custom_sw <- c("space[s]?", "spatial", "plan", "(plann)\\w{0,}", "public", "development",
-               "design", "process[es]?", "project[s]?", "figure[s]?", "map[s]?", "strateg(y|ies)", "intervention[s]?",
-               "create", "exist", "analys(i|e)s") |>
+custom_sw <- c("space[s]?", "spatial", "development",
+               "project[s]?", "figure[s]?", "map[s]?", "intervention[s]?",
+               "create", "exist", "main", "plan[s]?", "planning", "land", "design", "low") |>
   (\(.) paste0("\\b", ., "\\b"))() |>
   paste(collapse = "|")
 

### 2024-02-29 data-raw/emu_text.R
Load mintEMU functions in pre-processing script

@@ -1,5 +1,7 @@
 # Download and import the dataset -----------------------------------------
 
+devtools::load_all(".")
+
 # TODO Replace paths to emu_raw.csv and emu_metadata.csv with URL of published dataset
 
 ## Raw text data ----

### 2024-02-29 R/analyse.R
Fix object names

@@ -196,7 +196,7 @@ get_number_of_topics <- function(dtm,
                                  verbose = TRUE) {
 
   result <- ldatuning::FindTopicsNumber(
-    emu_dtm,
+    dtm,
     topics = topics,
     metrics = metrics,
     method = method,
@@ -296,6 +296,7 @@ count_docs_per_topic <- function(lda) {
              row.names = NULL)
 }
 
+# TODO check if this can be included in the Shiny dashboard
 #' Visualise LDA model
 #'
 #' @param phi   A matrix with each row representing the probability
@@ -367,7 +368,7 @@ vis_top_words_per_corpus <- function(lda, words, top_n = 20) {
     dplyr::filter(topic == max_topic) |>
     dplyr::ungroup()
 
-  emu_words_topics <- emu_words |>
+  emu_words_topics <- words |>
     left_join(terms, by = c("word" = "term"))
 
   top_terms <- mintEMU::get_top_words_per_corpus(words, top_n)

### 2024-03-01 README.Rmd
Update readme

@@
 
-This repository contains the data and code for our paper:
+<!-- This repository contains the data and code for the paper: -->
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
+<!-- > Forgaci, C., (2023). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx> -->
 
-Our pre-print is online here:
+<!-- The pre-print is online here: -->
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+<!-- > Forgaci, C., (2023). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->
 
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Forgaci, C., Wilczynska, A., (2023), (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
 
-  - [:file\_folder: paper](/analysis/paper): R Markdown source document
+  - [Paper](/analysis/paper): R Markdown source document
     for manuscript. Includes code to reproduce the figures and tables
@@
     and tables in this file)
-  - [:file\_folder: data](/analysis/data): Data used in the analysis.
-  - [:file\_folder: figures](/analysis/figures): Plots and other
+  - [Data](/analysis/data): Data used in the analysis.
+  - [Figures](/analysis/figures): Plots and other
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
+  - [Supplementary materials](/analysis/supplementary-materials):
     Supplementary materials including notes and other documents
@@
 
-**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
+**Data :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
 

### 2024-06-19 analysis/paper/paper.qmd
Remove unnecessary call to data
Not needed according to R Packages 2e, Ch.7, p.102
Update paper

@@
 abstract: |
-  This paper aims to describe the legacy of the European post-Master of Urbanism 
+  This paper aims to describe the legacy of the European post-Master in Urbanism 
   (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
@@
 
-# Load data ----
-data(emu_metadata)  # Thesis metadata
-data(emu_raw)       # Raw full text data
-data(emu_clean)     # Cleaned full text data
-
 # Join metadata and cleaned text data ----
@@
 
-The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021.
+The research presented in this paper was prompted by the closure of the European post-Master in Urbanism (EMU) at TU Delft in 2021.
 The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005.
@@
 All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism.
-Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.
+Although located in Europe and named "European", its geographic scope was global (see @fig-map-theses), visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.
 
@@
 
-## A text mining approach {#textmin}
+## A text mining approach {#tmapproach}
 
@@
 
-Acknowledging that opportunity, this paper leverages topic modeling, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
+Acknowledging that opportunity, this paper leverages topic modeling as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021].
 Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022].
@@
 
-# Methods {#methods}
+## Methods {#methods}
 
@@
 
-## Data collection {#datacol}
+### Data collection {#datacol}
 
@@
   mutate(grad_year = as.factor(grad_year),
-         `% of all theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
+         `% of theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
   bind_rows(data.frame(grad_year = "2007-2021", 
@@
                        n_total = sum(theses_all_per_year$n_total),
-                       `% of all theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
+                       `% of theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
                        check.names = FALSE)) |> 
@@
 Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published.
-Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually.
+Considering the relatively small size of the corpus and the possibility of directly contacting the alumni community, permissions were requested individually.
 As a result, of the total of `r n_theses_all` theses presented in the EMU program, full-text documents were obtained for `r n_theses_pdf` of which permissions were granted for `r n_theses`, that is, `r perc_theses_included` of all theses.
@@
 
+<!-- Table shown in html output -->
+
+::: {.content-visible when-format="html"}
+```{r}
+#| label: tbl-data-selection-html
+#| tbl-cap: Overview of theses included in the analysis
+#| tbl-cap-location: top
+
+theses_overview |>
+  mutate(`% of theses per period` = formattable::color_tile("lightpink", "lightgreen")(readr::parse_number(theses_overview$`% of theses per period`))) |>
+  rename("Count" = `No. theses included`, "Percentage" = `% of theses per period`) |> 
+  kableExtra::kable(format = "html", align = "lrrr", escape = FALSE) |>
+  kableExtra::column_spec(4, width = "3cm") |> 
+  kableExtra::row_spec(nrow(theses_overview), bold = TRUE) |>
+  kableExtra::kable_styling("hover", full_width = FALSE) |> 
+  kableExtra::add_header_above(c(" " = 2, "Included" = 2))
+```
+:::
+
+<!-- Table shown in non-html output -->
+
+::: {.content-hidden when-format="html"}
 ```{r}
@@
 theses_overview |> 
-  kableExtra::kable(align = "lrrr")
+  gt::gt() |> 
+  gt::tab_options(table.width = "300")
+
 ```
+:::
 
+```{=html}
+<!-- TODO Replace with interactive map when rendering HTML.
+This should be an argument in the `visualise_thesis_locations()` function. -->
+```
+::: {.content-visible when-format="html"}
 ```{r}
@@
 ```
+:::
 
-## Pre-processing {#prepro}
+::: {.content-hidden when-format="html"}
+```{r}
+#| label: fig-map-theses-html
+#| fig-cap: "Location of EMU theses included in the analysis"
+#| warning: false
+
+# Show thesis locations on world map
+visualize_thesis_locations(emu)
+```
+:::
+
+### Pre-processing {#prepro}
 
@@
 3.
-From the metadata only the title, graduation year, exchange semester and location of the case studied in the theses were kept.
+From the metadata only the title, graduation year, and location of the case studied in the theses were kept.
 Column names were shortened and updated to follow the naming conventions used in the dataset.
@@
   kableExtra::kable(align = "llr")
-
 ```
@@
 
-## Text mining {#infoext}
+### Text mining {#textmin}
 
-<!-- TODO Add intro to this section -->
+Text mining is a process of extracting implicit information from unstructured textual data [@jo2019].
+following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting a topic model and visualisation of the results.
 
-Text mining is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting a topic model and visualisation of the results.
-
-### Tokenization
+#### Tokenization {#tokenization}
 
@@
 
-### Document-term matrix
+#### Document-term matrix {#dtm}
 
@@
 
-### Topic model
+#### Topic model {#tomo}
 
@@
 | University    | Semester        | Topic                                            |
-|-----------------|-----------------|--------------------------------------|
+|------------------|------------------|------------------------------------|
 | TU Delft      | Fall semester   | Urban region networks                            |
@@
 ```{r}
+k.topics <- 2:9
+folding <- c(rep(1:5, each = 13), rep(5, 3))
+
+runonce <- function(k, fold) {
+    testing.dtm <- which(folding == fold)
+    training.dtm <- which(folding != fold)
+
+    training.model <- topicmodels::LDA(dtm[training.dtm, ], k = k)
+    test.model <- topicmodels::LDA(dtm[testing.dtm, ], model = training.model, control = list(estimate.beta = FALSE))
+
+    topicmodels::perplexity(test.model)
+}
+
+res <- NULL
+
+for (k in 2:9) {
+    for (fold in 1:5) {
+        res <- rbind(res, c(k, fold, runonce(k, fold)))
+    }
+}
+
+total.perp <- tapply(res[, 3], res[, 1], sum)
+round(total.perp)
+
+plot(2:9, total.perp, type = "b", xlab = "Number of topics", ylab = "Perplexity")
+```
+
+```{r}
 #| label: lda
@@
 ## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
-lda <- topicmodels::LDA(dtm, k = 3, method="Gibbs",
+lda <- topicmodels::LDA(dtm, k = 5, method="Gibbs",
                         control = list(seed = 2023, iter = 500, verbose = 100))
@@
 # Add pseudo-names to topics based on the top 5 words in each topic
-topic_names_5 <- name_topics(beta)
+topic_pseudonames <- name_topics(beta, n_words = 5)
 
@@
 
-# Results
+## Results {#results}
 
@@
 
-## Topics
+### Topics {#topics}
 
@@
 topic_prop <- as.data.frame(theta)
-colnames(topic_prop) <- paste0(paste0("Topic ", 1:length(topic_names_5), ": "), topic_names_5)
+colnames(topic_prop) <- paste0(paste0("Topic ", 1:length(topic_pseudonames), ": "), topic_pseudonames)
 
@@
 
-### Topic 1: `r topic_names_5[1]`
+#### Topic 1: `r topic_pseudonames[1]` {#topic1}
 
@@
 
-### Topic 2: `r topic_names_5[2]`
+#### Topic 2: `r topic_pseudonames[2]` {#topic2}
 
@@
 
-### Topic 3: `r topic_names_5[3]`
+#### Topic 3: `r topic_pseudonames[3]` {#topic3}
 
@@
 
-### Topic 4: `r topic_names_5[4]`
+#### Topic 4: `r topic_pseudonames[4]` {#topic4}
 
@@
 
-### Topic 5: `r topic_names_5[5]`
+#### Topic 5: `r topic_pseudonames[5]` {#topic5}
 
@@
 
-### Integrated projects
+#### Integrated projects {#intpro}
 
@@
 
-## The dynamics of thesis topics
+### The dynamics of thesis topics {#topdyn}
 
@@
 topic_prop_per_year <- theta[1:nrow(emu),]
-colnames(topic_prop_per_year) <- topic_names_5
+colnames(topic_prop_per_year) <- topic_pseudonames
 
@@
 # Extract documents as mixtures of topics from LDA model
-documents <- tidy(lda, matrix = "gamma")
+documents <- tidytext::tidy(lda, matrix = "gamma")
 
@@
 
-# Discussion
+## Discussion {#discussion}
 
@@
 
-# Conclusion {#conclusion}
+## Conclusion {#conclusion}
 
-# Acknowledgements
+## Acknowledgements
 
@@
 
-# References
+## References
 

### 2024-06-19 data-raw/emu_text.R
Update data

@@ -30,9 +30,9 @@ thesis_sw <- mintEMU::thesis_stopwords(
   add_stopwords = c("advisor", "prof", "professor", "fig", "ure", "figure", "ning", "dr", "ir", "drir"))
 # anonymisation_sw <- c("EMAIL_REMOVED", "AUTHOR_REMOVED", "FIRST_NAME_REMOVED", "LAST_NAME_REMOVED") |>
 #   paste(collapse = "|")
-custom_sw <- c("space[s]?", "spatial", "development",
+custom_sw <- c("development", "space", "spatial",
                "project[s]?", "figure[s]?", "map[s]?", "intervention[s]?",
-               "create", "exist", "main", "plan[s]?", "planning", "land", "design", "low") |>
+               "create", "exist", "main", "low") |>
   (\(.) paste0("\\b", ., "\\b"))() |>
   paste(collapse = "|")
 

### 2024-06-19 R/ngrams.R
Update column selection in ngram functions

@@ -20,7 +20,7 @@ get_ngrams <- function(data,
 
   # Create ngrams column and separate columns for each of the n terms
   data_ngrams <- data |>
-    dplyr::select(id_col, text_col) |>
+    dplyr::select(all_of(id_col), all_of(text_col)) |>
     tidytext::unnest_tokens(ngram, text_col, token = "ngrams", n = n) |>
     tidyr::separate(ngram, into = paste("w" , 1:n, sep = "_"),
                     sep = " ", remove = FALSE)
@@ -38,8 +38,8 @@ get_ngrams <- function(data,
 
   # TODO short words in a separate function
   short_words <- data |>
-    dplyr::select(text_raw) |>
-    tidytext::unnest_tokens(output = word, input = text_raw) |>
+    dplyr::select(all_of(text_col)) |>
+    tidytext::unnest_tokens(output = word, input = text_col) |>
     dplyr::anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
     dplyr::mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
     dplyr::filter(nchar(word) <= rm_n_short_words)

### 2024-06-19 R/clean.R
Update cleaning functions

@@ -105,7 +105,7 @@ find_meta_stopwords <- function(metadata,
 #' @export
 urbanism_stopwords <- function(add_stopwords = NULL, convert_to_regex = TRUE) {
   stop_words <- paste("\\b",
-                      c(add_stopwords, "city", "cities", "urban", "urbanism", "hab", "km"),
+                      c(add_stopwords |> tolower() |> trimws(), "city", "cities", "urban", "urbanism", "hab", "km"),
                       "\\b", sep = "")
 
   if (convert_to_regex) return(paste(stop_words, collapse = "|"))
@@ -149,7 +149,7 @@ short_words <-
            l_word = 4,
            l_list = 50,
            scope = 'word',
-           print_kable = TRUE) {
+           print_kable = FALSE) {
     if (scope == 'word') {
 
       out_table <-
@@ -244,7 +244,7 @@ word_vec_combos <- function(string1, string2, empty_replacement = "XXXXXXXXXXXXX
 
   # Find combinations with the condition
 
-  combinations <- mapply(word_combos,words1, words2, SIMPLIFY = F )
+  combinations <- mapply(word_combos, words1, words2, SIMPLIFY = F )
 
   combinations <- lapply(combinations, function(x) dplyr::if_else(x == " ", empty_replacement,x))
 
@@ -271,7 +271,7 @@ word_vec_combos <- function(string1, string2, empty_replacement = "XXXXXXXXXXXXX
 #'
 normalise_words <- function(string_vec) {
 
-   words <- strsplit(string_vec, "\\s+")
+  words <- strsplit(string_vec, "\\s+")
 
   # Remove Latin special characters
   normalised_words <- lapply(words, stringi::stri_trans_general, "Latin-ASCII")
@@ -282,24 +282,20 @@ normalise_words <- function(string_vec) {
   # Change all the first letters to upper case
   sentesized_words <- lapply(normalised_words, stringr::str_to_title)
 
-
   # Final list of words
-  final_words <- mapply(c, normalised_words, lowerised_words, sentesized_words, SIMPLIFY=FALSE)
+  final_words <- mapply(c, words, normalised_words, lowerised_words, sentesized_words, SIMPLIFY=FALSE)
 
   # Remove punctuation
-  final_words   <- lapply(final_words,
-                          stringr::str_remove_all, '[[:punct:]]+')
+  final_words <- lapply(final_words,
+                        stringr::str_remove_all, '[[:punct:]]+')
 
+  # add word boundary markers
[…]

### 2024-06-19 tests/testthat/test-convert.R
Make paths in conversion test relative

@@ -1,18 +1,17 @@
 test_that("convert function works", {
-setwd("C:/Users/skubilay/Desktop/mintEMU")
   # Test input
-  input_text <- "C:/Users/skubilay/Desktop/mintEMU/analysis/data/raw_data/2007-Umemura.pdf"
+  input_text <- here::here("analysis", "data", "raw_data", "2007-Umemura.pdf")
 
   # Expected output
-  expected_output <- "regional metropolis and public transport connection"
+  expected_output <- strsplit("re-generating the shrinking japanese city urban acupuncture junya umemura mentor luisa calabrese", split = " ") |>
+    unlist() |> sort()
 
   # Run the function
-  output <- convert_pdf_text(input_text)
-  matching_excerpt <-stringr::str_detect(output, expected_output)
+  output <- convert_pdf_text(list(input_text), start_pages = "1", end_pages = "1") |> mintEMU::clean_basic() |> strsplit(split = " ") |>
+    unlist() |> sort()
+  # matching_excerpt <- stringr::str_detect(output, expected_output)
 
   # Check the output against the expected result
-  expect_true(matching_excerpt, expected_output,
-              info = "Output should be a lowercase string with no punctuation, extra spaces or numbers."
-  )
+  expect_in(output, expected_output)
 
 })

### 2024-08-07 analysis/paper/paper.qmd
Update email in yaml header

@@
     correspondence: "yes"
-    email: C.Forgaci@tudelft.nl
+    email: c.forgaci@tudelft.nl
     orcid: 0000-0003-3218-5102

### 2024-08-07 R/analyse.R
Update topic overivew functions

@@ -259,58 +259,43 @@ get_tf_idf <- function(data, id_col = NULL, word_col = NULL) {
 
 #' Get topic proportions in an LDA model
 #'
-#' @param lda LDA model object
+#' Calculate the mean probability of each topic over all documents.
 #'
-#' @return Data frame with topics and their proportions
+#' @param theta Matrix with each row representing the probability
+#'              distribution of a document over topics
+#' @param digits Number of digits to round the results
+#'
+#' @return Data frame with topic ids and their proportions in the corpus
 #' @export
-get_topic_proportions <- function(lda, digits = 2) {
-
-  topic_prob <- modeltools::posterior(lda)$topics
-  term_prob <- modeltools::posterior(lda)$terms
-
-  ndocs <- length(unique(rownames(topic_prob)))
-
-  # mean probabilities over all documents
-  topic_proportions <- colSums(topic_prob) / ndocs
-  # assign topic names
-  names(topic_proportions) <- mintEMU::name_topics(term_prob)
-  # show summed proportions in decreased order
-  sorted_topic_proportions <- sort(topic_proportions, decreasing = TRUE) |>
-    round(digits = digits)
-  data.frame(topic = names(sorted_topic_proportions),
-             proportion = sorted_topic_proportions,
+get_topic_proportions <- function(theta, digits = 2) {
+  ndocs <- length(rownames(theta))
+  topic_proportions <- round(colSums(theta) / ndocs, digits)
+  data.frame(topic_id = 1:ncol(theta),
+             prop = topic_proportions,
              row.names = NULL)
 }
 
-#' Count documents per main topic
+#' Count documents by main topic
 #'
-#' @param lda LDA model
+#' For each topic, count how many documents have it as the main topic.
 #'
-#' @return Data frame with topics and count of documents per main topic
+#' @param theta Matrix with each row representing the probability
+#'              distribution of a document over topics
+#'
+#' @return Data frame with topics and count of documents per each topic
 #' @export
-count_docs_per_topic <- function(lda) {
-
-  topic_prob <- modeltools::posterior(lda)$topics
-  term_prob <- modeltools::posterior(lda)$terms
-
-  k <- ncol(topic_prob)
-  ndocs <- length(unique(rownames(topic_prob)))
-
-  counts_of_primary_topics <- rep(0, k)
-  names(counts_of_primary_topics) <- name_topics(term_prob)
[…]

### 2024-08-07 R/analyse.R
Add corpus overview functions in separate file

@@ -1,22 +1,3 @@
-#' Get word counts for corpus
-#'
-#' @param data A data frame comprising a corpus of texts
-#' @param text_col Name of the column in which the text is stored
-#'
-#' @return A numeric vector of word counts, one for each document in the corpus
-#' @export
-get_word_counts <- function(data, text_col = NULL) {
-  text_col <- rlang::sym(text_col)
-
-  word_counts <- vector(mode = "numeric", length = nrow(data))
-  for (i in 1:nrow(data)) {
-    word_counts[i] <-
-      tidytext::unnest_tokens(tbl = data[i,], output = word, input = !!text_col) |> nrow()
-  }
-
-  word_counts
-}
-
 #' Extract top n words with the highest count per document
 #'
 #' Given a data frame with a column storing the words found in a document

### 2024-08-07 R/corpus.R

@@ -1,0 +1,43 @@
+#' Get word counts for corpus
+#'
+#' @param data A data frame comprising a corpus of texts
+#' @param text_col Name of the column in which the text is stored
+#'
+#' @return A numeric vector of word counts, one for each document in the corpus
+#' @export
+get_word_counts <- function(data, text_col = NULL) {
+  text_col <- rlang::sym(text_col)
+
+  word_counts <- vector(mode = "numeric", length = nrow(data))
+  for (i in 1:nrow(data)) {
+    word_counts[i] <-
+      tidytext::unnest_tokens(tbl = data[i,], output = word, input = !!text_col) |> nrow()
+  }
+
+  word_counts
+}
+
+#' Get overview of theses in the corpus
+#'
+#' @param data A data frame comprising counts of theses included in the corpus
+#'
+#' @return A data frame with an overview of the theses
+#' @export
+get_theses_overview <- function(data) {
+  theses_overview <- data |>
+    dplyr::select(grad_year, n_total, n_permissions) |>
+    dplyr::mutate(
+      grad_year = as.factor(grad_year),
+      `% of theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
+    dplyr::bind_rows(
+      data.frame(
+        grad_year = "2007-2021",
+        n_permissions = nrow(emu),
+        n_total = sum(theses_all_per_year$n_total),
+        `% of theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
+        check.names = FALSE)) |>
+    dplyr::filter(!is.na(grad_year)) |>
+    dplyr::rename(`No. theses` = n_total,
+                  `Graduation year` = grad_year,
+                  `No. theses included` = n_permissions)
+}

### 2024-08-07 R/analyse.R
Update topic naming function

@@ -165,14 +165,14 @@ fit_lda <- function(dtm, k, ...) {
 #'
 name_topics <- function(beta, n_words = 5) {
 
-  topic_names <- c()
+  topic_pseudonames <- c()
 
   for (i in 1:nrow(beta)) {
     name <- paste(names(head(sort(beta[i, ], decreasing = TRUE), n_words)), collapse = " ")
-    topic_names <- c(topic_names, name)
+    topic_pseudonames <- c(topic_pseudonames, name)
   }
 
-  topic_names
+  tibble(topic_id = 1:nrow(beta), topic_pseudoname = topic_pseudonames)
 }
 
 #' Get number of topics for LDA model empirically

### 2025-02-03 analysis/paper/paper.qmd
Update abstract

@@
 abstract: |
-  This paper aims to describe the legacy of the European post-Master in Urbanism 
-  (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
-  analysis of the graduation theses produced over the duration of the program.
-  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
-  and evolutions thereof, found in the full texts of the theses. Results reveal 5 topics
-  that are to a large extent consistent with the set-up of the program in two thematic 
-  semesters at TU Delft and three possible exchange semester locations, each characterised 
-  by a specific approach. The paper discusses the relevance of the findings for the
-  state of the profession and education in the field. Set up as a research
-  compendium accompanied by a published FAIR dataset and an interactive dashboard, 
-  the research is fully reproducible and enables further analysis. 
+  After seventeen years of being part of the the European post-Master in 
+  Urbanism (EMU), Delft Univerisity of Technology (TU Delft) left the program in
+  2021. This paper aims to describe the legacy of TU Delft's contribution to the
+  EMU and the profession. It reports on a topic model of graduation theses
+  produced during the program. Results reveal topics that are consistent with
+  the set-up of the program in two thematic semesters and three possible
+  exchange semester locations. The paper discusses the relevance of the findings
+  for the state of the profession and education in the field.
+# TODO check sentences on results and discussion
+# TODO JPER abstracts should generally be no longer than 125 words and have one sentence to
+# answer each of the following questions:
+# 1. What is the issue addressed and why is it important for planning theory or
+# practice?
+# 2. What gaps or misunderstandings in the literature does the research fill or correct?
+# 3. How was the research conducted?
+# 4. What are the research findings?
+# 5. What are the conclusions for planning theory or practice and the implications for
+# further research? 
 keywords: |
